### 2022-08-03 Manuscript/manuscript.tex
add manuscript

\documentclass[aip,jcp,reprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,wrapfig,txfonts,siunitx}
\usepackage[version=4]{mhchem}
%\usepackage{natbib}
\bibliographystyle{achemso}

\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\alert}[1]{\textcolor{black}{#1}}
\usepackage[normalem]{ulem}
\newcommand{\fk}[1]{\textcolor{blue}{#1}}
\newcommand{\titou}[1]{\textcolor{red}{#1}}
\newcommand{\trashPFL}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\PFL}[1]{\titou{(\underline{\bf PFL}: #1)}}
\newcommand{\toto}[1]{\textcolor{green}{#1}}
\newcommand{\trashAS}[1]{\textcolor{green}{\sout{#1}}}
\newcommand{\AS}[1]{\toto{(\underline{\bf AS}: #1)}}
\newcommand{\ant}[1]{\textcolor{orange}{#1}}
\newcommand{\SupInf}{\textcolor{blue}{Supporting Information}}

\newcommand{\mc}{\multicolumn}
\newcommand{\fnm}{\footnotemark}
\newcommand{\fnt}{\footnotetext}
\newcommand{\tabc}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\QP}{\textsc{quantum package}}

\newcommand{\EHF}{E_\text{HF}}
\newcommand{\EDOCI}{E_\text{DOCI}}
\newcommand{\EFCI}{E_\text{FCI}}
\newcommand{\Ndet}{N_\text{det}}
\newcommand{\Nbas}{N}

\usepackage[
	colorlinks=true,
    citecolor=blue,
    breaklinks=true
	]{hyperref}
\urlstyle{same}

\begin{document}

\newcommand{\LCPQ}{Laboratoire de Chimie et Physique Quantiques (UMR 5626), Universit\'e de Toulouse, CNRS, UPS, France}

\title{Title}

\author{F\'abris Kossoski}
\email{fkossoski@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}
\author{Pierre-Fran\c{c}ois Loos}
\email{loos@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}

% Abstract
\begin{abstract}
{\bf Abstract:} 
\bigskip
\begin{center}
        \boxed{\includegraphics[keepaspectratio,width=2in]{TOC}}
\end{center}
\bigskip
\end{abstract}

% Title
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Introduction}
%\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Computational details}
%\label{sec:compdet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Results and discussion}
%\label{sec:res}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusion and perspectives}
%\label{sec:ccl}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgements}
This work was performed using HPC resources from CALMIP (Toulouse) under allocation 2021-18005.
This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~863481).
\end{acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Supporting information available}
\label{sec:SI}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section*{Data availability statement}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%The data that support the findings of this study are openly available in Zenodo at \href{http://doi.org/XX.XXXX/zenodo.XXXXXXX}{http://doi.org/XX.XXXX/zenodo.XXXXXXX}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{manuscript}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2022-08-04 CN_cc-pvdz/run_stat.sh
update

@@ -3,7 +3,7 @@
 norbitals=28
 nelectrons=13
 #xcut=8.0
-xcut=2.1
+xcut=2.8
 
 if [ -f "input" ]; then
   rm "input"

### 2022-08-04 vinyl_cc-pvdz/run_stat.sh

@@ -2,7 +2,7 @@
 
 norbitals=43
 nelectrons=15
-xcut=2.45
+xcut=2.9
 #xcut=12.0
 
 if [ -f "input" ]; then

### 2023-05-31 Manuscript/manuscript.tex
starting manuscript

@@
 
-\title{Title}
+\title{Seniority and hierarchy configuration interaction for radicals and excited states}
 
@@
 {\bf Abstract:} 
-\bigskip
-\begin{center}
-        \boxed{\includegraphics[keepaspectratio,width=2in]{TOC}}
-\end{center}
-\bigskip
+%\bigskip
+%\begin{center}
+%        \boxed{\includegraphics[keepaspectratio,width=2in]{TOC}}
+%\end{center}
+%\bigskip
 \end{abstract}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Introduction}
-%\label{sec:intro}
+\section{Introduction}
+\label{sec:intro}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+Seniority based methods.
+Limited studies for radicals and excited states
+
+Hierarchy CI \cite{Kossoski_2022}
+Similarly, only for closed-shell systems.
+
+We introduce single-reference CI approximations built on top of an open-shell reference.
+Seniority and hCI.
+How do these methods perform for ground state radicals?
+Impact of perturbative correction.
+
+Ground-state based CI vs. state-specific CI \cite{Kossoski_2023}
+Promising results for state-specific excitation-based CI.
+We introduce analogous state-specific seniority-based CI ($\Delta$sCI) and hierarchy-based CI ($\Delta$hCI).
+$\Delta$sCI2 
+$\Delta$hCI1
+How do these methods perform for excited states, from both closed-shell and open-shell systems?
+
+%For an arbitrary reference Slater determinant, with seniority $s_0$, the hierarchy $h$ associated to a another determinant
+With respect to a given reference Slater determinant, the hierarchy $h$ of a target determinant is defined as
+\begin{equation}
+        \label{eq:h}
+        %h = \frac{e+\Delta s/2}{2},
+	h = \frac{e+ (s-s_0)/2}{2},
+\end{equation}
+where $s$ and $s_0$ denote the seniority of target and reference determinants, respectively, and $e$ represents the excitation degree that connect the two determinants.
+This definition renders a degree of dissimilarity between two determinants (reference and target),
+which accounts for differences in orbital occupation (through the excitation degree $e$)
+and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
+The latter always assumes an integer value (for both even and odd numbers of electrons), such that $h$ becomes an integer or half-integer.
+Also, Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant, when $s_0 = 0$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Computational details}
-%\label{sec:compdet}
+\section{Computational details}
+\label{sec:compdet}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To gauge the performance of hCI against excitation-based and seniority-based CI for radical systems,
+we have calculated the potential energy curves (PECs) for the dissociation of four radical systems:
+\ce{OH}, \ce{CN}, vinyl (\ce{C2H3}), and \ce{H7}.
+%which display a variable number of bond breaking.
+The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020},
+%also reproduced in the \SupInf,
+and the PECs were computed along the \ce{C=C} double bond breaking coordinate, with the remaining internal coordinates kept frozen.
+For \ce{H7}, we considered equally spaced and linearly arranged hydrogen atoms, and the PECs were computed along the symmetric dissociation coordinate.
+From the PECs, we performed a similar analysis as in our paper on hCI for closed-shell systems. \cite{Kossoski_2022}
+Namely, for the different CI levels considered here, 
+we evaluated the convergence of the non-parallelity error (NPE), the distance error, the vibrational frequencies, and the equilibrium geometries, as functions of $\Ndet$.
+The NPE (distance error) of a given level of theory is defined as the maximum minus (plus) the minimum differences between its corresponding PEC and the FCI PEC.
+Details about how the vibrational frequencies and equilibrium geometries were obtained from the calculated PECs can be found in the \SupInf.
+
+To evaluate how the different levels of CI introduced here perform for excited states,
+we calculated vertical excitation energies for 60 electronic states, comprising 26 different molecular systems, with geometries extracted from the QUEST database. \cite{Veril_2021}
+% 18 closed-shells, 8 radicals
+We employed the aug-cc-pVDZ basis set for systems having up to three non-hydrogen atoms and the 6-31+G(d) basis set for the larger ones.
+We performed calculations following both the standard ground-state based CI route and the state-specific CI route. \cite{Kossoski_2023}
+For the latter, we employed the state-specific orbitals obtained in Ref.~\onlinecite{Kossoski_2023}.
+The computed excitation energies were compared against the reference values from the QUEST database. \cite{Veril_2021}
+The full set of excited states and calculated excitation energies, for the various methods considered here, are provided in the {\SupInf}.
+%For completeness, we also reproduce the geometries in the \SupInf.
+
+The hCI methods introduced here were implemented in {\QP} \cite{Garniron_2019} through a straightforward modification of the
+\textit{configuration interaction using a perturbative selection made iteratively} (CIPSI) algorithm. \cite{Huron_1973,Giner_2013,Giner_2015,Garniron_2018}
+By allowing only for the determinants that are connected with the reference determinant(s) up to a given maximum hierarchy $h$,
+the CIPSI algorithm is restricted to the truncated Hilbert space defined by the reference and the value of $h$.
+{\QP} \cite{Garniron_2019} was employed to perform all the CI calculations presented here.
+In a given calculation, the energies are considered to be converged when the (largest) Epstein-Nesbet second-order perturbation correction computed in the truncated Hilbert space 
+lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
+%by spanning the most important regions of the truncated Hilbert space
+This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space,
+while delievering fairly converged absolute and excitation energies.
+The ground- and excited-state CI energies are obtained with the Davidson iterative algorithm. \cite{Davidson_1975}
+
+%restricted open-shell HF orbitals
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results and discussion}
+\label{sec:res}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Results and discussion}
-%\label{sec:res}
+\subsection{hCI for radicals}
+\label{sec:res_A}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+PECs
+NPEs, distance errors
+equilibrium geometries and vibrational frequencies
+
+HF orbitals
+orbital optimization?
+
+perturbative correction for closed- and open-shell systems
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Conclusion and perspectives}
-%\label{sec:ccl}
+\subsection{hCI for excited states}
+\label{sec:res_B}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+hCI1/HF
+hCI1.5/HF
+hCI2/HF
+$\Delta$hCI1
+$\Delta$hCI1.5
+$\Delta$hCI2
+
+Higher-order levels for a few systems
+hCI2.5/HF
+hCI3/HF
+$\Delta$hCI2.5
+$\Delta$hCI3
+$\Delta$hCI3.5
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Seniority CI for excited states}
+\label{sec:res_C}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+sCI2
+$\Delta$sCI2
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+\label{sec:ccl}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-06-01 Manuscript/manuscript.tex
saving work

@@
 
+From here on, method names carrying the $\Delta$ symbol denote a state-specific approach,
+whereas those without the $\Delta$ mean that ground-state HF orbitals were employed.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-hCI1/HF
-hCI1.5/HF
-hCI2/HF
-$\Delta$hCI1
-$\Delta$hCI1.5
-$\Delta$hCI2
+In Tab.~\ref{tab:1} we present the statistical errors of the different methods, organized by their computational scaling.
+We found that $\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat more favourable MSE and RMSE of the latter.
+$\Delta$hCI2 still tends to underestimate the excitation energies, though to a much lesser extent than observed for the lower-order hCI methods.
 
-Higher-order levels for a few systems
-hCI2.5/HF
-hCI3/HF
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
+for the Set of 60 Excitation Energies Listed in the {\SupInf}.
+}
+\label{tab:1}
+\begin{ruledtabular}
+\begin{tabular}{lddd}
+method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
+\hline
+$\Delta$CSF       & -0.54 & 0.70 & 0.86 \\
+CIS               & +0.13 & 0.59 & 0.77 \\
+\hline
+hCI1              & +1.11 & 1.14 & 1.39 \\
+$\Delta$hCI1      & -0.90 & 1.17 & 1.40 \\
+\hline
+hCI1.5            & +1.88 & 1.88 & 2.20 \\
+$\Delta$hCI1.5    & -1.99 & 2.12 & 2.58 \\
+\hline
+$\Delta$CISD      & -0.08 & 0.17 & 0.22 \\
+$\Delta$hCI2      & -0.15 & 0.21 & 0.30 \\
+\hline
+sCI2              &  0.96 & 1.06 & 1.28 \\
+$\Delta$sCI2      &  0.39 & 0.60 & 0.73 \\
+%\hline
+%CIS(D)            & -0.03 & 0.35 & 0.40 \\
+%CC2               & -0.05 & 0.32 & 0.37 \\
+%CCSD              & +0.01 & 0.08 & 0.09 \\
+%CC3               & +0.01 & 0.03 & 0.06 \\
+%CCSDT             & -0.01 & 0.02 & 0.02 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+Despite the rather poor performance of lower-order hCI methods for excitation energies,
+we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
+For that, we performed calculations for a subset of 14 excited states of small molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
+with corresponding statistical errors shown in Tab.~\ref{tab:2}.
+Even though this is a small set, it should be enough to obtain the main trends on the statistics of the different methods.
+
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
+for the Set of 14 Excitation Energies Listed in the {\SupInf}.
+}
+\label{tab:2}
+\begin{ruledtabular}
+\begin{tabular}{lddd}
+method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
+\hline
+%CIS               & +0.16 & 0.82 & 0.96  \\
+%$\Delta$CSF       & -0.84 & 0.91 & 1.03  \\
+%\hline
+CISD              & +4.09 & 4.09 & 4.18  \\
+hCI2              & +2.89 & 3.00 & 3.28  \\
+$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
+$\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
+\hline
+hCI2.5            & +1.93 & 1.93 & 2.05  \\
+$\Delta$hCI2.5    & -0.27 & 0.27 & 0.30  \\
+\hline
+CISDT             & +0.09 & 0.18 & 0.20  \\
+hCI3              & +0.18 & 0.18 & 0.19  \\
+$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
+$\Delta$hCI3      & -0.21 & 0.21 & 0.24  \\
+\hline
+$\Delta$hCI3.5    & -0.08 & 0.08 & 0.09  \\
+\hline
+CISDTQ            & +0.14 & 0.14 & 0.16  \\
+$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+We found that the errors of both ground-state based hCI and state-specific $\Delta$hCI to systematically decrease when going toward higher orders of $h$, which would be expected.
+%Unfortunately, the pace of decrease is too slow to render these approaches very useful.
+%In line with what has been found for excitation-based CI \cite{Kossoski_2023},
+
+The case $h=2$ was discussed above for the larger set of excitation energies.
+Here we just notice that hCI2 is found to be somewhat more accurate than CISD, even though the errors are still too large.
+However, this improvement does not reflect on the state specific route (compare $\Delta$hCI2 and $\Delta$CISD in Tabs.~\ref{tab:1} and ~\ref{tab:2}).
+For $h=2.5$, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
+it is actually less accurate then $\Delta$hCI2, and therefore not useful.
+Moving to $h=3$, we
+
+hCI2.5
+hCI3
 $\Delta$hCI2.5

### 2023-06-01 Manuscript/manuscript.tex
saving work

@@
 
-We found that the errors of both ground-state based hCI and state-specific $\Delta$hCI to systematically decrease when going toward higher orders of $h$, which would be expected.
+%We found that the errors of both ground-state based hCI and state-specific $\Delta$hCI to systematically decrease when going toward higher orders of $h$, which would be expected.
 %Unfortunately, the pace of decrease is too slow to render these approaches very useful.
@@
 
-The case $h=2$ was discussed above for the larger set of excitation energies.
-Here we just notice that hCI2 is found to be somewhat more accurate than CISD, even though the errors are still too large.
+The case of $\Delta$hCI2 was discussed above for the larger set of excitation energies.
+Here we further notice that hCI2 was found to be somewhat more accurate than CISD, even though the errors are still too large.
 However, this improvement does not reflect on the state specific route (compare $\Delta$hCI2 and $\Delta$CISD in Tabs.~\ref{tab:1} and ~\ref{tab:2}).
-For $h=2.5$, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
-it is actually less accurate then $\Delta$hCI2, and therefore not useful.
-Moving to $h=3$, we
-
-hCI2.5
-hCI3
-$\Delta$hCI2.5
-$\Delta$hCI3
-$\Delta$hCI3.5
+Moving one step further, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
+it is actually less accurate then $\Delta$hCI2.
+An improved performance is found for the methods displaying an $\order*{N^8}$ computational scaling.
+The four methods, hCI3, $\Delta$hCI3, CISDT, and $\Delta$CISDT, display quite similar MAEs and RMSEs.
+Therefore, there is no advantage by moving from excitation-based to hierarchy-based CI at this order, just as discussed above for the $\Delta$hCI2 and $\Delta$CISD.
+Similarly, the $\Delta$hCI3 does not improve with respect to hCI3, as also observed for $\Delta$CISDT and CISDT. \cite{Kossoski_2023}
+We notice, however, the negative MSEs for the state-specific methods, in contrast to the positives values obtained for the ground-state based route.
+At the next order, $\Delta$hCI3.5 has significantly smaller errors, but at a costly $\order*{N^9}$ scaling.
 
@@
 
-sCI2
-$\Delta$sCI2
+We refer back to Tab.~\ref{tab:1} to discuss the performance of seniority-based CI for excited states.
+sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
+This improvement, however, is not as expressive as typically observed for excitataion-based and hierarchy-based methods.
+(Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:2}).
+The large errors, even at the state-specific route, combined with the formal $\order*{e^N}$ scaling of these methods,
+render them unsuitable for calculations of accurate excitation energies.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We conclude that 
+Meanwhile, 
+, and therefore not useful.
 

### 2023-06-02 Manuscript/manuscript.tex
saving work

@@
 
-Seniority based methods.
+Configuration interaction (CI) offers a systematic way to solve the many-body electronic structure problem.
+By including progressively more determinants in the Hilbert space, the wave function becomes increasingly closer to the exact one, and so does the electronic energy.
+In full CI (FCI), all determinants are accounted for and the problem is solved exactly (for a given basis set).
+In practice, however, one resorts to approximate CI methods, where only the determinants that satisfy a given criterion are included in the truncated Hilbert space.
+For small systems, FCI is still attainable and one can gauge it against approximate CI methods to assesss their performance.
+
+Excitation-based CI
+Seniority based CI (sCI)
 Limited studies for radicals and excited states
+We have recently introduced hierarchy CI (hCI), \cite{Kossoski_2022}
+where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$.
+For different properties and molecular systems, hCI showed an improved convergence to the FCI results when compared to the traditional excitation-based CI. \cite{Kossoski_2022}
+However, hCI has only been defined for a closed-shell reference, thus systems with an even number of electrons.
+A generalization of the method for radicals has not yet been proposed.
 
-Hierarchy CI \cite{Kossoski_2022}
-Similarly, only for closed-shell systems.
+The higher-lying CI roots give access to excited states.
+Whereas the performance of excitation-based CI for excited states is well stablished, \cite{} this remains an open question for sCI methods,
+despite the growing number of methods targeting excited states that exploit the concept of seniority number. \cite{} 
+Similarly, the favourable results of hCI when compared to excitation-based CI for ground states raises the question on how well it can describe excitation energies.
 
-We introduce single-reference CI approximations built on top of an open-shell reference.
-Seniority and hCI.
-How do these methods perform for ground state radicals?
-Impact of perturbative correction.
+When targeting excited states with CI methods, one can employ the ground-state HF orbitals or instead optimize the orbitals for each state of interest, in a so-called state-specific approach.
+Promising results for state-specific excitation-based CI.
+\cite{Kossoski_2023}
+Here we introduce analogous state-specific seniority-based CI ($\Delta$sCI) and hierarchy-based CI ($\Delta$hCI).
+How do these methods perform for excited states, from both closed-shell and open-shell systems?
 
-Ground-state based CI vs. state-specific CI \cite{Kossoski_2023}
-Promising results for state-specific excitation-based CI.
-We introduce analogous state-specific seniority-based CI ($\Delta$sCI) and hierarchy-based CI ($\Delta$hCI).
-$\Delta$sCI2 
-$\Delta$hCI1
-How do these methods perform for excited states, from both closed-shell and open-shell systems?
+Here we address the knowledge gaps outlined above by setting four goals.
+Our first goal is to generalize hCI \cite{Kossoski_2022} for the case of an open-shell reference determinant, thus extending its domain to radical species.
+The second is to assess the performance of this method for ground state radicals, similarly to what has been done for closed-shell systems. \cite{Kossoski_2022}
+%Seniority hCI
+The third goal is to assess the improvement from perturbative corrections to ground-state radicals and closed-shell systems, for both excitation-based and hierarchy-based CI methods.
+Finally, we introduce and gauge seniority-based and hierarchy-based CI methods for excited states, along both the standard ground-state based route and the state-specific route. \cite{Kossoski_2023}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Hierarchy configuration interaction}
+\label{sec:hCI}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+state-specific methods
+$\Delta$sCI2 
+$\Delta$hCI1
 %restricted open-shell HF orbitals

### 2023-06-04 CN_cc-pvdz/rpt2/run_stat.sh
update hCI for ground states

@@ -2,7 +2,7 @@
 
 norbitals=28
 nelectrons=13
-xcut=7.0
+xcut=5.5
 
 if [ -f "input" ]; then
   rm "input"

### 2023-06-04 F2_cc-pvdz/plot_pes_hCI.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+gnuplot plot_pes_hCI.gnu
+epspdf plot_pes_hCI.eps
+okular plot_pes_hCI.pdf

### 2023-06-04 F2_cc-pvdz/rpt2/cp_to_manuscript.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+mol='f2'
+
+path='/home/fabris/ongoing_projects/seniority/Manuscript'
+
+cp plot_pes.pdf   $path/${mol}_pes.pdf
+cp plot_error.pdf $path/${mol}_pes_error.pdf
+cp plot_stat.pdf  $path/${mol}_pes_stat.pdf
+cp xe.pdf         $path/${mol}_xe.pdf

### 2023-06-04 F2_cc-pvdz/rpt2/fit_eq.sh

@@ -1,0 +1,295 @@
+#!/bin/bash
+
+#######################################################################
+
+if [ -f "fit.log" ]; then
+  rm "fit.log"
+fi
+
+gnuplot fit_eq.gnu
+
+#######################################################################
+grep ^a fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/a0 /FCI      /g' |
+        sed 's/a1 /RHF      /g' |
+        sed 's/a2 /hfCISD   /g' |
+        sed 's/a3 /hfCISDT  /g' |
+        sed 's/a4 /hfCISDTQ /g' |
+        sed 's/a5 /hfCIs0   /g' |
+        sed 's/a6 /hfCIs2   /g' |
+        sed 's/a7 /hfCIs4   /g' |
+        sed 's/a8 /hfCIo1   /g' |
+        sed 's/a9 /hfCIo1.5 /g' |
+        sed 's/a10/hfCIo2   /g' |
+        sed 's/a11/hfCIo2.5 /g' |
+        sed 's/a12/hfCIo3   /g' |
+        sed 's/a13/hfCIo3.5 /g' |
+        sed 's/a14/hfCIo4   /g' |
+        sed 's/a15/ooCIS    /g' |
+        sed 's/a16/ooCISD   /g' |
+        sed 's/a17/ooCISDT  /g' |
+        sed 's/a18/ooCIs0   /g' |
+        sed 's/a19/ooCIs2   /g' |
+        sed 's/a20/ooCIo1   /g' |
+        sed 's/a21/ooCIo1.5 /g' |
+        sed 's/a22/ooCIo2   /g' |
+        sed 's/a23/ooCIo2.5 /g' |
+        sed 's/a24/ooCIo3   /g' > a.dat
+
+grep ^D fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/D0 /FCI      /g' |
+        sed 's/D1 /RHF      /g' |
+        sed 's/D2 /hfCISD   /g' |
+        sed 's/D3 /hfCISDT  /g' |
[…]

### 2023-06-04 F2_cc-pvdz/rpt2/rename_pes_files.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+
+if false;
+then
+
+mv pes_e2_pt2.dat pes_CISD_pt2.dat
+mv pes_e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_e4_pt2.dat pes_CISDTQ_pt2.dat
+
+mv pes_e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+fi
+
+if true;
+then
+
+mv pes_s4.e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_s6.e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_s8.e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+mv pes_s4.e2_pt2.dat pes_CISD_pt2.dat
+mv pes_s6.e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_s8.e4_pt2.dat pes_CISDTQ_pt2.dat
+
+fi
+
+if true;
+then
+
+mv pes_g4_pt2.dat pes_CIo1_pt2.dat
+mv pes_g6_pt2.dat pes_CIo1.5_pt2.dat
+mv pes_g8_pt2.dat pes_CIo2_pt2.dat
+mv pes_g10_pt2.dat pes_CIo2.5_pt2.dat
+mv pes_g12_pt2.dat pes_CIo3_pt2.dat
+mv pes_g14_pt2.dat pes_CIo3.5_pt2.dat
+mv pes_g16_pt2.dat pes_CIo4_pt2.dat
+
+mv pes_g4_rpt2.dat pes_CIo1_rpt2.dat
+mv pes_g6_rpt2.dat pes_CIo1.5_rpt2.dat
+mv pes_g8_rpt2.dat pes_CIo2_rpt2.dat
+mv pes_g10_rpt2.dat pes_CIo2.5_rpt2.dat
+mv pes_g12_rpt2.dat pes_CIo3_rpt2.dat
+mv pes_g14_rpt2.dat pes_CIo3.5_rpt2.dat
+mv pes_g16_rpt2.dat pes_CIo4_rpt2.dat
+
+fi

### 2023-06-04 F2_cc-pvdz/rpt2/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=26
+nelectrons=14
+xcut=8.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 F2_cc-pvdz/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=26
+nelectrons=14
+xcut=8.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 H4_cc-pvdz/plot_eq.sh

@@ -1,0 +1,23 @@
+#!/bin/bash
+
+grep ^xe fit.log | grep = | sed 's/=//g' | cut -c -30 | 
+	sed 's/xe0 /FCI   /g' |
+	sed 's/xe1 /RHF   /g' |
+	sed 's/xe2 /SD    /g' |
+	sed 's/xe3 /SDT   /g' |
+	sed 's/xe4 /SDTQ  /g' |
+	sed 's/xe5 /o1    /g' |
+	sed 's/xe6 /o1.5  /g' |
+	sed 's/xe7 /o2    /g' |
+	sed 's/xe8 /o2.5  /g' |
+	sed 's/xe9 /o3    /g' |
+	sed 's/xe10/o3.5  /g' |
+	sed 's/xe11/o4    /g' |
+	sed 's/xe12/ooSD  /g' |
+	sed 's/xe13/ooSDT /g' |
+	sed 's/xe14/ooo1  /g' |
+	sed 's/xe15/ooo1.5/g' |
+	sed 's/xe16/ooo2  /g' |
+	sed 's/xe17/ooo2.5/g' |
+	sed 's/xe18/ooo3  /g' #|
+

### 2023-06-04 H4_cc-pvdz/rpt2/cp_to_manuscript.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+mol='h4'
+
+path='/home/fabris/ongoing_projects/seniority/Manuscript'
+
+cp plot_pes.pdf   $path/${mol}_pes.pdf
+cp plot_error.pdf $path/${mol}_pes_error.pdf
+cp plot_stat.pdf  $path/${mol}_pes_stat.pdf
+cp xe.pdf         $path/${mol}_xe.pdf

### 2023-06-04 H4_cc-pvdz/rpt2/rename_pes_files.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+
+mv pes_e2_pt2.dat pes_CISD_pt2.dat
+mv pes_e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_e4_pt2.dat pes_CISDTQ_pt2.dat
+
+mv pes_e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+mv pes_g4_pt2.dat pes_CIo1_pt2.dat
+mv pes_g6_pt2.dat pes_CIo1.5_pt2.dat
+mv pes_g8_pt2.dat pes_CIo2_pt2.dat
+mv pes_g10_pt2.dat pes_CIo2.5_pt2.dat
+mv pes_g12_pt2.dat pes_CIo3_pt2.dat
+mv pes_g14_pt2.dat pes_CIo3.5_pt2.dat
+mv pes_g16_pt2.dat pes_CIo4_pt2.dat
+
+mv pes_g4_rpt2.dat pes_CIo1_rpt2.dat
+mv pes_g6_rpt2.dat pes_CIo1.5_rpt2.dat
+mv pes_g8_rpt2.dat pes_CIo2_rpt2.dat
+mv pes_g10_rpt2.dat pes_CIo2.5_rpt2.dat
+mv pes_g12_rpt2.dat pes_CIo3_rpt2.dat
+mv pes_g14_rpt2.dat pes_CIo3.5_rpt2.dat
+mv pes_g16_rpt2.dat pes_CIo4_rpt2.dat
+

### 2023-06-04 H4_cc-pvdz/rpt2/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=20
+nelectrons=4
+xcut=10.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 H4_cc-pvdz/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=20
+nelectrons=4
+xcut=10.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 H8_cc-pvdz/compute_NPE.pl

@@ -1,0 +1,125 @@
+#!/usr/bin/perl
+
+use warnings;
+use strict;
+use List::Util qw(min max);
+
+my $case = $ARGV[0];
+my $xcut = $ARGV[1];
+
+my(@files,@index,$detfile,$outfile);
+
+if($case==1){
+@files = qw( pes_rhf.dat pes_CIo1.dat pes_CIo1.5.dat pes_CIo2.dat pes_CIo2.5.dat pes_CIo3.dat pes_CIo3.5.dat pes_CIo4.dat);
+@index = qw( 0 1 1.5 2 2.5 3 3.5 4);
+$detfile = "det_CIo.dat";
+$outfile = "stat_CIo.dat";
+}
+elsif($case==2){
+@files = qw( pes_rhf.dat pes_ooCIo1.dat pes_ooCIo1.5.dat pes_ooCIo2.dat pes_ooCIo2.5.dat pes_ooCIo3.dat pes_ooCIo3.5.dat pes_ooCIo4.dat);
+@index = qw( 0 1 1.5 2 2.5 3 3.5 4);
+$detfile = "det_CIo.dat";
+$outfile = "stat_ooCIo.dat";
+}
+elsif($case==3){
+@files = qw( pes_rhf.dat pes_rhf.dat pes_CISD.dat pes_CISDT.dat pes_CISDTQ.dat);
+@index = qw( 0 1 2 3 4);
+$detfile = "det_CI.dat";
+$outfile = "stat_CI.dat";
+}
+elsif($case==4){
+@files = qw( pes_rhf.dat pes_ooCIS.dat pes_ooCISD.dat pes_ooCISDT.dat pes_ooCISDTQ.dat);
+@index = qw( 0 1 2 3 4);
+$detfile = "det_CI.dat";
+$outfile = "stat_ooCI.dat";
+}
+elsif($case==5){
+@files = qw( pes_s0.dat pes_s2.dat pes_s4.dat);
+@index = qw( 0 2 4);
+$detfile = "det_CIs.dat";
+$outfile = "stat_CIs.dat";
+}
+elsif($case==6){
+@files = qw( pes_ooCIs0.dat pes_ooCIs2.dat pes_ooCIs4.dat);
+@index = qw( 0 2 4);
+$detfile = "det_CIs.dat";
+$outfile = "stat_ooCIs.dat";
+}
+else{
+print "case variable not recognized \n";
+die;
+}
+
[…]

### 2023-06-04 H8_cc-pvdz/fit_eq.sh

@@ -1,0 +1,295 @@
+#!/bin/bash
+
+#######################################################################
+
+if [ -f "fit.log" ]; then
+  rm "fit.log"
+fi
+
+gnuplot fit_eq.gnu
+
+#######################################################################
+grep ^a fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/a0 /FCI      /g' |
+        sed 's/a1 /RHF      /g' |
+        sed 's/a2 /hfCISD   /g' |
+        sed 's/a3 /hfCISDT  /g' |
+        sed 's/a4 /hfCISDTQ /g' |
+        sed 's/a5 /hfCIs0   /g' |
+        sed 's/a6 /hfCIs2   /g' |
+        sed 's/a7 /hfCIs4   /g' |
+        sed 's/a8 /hfCIo1   /g' |
+        sed 's/a9 /hfCIo1.5 /g' |
+        sed 's/a10/hfCIo2   /g' |
+        sed 's/a11/hfCIo2.5 /g' |
+        sed 's/a12/hfCIo3   /g' |
+        sed 's/a13/hfCIo3.5 /g' |
+        sed 's/a14/hfCIo4   /g' |
+        sed 's/a15/ooCIS    /g' |
+        sed 's/a16/ooCISD   /g' |
+        sed 's/a17/ooCISDT  /g' |
+        sed 's/a18/ooCIs0   /g' |
+        sed 's/a19/ooCIs2   /g' |
+        sed 's/a20/ooCIo1   /g' |
+        sed 's/a21/ooCIo1.5 /g' |
+        sed 's/a22/ooCIo2   /g' |
+        sed 's/a23/ooCIo2.5 /g' |
+        sed 's/a24/ooCIo3   /g' > a.dat
+
+grep ^D fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/D0 /FCI      /g' |
+        sed 's/D1 /RHF      /g' |
+        sed 's/D2 /hfCISD   /g' |
+        sed 's/D3 /hfCISDT  /g' |
[…]

### 2023-06-04 H8_cc-pvdz/rpt2/cp_to_manuscript.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+mol='h8'
+
+path='/home/fabris/ongoing_projects/seniority/Manuscript'
+
+cp plot_pes.pdf   $path/${mol}_pes.pdf
+cp plot_error.pdf $path/${mol}_pes_error.pdf
+cp plot_stat.pdf  $path/${mol}_pes_stat.pdf
+cp xe.pdf         $path/${mol}_xe.pdf

### 2023-06-04 H8_cc-pvdz/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=40
+nelectrons=8
+xcut=10.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 HF_cc-pvdz/rename_pes_files.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+
+if false;
+then
+
+mv pes_e2_pt2.dat pes_CISD_pt2.dat
+mv pes_e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_e4_pt2.dat pes_CISDTQ_pt2.dat
+
+mv pes_e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+fi
+
+if false;
+then
+
+mv pes_s4.e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_s6.e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_s8.e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+mv pes_s4.e2_pt2.dat pes_CISD_pt2.dat
+mv pes_s6.e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_s8.e4_pt2.dat pes_CISDTQ_pt2.dat
+
+fi
+
+if true;
+then
+
+mv pes_g4_pt2.dat pes_CIo1_pt2.dat
+mv pes_g6_pt2.dat pes_CIo1.5_pt2.dat
+mv pes_g8_pt2.dat pes_CIo2_pt2.dat
+mv pes_g10_pt2.dat pes_CIo2.5_pt2.dat
+mv pes_g12_pt2.dat pes_CIo3_pt2.dat
+mv pes_g14_pt2.dat pes_CIo3.5_pt2.dat
+mv pes_g16_pt2.dat pes_CIo4_pt2.dat
+
+mv pes_g4_rpt2.dat pes_CIo1_rpt2.dat
+mv pes_g6_rpt2.dat pes_CIo1.5_rpt2.dat
+mv pes_g8_rpt2.dat pes_CIo2_rpt2.dat
+mv pes_g10_rpt2.dat pes_CIo2.5_rpt2.dat
+mv pes_g12_rpt2.dat pes_CIo3_rpt2.dat
+mv pes_g14_rpt2.dat pes_CIo3.5_rpt2.dat
+mv pes_g16_rpt2.dat pes_CIo4_rpt2.dat
+
+fi

### 2023-06-04 HF_cc-pvdz/rpt2/cp_to_manuscript.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+mol='hf'
+
+path='/home/fabris/ongoing_projects/seniority/Manuscript'
+
+cp plot_pes.pdf      $path/${mol}_pes.pdf
+cp plot_error.pdf    $path/${mol}_pes_error.pdf
+
+cp plot_stat.pdf     $path/${mol}_npe.pdf
+cp plot_distance.pdf $path/${mol}_distance.pdf
+cp freq.pdf          $path/${mol}_freq.pdf
+cp xe.pdf            $path/${mol}_xe.pdf

### 2023-06-04 HF_cc-pvdz/rpt2/plot.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+gnuplot plot.gnu
+epspdf plot.eps
+okular plot.pdf

### 2023-06-04 HF_cc-pvdz/rpt2/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=18
+nelectrons=8
+xcut=6.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 HF_cc-pvdz/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=18
+nelectrons=8
+xcut=6.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 N2_cc-pvdz/plot_r4.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+gnuplot plot_r4.gnu
+epspdf plot_r4.eps
+okular plot_r4.pdf

### 2023-06-04 N2_cc-pvdz/rename_pes_files.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+
+if true;
+then
+
+mv pes_e2_pt2.dat pes_CISD_pt2.dat
+mv pes_e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_e4_pt2.dat pes_CISDTQ_pt2.dat
+
+mv pes_e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+fi
+
+if false;
+then
+
+mv pes_s4.e2_rpt2.dat pes_CISD_rpt2.dat
+mv pes_s6.e3_rpt2.dat pes_CISDT_rpt2.dat
+mv pes_s8.e4_rpt2.dat pes_CISDTQ_rpt2.dat
+
+mv pes_s4.e2_pt2.dat pes_CISD_pt2.dat
+mv pes_s6.e3_pt2.dat pes_CISDT_pt2.dat
+mv pes_s8.e4_pt2.dat pes_CISDTQ_pt2.dat
+
+fi
+
+if true;
+then
+
+mv pes_g4_pt2.dat pes_CIo1_pt2.dat
+mv pes_g6_pt2.dat pes_CIo1.5_pt2.dat
+mv pes_g8_pt2.dat pes_CIo2_pt2.dat
+mv pes_g10_pt2.dat pes_CIo2.5_pt2.dat
+mv pes_g12_pt2.dat pes_CIo3_pt2.dat
+mv pes_g14_pt2.dat pes_CIo3.5_pt2.dat
+mv pes_g16_pt2.dat pes_CIo4_pt2.dat
+
+mv pes_g4_rpt2.dat pes_CIo1_rpt2.dat
+mv pes_g6_rpt2.dat pes_CIo1.5_rpt2.dat
+mv pes_g8_rpt2.dat pes_CIo2_rpt2.dat
+mv pes_g10_rpt2.dat pes_CIo2.5_rpt2.dat
+mv pes_g12_rpt2.dat pes_CIo3_rpt2.dat
+mv pes_g14_rpt2.dat pes_CIo3.5_rpt2.dat
+mv pes_g16_rpt2.dat pes_CIo4_rpt2.dat
+
+fi

### 2023-06-04 N2_cc-pvdz/rpt2/cp_to_manuscript.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+
+mol='n2'
+
+#path='/home/fabris/ongoing_projects/seniority/Manuscript'
+path='/home/fabris/seniority/Manuscript'
+
+cp plot_pes.pdf   $path/${mol}_pes.pdf
+cp plot_error.pdf $path/${mol}_pes_error.pdf
+cp plot_stat.pdf  $path/${mol}_pes_stat.pdf
+cp plot_closeness.pdf $path/${mol}_pes_closeness.pdf
+cp freq.pdf       $path/${mol}_freq.pdf
+cp xe.pdf         $path/${mol}_xe.pdf

### 2023-06-04 N2_cc-pvdz/rpt2/plot_distance.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+gnuplot plot_distance.gnu
+epspdf plot_distance.eps
+okular plot_distance.pdf

### 2023-06-04 N2_cc-pvdz/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=26
+nelectrons=10
+xcut=4.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 ethylene_cc-pvdz/fit_eq.sh

@@ -1,0 +1,295 @@
+#!/bin/bash
+
+#######################################################################
+
+if [ -f "fit.log" ]; then
+  rm "fit.log"
+fi
+
+gnuplot fit_eq.gnu
+
+#######################################################################
+grep ^a fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/a0 /FCI      /g' |
+        sed 's/a1 /RHF      /g' |
+        sed 's/a2 /hfCISD   /g' |
+        sed 's/a3 /hfCISDT  /g' |
+        sed 's/a4 /hfCISDTQ /g' |
+        sed 's/a5 /hfCIs0   /g' |
+        sed 's/a6 /hfCIs2   /g' |
+        sed 's/a7 /hfCIs4   /g' |
+        sed 's/a8 /hfCIo1   /g' |
+        sed 's/a9 /hfCIo1.5 /g' |
+        sed 's/a10/hfCIo2   /g' |
+        sed 's/a11/hfCIo2.5 /g' |
+        sed 's/a12/hfCIo3   /g' |
+        sed 's/a13/hfCIo3.5 /g' |
+        sed 's/a14/hfCIo4   /g' |
+        sed 's/a15/ooCIS    /g' |
+        sed 's/a16/ooCISD   /g' |
+        sed 's/a17/ooCISDT  /g' |
+        sed 's/a18/ooCIs0   /g' |
+        sed 's/a19/ooCIs2   /g' |
+        sed 's/a20/ooCIo1   /g' |
+        sed 's/a21/ooCIo1.5 /g' |
+        sed 's/a22/ooCIo2   /g' |
+        sed 's/a23/ooCIo2.5 /g' |
+        sed 's/a24/ooCIo3   /g' > a.dat
+
+grep ^D fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/D0 /FCI      /g' |
+        sed 's/D1 /RHF      /g' |
+        sed 's/D2 /hfCISD   /g' |
+        sed 's/D3 /hfCISDT  /g' |
[…]

### 2023-06-04 ethylene_cc-pvdz/rpt2/cp_to_manuscript.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+mol='ethylene'
+
+path='/home/fabris/ongoing_projects/seniority/Manuscript'
+
+cp plot_pes.pdf   $path/${mol}_pes.pdf
+cp plot_error.pdf $path/${mol}_pes_error.pdf
+cp plot_stat.pdf  $path/${mol}_pes_stat.pdf
+cp xe.pdf         $path/${mol}_xe.pdf

### 2023-06-04 ethylene_cc-pvdz/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=40
+nelectrons=12
+xcut=16.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-06-04 vinyl_cc-pvdz/rpt2/fit_eq.sh

@@ -156,7 +156,7 @@ if [[ $(grep 'a25 ' a0.dat) ]]; then grep '3   ' det_CIo.dat | cut -c 4- >> det.
 paste det.dat a.dat D.dat > det_aD.dat
 
 grep 'FCI'      det_aD.dat | sed 's/FCI/1 /g'     >  det_aD_FCI.dat
-grep 'FCI'      det_aD.dat | sed 's/FCI/1.0E9 /g' >> det_aD_FCI.dat
+grep 'FCI'      det_aD.dat | sed 's/FCI/1.0E12 /g' >> det_aD_FCI.dat
 
 grep 'ROHF'     det_aD.dat | sed 's/ROHF/0 /g'     >  det_aD_CI.dat
 grep 'hfCIS '   det_aD.dat | sed 's/hfCIS /1 /g'   >> det_aD_CI.dat
@@ -256,7 +256,7 @@ grep 'xe25 ' xe0.dat | sed 's/xe25/ooCIo3   /g' >> xe.dat
 paste det.dat xe.dat D.dat > det_xe.dat
 
 grep 'FCI' det_xe.dat | sed 's/FCI/1 /g'   >  det_xe_FCI.dat
-grep 'FCI' det_xe.dat | sed 's/FCI/1E9 /g' >> det_xe_FCI.dat
+grep 'FCI' det_xe.dat | sed 's/FCI/1E12 /g' >> det_xe_FCI.dat
 
 grep 'ROHF'     det_xe.dat | sed 's/ROHF/0 /g'     >  det_xe_CI.dat
 grep 'hfCIS '   det_xe.dat | sed 's/hfCIS /1 /g'   >> det_xe_CI.dat

### 2023-06-04 Manuscript/manuscript.tex
saving work

@@
 
-Configuration interaction (CI) offers a systematic way to solve the many-body electronic structure problem.
+Configuration interaction (CI) offers a systematic way to solve the many-body electronic structure problem. \cite{SzaboBook,Helgakerbook}
 By including progressively more determinants in the Hilbert space, the wave function becomes increasingly closer to the exact one, and so does the electronic energy.
@@
 In practice, however, one resorts to approximate CI methods, where only the determinants that satisfy a given criterion are included in the truncated Hilbert space.
-For small systems, FCI is still attainable and one can gauge it against approximate CI methods to assesss their performance.
+For small systems, FCI is still attainable and one can thus gauge the performance of approximate CI methods.
+Excitation-based CI
+Seniority based CI (sCI) \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}
+sCI works well to describe molecular dissociation \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
+Limited studies for radicals and excited states
 
-Excitation-based CI
-Seniority based CI (sCI)
-Limited studies for radicals and excited states
 We have recently introduced hierarchy CI (hCI), \cite{Kossoski_2022}
@@
 
-PECs
-NPEs, distance errors
-equilibrium geometries and vibrational frequencies
+% redefine the distance errors because of negative values?
+% vinyl missing point at 5.2 for s3i+pt2 and at 2.0 for s5+pt2
+% CN correct dissociating part for h4 and h3.5
+% H8 missing points for Q+pt2
+% H8 missing 9.0 for T+pt2
 
-HF orbitals
-orbital optimization?
+The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf},
+along with the NPEs, distance errors, equilibrium geometries and vibrational frequencies, plotted as a function of $\Ndet$ for the different CI methods.
+The present results for open-shell systems are discussed in detail in the following,
+and display similar trends to the ones previously reported for closed-shell systems. \cite{Kossoski_2022}
+hCI typically improves the convergence of the different properties when compared to excitation-based CI.
+Concerning the NPEs, such improvement is clear for \ce{OH} and \ce{CN} (involving a single bond breaking), less so for vinyl (double bond breaking),
+whereas for \ce{H7} (multiple bond breaking), hCI and excitation-based CI perform similarly.
+This behavior had also been observed for dissociation of closed-shell systems. \cite{Kossoski_2022}
+The same trend is observed for the distance error, although the improvement from excitation-based CI to hCI is less marked than for the NPEs.
+hCI further leads to a faster convergence of the equilibrium geometry of \ce{CN}, while comparable results are found for the geometries of the other systems.
+Meanwhile, a clear improvement from excitation-based CI to hCI is observed for the vibrational frequencies, except for \ce{H7}, where no difference is found.
 
-perturbative correction for closed- and open-shell systems
+%HF orbitals
+%orbital optimization?
+
+% rPT2 vs PT2
+
+We further assesssed the effect of the EN2 perturbative correction for both the open-shell systems studied here 
+and the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) addressed in our first report on hCI. \cite{Kossoski_2022}
+The full set of PECSs is shown in the {\SupInf}.
+At stretched geometries, the EN2 correction can lead to kinks and large jumps in the PECSs.
+Such problems were encountered at some CI methods for \ce{CN}, vinyl, \ce{H8}, and \ce{H7}, and to a less extent, for \ce{N2} (only for hCI1).
+For the remaining five systems, the EN2 correction did not produce problematic PECs.
+
+Around the equilibrium geometry, the EN2 correction produced well-behaved PECs at all levels of CI and for all systems, 
+except for \ce{CN}, which is related to its different states around equilibrium.
+%
+xe
+\ce{OH} massive improvement
+\ce{CN} \ce{H4} \ce{N2} \ce{F2} \ce{HF} big improvement
+\ce{H7} \ce{H8} vinyl ethylene some improvement
+%
+freq
+\ce{OH} massive improvement
+\ce{HF} \ce{H8} vinyl ethylene some improvement
+\ce{CN} \ce{F2} \ce{N2} \ce{H4} \ce{H7} big improvement
 

### 2023-06-15 Manuscript/manuscript.tex
saving work

@@
 {\QP} \cite{Garniron_2019} was employed to perform all the CI calculations presented here.
-In a given calculation, the energies are considered to be converged when the (largest) Epstein-Nesbet second-order perturbation correction computed in the truncated Hilbert space 
+In a given calculation, the energies are considered to be converged when the (largest) renormalized second-order Epstein-Nesbet (EN2) perturbation correction computed in the truncated Hilbert space 
 lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
@@
 
-% rPT2 vs PT2
-
 We further assesssed the effect of the EN2 perturbative correction for both the open-shell systems studied here 
@@
 The full set of PECSs is shown in the {\SupInf}.
-At stretched geometries, the EN2 correction can lead to kinks and large jumps in the PECSs.
-Such problems were encountered at some CI methods for \ce{CN}, vinyl, \ce{H8}, and \ce{H7}, and to a less extent, for \ce{N2} (only for hCI1).
-For the remaining five systems, the EN2 correction did not produce problematic PECs.
+At stretched geometries, the perturbative correction can lead to kinks and large jumps in the PECs.
+Such problems were encountered at some levels of CI for \ce{CN}, vinyl, \ce{H8}, and \ce{H7}, and to a less extent, for \ce{N2} (only for hCI1).
+For the remaining five systems (\ce{HF}, \ce{F2}, ethylene, \ce{H4}, and \ce{OH}),
+the EN2 correction produced smooth PECs at dissociation
+Around the equilibrium geometry, the perturbative correction produced well-behaved PECs at all levels of CI and for all systems, except for \ce{CN},
+which is related to the close-lying states around equilibrium.
 
-Around the equilibrium geometry, the EN2 correction produced well-behaved PECs at all levels of CI and for all systems, 
-except for \ce{CN}, which is related to its different states around equilibrium.
-%
-xe
-\ce{OH} massive improvement
-\ce{CN} \ce{H4} \ce{N2} \ce{F2} \ce{HF} big improvement
-\ce{H7} \ce{H8} vinyl ethylene some improvement
-%
-freq
-\ce{OH} massive improvement
-\ce{HF} \ce{H8} vinyl ethylene some improvement
-\ce{CN} \ce{F2} \ce{N2} \ce{H4} \ce{H7} big improvement
+We found that the EN2 perturbative correction generally reduces the errors of all observables considered here, for a given level of excitation-based CI and hCI routes.
+As such, it accelerates the rate of convergence as a function of the excitation degree or hierarchy parameter.
+Importantly, the perturbative corrected hCI route still outperforms the excitation-based counterpart, just as for the uncorrected comparison.
+Furthermore, the reduction of the errors is overall more monotonic than without the perturbative correction.
+This is illustrated in Fig.\ref{} for the particular case of \ce{F2}.
+For the sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation and hierarchy-based alternatives.
+
+% rPT2 vs PT2 there are important differences
+
+%xe
+%\ce{OH} massive improvement
+%\ce{CN} \ce{H4} \ce{N2} \ce{F2} \ce{HF} big improvement
+%\ce{H7} \ce{H8} vinyl ethylene some improvement
+
+%freq
+%\ce{OH} massive improvement
+%\ce{CN} \ce{F2} \ce{N2} \ce{H4} \ce{H7} big improvement
+%\ce{HF} \ce{H8} vinyl ethylene some improvement
 
@@
 In Tab.~\ref{tab:1} we present the statistical errors of the different methods, organized by their computational scaling.
+% hCI1 and hCI1.5
 We found that $\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat more favourable MSE and RMSE of the latter.
-$\Delta$hCI2 still tends to underestimate the excitation energies, though to a much lesser extent than observed for the lower-order hCI methods.
+This means that the additional determinants present in the former method recover a larger fraction of the correlation energy for excited state than for the ground state,
+thus creating a bias toward underestimated excitation energies.
+Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
 
@@
 with corresponding statistical errors shown in Tab.~\ref{tab:2}.
-Even though this is a small set, it should be enough to obtain the main trends on the statistics of the different methods.
+Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order methods.
 
@@
 An improved performance is found for the methods displaying an $\order*{N^8}$ computational scaling.
-The four methods, hCI3, $\Delta$hCI3, CISDT, and $\Delta$CISDT, display quite similar MAEs and RMSEs.
-Therefore, there is no advantage by moving from excitation-based to hierarchy-based CI at this order, just as discussed above for the $\Delta$hCI2 and $\Delta$CISD.
-Similarly, the $\Delta$hCI3 does not improve with respect to hCI3, as also observed for $\Delta$CISDT and CISDT. \cite{Kossoski_2023}
-We notice, however, the negative MSEs for the state-specific methods, in contrast to the positives values obtained for the ground-state based route.
+The four methods, hCI3, $\Delta$hCI3, CISDT, and $\Delta$CISDT, present quite similar MAEs and RMSEs.
+Therefore, there is no real advantage in moving from excitation-based to hierarchy-based CI at this order, just as discussed above for the $\Delta$hCI2 and $\Delta$CISD.
+Similarly, the $\Delta$hCI3 is not more accurate than hCI3, as also observed for $\Delta$CISDT and CISDT. \cite{Kossoski_2023}
+We notice, however, that the state-specific methods present negative MSEs, in contrast to the positives values obtained in the ground-state based route.
+This reflects an umbalanced description of correlation effects, favoring the excited states.
 At the next order, $\Delta$hCI3.5 has significantly smaller errors, but at a costly $\order*{N^9}$ scaling.
@@
 sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
-This improvement, however, is not as expressive as typically observed for excitataion-based and hierarchy-based methods.
-(Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:2}).
-The large errors, even at the state-specific route, combined with the formal $\order*{e^N}$ scaling of these methods,
-render them unsuitable for calculations of accurate excitation energies.
+This improvement, however, is not as expressive as observed for excitataion-based and hierarchy-based methods.
+Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:2}.
+The large errors, even following the state-specific approach, combined with the formal $\order*{e^N}$ scaling of these methods,
+render seniority-based CI methods unsuitable for calculations of excitation energies.
 
@@
 
-We conclude that 
-Meanwhile, 
-, and therefore not useful.
+%three CI routes (excitation, hierarchy, and seniority-based).
+
+We have considered both closed and open shell systems, considering ground-state properties and excitation energies, for a diverse set of molecular systems.
+Extending our previous findings for closed-shell \cite{Kossoski_2022} to open-shell systems,
+we conclude that the hCI route has an edge in comparison to excitation-based CI,
+and that seniority-based CI methods lead to far less accurate results than the two former alternatives, for a given computational cost.
+We also found that, for closed- and open-shell systems,
+%EN2 perturbative corrections lead to faster and more monotonic convergence of different properties
+EN2 perturbative corrections lead to substantial improvements for hCI and excitation-based CI routes (while keeping the upper hand for the former), but do not help much with the deficiencies of the sCI route.
+
+We further assessed the performance of sCI and hCI methods, based on either HF ground-state orbitals or state-specific orbitals, to describe excited states of closed-shell systems and radicals.
+Overall, sCI methods present relatively poor results considering their high computational costs, and thus should not be useful in practice.
+%it is fair to conclude that
 

### 2023-06-15 determinants/determinants.f90

@@ -1,0 +1,244 @@
+program determinants
+implicit none
+
+integer*16 :: o, v
+integer*16 :: e, k
+integer*16 :: n, m
+
+integer*16 :: s0d0
+integer*16 ::       s2d1
+integer*16 :: s0d2, s2d2, s4d2
+integer*16 ::       s2d3, s4d3, s6d3
+integer*16 :: s0d4, s2d4, s4d4, s6d4, s8d4
+integer*16 ::       s2d5, s4d5, s6d5
+integer*16 :: s0d6, s2d6, s4d6
+integer*16 ::       s2d7
+integer*16 :: s0d8
+
+integer*16 :: rhf, fci
+integer*16 :: cis0, cis2, cis4, cis6
+integer*16 :: cis, cisd, cisdt, cisdtq
+integer*16 :: cio1, cio1p5, cio2, cio2p5, cio3, cio3p5, cio4
+
+integer :: nunit
+character(len=8), parameter :: fmt1 = "(*(i16))"
+character(len=8), parameter :: fmt2 = "(a,i16)"
+
+character(len=32) :: filename
+
+!!!!!!!!!!!!!!!!!!!!!!!
+read(*,*) e, k
+!!!!!!!!!!!!!!!!!!!!!!!
+
+o = e/2
+v = k - o
+
+!!!!!!!!!!!!!!!!!!!!!!!
+call binomial_coefficient(k,o,m)
+! CIs0
+cis0 = m
+! CIs2
+cis2 = cis0 + m*o*v
+! CIs4
+cis4 = cis2 + m*o*(o-1)*v*(v-1)/4
+! CIs6
+cis6 = cis4 + m*o*(o-1)*(o-2)*v*(v-1)*(v-2)/36
+!!!!!!!!!!!!!!!!!!!!!!!
+
+filename = 'det_CIs.dat'
+open(newunit=nunit,file=filename,status='unknown')
+  write(nunit,fmt2) '0   ', cis0
+  write(nunit,fmt2) '2   ', cis2
+  write(nunit,fmt2) '4   ', cis4
+  write(nunit,fmt2) '6   ', cis6
+close(nunit)
+
+!!!!!!!!!!!!!!!!!!!!!!!
+
+s0d0 = 1
+
+s0d2 = o*v
+
[…]

### 2023-06-15 determinants/determinants_odd.f90

@@ -1,0 +1,245 @@
+program determinants_odd
+implicit none
+
+integer*16 :: o, v
+integer*16 :: e, k
+integer*16 :: n, m
+
+integer*16 :: s1d0
+integer*16 ::       s3d1
+integer*16 :: s1d2, s3d2, s5d2
+integer*16 ::       s3d3, s5d3, s7d3
+integer*16 :: s1d4, s3d4, s5d4, s7d4, s9d4
+integer*16 ::       s3d5, s5d5, s7d5
+integer*16 :: s1d6, s3d6, s5d6
+integer*16 ::       s3d7
+integer*16 :: s1d8
+
+integer*16 :: rhf, fci
+integer*16 :: cis1, cis3, cis5, cis7
+integer*16 :: cis, cisd, cisdt, cisdtq
+integer*16 :: cio1, cio1p5, cio2, cio2p5, cio3, cio3p5, cio4
+
+integer :: nunit
+character(len=8), parameter :: fmt1 = "(*(i16))"
+character(len=8), parameter :: fmt2 = "(a,i16)"
+
+character(len=32) :: filename
+
+!!!!!!!!!!!!!!!!!!!!!!!
+read(*,*) e, k
+!!!!!!!!!!!!!!!!!!!!!!!
+
+!o = e/2
+o = (e-1)/2
+v = k - o
+
+!!!!!!!!!!!!!!!!!!!!!!!
+call binomial_coefficient(k,o,m)
+! CIs1
+cis1 = m
+! CIs3
+cis3 = cis1 + m*o*v
+! CIs5
+cis5 = cis3 + m*o*(o-1)*v*(v-1)/4
+! CIs7
+cis7 = cis5 + m*o*(o-1)*(o-2)*v*(v-1)*(v-2)/36
+!!!!!!!!!!!!!!!!!!!!!!!
+
+filename = 'det_CIs.dat'
+open(newunit=nunit,file=filename,status='unknown')
+  write(nunit,fmt2) '1   ', cis1
+  write(nunit,fmt2) '3   ', cis3
+  write(nunit,fmt2) '5   ', cis5
+  write(nunit,fmt2) '7   ', cis7
+close(nunit)
+
+!!!!!!!!!!!!!!!!!!!!!!!
+
[…]

### 2023-06-16 Manuscript/manuscript.tex
saving work

@@
 In practice, however, one resorts to approximate CI methods, where only the determinants that satisfy a given criterion are included in the truncated Hilbert space.
-For small systems, FCI is still attainable and one can thus gauge the performance of approximate CI methods.
-Excitation-based CI
-Seniority based CI (sCI) \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}
-sCI works well to describe molecular dissociation \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
-Limited studies for radicals and excited states
 
-We have recently introduced hierarchy CI (hCI), \cite{Kossoski_2022}
-where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$.
-For different properties and molecular systems, hCI showed an improved convergence to the FCI results when compared to the traditional excitation-based CI. \cite{Kossoski_2022}
+The most well-known CI route is excitation-based.
+Starting from a reference determinant, typically the Hartree-Fock (HF) determinant,
+one generates all determinants that are connected to it by excitating at most $e$ electrons.
+The excitation degree $e$ thus defines the order of the approximate CI model:
+CI with single excitations (CIS); CI with single and double excitations (CISD), CI with single, double and triple excitations (CISDT), and so on.
+A different CI route is based on the seniority number $s$ (the number of unpaired electrons in a given determinant).
+In seniority-based CI (sCI) \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}, there is no reference determinant, and one accounts for all determinants having seniority equal or less than $s$.
+For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually refered to as doubly-occupied CI (DOCI), \cite{}
+which is followed by the higher-order models sCI2, sCI4, and so on.
+For an odd number of electrons, the seniority route follows along odd numbers of $s$: sCI1, sCI3, and so on.
+%Limited studies for radicals and excited states
+
+%sCI works well to describe molecular dissociation \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
+%For small systems, FCI is still attainable and one can thus gauge the performance of approximate CI methods.
+We have recently introduced a third CI route, hierarchy CI (hCI), \cite{Kossoski_2022}
+where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$ according to $h = (e+s/2)/2$.
+For different properties and closed-shell molecular systems, 
+hCI was found to display an overall faster convergence to the FCI results than the traditional excitation-based CI, as a function of the number of determinants $\Ndet$. \cite{Kossoski_2022}
 However, hCI has only been defined for a closed-shell reference, thus systems with an even number of electrons.
-A generalization of the method for radicals has not yet been proposed.
+Here, our first goal is to generalize hCI for open-shell systems and to assess the performance of this class of methods by performing calculations for various properties of several radicals.
+%A generalization of the method for open-shell systems has not yet been proposed.
+%excitations of increasingly higher order give rise to a systematically improvable sequence of CI methods:
+The second goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction, \cite{}
+impacts different properties of ground-state closed-shell systems and radicals, for excitation-based, sCI and hCI methods.
 
-The higher-lying CI roots give access to excited states.
-Whereas the performance of excitation-based CI for excited states is well stablished, \cite{} this remains an open question for sCI methods,
-despite the growing number of methods targeting excited states that exploit the concept of seniority number. \cite{} 
-Similarly, the favourable results of hCI when compared to excitation-based CI for ground states raises the question on how well it can describe excitation energies.
-
-When targeting excited states with CI methods, one can employ the ground-state HF orbitals or instead optimize the orbitals for each state of interest, in a so-called state-specific approach.
-Promising results for state-specific excitation-based CI.
+%Whereas the first CI root represents the electronic ground state, higher-lying roots give access to excited states.
+%Similarly, the favourable results of hCI when compared to excitation-based CI for ground states raises the question on how well it can describe excitation energies.
+%The performance of excitation-based CI for excited states is well stablished, \cite{} but remains an open question for sCI methods,
+Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022} our third goal is to define and assess the accuracy of hCI methods for excited states.
+Finally, to the best of our knowledge, sCI methods have not yet been directly used to target excited states,
+despite the growing number of excited state methods that exploit the concept of seniority number. \cite{} 
+Our fourth goal is therefore to introduce and gauge sCI methods for excited states.
+There are two approaches to target excited states with CI methods.
+One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix.
+Instead, one may optimize the orbitals for the excited state of interest (described with an appropriate reference), followed by a separate CI calculation,
+in a so-called state-specific approach ($\Delta$CI).
+There has been a recent surge in the development of state-specific methods, covering 
+single-reference and multiconfigurational self-consistent field,
+\cite{Ziegler_1977,Burton_2021,Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Hanscam_2022,Kossoski_2023,Marie_2023}
+% add references for SCF
+%\cite{Ziegler_1977,Burton_2021}
+%multiconfigurational SCF, \cite{Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Kossoski_2023,Marie_2023}
+density functional theory, 
+\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022}
+and coupled-cluster 
+\cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023}
+methods.
+By employing a minimal configuration state function reference, we have recently showed that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues.
 \cite{Kossoski_2023}
-Here we introduce analogous state-specific seniority-based CI ($\Delta$sCI) and hierarchy-based CI ($\Delta$hCI).
-How do these methods perform for excited states, from both closed-shell and open-shell systems?
-
-Here we address the knowledge gaps outlined above by setting four goals.
-Our first goal is to generalize hCI \cite{Kossoski_2022} for the case of an open-shell reference determinant, thus extending its domain to radical species.
-The second is to assess the performance of this method for ground state radicals, similarly to what has been done for closed-shell systems. \cite{Kossoski_2022}
-%Seniority hCI
-The third goal is to assess the improvement from perturbative corrections to ground-state radicals and closed-shell systems, for both excitation-based and hierarchy-based CI methods.
-Finally, we introduce and gauge seniority-based and hierarchy-based CI methods for excited states, along both the standard ground-state based route and the state-specific route. \cite{Kossoski_2023}
+Based on this promising set of results for excitation-based CI, here we explore both ground-state based and state-specific approaches for sCI and hCI models.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+along the same lines of our previous
 
@@
 Importantly, the perturbative corrected hCI route still outperforms the excitation-based counterpart, just as for the uncorrected comparison.
-Furthermore, the reduction of the errors is overall more monotonic than without the perturbative correction.
-This is illustrated in Fig.\ref{} for the particular case of \ce{F2}.
-For the sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation and hierarchy-based alternatives.
+Furthermore, the reduction of the errors is overall more monotonic than without the perturbative correction,
+which is illustrated in Fig.\ref{} for the particular case of \ce{F2}.
+For the sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation-based and hCI alternatives.
 
@@
 %three CI routes (excitation, hierarchy, and seniority-based).
+%Here we address the knowledge gaps outlined above by setting four goals.
+%Extending our previous findings for closed-shell \cite{Kossoski_2022} to open-shell systems,
+%We further explored Epstein-Nesbet second-order perturbative corrections to the various approximate CI methods.
+%The second is to assess the performance of this method for ground state radicals, similarly to what has been done for closed-shell systems. \cite{Kossoski_2022}
 
-We have considered both closed and open shell systems, considering ground-state properties and excitation energies, for a diverse set of molecular systems.
-Extending our previous findings for closed-shell \cite{Kossoski_2022} to open-shell systems,
-we conclude that the hCI route has an edge in comparison to excitation-based CI,
-and that seniority-based CI methods lead to far less accurate results than the two former alternatives, for a given computational cost.
+%The third goal is to assess the improvement from perturbative corrections to ground-state radicals and closed-shell systems, for both excitation-based and hierarchy-based CI methods.
+%For hCI and excitation-based CI, the perturbative correction tends to accelerate and stabilize the convergence of different properties toward the FCI results.
+%EN2 perturbative corrections lead to faster and more monotonic convergence of different properties
+%We have considered both closed and open shell systems, considering ground-state properties and excitation energies, for a diverse set of molecular systems.
+%Finally, we introduce and gauge seniority-based and hierarchy-based CI methods for excited states, along both the standard ground-state based route and the state-specific route. \cite{Kossoski_2023}
+
+Here we have generalized hCI \cite{Kossoski_2022} for the case of an open-shell reference determinant, thus extending its applicability to radical species and excited states.
+By surveying the dissociation of several open-shell systems, 
+we found that the hCI route has an edge in comparison to excitation-based CI, for both weakly correlated (equilibrium geometries) and strongly correlating (dissociation) regimes.
+Meanwhile, seniority-based CI leads to far less accurate results in comparison to hCI or excitation-based CI, for a given computational cost.
+These findings are in line with what had been found for closed-shell systems. \cite{Kossoski_2022}
 We also found that, for closed- and open-shell systems,
-%EN2 perturbative corrections lead to faster and more monotonic convergence of different properties
-EN2 perturbative corrections lead to substantial improvements for hCI and excitation-based CI routes (while keeping the upper hand for the former), but do not help much with the deficiencies of the sCI route.
+the EN2 perturbative corrections lead to substantial improvements for hCI and excitation-based CI routes (while keeping the upper hand for the former), but do not help much with the deficiencies of the sCI route.
 
-We further assessed the performance of sCI and hCI methods, based on either HF ground-state orbitals or state-specific orbitals, to describe excited states of closed-shell systems and radicals.
-Overall, sCI methods present relatively poor results considering their high computational costs, and thus should not be useful in practice.
+We further gauged the performance of hCI and sCI methods, based on either HF ground-state orbitals or state-specific orbitals, to describe excited states of closed-shell systems and radicals.
+Despite their elevated computational cost, sCI methods produced relatively poor results (even in a state-specific approach), and thus should not be useful in practice.
+Similarly, for lower orders of hCI (hCI1 and hCI1.5), the performance is very poor,
+with ground-state (state-specific) orbitals leading to largely overestimated (underestimated) excitation energies.
+The situation improves at the state-specific $\Delta$hCI2 order, even though its accuracy is somewhat worse than $\Delta$CISD, both methods sharing the same computational scaling. 
+Unfortunately, $\Delta$hCI2.5 does not lead to more accurate results, whereas all the CI methods at next next order (CISDT, $\Delta$CISDT, hCI3, and $\Delta$hCI3) show comparable performance.
+Overall, hCI methods for excited states are either innacurate given their computational cost ($h=1$, $h=1.5$ and $h=2.5$) or comparable to excitation-based CI at the same order ($h=2$ and $h=3$).
 %it is fair to conclude that

### 2023-06-18 Manuscript/manuscript.tex
saving work

@@
 In full CI (FCI), all determinants are accounted for and the problem is solved exactly (for a given basis set).
-In practice, however, one resorts to approximate CI methods, where only the determinants that satisfy a given criterion are included in the truncated Hilbert space.
+In practice, however, one resorts to approximate CI models, where only the determinants that satisfy a given criterion are included in the truncated Hilbert space.
 
@@
 In seniority-based CI (sCI) \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}, there is no reference determinant, and one accounts for all determinants having seniority equal or less than $s$.
-For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually refered to as doubly-occupied CI (DOCI), \cite{}
+For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually refered to as doubly-occupied CI (DOCI),
 which is followed by the higher-order models sCI2, sCI4, and so on.
@@
 %sCI works well to describe molecular dissociation \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
-%For small systems, FCI is still attainable and one can thus gauge the performance of approximate CI methods.
+%For small systems, FCI is still attainable and one can thus gauge the performance of approximate CI models.
 We have recently introduced a third CI route, hierarchy CI (hCI), \cite{Kossoski_2022}
@@
 %excitations of increasingly higher order give rise to a systematically improvable sequence of CI methods:
-The second goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction, \cite{}
+The second goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction, \cite{Garniron_2019}
 impacts different properties of ground-state closed-shell systems and radicals, for excitation-based, sCI and hCI methods.
@@
 %The performance of excitation-based CI for excited states is well stablished, \cite{} but remains an open question for sCI methods,
-Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022} our third goal is to define and assess the accuracy of hCI methods for excited states.
-Finally, to the best of our knowledge, sCI methods have not yet been directly used to target excited states,
-despite the growing number of excited state methods that exploit the concept of seniority number. \cite{} 
-Our fourth goal is therefore to introduce and gauge sCI methods for excited states.
-There are two approaches to target excited states with CI methods.
-One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix.
+Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022} our third goal is to define and assess the accuracy of hCI models for excited states.
+Finally, to the best of our knowledge, sCI models have not yet been directly used to target excited states,
+despite the growing number of methods that exploit the concept of seniority number,
+for both ground
+\cite{Limacher_2013,Limacher_2014,Tecmer_2014,Boguslawski_2014a,Boguslawski_2015,Boguslawski_2014b,Boguslawski_2014c,Johnson_2020,Henderson_2014,Stein_2014,Henderson_2015,Chen_2015,Bytautas_2018,Marie_2021,Boguslawski_2021,Tecmer_2022,Mamache_2023}
+and excited states.
+\cite{Boguslawski_2016b,Boguslawski_2016c,Boguslawski_2019,Nowak_2019,Kossoski_2021,Marie_2021,Tecmer_2022,Rishi_2023,Nowak_2023} 
+Our fourth goal is therefore to introduce and gauge sCI models for excited states.
+There are two possible approaches to target excited states with CI.
+One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix, which we refer to as ground-state based approach.
 Instead, one may optimize the orbitals for the excited state of interest (described with an appropriate reference), followed by a separate CI calculation,
@@
 methods.
-By employing a minimal configuration state function reference, we have recently showed that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues.
-\cite{Kossoski_2023}
+In particular, by employing a minimal configuration state function reference, 
+we have recently shown that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues. \cite{Kossoski_2023}
 Based on this promising set of results for excitation-based CI, here we explore both ground-state based and state-specific approaches for sCI and hCI models.
@@
 
-From here on, method names carrying the $\Delta$ symbol denote a state-specific approach,
-whereas those without the $\Delta$ mean that ground-state HF orbitals were employed.
+state-specific methods
+$\Delta$sCI2 
+$\Delta$hCI1
+From here on, labels carrying the $\Delta$ symbol denote a state-specific approach,
+whereas those without imply the ground-state based approach.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-along the same lines of our previous
+
+The hCI models introduced here were implemented in {\QP} \cite{Garniron_2019} through a straightforward modification of the
+\textit{configuration interaction using a perturbative selection made iteratively} (CIPSI) algorithm. \cite{Huron_1973,Giner_2013,Giner_2015,Garniron_2018}
+By allowing only for the determinants that are connected with the reference determinant(s) up to a given maximum hierarchy $h$,
+the CIPSI algorithm is restricted to the truncated Hilbert space defined by the reference determinant(s) and the value of $h$.
+{\QP} \cite{Garniron_2019} was also employed to perform all the CI calculations presented here.
+In a given calculation, the energies are considered to be converged when the (largest) renormalized EN2 correction computed in the truncated Hilbert space 
+lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
+%by spanning the most important regions of the truncated Hilbert space
+This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space,
+while delievering fairly converged absolute and excitation energies.
+The ground- and excited-state CI energies are obtained with the Davidson iterative algorithm. \cite{Davidson_1975}
+For a given approximate CI model, we futher evaluated the bare and the renormalized EN2 perturbative correction. \cite{Garniron_2019} 
+%associated with the determinants left outside the truncated space.
+This calculation involves a single loop over the determinants left outside the truncated space but connected to it via at most double excitations.
+Looping over these external doubly excited determinants thus represents an additional $\order*{N^4}$ factor on top of the computational scaling of the CI calculation.
+For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^{10}}$,
+though with a small prefactor stemming from the EN2 calculation.
 
 To gauge the performance of hCI against excitation-based and seniority-based CI for radical systems,
-we have calculated the potential energy curves (PECs) for the dissociation of four radical systems:
+we have calculated the ground-state potential energy curves (PECs) for the dissociation of four radical systems:
 \ce{OH}, \ce{CN}, vinyl (\ce{C2H3}), and \ce{H7}.
+These calculations employed the ground-state restricted open-shell HF orbitals and the frozen core approximation.
 %which display a variable number of bond breaking.
@@
 For \ce{H7}, we considered equally spaced and linearly arranged hydrogen atoms, and the PECs were computed along the symmetric dissociation coordinate.
-From the PECs, we performed a similar analysis as in our paper on hCI for closed-shell systems. \cite{Kossoski_2022}
-Namely, for the different CI levels considered here, 
+The results were analyzed along the same lines of our report on hCI for closed-shell systems. \cite{Kossoski_2022}
+Namely, for the different CI models considered here, 
 we evaluated the convergence of the non-parallelity error (NPE), the distance error, the vibrational frequencies, and the equilibrium geometries, as functions of $\Ndet$.
-The NPE (distance error) of a given level of theory is defined as the maximum minus (plus) the minimum differences between its corresponding PEC and the FCI PEC.
-Details about how the vibrational frequencies and equilibrium geometries were obtained from the calculated PECs can be found in the \SupInf.
+The NPE (distance error) of a given level of theory is defined as the maximum minus (plus) the minimum differences between its corresponding PEC and the FCI one, for a given range of coordinates.
+Details about how the vibrational frequencies and equilibrium geometries were obtained from the calculated PECs,
+and the ranges defining the NPE and distance errors can be found in the \SupInf.
 
-To evaluate how the different levels of CI introduced here perform for excited states,
-we calculated vertical excitation energies for 60 electronic states, comprising 26 different molecular systems, with geometries extracted from the QUEST database. \cite{Veril_2021}
-% 18 closed-shells, 8 radicals
+The various CI models introduced here were further assessed based on calculated vertical excitation energies for 60 electronic states,
+from 18 closed-shell systems and 8 radicals, with geometries extracted from the QUEST database. \cite{Veril_2021}
+The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.
 We employed the aug-cc-pVDZ basis set for systems having up to three non-hydrogen atoms and the 6-31+G(d) basis set for the larger ones.
+Core orbitals were frozen.
 We performed calculations following both the standard ground-state based CI route and the state-specific CI route. \cite{Kossoski_2023}
 For the latter, we employed the state-specific orbitals obtained in Ref.~\onlinecite{Kossoski_2023}.
-The computed excitation energies were compared against the reference values from the QUEST database. \cite{Veril_2021}
-The full set of excited states and calculated excitation energies, for the various methods considered here, are provided in the {\SupInf}.
+In contrast to excitation-based CI, energies obtained with sCI and hCI models are not invariant under rotations withing the occupied and virtual subspaces.
+The computed excitation energies were compared against the reference values provided in the QUEST database. \cite{Veril_2021}
 %For completeness, we also reproduce the geometries in the \SupInf.
-
-The hCI methods introduced here were implemented in {\QP} \cite{Garniron_2019} through a straightforward modification of the
-\textit{configuration interaction using a perturbative selection made iteratively} (CIPSI) algorithm. \cite{Huron_1973,Giner_2013,Giner_2015,Garniron_2018}
-By allowing only for the determinants that are connected with the reference determinant(s) up to a given maximum hierarchy $h$,
-the CIPSI algorithm is restricted to the truncated Hilbert space defined by the reference and the value of $h$.
-{\QP} \cite{Garniron_2019} was employed to perform all the CI calculations presented here.
-In a given calculation, the energies are considered to be converged when the (largest) renormalized second-order Epstein-Nesbet (EN2) perturbation correction computed in the truncated Hilbert space 
-lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
-%by spanning the most important regions of the truncated Hilbert space
-This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space,
-while delievering fairly converged absolute and excitation energies.
-The ground- and excited-state CI energies are obtained with the Davidson iterative algorithm. \cite{Davidson_1975}
-
-state-specific methods
-$\Delta$sCI2 
-$\Delta$hCI1
-%restricted open-shell HF orbitals
 
@@
 The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf},
-along with the NPEs, distance errors, equilibrium geometries and vibrational frequencies, plotted as a function of $\Ndet$ for the different CI methods.
+along with the NPEs, distance errors, equilibrium geometries and vibrational frequencies, plotted as a function of $\Ndet$ for the different CI models.
 The present results for open-shell systems are discussed in detail in the following,
@@
 
-%HF orbitals
 %orbital optimization?
@@
 For the remaining five systems (\ce{HF}, \ce{F2}, ethylene, \ce{H4}, and \ce{OH}),
-the EN2 correction produced smooth PECs at dissociation
-Around the equilibrium geometry, the perturbative correction produced well-behaved PECs at all levels of CI and for all systems, except for \ce{CN},
-which is related to the close-lying states around equilibrium.
+the EN2 correction produced smooth PECs at dissociation.
+Around the equilibrium geometry, the perturbative correction produced well-behaved PECs at all levels of CI and for all systems, 
+except for \ce{CN}, which is related to crossing of states around equilibrium.
 
@@
 We found that $\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat more favourable MSE and RMSE of the latter.
-This means that the additional determinants present in the former method recover a larger fraction of the correlation energy for excited state than for the ground state,
+This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state,
 thus creating a bias toward underestimated excitation energies.
@@
 \begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
-for the Set of 60 Excitation Energies Listed in the {\SupInf}.
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of 
+38 Singly Excited States of Closed-Shell Systems and 
+Listed in the {\SupInf}.
 }
@@
 \hline
-$\Delta$CSF       & -0.54 & 0.70 & 0.86 \\
-CIS               & +0.13 & 0.59 & 0.77 \\
+$\Delta$CSF       & -0.75 & 0.81 & 0.95 \\
+CIS               & +0.03 & 0.62 & 0.63 \\
 \hline
-hCI1              & +1.11 & 1.14 & 1.39 \\
-$\Delta$hCI1      & -0.90 & 1.17 & 1.40 \\
+hCI1              & +1.08 & 1.12 & 1.25 \\
+$\Delta$hCI1      & -1.37 & 1.41 & 1.52 \\
+hCI1.5            & +1.95 & 1.95 & 2.03 \\
+$\Delta$hCI1.5    & -2.95 & 2.95 & 3.14 \\
+$\Delta$CISD      & -0.13 & 0.17 & 0.22 \\
+$\Delta$hCI2      & -0.19 & 0.20 & 0.25 \\
 \hline
-hCI1.5            & +1.88 & 1.88 & 2.20 \\
-$\Delta$hCI1.5    & -1.99 & 2.12 & 2.58 \\
-\hline
-$\Delta$CISD      & -0.08 & 0.17 & 0.22 \\
-$\Delta$hCI2      & -0.15 & 0.21 & 0.30 \\
-\hline
-sCI2              &  0.96 & 1.06 & 1.28 \\
-$\Delta$sCI2      &  0.39 & 0.60 & 0.73 \\
-%\hline
-%CIS(D)            & -0.03 & 0.35 & 0.40 \\
-%CC2               & -0.05 & 0.32 & 0.37 \\
-%CCSD              & +0.01 & 0.08 & 0.09 \\
-%CC3               & +0.01 & 0.03 & 0.06 \\
-%CCSDT             & -0.01 & 0.02 & 0.02 \\
+sCI2/sCI2         & +1.40 & 1.40 & 1.53 \\
+sCI2/sCI0         & -0.22 & 0.42 & 0.57 \\
+$\Delta$sCI2/sCI2 & +0.74 & 0.77 & 0.85 \\
+$\Delta$sCI2/sCI0 & -0.88 & 0.88 & 0.96 \\
 \end{tabular}
@@
 
-Despite the rather poor performance of lower-order hCI methods for excitation energies,
-we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
-For that, we performed calculations for a subset of 14 excited states of small molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
-with corresponding statistical errors shown in Tab.~\ref{tab:2}.
-Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order methods.
-
 \begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
-for the Set of 14 Excitation Energies Listed in the {\SupInf}.
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of 
+16 Singly-Excited States from Open-Shell Doublets 
+Listed in the {\SupInf}.
 }
@@
 \hline
-%CIS               & +0.16 & 0.82 & 0.96  \\
-%$\Delta$CSF       & -0.84 & 0.91 & 1.03  \\
-%\hline
+$\Delta$CSF       & -0.09 & 0.50 & 0.67 \\
+CIS               & +0.30 & 0.41 & 0.55 \\
+\hline
+hCI1              & +0.76 & 0.80 & 1.02 \\
+$\Delta$hCI1      & -0.09 & 0.58 & 0.94 \\
+hCI1.5            & +1.22 & 1.22 & 1.45 \\
+$\Delta$hCI1.5    & -0.26 & 0.66 & 0.92 \\
+$\Delta$CISD      & +0.00 & 0.16 & 0.23 \\
+$\Delta$hCI2      & -0.03 & 0.14 & 0.19 \\
+\hline
+sCI1              & +0.27 & 0.44 & 0.66 \\
+$\Delta$sCI1      & -0.15 & 0.28 & 0.40 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+Despite the rather poor performance of lower-order hCI models for excitation energies,
+we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
+For that, we performed calculations for a subset of 14 excited states of small molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
+with corresponding statistical errors shown in Tab.~\ref{tab:2}.
+Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order models.
+
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
+for the Set of 14 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+}
+\label{tab:2}
+\begin{ruledtabular}
+\begin{tabular}{lddd}
+method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
+\hline
 CISD              & +4.09 & 4.09 & 4.18  \\
+CISDT             & +0.09 & 0.18 & 0.20  \\
+CISDTQ            & +0.14 & 0.14 & 0.16  \\
+%
+$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
+$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
+$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
+%
 hCI2              & +2.89 & 3.00 & 3.28  \\
-$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
+hCI2.5            & +1.93 & 1.93 & 2.05  \\
+hCI3              & +0.18 & 0.18 & 0.19  \\
+%
 $\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
-\hline
-hCI2.5            & +1.93 & 1.93 & 2.05  \\
 $\Delta$hCI2.5    & -0.27 & 0.27 & 0.30  \\
-\hline
-CISDT             & +0.09 & 0.18 & 0.20  \\
-hCI3              & +0.18 & 0.18 & 0.19  \\
-$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
 $\Delta$hCI3      & -0.21 & 0.21 & 0.24  \\
-\hline
 $\Delta$hCI3.5    & -0.08 & 0.08 & 0.09  \\
-\hline
-CISDTQ            & +0.14 & 0.14 & 0.16  \\
-$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
 \end{tabular}
@@
 However, this improvement does not reflect on the state specific route (compare $\Delta$hCI2 and $\Delta$CISD in Tabs.~\ref{tab:1} and ~\ref{tab:2}).
-Moving one step further, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
+Moving to the next order, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
 it is actually less accurate then $\Delta$hCI2.
-An improved performance is found for the methods displaying an $\order*{N^8}$ computational scaling.
-The four methods, hCI3, $\Delta$hCI3, CISDT, and $\Delta$CISDT, present quite similar MAEs and RMSEs.
-Therefore, there is no real advantage in moving from excitation-based to hierarchy-based CI at this order, just as discussed above for the $\Delta$hCI2 and $\Delta$CISD.
-Similarly, the $\Delta$hCI3 is not more accurate than hCI3, as also observed for $\Delta$CISDT and CISDT. \cite{Kossoski_2023}
-We notice, however, that the state-specific methods present negative MSEs, in contrast to the positives values obtained in the ground-state based route.
+An improved performance is found for hCI3, $\Delta$hCI3, CISDT, and $\Delta$CISDT, which present quite similar MAEs and RMSEs.
+There is thus no advantage in moving from excitation-based to hCI at this order, similarly to what discussed before for the $\Delta$hCI2 and $\Delta$CISD models.
+Moreover, $\Delta$hCI3 is not more accurate than hCI3, as also observed for $\Delta$CISDT and CISDT. \cite{Kossoski_2023}
+We notice, however, that the state-specific models present negative MSEs, in contrast to the positives values obtained in the ground-state based route.
 This reflects an umbalanced description of correlation effects, favoring the excited states.
-At the next order, $\Delta$hCI3.5 has significantly smaller errors, but at a costly $\order*{N^9}$ scaling.
+At the next order, $\Delta$hCI3.5 has significantly smaller errors, but at a much more elevated computational cost.
 
@@
 
-We refer back to Tab.~\ref{tab:1} to discuss the performance of seniority-based CI for excited states.
+We refer to Tab.~\ref{tab:1} and ~\ref{tab:2} to discuss the performance of seniority-based CI for excited states.
 sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
 This improvement, however, is not as expressive as observed for excitataion-based and hierarchy-based methods.
-Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:2}.
+Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:1}.
 The large errors, even following the state-specific approach, combined with the formal $\order*{e^N}$ scaling of these methods,
-render seniority-based CI methods unsuitable for calculations of excitation energies.
+render seniority-based CI unsuitable for calculations of excitation energies.
+
+pCCD-CI for excited states \cite{Nowak_2023}
 
@@
 We further gauged the performance of hCI and sCI methods, based on either HF ground-state orbitals or state-specific orbitals, to describe excited states of closed-shell systems and radicals.
-Despite their elevated computational cost, sCI methods produced relatively poor results (even in a state-specific approach), and thus should not be useful in practice.
+Despite their elevated computational cost, sCI models produced relatively poor results (even in a state-specific approach), and thus should not be useful in practice.
 Similarly, for lower orders of hCI (hCI1 and hCI1.5), the performance is very poor,
 with ground-state (state-specific) orbitals leading to largely overestimated (underestimated) excitation energies.
-The situation improves at the state-specific $\Delta$hCI2 order, even though its accuracy is somewhat worse than $\Delta$CISD, both methods sharing the same computational scaling. 
-Unfortunately, $\Delta$hCI2.5 does not lead to more accurate results, whereas all the CI methods at next next order (CISDT, $\Delta$CISDT, hCI3, and $\Delta$hCI3) show comparable performance.
-Overall, hCI methods for excited states are either innacurate given their computational cost ($h=1$, $h=1.5$ and $h=2.5$) or comparable to excitation-based CI at the same order ($h=2$ and $h=3$).
+The situation improves at the state-specific $\Delta$hCI2 order, even though its accuracy is somewhat worse than $\Delta$CISD, both models sharing the same computational scaling. 
+Unfortunately, $\Delta$hCI2.5 does not lead to more accurate results, whereas all the CI models at next next order (CISDT, $\Delta$CISDT, hCI3, and $\Delta$hCI3) show comparable performance.
+Overall, hCI models for excited states are either innacurate given their computational cost ($h=1$, $h=1.5$ and $h=2.5$) or comparable to excitation-based CI at the same order ($h=2$ and $h=3$).
 %it is fair to conclude that

### 2023-06-19 Manuscript/manuscript.tex
saving work

@@
 Namely, for the different CI models considered here, 
-we evaluated the convergence of the non-parallelity error (NPE), the distance error, the vibrational frequencies, and the equilibrium geometries, as functions of $\Ndet$.
+we evaluated the convergence of the non-parallelity error (NPE), the distance error, the harmonic vibrational frequencies, and the equilibrium bond lengths, as functions of $\Ndet$.
 The NPE (distance error) of a given level of theory is defined as the maximum minus (plus) the minimum differences between its corresponding PEC and the FCI one, for a given range of coordinates.
-Details about how the vibrational frequencies and equilibrium geometries were obtained from the calculated PECs,
+Details about how the harmonic vibrational frequencies and equilibrium geometries were obtained from the calculated PECs,
 and the ranges defining the NPE and distance errors can be found in the \SupInf.
@@
 The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf},
-along with the NPEs, distance errors, equilibrium geometries and vibrational frequencies, plotted as a function of $\Ndet$ for the different CI models.
+along with the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies, plotted as a function of $\Ndet$ for the different CI models.
 The present results for open-shell systems are discussed in detail in the following,
@@
 
-In Tab.~\ref{tab:1} we present the statistical errors of the different methods, organized by their computational scaling.
+The statistical errors on the excitation energies, for the different CI models considered here,
+are presented in Tab.~\ref{tab:1} and Tab.~\ref{tab:2},
+%organized by their computational scaling,
+for singly excited states of closed-shell and open-shell systems, respectively.
+
 % hCI1 and hCI1.5
@@
 Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
+
+%This provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models.
 
@@
 we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
-For that, we performed calculations for a subset of 14 excited states of small molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
-with corresponding statistical errors shown in Tab.~\ref{tab:2}.
+For that, we performed calculations for a subset of 14 singly-excited states of small closed-shell molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
+with corresponding statistical errors shown in Tab.~\ref{tab:3}.
 Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order models.
@@
 }
-\label{tab:2}
+\label{tab:3}
 \begin{ruledtabular}
@@
 
-We refer to Tab.~\ref{tab:1} and ~\ref{tab:2} to discuss the performance of seniority-based CI for excited states.
-sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
-This improvement, however, is not as expressive as observed for excitataion-based and hierarchy-based methods.
-Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:1}.
-The large errors, even following the state-specific approach, combined with the formal $\order*{e^N}$ scaling of these methods,
-render seniority-based CI unsuitable for calculations of excitation energies.
+We refer back to Tab.~\ref{tab:1} to discuss the performance of seniority-based CI for excited states of closed-shell systems.
+The first model, sCI2/sCI2, systematically overestimates the excitation energies, with MAEs of \SI{1.40}{\eV}.
+There are two factors at play that explain such large errors.
+First, the ground state description is favored due to the use of ground-state HF orbitals.
+Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAEs to \SI{0.77}{\eV}, though still overestimating the excitation energies.
+The second factor is due to an unbalanced description of correlation for ground and excited states.
+%While the ground state reference (which has $s=0$) can access determinants with unpaired electrons (in the $s=2$ sector)
+While sCI2/sCI2 accounts for an additional seniority sector ($s=2$) for the ground state reference (which has $s=0$), 
+the excited state reference already lives in the same $s=2$ seniority sector.
+In other words, unpaired excitations are allowed for the ground state but not for the excited state, relative to their respective reference determinants, which creates a bias toward the former.
+A possible solution to this unbalance is to restrict the determinants to the maximum seniority number of the reference.
+This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.42}{\eV}, compared to \SI{1.40}{\eV} of sCI2/sCI2.
+However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{0.88}{\eV}.
+Ultimately, we did not find a combination of sCI model and orbitals that produced reasonable excitation energies for closed-shell systems.
+This highlights the challenge in describing on an equal footing states with qualitatively different characters (a ground state closed-shell and an excited-state open-shell).
+%sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
+%This improvement, however, is not as expressive as observed for excitataion-based and hierarchy-based methods.
+%Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:1}.
+%The large errors, even following the state-specific approach, combined with the formal $\order*{e^N}$ scaling of these methods, render seniority-based CI unsuitable for calculations of excitation energies.
+pCCD-CI for excited states \cite{Nowak_2023}
 
-pCCD-CI for excited states \cite{Nowak_2023}
+The situation for open-shell systems is quite different.
+As shown in Tab.~\ref{tab:2}, sCI delivers considerably more accurate excitation energies than for open-shell than for closed-shell systems.
+This general observation can be understood from the fact that both ground and excited states of open-shells share the same type of reference (a single $s=1$ open-shell determinant),
+which is not the case for closed-shells.
+The same type of reference provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models.
+% check for s=3 reference
+sCI1 presents a MAE of \SI{0.44}{\eV} and overestimates the excitation energy (MSE of \SI{+0.27}{\eV}), reflecting the bias toward the ground state due to ground-state orbitals.
+With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.28}{\eV} and the MSE to \SI{-0.15}{\eV}, which is smaller in absolute value than in sCI1.
+
+Even though $\Delta$sCI1 is less accurate and more expensive than CI models like $\Delta$CISD and $\Delta$hCI2, 
+its descent errors are encouraging for the development of polynomial scaling coupled cluster methods based on the concept of seniority.
+For closed-shell systems, sCI0 energies can be very well reproduced with pair coupled cluster doubles (pCCD), \cite{Kossoski_2021} a method of mean-field cost.
+Similarly, some formulation of pCCD to open-shell systems could share a connection to low-order seniority-based CI models like sCI1.
+If that is the case, then a state-specific approach of such pCCD open-shell formulation could approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.28}{\eV}) at a mean-field cost,
+serving as a better starting point to methods targeting the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.50}{\eV}).
+To the best of our knoweledge, there has been a single proposed extension of pCCD to open-shell systems,
+based on the ionization-potential equation-of-motion formalism. \cite{Boguslawski_2021}
 

### 2023-06-20 Manuscript/manuscript.tex
saving work

@@
 Core orbitals were frozen.
+For the radicals, we further impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies accounting for a set of appropriate spin-flipped determinants.
+In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS, \cite{Maurice_1996} and similarly for the other hCI and excitation-based CI models.
 We performed calculations following both the standard ground-state based CI route and the state-specific CI route. \cite{Kossoski_2023}
@@
 %For completeness, we also reproduce the geometries in the \SupInf.
+
+% comments on the different references
 
@@
 
-The statistical errors on the excitation energies, for the different CI models considered here,
+The statistical errors for the excitation energies, for the different CI models considered here,
 are presented in Tab.~\ref{tab:1} and Tab.~\ref{tab:2},
@@
 for singly excited states of closed-shell and open-shell systems, respectively.
+There are two key elements that explain the main trends on the observed statistical errors.
+The first concerns the set of molecular orbitals.
 
-% hCI1 and hCI1.5
-We found that $\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat more favourable MSE and RMSE of the latter.
-This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state,
-thus creating a bias toward underestimated excitation energies.
-Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
+hCI1 and hCI1.5 largely overestimate the excitation energies of closed-shell systems (see Tab.~\ref{tab:1}).
+The excitations in these low-order hCI models already account for some correlation of the ground state, but represent orbital relaxation of the excited state, which in turn remains less correlated.
+This favors the description of the ground state and explains the overestimated excitation energies, similarly to what is found for CISD. \cite{}
+%As will be further discussed below, this effect parallels
+A state-specific approach would be expected to eliminate this bias and reduce the errors, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
+However, we found that $\Delta$hCI1 and $\Delta$hCI1.5 produced even larger MAEs than with the ground-state based approach, though this time by largely underestimating the excitation energies.
+This can be understood from the different classes of determinants accessed from the ground state reference (closed-shell determinant) and excited state reference (single open-shell CSF).
+Lower orders of hCI therefore capture a larger fraction of correlation for the latter type of reference, and are thus not suitable for calculations of excitation energies for closed-shell systems.
+
+Ramping up to $\Delta$hCI2, this inbalance is largely reduced, even though the excitation energies are still underestimated.
+$\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat smaller absolute MSE and RMSE of the latter.
+This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state, thus creating a bias toward underestimated excitation energies.
+%Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
 
@@
 $\Delta$hCI1.5    & -2.95 & 2.95 & 3.14 \\
+$\Delta$hCI2      & -0.19 & 0.20 & 0.25 \\
 $\Delta$CISD      & -0.13 & 0.17 & 0.22 \\
-$\Delta$hCI2      & -0.19 & 0.20 & 0.25 \\
 \hline
@@
 \end{table}
+
+The statistical errors for the excitation energies of open-shell systems are shown in Tab.~\ref{tab:2}.
 
@@
 $\Delta$hCI1.5    & -0.26 & 0.66 & 0.92 \\
+$\Delta$hCI2      & -0.03 & 0.14 & 0.19 \\
 $\Delta$CISD      & +0.00 & 0.16 & 0.23 \\
-$\Delta$hCI2      & -0.03 & 0.14 & 0.19 \\
 \hline
@@
 %
+%hCI1              &       &      &      \\
+%hCI1.5            &       &      &      \\
 hCI2              & +2.89 & 3.00 & 3.28  \\
@@
 %
+%$\Delta$hCI1      &       &      &      \\
+%$\Delta$hCI1.5    &       &      &      \\
 $\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
@@
 We refer back to Tab.~\ref{tab:1} to discuss the performance of seniority-based CI for excited states of closed-shell systems.
-The first model, sCI2/sCI2, systematically overestimates the excitation energies, with MAEs of \SI{1.40}{\eV}.
-There are two factors at play that explain such large errors.
+The first model, sCI2/sCI2, systematically overestimates the excitation energies, and presents a MAEs of \SI{1.40}{\eV}.
+The same two factors discussed above for hCI explain such large errors.
 First, the ground state description is favored due to the use of ground-state HF orbitals.
 Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAEs to \SI{0.77}{\eV}, though still overestimating the excitation energies.
-The second factor is due to an unbalanced description of correlation for ground and excited states.
+Second, there is an unbalanced description of correlation for ground and excited states.
 %While the ground state reference (which has $s=0$) can access determinants with unpaired electrons (in the $s=2$ sector)
@@
 The situation for open-shell systems is quite different.
-As shown in Tab.~\ref{tab:2}, sCI delivers considerably more accurate excitation energies than for open-shell than for closed-shell systems.
+As shown in Tab.~\ref{tab:2}, sCI delivers considerably more accurate excitation energies for open-shell than for closed-shell systems.
 This general observation can be understood from the fact that both ground and excited states of open-shells share the same type of reference (a single $s=1$ open-shell determinant),
-which is not the case for closed-shells.
-The same type of reference provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models.
+thus significantly minimizing an unbalanced account of correlation, encountered for excitations from closed-shells.
+The same type of reference provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models applied to excitations of open-shell systems.
 % check for s=3 reference
 sCI1 presents a MAE of \SI{0.44}{\eV} and overestimates the excitation energy (MSE of \SI{+0.27}{\eV}), reflecting the bias toward the ground state due to ground-state orbitals.
-With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.28}{\eV} and the MSE to \SI{-0.15}{\eV}, which is smaller in absolute value than in sCI1.
+With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.28}{\eV} and the MSE to \SI{-0.15}{\eV}, which is smaller in absolute value than found for sCI1.
 

### 2023-06-27 Manuscript/manuscript.tex
update
saving work

@@
 
-The statistical errors for the excitation energies, for the different CI models considered here,
-are presented in Tab.~\ref{tab:1} and Tab.~\ref{tab:2},
-%organized by their computational scaling,
+For each CI model considered here, we evaluated the mean signed error (MSE), mean absolute error (MAE), and root-mean square error (RMSE) associated with the excitation energy.
+The statistical errors are presented in Tab.~\ref{tab:1} and Tab.~\ref{tab:2},
 for singly excited states of closed-shell and open-shell systems, respectively.
-There are two key elements that explain the main trends on the observed statistical errors.
-The first concerns the set of molecular orbitals.
+For the lower order models, the calculations were performed for 46 (closed-shell) and 20 (open-shell) excited states.
+In turn, a subset of 14 (closed-shell) and 5 (open-shell) excited states was considered for the higher order models, given their more intensive computational cost.
+Even though these subsets are too small for meaningful absolute statistics, they can still show the main trends on the higher order CI models.
+%A detailed comparison of excitation-based CI along the ground-state and the state-specific routes can be found elsewhere. \cite{Kossoski_2023}
+%The focus of the present discussion lies on the comparison between hCI, sCI, and excitation-based CI, for both routes,
+%
+%There are two key elements that explain the main trends on the observed statistical errors. The first concerns the set of molecular orbitals.
 
-hCI1 and hCI1.5 largely overestimate the excitation energies of closed-shell systems (see Tab.~\ref{tab:1}).
-The excitations in these low-order hCI models already account for some correlation of the ground state, but represent orbital relaxation of the excited state, which in turn remains less correlated.
-This favors the description of the ground state and explains the overestimated excitation energies, similarly to what is found for CISD. \cite{}
-%As will be further discussed below, this effect parallels
-A state-specific approach would be expected to eliminate this bias and reduce the errors, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
-However, we found that $\Delta$hCI1 and $\Delta$hCI1.5 produced even larger MAEs than with the ground-state based approach, though this time by largely underestimating the excitation energies.
-This can be understood from the different classes of determinants accessed from the ground state reference (closed-shell determinant) and excited state reference (single open-shell CSF).
-Lower orders of hCI therefore capture a larger fraction of correlation for the latter type of reference, and are thus not suitable for calculations of excitation energies for closed-shell systems.
+We start the discussion with the excitations from closed-shell systems, with corresponding statistics shown in Tab.~\ref{tab:1}.
+The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.07}{\eV}.
+This model shares the same computational scaling with CIS, but includes the paired double excitations, 
+which clearly do not improve the descent CIS results (MAE of \SI{0.61}{\eV}).
+Moving one rank up, to hCI1.5, the MAE increases to \SI{1.91}{\eV}, and reaches a maxium of \SI{3.53}{\eV} at the hCI2 level.
+From that point on, the errors systematically decrease.
+hCI2.5 is stil as inaccurate (MAE of \SI{1.95}{\eV}) as hCI1.5.
+Finally, at the hCI3 level, significantly smaller errors are achieved (MAE of \SI{0.19}{\eV}),
+which are further reduced at the hCI3.5 level (MAE of \SI{0.11}{\eV}).
 
-Ramping up to $\Delta$hCI2, this inbalance is largely reduced, even though the excitation energies are still underestimated.
-$\Delta$hCI2 and $\Delta$CISD display comparable MAEs, despite the somewhat smaller absolute MSE and RMSE of the latter.
-This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state, thus creating a bias toward underestimated excitation energies.
+The errors of the hCI models thus increase as one augments the hierarchy parameter $h$, reaching a maximum at hCI2, and then decrease towards higher orders.
+This behavior paralles what is found with excitation-based CI, where CISD is far less accurate than CIS, which in turn is less accurate than CISDT and CISDTQ.
+This can be understood based on the role of the excited determinants for ground and excited states.
+%
+The excited determinants of the low-order hCI models (hCI1 and hCI1.5) already account for some correlation of the ground state, but represent mostly orbital relaxation of the excited state, which thus remains less correlated.
+This favored description of the ground state explains the overestimated excitation energies.
+The effect becomes exagerated at the hCI2 (and CISD) models, since it captures most of the ground state correlation through the unpaired double excitations.
+Higher orders models are needed to recover most of the excited state correlation and the excitation energy errors thus decrease.
+%(though always overestimating the reference values).
+The same argument also explains why hCI systematically overestimates the excitation energies, just as excitation-based CI does.
+
+The trends discussed above are directly linked with the use of ground-state HF orbitals.
+Employing state-specific orbitals would be expected to suppress this bias and improve the accuracy, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
+For the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, though by underestimating the excitation energies.
+Besides the set of orbitals, a second key aspect must be considered.
+For a given hierarchy parameter $h$, different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF).
+The case of $\Delta$hCI1 is illustrated in Fig.~\ref{fig:DhCI1}.
+%For the ground state calculation, it accounts for two classes of excited determinants, generated by single excitation and by paired double excitation from the closed-shell reference.
+%For the singly-excited state, four types of determinants are generated. On top of the single and paired double excitations from the doubly-occupied orbitals, 
+There is arguably more diversity in the types of determinants employed for the excited-state calculation than for the ground-state one.
+Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the former calculation.
+%and are thus not suitable for calculations of excitation energies for closed-shell systems.
+Although hCI is constructed to account for all classes of determinants whose number share the same computational scaling,
+at the lower orders this procedure does not lead to a balanced description of correlation, at least for excitations of closed-shell systems.
+As will be discussed later for the excitations of open-shell systems, this inbalance probably arises from the distinct reference determinants employed for the closed-shell excitations.
+%
+Another important point to realize is that a state-specific approach is not necessarily more accurate than a ground-state based approach.
+Both the orbitals and the classes of determinants of a given CI model play an important role in the fine balance of ground- and excited-state correlation effects.
+
+Ramping up to $\Delta$hCI2, this inbalance is reduced to a great extent, even though the excitation energies are still underestimated.
 %Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
+$\Delta$hCI2 display a MAE of \SI{0.20}{\eV}, comparable to that of $\Delta$CISD (\SI{0.17}{\eV}) display comparable MAEs.
+Albeit small, this difference is probably statistically significant, and $\Delta$CISD should still outperform $\Delta$hCI2 by some margin.
+Similarly, $\Delta$hCI2 has a more negative MSE (\SI{-0.19}{\eV} than $\Delta$CISD (\SI{-0.13}{\eV})
+%This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state, thus creating a bias toward underestimated excitation energies.
+Ramping up to $\Delta$hCI2.5 increases the errors (MAE of \SI{0.27}{\eV}),
+which is probably due to another layer of inbalance concerning the (ground-state and excited-state) determinants that enter at this stage.
+(Despite the different number of states considered, the observed difference between $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant.)
+The situation improves at $\Delta$hCI3 (MAE of \SI{0.21}{\eV}),
+while $\Delta$hCI3.5 generates considerably more accurate results (MAE of \SI{0.08}{\eV}),
 
-%This provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models.
-
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of 
-38 Singly Excited States of Closed-Shell Systems and 
-Listed in the {\SupInf}.
-}
-\label{tab:1}
-\begin{ruledtabular}
-\begin{tabular}{lddd}
-method            & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
-\hline
-$\Delta$CSF       & -0.75 & 0.81 & 0.95 \\
-CIS               & +0.03 & 0.62 & 0.63 \\
-\hline
-hCI1              & +1.08 & 1.12 & 1.25 \\
-$\Delta$hCI1      & -1.37 & 1.41 & 1.52 \\
-hCI1.5            & +1.95 & 1.95 & 2.03 \\
-$\Delta$hCI1.5    & -2.95 & 2.95 & 3.14 \\
-$\Delta$hCI2      & -0.19 & 0.20 & 0.25 \\
-$\Delta$CISD      & -0.13 & 0.17 & 0.22 \\
-\hline
-sCI2/sCI2         & +1.40 & 1.40 & 1.53 \\
-sCI2/sCI0         & -0.22 & 0.42 & 0.57 \\
-$\Delta$sCI2/sCI2 & +0.74 & 0.77 & 0.85 \\
-$\Delta$sCI2/sCI0 & -0.88 & 0.88 & 0.96 \\
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-
-The statistical errors for the excitation energies of open-shell systems are shown in Tab.~\ref{tab:2}.
-
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of 
-16 Singly-Excited States from Open-Shell Doublets 
-Listed in the {\SupInf}.
-}
-\label{tab:2}
-\begin{ruledtabular}
-\begin{tabular}{lddd}
-method            & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
-\hline
-$\Delta$CSF       & -0.09 & 0.50 & 0.67 \\
-CIS               & +0.30 & 0.41 & 0.55 \\
-\hline
-hCI1              & +0.76 & 0.80 & 1.02 \\
-$\Delta$hCI1      & -0.09 & 0.58 & 0.94 \\
-hCI1.5            & +1.22 & 1.22 & 1.45 \\
-$\Delta$hCI1.5    & -0.26 & 0.66 & 0.92 \\
-$\Delta$hCI2      & -0.03 & 0.14 & 0.19 \\
-$\Delta$CISD      & +0.00 & 0.16 & 0.23 \\
-\hline
-sCI1              & +0.27 & 0.44 & 0.66 \\
-$\Delta$sCI1      & -0.15 & 0.28 & 0.40 \\
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-
-Despite the rather poor performance of lower-order hCI models for excitation energies,
-we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
-For that, we performed calculations for a subset of 14 singly-excited states of small closed-shell molecular systems (\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
-with corresponding statistical errors shown in Tab.~\ref{tab:3}.
-Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order models.
-
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
-for the Set of 14 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
-}
-\label{tab:3}
-\begin{ruledtabular}
-\begin{tabular}{lddd}
-method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
-\hline
-CISD              & +4.09 & 4.09 & 4.18  \\
-CISDT             & +0.09 & 0.18 & 0.20  \\
-CISDTQ            & +0.14 & 0.14 & 0.16  \\
-%
-$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
-$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
-$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
-%
-%hCI1              &       &      &      \\
-%hCI1.5            &       &      &      \\
-hCI2              & +2.89 & 3.00 & 3.28  \\
-hCI2.5            & +1.93 & 1.93 & 2.05  \\
-hCI3              & +0.18 & 0.18 & 0.19  \\
-%
-%$\Delta$hCI1      &       &      &      \\
-%$\Delta$hCI1.5    &       &      &      \\
-$\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
-$\Delta$hCI2.5    & -0.27 & 0.27 & 0.30  \\
-$\Delta$hCI3      & -0.21 & 0.21 & 0.24  \\
-$\Delta$hCI3.5    & -0.08 & 0.08 & 0.09  \\
-\end{tabular}
-\end{ruledtabular}
-\end{table}
+%Despite the rather poor performance of lower-order hCI models for excitation energies,
+%we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
+%For that, we performed calculations for a subset of 14 singly-excited states of small closed-shell molecular systems (indicated in the {\SupInf}).
+%(\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
+%with corresponding statistical errors shown in Tab.~\ref{tab:1}.
+%Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order models.
 
@@
 
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of 
+Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+}
+\label{tab:1}
+\begin{ruledtabular}
+\begin{tabular}{ldddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
+\hline
+CIS               & 46  & +0.04 & 0.61 & 0.60 \\
+CISD              & 14  & +4.09 & 4.09 & 4.18 \\
+CISDT             & 14  & +0.10 & 0.19 & 0.20 \\
+CISDTQ            & 14  & +0.15 & 0.15 & 0.17 \\
+%
+\hline
+hCI1              & 46  & +0.97 & 1.07 & 1.20 \\
+hCI1.5            & 46  & +1.91 & 1.91 & 2.00 \\
+hCI2              & 14  & +2.99 & 3.53 & 3.61 \\
+hCI2.5            & 14  & +1.95 & 1.95 & 2.06 \\
+hCI3              & 14  & +0.19 & 0.19 & 0.21 \\
+hCI3.5            & 14  & +0.11 & 0.11 & 0.13 \\
+%
+\hline
+$\Delta$CSF       & 46  & -0.73 & 0.80 & 0.94 \\
+$\Delta$CISD      & 46  & -0.13 & 0.17 & 0.22 \\
+$\Delta$CISDT     & 14  & -0.20 & 0.20 & 0.22 \\
+$\Delta$CISDTQ    & 14  & -0.02 & 0.02 & 0.02 \\
+%
+\hline
+$\Delta$hCI1      & 46  & -1.36 & 1.40 & 1.50 \\
+$\Delta$hCI1.5    & 46  & -2.91 & 2.91 & 3.10 \\
+$\Delta$hCI2      & 46  & -0.19 & 0.20 & 0.25 \\
+$\Delta$hCI2.5    & 14  & -0.27 & 0.27 & 0.30 \\
+$\Delta$hCI3      & 14  & -0.21 & 0.21 & 0.24 \\
+$\Delta$hCI3.5    & 14  & -0.08 & 0.08 & 0.09 \\
+%
+\hline
+sCI2/sCI2         & 46  & +1.26 & 1.34 & 1.49 \\
+sCI2/sCI0         & 46  & -0.41 & 0.57 & 0.89 \\
+$\Delta$sCI2/sCI2 & 46  & +0.69 & 0.79 & 0.90 \\
+$\Delta$sCI2/sCI0 & 46  & -0.99 & 0.99 & 1.15 \\
+%
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+We now move to the discussion about the excitation energies of open-shell systems.
+The statistical errors are shown in Tab.~\ref{tab:2}.
+%
 The case of $\Delta$hCI2 was discussed above for the larger set of excitation energies.
@@
 
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of 
+Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
+}
+\label{tab:2}
+\begin{ruledtabular}
+\begin{tabular}{ldddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
+\hline
+CIS               & 20 & +0.33 & 0.36 & 0.50 \\
+CISD              & 5  & +1.98 & 1.98 & 2.16 \\
+CISDT             & 5  & +0.02 & 0.03 & 0.05 \\
+CISDTQ            & 5  & +0.03 & 0.03 & 0.05 \\
+%
+\hline
+hCI1              & 20 & +0.70 & 0.77 & 0.96 \\
+hCI1.5            & 20 & +1.11 & 1.11 & 1.30 \\
+hCI2              & 5  & +0.69 & 0.69 & 0.85 \\
+hCI2.5            & 5  & +0.01 & 0.03 & 0.04 \\
+hCI3              & 5  & -0.00 & 0.02 & 0.03 \\
+hCI3.5            & 5  & -0.00 & 0.01 & 0.02 \\
+hCI4              & 5  &       &      &      \\
+%
+\hline
+$\Delta$CSF       & 20 & -0.04 & 0.41 & 0.58 \\
+$\Delta$CISD      & 20 & +0.01 & 0.12 & 0.21 \\
+$\Delta$CISDT     & 5  & -0.05 & 0.05 & 0.09 \\
+$\Delta$CISDTQ    & 5  & -0.02 & 0.02 & 0.02 \\
+%
+\hline
+$\Delta$hCI1      & 20 & -0.19 & 0.34 & 0.48 \\
+$\Delta$hCI1.5    & 20 & -0.10 & 0.46 & 0.70 \\
+$\Delta$hCI2      & 20 & -0.01 & 0.11 & 0.16 \\
+$\Delta$hCI2.5    & 5  & -0.06 & 0.06 & 0.12 \\
+$\Delta$hCI3      & 5  & -0.03 & 0.03 & 0.06 \\
+$\Delta$hCI3.5    & 5  & -0.02 & 0.02 & 0.02 \\
+$\Delta$hCI4      & 5  &       &      &      \\
+%
+\hline
+sCI1              & 20 & +0.24 & 0.39 & 0.61 \\
+$\Delta$sCI1      & 20 & -0.11 & 0.27 & 0.38 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+%\begin{table}[ht!]
+%\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
+%for the Set of 14 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+%}
+%\label{tab:3}
+%\begin{ruledtabular}
+%\begin{tabular}{lddd}
+%method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
+%\hline
+%CISD              & +4.09 & 4.09 & 4.18  \\
+%CISDT             & +0.09 & 0.18 & 0.20  \\
+%CISDTQ            & +0.14 & 0.14 & 0.16  \\
+%$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
+%$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
+%$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
+%\hline
+%hCI1              & +1.62 & 1.62 & 1.65 \\
+%hCI1.5            & +2.10 & 2.10 & 2.11 \\
+%hCI2              & +2.89 & 3.00 & 3.28  \\
+%hCI2.5            & +1.93 & 1.93 & 2.05  \\
+%hCI3              & +0.18 & 0.18 & 0.19  \\
+%$\Delta$hCI1      & -1.37 & 1.50 & 1.66 \\
+%$\Delta$hCI1.5    & -2.62 & 2.62 & 2.72 \\
+%$\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
+%$\Delta$hCI2.5    & -0.27 & 0.27 & 0.30  \\
+%$\Delta$hCI3      & -0.21 & 0.21 & 0.24  \\
+%$\Delta$hCI3.5    & -0.08 & 0.08 & 0.09  \\
+%\end{tabular}
+%\end{ruledtabular}
+%\end{table}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%This provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models.
 

### 2023-06-28 Manuscript/manuscript.tex
saving work

@@
 \cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022}
-and coupled-cluster 
+and coupled-cluster (CC)
 \cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023}
@@
 Furthermore, the reduction of the errors is overall more monotonic than without the perturbative correction,
-which is illustrated in Fig.\ref{} for the particular case of \ce{F2}.
+which is outlined in Fig.\ref{} for the particular case of \ce{F2}.
 For the sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation-based and hCI alternatives.
@@
 
-For each CI model considered here, we evaluated the mean signed error (MSE), mean absolute error (MAE), and root-mean square error (RMSE) associated with the excitation energy.
+% intro
+
+For each CI model considered here, we evaluated the mean signed error (MSE), mean absolute error (MAE), and root-mean square error (RMSE) 
+with respect to reference theoretical values for the excitation energies.
 The statistical errors are presented in Tab.~\ref{tab:1} and Tab.~\ref{tab:2},
@@
 %The focus of the present discussion lies on the comparison between hCI, sCI, and excitation-based CI, for both routes,
+
+% ground-state based hCI
+
+We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Tab.~\ref{tab:1}.
 %
-%There are two key elements that explain the main trends on the observed statistical errors. The first concerns the set of molecular orbitals.
-
-We start the discussion with the excitations from closed-shell systems, with corresponding statistics shown in Tab.~\ref{tab:1}.
 The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.07}{\eV}.
@@
 Moving one rank up, to hCI1.5, the MAE increases to \SI{1.91}{\eV}, and reaches a maxium of \SI{3.53}{\eV} at the hCI2 level.
-From that point on, the errors systematically decrease.
-hCI2.5 is stil as inaccurate (MAE of \SI{1.95}{\eV}) as hCI1.5.
-Finally, at the hCI3 level, significantly smaller errors are achieved (MAE of \SI{0.19}{\eV}),
-which are further reduced at the hCI3.5 level (MAE of \SI{0.11}{\eV}).
+Here we notice that hCI2 delivers somewhat smaller errors than CISD (MAE of \SI{4.09}{\eV}), even though both are way too large.
+From that point on, the next hCI models generate progressively better results.
+Despite the improvement with respect to hCI2, hCI2.5 is stil as inaccurate (MAE of \SI{1.95}{\eV}) as hCI1.5.
+At the hCI3 level, significantly smaller errors are finally achieved (MAE of \SI{0.19}{\eV}).
+This is quite close to that obtained with the excitation-based model of same order, CISDT (\SI{0.17}{\eV}).
+Even smaller errors are produced at the hCI3.5 level (MAE of \SI{0.11}{\eV}), but at a considerable computational cost.
 
-The errors of the hCI models thus increase as one augments the hierarchy parameter $h$, reaching a maximum at hCI2, and then decrease towards higher orders.
+% first origin of the unbalance hCI
+
+%There are two key elements that explain the main trends on the observed statistical errors. The first concerns the set of molecular orbitals.
+The errors of the hCI models first increase as one augments the hierarchy parameter $h$, reaching a maximum at hCI2, and then decrease towards higher orders.
 This behavior paralles what is found with excitation-based CI, where CISD is far less accurate than CIS, which in turn is less accurate than CISDT and CISDTQ.
 This can be understood based on the role of the excited determinants for ground and excited states.
-%
-The excited determinants of the low-order hCI models (hCI1 and hCI1.5) already account for some correlation of the ground state, but represent mostly orbital relaxation of the excited state, which thus remains less correlated.
+The excited determinants of the low-order hCI models (hCI1 and hCI1.5) already account for some correlation of the ground state,
+but represent mostly orbital relaxation of the excited state, which thus remains less correlated.
 This favored description of the ground state explains the overestimated excitation energies.
-The effect becomes exagerated at the hCI2 (and CISD) models, since it captures most of the ground state correlation through the unpaired double excitations.
-Higher orders models are needed to recover most of the excited state correlation and the excitation energy errors thus decrease.
+The effect becomes exagerated at the hCI2 (and CISD) models, since it captures most of the ground state correlation through the unpaired double excitations,
+but still little correlation for the excited state.
+Higher orders models are needed to recover most of the excited state correlation, and at this point the errors on the excitation energies start to decrease.
 %(though always overestimating the reference values).
-The same argument also explains why hCI systematically overestimates the excitation energies, just as excitation-based CI does.
+The same argument also explains why hCI systematically overestimates the excitation energies, just as for excitation-based CI.
 
-The trends discussed above are directly linked with the use of ground-state HF orbitals.
-Employing state-specific orbitals would be expected to suppress this bias and improve the accuracy, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
-For the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, though by underestimating the excitation energies.
-Besides the set of orbitals, a second key aspect must be considered.
+% state-specific hCI and second origin of the unbalance
+
+The trend discussed above stems from the use of ground-state HF orbitals.
+Employing state-specific orbitals would be expected to suppress the bias toward the ground state and lead to improved results, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
+However, for the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, though by underestimating the excitation energies.
+%Besides the set of orbitals, a second key aspect must be considered.
+Besides the set of orbitals, the types of determinants included at each order play an equally important role and explains our observation.
 For a given hierarchy parameter $h$, different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF).
-The case of $\Delta$hCI1 is illustrated in Fig.~\ref{fig:DhCI1}.
+The case of $\Delta$hCI1 is portrayed in Fig.~\ref{fig:DhCI1}.
 %For the ground state calculation, it accounts for two classes of excited determinants, generated by single excitation and by paired double excitation from the closed-shell reference.
@@
 There is arguably more diversity in the types of determinants employed for the excited-state calculation than for the ground-state one.
-Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the former calculation.
+Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the excited state than for ground state.
 %and are thus not suitable for calculations of excitation energies for closed-shell systems.
@@
 Another important point to realize is that a state-specific approach is not necessarily more accurate than a ground-state based approach.
-Both the orbitals and the classes of determinants of a given CI model play an important role in the fine balance of ground- and excited-state correlation effects.
+Both the orbitals and the classes of determinants of a given CI model control the fine balance of ground- and excited-state correlation effects.
 
-Ramping up to $\Delta$hCI2, this inbalance is reduced to a great extent, even though the excitation energies are still underestimated.
+%Ramping up to $\Delta$hCI2, this inbalance is reduced to a great extent.
+%even though the excitation energies are still underestimated.
 %Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
-$\Delta$hCI2 display a MAE of \SI{0.20}{\eV}, comparable to that of $\Delta$CISD (\SI{0.17}{\eV}) display comparable MAEs.
-Albeit small, this difference is probably statistically significant, and $\Delta$CISD should still outperform $\Delta$hCI2 by some margin.
-Similarly, $\Delta$hCI2 has a more negative MSE (\SI{-0.19}{\eV} than $\Delta$CISD (\SI{-0.13}{\eV})
+
+% state-specific vs. ground-state hCI
+
+The state-specific advantage only shows up at the $\Delta$hCI2 level, which is substantially more accurate than hCI2 (MAEs of \SI{0.20}{\eV} and \SI{3.53}{\eV}).
+This improvement parallels our previous finding on the excitation-based CI models of the same order, CISD and $\Delta$CISD. \cite{Kossoski_2023}
+Even though hCI2 is somewhat more accurate than CISD (both have large errors), their state-specific versions share a comparable performance.
+Actually, $\Delta$hCI2 is slightly less accurate than $\Delta$CISD (MAEs of \SI{0.20}{\eV} and \SI{0.17}{\eV}).
+Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.19}{\eV} against \SI{-0.13}{\eV}).
+Albeit small, these differences are statistically significant, and $\Delta$CISD would still be preferable to $\Delta$hCI2 by some margin.
+
 %This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state, thus creating a bias toward underestimated excitation energies.
 Ramping up to $\Delta$hCI2.5 increases the errors (MAE of \SI{0.27}{\eV}),
-which is probably due to another layer of inbalance concerning the (ground-state and excited-state) determinants that enter at this stage.
-(Despite the different number of states considered, the observed difference between $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant.)
+which probably reflects another set of unbalanced state-specific determinants that enters at this stage.
+(Despite the differing number of states considered, the observed variation on the accuracy of $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant.)
+Even though $\Delta$hCI2.5 is much more accurate than hCI2.5, $\Delta$hCI2 or $\Delta$CISD remain cheaper and more accurate.
+
 The situation improves at $\Delta$hCI3 (MAE of \SI{0.21}{\eV}),
-while $\Delta$hCI3.5 generates considerably more accurate results (MAE of \SI{0.08}{\eV}),
+while $\Delta$hCI3.5 generates considerably more accurate results (MAE of \SI{0.08}{\eV}).
+There is no gain in going from hCI3 to $\Delta$hCI3 and from hCI3.5 to $\Delta$hCI3.5,
+similarly to what had been found for CISDT and $\Delta$CISDT. \cite{Kossoski_2023}
+%which also parallels the comparison between CISDT and $\Delta$CISDT. \cite{Kossoski_2023}
+We notice, however, that the state-specific route presents negative MSEs, in contrast to the positives values obtained with the ground-state based route.
+Furthermore, excitation-based CI and hCI perform similarly at this order (the former perhaps slightly better),
+as also observed for $\Delta$CISD and $\Delta$hCI2.
 
-%Despite the rather poor performance of lower-order hCI models for excitation energies,
+%As discussed above, adopting the state-specific approach at lower-order hCI models ($\Delta$hCI1 and $\Delta$hCI1.5)
+%actually worsens the results with respect to the ground-state approach (hCI1 and hCI1.5).
+
 %we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
@@
 CISD              & 14  & +4.09 & 4.09 & 4.18 \\
-CISDT             & 14  & +0.10 & 0.19 & 0.20 \\
+CISDT             & 14  & +0.12 & 0.17 & 0.18 \\
 CISDTQ            & 14  & +0.15 & 0.15 & 0.17 \\
@@
 $\Delta$hCI2.5    & 14  & -0.27 & 0.27 & 0.30 \\
-$\Delta$hCI3      & 14  & -0.21 & 0.21 & 0.24 \\
+$\Delta$hCI3      & 14  & -0.22 & 0.22 & 0.24 \\
 $\Delta$hCI3.5    & 14  & -0.08 & 0.08 & 0.09 \\
@@
 
-We now move to the discussion about the excitation energies of open-shell systems.
+We now shift to excitations for open-shell systems.
 The statistical errors are shown in Tab.~\ref{tab:2}.
-%
-The case of $\Delta$hCI2 was discussed above for the larger set of excitation energies.
-Here we further notice that hCI2 was found to be somewhat more accurate than CISD, even though the errors are still too large.
-However, this improvement does not reflect on the state specific route (compare $\Delta$hCI2 and $\Delta$CISD in Tabs.~\ref{tab:1} and ~\ref{tab:2}).
-Moving to the next order, even though the state-specific route ($\Delta$hCI2.5) is far more accurate than the ground-state based route (hCI2.5),
-it is actually less accurate then $\Delta$hCI2.
-An improved performance is found for hCI3, $\Delta$hCI3, CISDT, and $\Delta$CISDT, which present quite similar MAEs and RMSEs.
-There is thus no advantage in moving from excitation-based to hCI at this order, similarly to what discussed before for the $\Delta$hCI2 and $\Delta$CISD models.
-Moreover, $\Delta$hCI3 is not more accurate than hCI3, as also observed for $\Delta$CISDT and CISDT. \cite{Kossoski_2023}
-We notice, however, that the state-specific models present negative MSEs, in contrast to the positives values obtained in the ground-state based route.
-This reflects an umbalanced description of correlation effects, favoring the excited states.
-At the next order, $\Delta$hCI3.5 has significantly smaller errors, but at a much more elevated computational cost.
+%in the following comparison to the results for closed-shell excitations,
+%In comparison to the previous discussion,
+The key difference on the calculations for open-shell excitations is that the same type of reference determinant were employed for ground- and excited-state calculations,
+whereas the closed-shell excitations relied on different types of reference determinants.
+This accounts for the more accurate results for open-shell excitations and explains most of the differences with respect to excitations from closed-shell systems, outlined in detail below.
+
+%We first point out the similarities and later the key differences with respect to the closed-shell discussion.
+
+When using ground-state orbitals, we found the performance initially degrates and later improves as the order increases, for both excitation-based CI and hCI.
+This is similar to what was observed for the closed-shell excitations, 
+and can be traced back to a bias towards the ground state due to the choice of orbitals.
+%more ground-state correlation being recovered at the lower orders since one starts from ground-state orbitals.
+An important difference, however, is that the maximum error appears at a lower order (hCI1.5) and is significantly smaller (MAE of \SI{1.11}{\eV}) for open-shells
+than for closed-shells (at hCI2 with a MAE of \SI{3.53}{\eV}).
+This reflects the choice of reference, since both ground and excited states of the open shells considered here are qualitatively described with the same type of determinant.
+%This reflects the fact that ground and excited states of the open shells considered here are qualitatively described with the same type of determinant.
+For the same reason, the MSEs are systematically smaller for the open-shell excitations.
+
+Once state-specific orbitals are employed, there is no longer a bias towards the ground state related to the orbitals.
+In a big contrast to the closed-shell case, here the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.34}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.46}{\eV}), 
+become more accurate than their ground-state based counterparts, hCI1 (MAE of \SI{0.77}{\eV}) and hCI1.5 (MAE of \SI{1.11}{\eV}).
+Again, thanks to the same type of reference in ground- and excited-state calculations.
+The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:DhCI1}, in comparison to the more unbalanced case of excitation from closed-shells.
+We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5).
+%and that $\Delta$hCI1 is slightly more accurate than $\Delta$hCI1.5.
+
+This residual bias disappears at the next order, $\Delta$hCI2, whereas for the closed-shell systems even $\Delta$hCI3.5 still understimated excitation energies. 
+The same holds for excitation-based CI.
+For open-shells, $\Delta$hCI2 is arguably slightly more accurate (MAE of \SI{0.11}{\eV}) than $\Delta$CISD (MAE of \SI{0.12}{\eV}), also based on the larger differences on the RMSEs.
+In further contrast to the closed-shell case,
+$\Delta$hCI2.5 is more accurate than $\Delta$hCI2, producing a MAE of \SI{0.06}{\eV}.
+(Accounting only for the 5 states considered for $\Delta$hCI2.5, $\Delta$hCI2 would have a MAE of \SI{0.09}{\eV}.)
+The statistics at the next orders ($h=3$ and $h=3.5$) is limited, 
+but suggests a possible advantage of the hCI models against the excitation-based ones,
+and that the ground-state routes perform slightly better than the state-specific ones.
 
@@
 hCI3.5            & 5  & -0.00 & 0.01 & 0.02 \\
-hCI4              & 5  &       &      &      \\
+%hCI4              & 5  &       &      &      \\
 %
@@
 $\Delta$hCI3.5    & 5  & -0.02 & 0.02 & 0.02 \\
-$\Delta$hCI4      & 5  &       &      &      \\
+%$\Delta$hCI4      & 5  &       &      &      \\
 %
@@
 \end{table}
+
+Methods that make use of the concept of seniority to generate excited determinants 
+are not invariant under orbital rotations within the occupied and virtual subspaces. \cite{Bytautas_2011,Limacher_2014,Stein_2014}
+This applies for the hCI and sCI models considered here.
+Here we employed HF ground-state orbitals or excited-state orbitals optimized for a minimal reference. \cite{Kossoski_2023}
+Further exploiting the rotations within occupied and virtual subspaces could render more suitable orbitals for hCI and sCI calculations.
+Alternatively, one could fully optimize the orbitals at a given CI model.
 
@@
 We refer back to Tab.~\ref{tab:1} to discuss the performance of seniority-based CI for excited states of closed-shell systems.
-The first model, sCI2/sCI2, systematically overestimates the excitation energies, and presents a MAEs of \SI{1.40}{\eV}.
+The first model, sCI2/sCI2, systematically overestimates the excitation energies, and presents a MAEs of \SI{1.34}{\eV}.
 The same two factors discussed above for hCI explain such large errors.
 First, the ground state description is favored due to the use of ground-state HF orbitals.
-Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAEs to \SI{0.77}{\eV}, though still overestimating the excitation energies.
+Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAE to \SI{0.79}{\eV}, though still overestimating the excitation energies.
 Second, there is an unbalanced description of correlation for ground and excited states.
@@
 While sCI2/sCI2 accounts for an additional seniority sector ($s=2$) for the ground state reference (which has $s=0$), 
-the excited state reference already lives in the same $s=2$ seniority sector.
-In other words, unpaired excitations are allowed for the ground state but not for the excited state, relative to their respective reference determinants, which creates a bias toward the former.
+no further electron unpaired is allowed for the excited state, whose reference already lives in the $s=2$ seniority sector.
+In other words, unpaired excitations are allowed for the ground state but not for the excited state, relative to their respective reference determinants, which creates a bias towards the former.
 A possible solution to this unbalance is to restrict the determinants to the maximum seniority number of the reference.
-This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.42}{\eV}, compared to \SI{1.40}{\eV} of sCI2/sCI2.
-However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{0.88}{\eV}.
+This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.57}{\eV}, compared to \SI{1.34}{\eV} of sCI2/sCI2.
+However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{0.99}{\eV}.
+Clearly, the $s=2$ sector captures more correlation for the excited states than the $s=0$ does for the ground state.
 Ultimately, we did not find a combination of sCI model and orbitals that produced reasonable excitation energies for closed-shell systems.
-This highlights the challenge in describing on an equal footing states with qualitatively different characters (a ground state closed-shell and an excited-state open-shell).
 %sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
@@
 
-The situation for open-shell systems is quite different.
-As shown in Tab.~\ref{tab:2}, sCI delivers considerably more accurate excitation energies for open-shell than for closed-shell systems.
-This general observation can be understood from the fact that both ground and excited states of open-shells share the same type of reference (a single $s=1$ open-shell determinant),
-thus significantly minimizing an unbalanced account of correlation, encountered for excitations from closed-shells.
-The same type of reference provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models applied to excitations of open-shell systems.
+The situation for open-shell systems is quite different, in close parallel to the previous discussion of hCI.
+As shown in Tab.~\ref{tab:2}, sCI provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
+The reason should not be surprising at this point:
+both ground and excited states of open-shells share the same type of reference (a single $s=1$ open-shell determinant).
 % check for s=3 reference
-sCI1 presents a MAE of \SI{0.44}{\eV} and overestimates the excitation energy (MSE of \SI{+0.27}{\eV}), reflecting the bias toward the ground state due to ground-state orbitals.
-With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.28}{\eV} and the MSE to \SI{-0.15}{\eV}, which is smaller in absolute value than found for sCI1.
+Moreover, the state-specific approach is clearly superior.
+sCI1 presents a MAE of \SI{0.39}{\eV} and overestimates the excitation energy (MSE of \SI{+0.24}{\eV}), reflecting the bias toward the ground state due to ground-state orbitals.
+With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.27}{\eV} and the MSE to \SI{-0.11}{\eV}, which is smaller in absolute value than found for sCI1.
 
-Even though $\Delta$sCI1 is less accurate and more expensive than CI models like $\Delta$CISD and $\Delta$hCI2, 
-its descent errors are encouraging for the development of polynomial scaling coupled cluster methods based on the concept of seniority.
-For closed-shell systems, sCI0 energies can be very well reproduced with pair coupled cluster doubles (pCCD), \cite{Kossoski_2021} a method of mean-field cost.
-Similarly, some formulation of pCCD to open-shell systems could share a connection to low-order seniority-based CI models like sCI1.
-If that is the case, then a state-specific approach of such pCCD open-shell formulation could approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.28}{\eV}) at a mean-field cost,
-serving as a better starting point to methods targeting the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.50}{\eV}).
-To the best of our knoweledge, there has been a single proposed extension of pCCD to open-shell systems,
-based on the ionization-potential equation-of-motion formalism. \cite{Boguslawski_2021}
+Even though $\Delta$sCI1 is less accurate and has less favourable computational scaling than CI models like $\Delta$CISD and $\Delta$hCI2,
+its descent errors are encouraging for another reason.
+the development of polynomial scaling CC methods based on the concept of seniority.
+For closed-shell systems, sCI0 energies can be very well reproduced with pair coupled cluster doubles (pCCD), \cite{Kossoski_2021} a method that has mean-field cost.
+Similarly, a formulation of pCCD to open-shell systems might share a analogous connection to low-order seniority-based CI models like sCI1.
+If that is the case, then a state-specific approach of such pCCD formulation adapted to open-shells may approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.27}{\eV}) at a mean-field cost.
+This method could then provide a better starting point to recover the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.41}{\eV}).
+To the best of our knoweledge, there has been a single proposed extension of pCCD to open-shell systems, \cite{Boguslawski_2021}
+based on the ionization-potential equation-of-motion (EOM) formalism.
 
@@
 Here we have generalized hCI \cite{Kossoski_2022} for the case of an open-shell reference determinant, thus extending its applicability to radical species and excited states.
-By surveying the dissociation of several open-shell systems, 
-we found that the hCI route has an edge in comparison to excitation-based CI, for both weakly correlated (equilibrium geometries) and strongly correlating (dissociation) regimes.
+In the first part,
+by surveying the dissociation of several open-shell systems, 
+we found that the hCI route generally outperforms excitation-based CI, for both weakly correlated (equilibrium geometries) and strongly correlating (dissociation) regimes.
 Meanwhile, seniority-based CI leads to far less accurate results in comparison to hCI or excitation-based CI, for a given computational cost.
 These findings are in line with what had been found for closed-shell systems. \cite{Kossoski_2022}
-We also found that, for closed- and open-shell systems,
-the EN2 perturbative corrections lead to substantial improvements for hCI and excitation-based CI routes (while keeping the upper hand for the former), but do not help much with the deficiencies of the sCI route.
+We further evaluated, for closed- and open-shell systems, the impact of EN2 perturbative corrections.
+They lead to substantial improvements for hCI and excitation-based CI routes (while keeping the advantage of the former), 
+but do not help much with the deficiencies of the sCI route.
 
-We further gauged the performance of hCI and sCI methods, based on either HF ground-state orbitals or state-specific orbitals, to describe excited states of closed-shell systems and radicals.
-Despite their elevated computational cost, sCI models produced relatively poor results (even in a state-specific approach), and thus should not be useful in practice.
-Similarly, for lower orders of hCI (hCI1 and hCI1.5), the performance is very poor,
+In the second part,
+we gauged the performance of hCI to describe excited states of closed-shell and open-shell systems, based on either HF ground-state orbitals or state-specific orbitals.
+We discussed the 
+%
+For lower orders of hCI (hCI1 and hCI1.5), the performance is poor,
 with ground-state (state-specific) orbitals leading to largely overestimated (underestimated) excitation energies.
 The situation improves at the state-specific $\Delta$hCI2 order, even though its accuracy is somewhat worse than $\Delta$CISD, both models sharing the same computational scaling. 
-Unfortunately, $\Delta$hCI2.5 does not lead to more accurate results, whereas all the CI models at next next order (CISDT, $\Delta$CISDT, hCI3, and $\Delta$hCI3) show comparable performance.
+Unfortunately, $\Delta$hCI2.5 does not lead to more accurate results, whereas all the CI models at next next order (CISDT, $\Delta$CISDT, hCI3, and $\Delta$hCI3) typically show comparable performances.
 Overall, hCI models for excited states are either innacurate given their computational cost ($h=1$, $h=1.5$ and $h=2.5$) or comparable to excitation-based CI at the same order ($h=2$ and $h=3$).
 %it is fair to conclude that 
+The present findings highlight the challenge in describing on an equal footing states with qualitatively different characters (like ground and singly-excited states of closed-shell systems).
+%, and thus should not be useful in practice.
+
+Here we put forward the interesting perspective of developing hierarchy-based CC (hCC) methods for excited states.
+For instance, EOM-CC with single and double excitations (EOM-CCSD) \cite{} is very accurate for excitation energies of singly-excited states, \cite{Loos_2018,Loos_2020,Loos_2020a},
+whereas CISD provides too big excitation energies, 
+even though both methods use ground-state orbitals and span the same excited determinants.
+Here we found that hCI2 is more accurate than CISD to some extent. In particular, the overestimated energies are slightly less exaggerated in the former.
+We wonder whether this improvement would be transferable to an analogous EOM formulation built on top of hCC truncated at $h=2$, EOM-hCC2.
+Importantly, it is not obvious what the computational scaling of such a method would be.
+Along the same line, it would be interesting to develop and test the performance of cheaper orders like EOM-hCC1.
+Alternatively, one could develop hCC methods to target excited states along the state-specific route.
+
+%This general observation can be understood from the fact that 
+%The same type of reference provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models applied to excitations of open-shell systems.
+%thus significantly minimizing an unbalanced account of correlation, encountered for excitations from closed-shells.
+
+Finally, in the third part of this contribution we employed different sCI models to compute excitation energies of closed-shell and open-shell systems.
+In spite of their formal exponential computational scaling, the sCI models did not produce attractive results compared with more competitive CI alternatives.
+The relevance of these calculations is related to the connection between seniority-based CI and pCCD methods.
+We raised the question whether there is an extension of pCCD to open-shell systems that can provide very similar energies to sCI1 (but with a polynomial scaling),
+which is already established (pCCD and sCI0) for closed-shell systems. \cite{Kossoski_2021}
+In this sense, the present results obtained with sCI1 for open-shell excitations are appealing and encourage the development of such a generalized pCCD.
 

### 2023-07-02 Manuscript/manuscript.tex
update

@@
 For a given approximate CI model, we futher evaluated the bare and the renormalized EN2 perturbative correction. \cite{Garniron_2019} 
-%associated with the determinants left outside the truncated space.
-This calculation involves a single loop over the determinants left outside the truncated space but connected to it via at most double excitations.
-Looping over these external doubly excited determinants thus represents an additional $\order*{N^4}$ factor on top of the computational scaling of the CI calculation.
-For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^{10}}$,
+This calculation involves a single loop over the determinants left outside the truncated (internal) space but connected to it via at most double excitations.
+Looping over these external doubly excited determinants has a computational scaling equal to $\Ndet$,
+thus $\order*{2e+4}$ for excitation-based CI and $\order*{2h+4}$ for hCI, where $e$ or $h$ define the internal CI space.
+For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^8}$,
 though with a small prefactor stemming from the EN2 calculation.

### 2023-07-03 Manuscript/manuscript.tex
saving work

@@
 where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$ according to $h = (e+s/2)/2$.
+This definition ensures that all types of determinants whose number share the same scaling with system size are included at the same hierarchy $h$.
 For different properties and closed-shell molecular systems, 
-hCI was found to display an overall faster convergence to the FCI results than the traditional excitation-based CI, as a function of the number of determinants $\Ndet$. \cite{Kossoski_2022}
+hCI was found to display an overall faster convergence to the FCI results than the traditional excitation-based CI. \cite{Kossoski_2022}
 However, hCI has only been defined for a closed-shell reference, thus systems with an even number of electrons.
@@
 Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022} our third goal is to define and assess the accuracy of hCI models for excited states.
-Finally, to the best of our knowledge, sCI models have not yet been directly used to target excited states,
+Furthermore, to the best of our knowledge, sCI models have not yet been directly used to target excited states,
 despite the growing number of methods that exploit the concept of seniority number,
@@
 methods.
-In particular, by employing a minimal configuration state function reference, 
+In particular, by employing a minimal configuration state function (CSF) reference, 
 we have recently shown that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues. \cite{Kossoski_2023}
@@
 
-%For an arbitrary reference Slater determinant, with seniority $s_0$, the hierarchy $h$ associated to a another determinant
-With respect to a given reference Slater determinant, the hierarchy $h$ of a target determinant is defined as
+%We introduced hCI for closed-shell systems, \cite{Kossoski_2022}
+We introduced hCI \cite{Kossoski_2022} as a truncation of the Hilbert space based on a two-dimensional map of determinants (see Fig.~\ref{fig:hci})
+built from their seniority $s$ and their excitation degree $e$ with respect to a reference determinant.
+It is built such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ enter at the same order of truncation,
+which means moving diagonaly in this seniority-excitation map.
+
+The first order for a closed-shell reference determinant is hCI1, which accounts for all single excitations (as CIS) 
+plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital).
+For both types of determinants, their number scales as $N^2$, and thus should be taken into account at the same hierarchy of hCI ($h=1$ in this case).
+At the next integer order, hCI2 accounts for all types of determinants where $\Ndet$ scale as $N^4$ or less.
+That means all single and double excitations (as CISD), plus the subset of triple excitations that leave only two unpaired electrons, 
+plus the subset of quadruple excitations where no electrons are unpaired.
+More generally, in hCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2h}$.
+Similarly, for excitation-based CI, $\Ndet$ scales as $N^{2e}$, where $e$ is the excitation degree.
+
+Here we generalize hCI for an arbitrary type of reference Slater determinant and for odd number of electrons.
+With respect to a given reference determinant, we define the hierarchy $h$ of a target determinant as
 \begin{equation}
   \label{eq:h}
-        %h = \frac{e+\Delta s/2}{2},
   h = \frac{e+ (s-s_0)/2}{2},
 \end{equation}
-where $s$ and $s_0$ denote the seniority of target and reference determinants, respectively, and $e$ represents the excitation degree that connect the two determinants.
-This definition renders a degree of dissimilarity between two determinants (reference and target),
-which accounts for differences in orbital occupation (through the excitation degree $e$)
+where $s$ and $s_0$ denote the seniority of target and reference determinants, respectively, and $e$ represents the excitation degree that connects the two determinants.
+The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with an even (odd) numbers of electrons, whereas $e$ is also an integer.
+Therefore, $h$ can assume integer or half-integer values, for any type of reference determinant.
+For a given hCI model defined by the hierarchy $h$, all determinants having a hierarchy less than or equal to $h$ are selected.
+Notice that Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant, when $s_0 = 0$.
+
+The above definition ensures that all types of determinants with a given scaling enter at the same hierarchy $h$.
+% computational scaling
+We show in Fig.~\ref{fig:hci} how the seniority-excitation map of determinants is partitioned for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
+%the different types of determinants that are accessed for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
+hCI can also be interpreted as the CI model that provides increasingly more dissimilar determinants from the given reference determinant.
+This dissimilarity is represented by the hierarchy $h$, which accounts for differences in orbital occupation (through the excitation degree $e$)
 and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
-The latter always assumes an integer value (for both even and odd numbers of electrons), such that $h$ becomes an integer or half-integer.
-Also, Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant, when $s_0 = 0$.
-
-state-specific methods
-$\Delta$sCI2 
-$\Delta$hCI1
-From here on, labels carrying the $\Delta$ symbol denote a state-specific approach,
-whereas those without imply the ground-state based approach.
 
@@
 Looping over these external doubly excited determinants has a computational scaling equal to $\Ndet$,
-thus $\order*{2e+4}$ for excitation-based CI and $\order*{2h+4}$ for hCI, where $e$ or $h$ define the internal CI space.
+thus $\order*{N^{2e+4}}$ for excitation-based CI and $\order*{N^{2h+4}}$ for hCI, where $e$ or $h$ define the internal CI space.
 For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^8}$,
@@
 For the latter, we employed the state-specific orbitals obtained in Ref.~\onlinecite{Kossoski_2023}.
-In contrast to excitation-based CI, energies obtained with sCI and hCI models are not invariant under rotations withing the occupied and virtual subspaces.
+Notice that, in contrast to excitation-based CI, the energies obtained with sCI and hCI models are not invariant under rotations within the occupied and virtual subspaces.
 The computed excitation energies were compared against the reference values provided in the QUEST database. \cite{Veril_2021}
@@
 
+The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as reference determinant for the hCI ground state calculations.
+For the state-specific hCI calculations, we employed a minimal CSF reference, \cite{Kossoski_2023}
+meaning a single open-shell determinant for the doublet excited states and a single CSF for the singly-excited states from closed-shells.
+As for the seniority-based CI models, 
+we considered sCI1 for both ground and (state-specific) excited state calculations for the open-shell systems (odd number of electrons).
+For the closed-shell systems (even number of electrons), we explored two models.
+In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal seniority-based CI model for computing excitation energies for closed-shells.
+Further including the $s=2$ sector for the ground state calculations gives rise to the sCI2/sCI2 model.
+CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it imply a ground-state based approach.
 % comments on the different references

### 2023-07-09 Manuscript/manuscript.tex
saving work

@@
 
-%We introduced hCI for closed-shell systems, \cite{Kossoski_2022}
-We introduced hCI \cite{Kossoski_2022} as a truncation of the Hilbert space based on a two-dimensional map of determinants (see Fig.~\ref{fig:hci})
-built from their seniority $s$ and their excitation degree $e$ with respect to a reference determinant.
+We introduced hCI \cite{Kossoski_2022} as a truncation of the Hilbert space based on a two-dimensional map of determinants 
+based on their seniority $s$ and their excitation degree $e$ with respect to a reference determinant (see Fig.~\ref{fig:hCI}).
 It is built such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ enter at the same order of truncation,
@@
 
-The first order for a closed-shell reference determinant is hCI1, which accounts for all single excitations (as CIS) 
-plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital).
-For both types of determinants, their number scales as $N^2$, and thus should be taken into account at the same hierarchy of hCI ($h=1$ in this case).
-At the next integer order, hCI2 accounts for all types of determinants where $\Ndet$ scale as $N^4$ or less.
-That means all single and double excitations (as CISD), plus the subset of triple excitations that leave only two unpaired electrons, 
-plus the subset of quadruple excitations where no electrons are unpaired.
-More generally, in hCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2h}$.
-Similarly, for excitation-based CI, $\Ndet$ scales as $N^{2e}$, where $e$ is the excitation degree.
+%%% FIG 1 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=0.3\linewidth]{table_hCI_closed_shell}
+\hspace{0.02\linewidth}
+\includegraphics[width=0.3\linewidth]{table_hCI_one_open_shell}
+\hspace{0.02\linewidth}
+\includegraphics[width=0.3\linewidth]{table_hCI_two_open_shell}
+\caption{Partitioning of the Hilbert space according to the excitation degree $e$ with respect to a given reference determinant and the seniority number $s$.
+The color tones represent the determinants that are included at a given hCI model,
+for a closed-shell reference (left), an open-shell reference with one unpaired electron (center) and an open-shell reference with two unpaired electrons (right).}
+\label{fig:hCI}
+\end{figure*}
+%%% %%% %%%
 
-Here we generalize hCI for an arbitrary type of reference Slater determinant and for odd number of electrons.
+hCI was initially defined for a closed-shell reference. \cite{Kossoski_2022}
+Here we generalize hCI for an arbitrary type of reference Slater determinant and for systems with an odd number of electrons.
 With respect to a given reference determinant, we define the hierarchy $h$ of a target determinant as
@@
 Therefore, $h$ can assume integer or half-integer values, for any type of reference determinant.
-For a given hCI model defined by the hierarchy $h$, all determinants having a hierarchy less than or equal to $h$ are selected.
+For a given hCI model defined by $h$, we include all the determinants having a hierarchy less than or equal to $h$ with respect to any determinant in the reference.
 Notice that Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant, when $s_0 = 0$.
@@
 % computational scaling
-We show in Fig.~\ref{fig:hci} how the seniority-excitation map of determinants is partitioned for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
+We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
 %the different types of determinants that are accessed for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
@@
 and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
+
+Fig.~\ref{fig:hCI} shows the classes of determinants included at the lower orders of hCI (up to hCI2),
+for a close-shell reference and for an open-shell reference with one and two unpaired electrons.
+The first order for a closed-shell reference determinant is hCI1, which accounts for all single excitations (as CIS) 
+plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital).
+For both types of determinants, their number scales as $N^2$, and are thus taken into account at the same hierarchy of hCI ($h=1$ in this case).
+At the next integer order, hCI2 accounts for all types of determinants where $\Ndet$ scale as $N^4$ or less.
+That means all single and double excitations (as CISD), plus the subset of triple excitations that leave only two unpaired electrons, 
+plus the subset of quadruple excitations where no electrons are unpaired.
+More generally, in hCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2h}$.
+Similarly, for excitation-based CI, $\Ndet$ scales as $N^{2e}$, where $e$ is the excitation degree.
 
@@
 For the state-specific hCI calculations, we employed a minimal CSF reference, \cite{Kossoski_2023}
-meaning a single open-shell determinant for the doublet excited states and a single CSF for the singly-excited states from closed-shells.
+meaning, for the states considered here, a single open-shell determinant for the doublet excited states and a single CSF for the excited states from closed-shells.
 As for the seniority-based CI models, 
@@
 For the closed-shell systems (even number of electrons), we explored two models.
-In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal seniority-based CI model for computing excitation energies for closed-shells.
+In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal seniority-based CI model for computing singly-excited states of closed-shell systems.
 Further including the $s=2$ sector for the ground state calculations gives rise to the sCI2/sCI2 model.
@@
 For a given hierarchy parameter $h$, different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF).
-The case of $\Delta$hCI1 is portrayed in Fig.~\ref{fig:DhCI1}.
+The case of $\Delta$hCI1 is portrayed in Fig.~\ref{fig:hCI1}.
 %For the ground state calculation, it accounts for two classes of excited determinants, generated by single excitation and by paired double excitation from the closed-shell reference.
@@
 Again, thanks to the same type of reference in ground- and excited-state calculations.
-The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:DhCI1}, in comparison to the more unbalanced case of excitation from closed-shells.
+The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:hCI1}, in comparison to the more unbalanced case of excitation from closed-shells.
 We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5).

### 2023-07-10 Manuscript/manuscript.tex
saving work

@@
 
+%%% FIG 2 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{determinants}
+\caption{}
+\label{fig:determinants}
+\end{figure*}
+%%% %%% %%%
+
 Fig.~\ref{fig:hCI} shows the classes of determinants included at the lower orders of hCI (up to hCI2),
@@
 For both types of determinants, their number scales as $N^2$, and are thus taken into account at the same hierarchy of hCI ($h=1$ in this case).
+An odd number of excitations from a closed-shell reference always leave unpaired electrons, thus the empty blocks at $s=0$ and $e$ odd.
+hCI1.5 includes to the hCI1 determinants the set of double excitations where two electrons are unpaired.
 At the next integer order, hCI2 accounts for all types of determinants where $\Ndet$ scale as $N^4$ or less.
@@
 Similarly, for excitation-based CI, $\Ndet$ scales as $N^{2e}$, where $e$ is the excitation degree.
+
+For an odd number of electrons, $s$ assumes odd values starting from $s=1$.
+The simplest reference is an open-shell determinant with one unpaired electron (center pannel of Fig.~\ref{fig:hCI}).
+In contrast to the closed-shell reference case, here there are no empty blocks in the excitation-seniority map.
+Actually, the hCI series starts at the hCI0.5 level, accounting for the single excitations from and into the singly-occupied orbital only.
+More generally, since there are more classes of orbitals for the open-shell case (doubly-occupied, singly-occupied and unnocupied) than for the closed-shell case (doubly-occpupied or unnocupied),
+there is correspondingly more classes of determinants at a given level of hCI.
+The case hCI1 is shown in Fig.~\ref{fig:determinants}.
+
+Finally, one can employ a high-spin reference determinant to define the hCI models.
+The case of two unpaired electrons is shown in the right pannel of Fig.~\ref{fig:hCI}.
+The key difference with respect to the closed-shell case is that the hCI sequence is displaced one block to the right in the seniority-excitation map,
+reflecting the shift from the $s_0=0$ to the $s_0=2$ reference.
+In the two previous cases, the hCI0 level only includes the reference determinant.
+For a $s=2$ reference, however, hCI0 not only accounts for the reference determinant,
+but also the two closed-shell determinants produced by the single excitation that pairs the two unpaired electrons.
+One can adopt even higher-spin references, further displacing the hCI sequence to the right.
+In this case, hCI0 includes all the lower spin determinants obtained by partially or totally pairing the unpaired electrons.
+Notice that the number of such determinants do not depend on the size of the system (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
+hCI can be built in a similar way on top of a high-spin reference for odd number of electrons.
 

### 2023-07-11 Manuscript/manuscript.tex
added determinants fig

@@
 We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
-%the different types of determinants that are accessed for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
+For these three types of reference, Fig.~\ref{fig:determinants} shows the classes of determinants included up to hCI1.
 hCI can also be interpreted as the CI model that provides increasingly more dissimilar determinants from the given reference determinant.
@@
 \includegraphics[width=1.0\linewidth]{determinants}
-\caption{}
+\caption{Classes of determinants generated in hCI0, hCI0.5 and hCI1, for three different reference determinants:
+closed-shell (top left), open-shell with one unpaired electron (top right) and open-shell with two unpaired electrons (bottom).
+Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied and unoccupied).
+}
 \label{fig:determinants}
@@
 
-Fig.~\ref{fig:hCI} shows the classes of determinants included at the lower orders of hCI (up to hCI2),
-for a close-shell reference and for an open-shell reference with one and two unpaired electrons.
-The first order for a closed-shell reference determinant is hCI1, which accounts for all single excitations (as CIS) 
+The first order for a closed-shell reference determinant is hCI1 (see top left of Fig.~\ref{fig:determinants}), which accounts for all single excitations (as CIS) 
 plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital).
@@
 For an odd number of electrons, $s$ assumes odd values starting from $s=1$.
-The simplest reference is an open-shell determinant with one unpaired electron (center pannel of Fig.~\ref{fig:hCI}).
+The simplest reference is an open-shell determinant with one unpaired electron (see center pannel of Fig.~\ref{fig:hCI} and top right of Fig.~\ref{fig:determinants}).
 In contrast to the closed-shell reference case, here there are no empty blocks in the excitation-seniority map.
-Actually, the hCI series starts at the hCI0.5 level, accounting for the single excitations from and into the singly-occupied orbital only.
-More generally, since there are more classes of orbitals for the open-shell case (doubly-occupied, singly-occupied and unnocupied) than for the closed-shell case (doubly-occpupied or unnocupied),
+Actually, the hCI series presents the hCI0.5 level, accounting for the single excitations from and into the singly-occupied orbital only.
+More generally, since there are more types of orbitals for the open-shell case (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell case (doubly-occpupied or unoccupied),
 there is correspondingly more classes of determinants at a given level of hCI.
-The case hCI1 is shown in Fig.~\ref{fig:determinants}.
+This can be appreciated from the comparison of hCI1 for closed- and open-shell references, shown in Fig.~\ref{fig:determinants}.
 
@@
 but also the two closed-shell determinants produced by the single excitation that pairs the two unpaired electrons.
-One can adopt even higher-spin references, further displacing the hCI sequence to the right.
-In this case, hCI0 includes all the lower spin determinants obtained by partially or totally pairing the unpaired electrons.
+The lower order classes of determinants can be seen in the bottom of Fig.~\ref{fig:determinants},
+which clearly outnumber the fewer classes associated with a closed-shell reference.
+One could adopt even higher-spin references, further displacing the hCI sequence to the right.
+In this case, hCI0 would include all the lower spin determinants obtained by partially or totally pairing the unpaired electrons.
 Notice that the number of such determinants do not depend on the size of the system (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
@@
 However, for the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, though by underestimating the excitation energies.
-%Besides the set of orbitals, a second key aspect must be considered.
 Besides the set of orbitals, the types of determinants included at each order play an equally important role and explains our observation.
 For a given hierarchy parameter $h$, different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF).
-The case of $\Delta$hCI1 is portrayed in Fig.~\ref{fig:hCI1}.
+The case of $\Delta$hCI1 can be understood by comparing the hCI1 determinants for these two references, shown in Fig.~\ref{fig:determinants}.
 %For the ground state calculation, it accounts for two classes of excited determinants, generated by single excitation and by paired double excitation from the closed-shell reference.
 %For the singly-excited state, four types of determinants are generated. On top of the single and paired double excitations from the doubly-occupied orbitals, 
-There is arguably more diversity in the types of determinants employed for the excited-state calculation than for the ground-state one.
+There is clearly more diversity in the classes of determinants employed for the excited-state calculation than for the ground-state one, reflecting the different references employed.
 Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the excited state than for ground state.
@@
 Again, thanks to the same type of reference in ground- and excited-state calculations.
-The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:hCI1}, in comparison to the more unbalanced case of excitation from closed-shells.
+The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:determinants}, in comparison to the more unbalanced case of excitation from closed-shells.
 We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5).

### 2023-07-12 Manuscript/sup.tex
added SupInf

\documentclass[aip,jcp,preprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,wrapfig,txfonts,setspace}
\usepackage{siunitx}[=v2]
\usepackage[version=4]{mhchem}
%\usepackage{natbib}
%\bibliographystyle{achemso}

\maxdeadcycles=200

\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\alert}[1]{\textcolor{black}{#1}}
\usepackage[normalem]{ulem}
\newcommand{\titou}[1]{\textcolor{red}{#1}}
\newcommand{\trashPFL}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\PFL}[1]{\titou{(\underline{\bf PFL}: #1)}}
\newcommand{\toto}[1]{\textcolor{green}{#1}}
\newcommand{\trashAS}[1]{\textcolor{green}{\sout{#1}}}
\newcommand{\AS}[1]{\toto{(\underline{\bf AS}: #1)}}
\newcommand{\ant}[1]{\textcolor{orange}{#1}}
\newcommand{\SupInf}{\textcolor{blue}{Supporting Information}}

\newcommand{\mc}{\multicolumn}
\newcommand{\fnm}{\footnotemark}
\newcommand{\fnt}{\footnotetext}
\newcommand{\tabc}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\QP}{\textsc{quantum package}}

\newcommand{\EHF}{E_\text{HF}}
\newcommand{\EDOCI}{E_\text{DOCI}}
\newcommand{\EFCI}{E_\text{FCI}}

\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\theequation}{S\arabic{equation}}

\usepackage[
	colorlinks=true,
    citecolor=blue,
    breaklinks=true
	]{hyperref}
\urlstyle{same}

\begin{document}

\newcommand{\LCPQ}{Laboratoire de Chimie et Physique Quantiques (UMR 5626), Universit\'e de Toulouse, CNRS, UPS, France}

\title{Supporting Information for ``Seniority and hierarchy configuration interaction for radicals and excited states''}

\author{F\'abris Kossoski}
\email{fkossoski@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}
\author{Pierre-Fran\c{c}ois Loos}
\email{loos@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}

% Abstract
\begin{abstract}
%Here comes the abstract.
%\bigskip
%\begin{center}
%        \boxed{\includegraphics[width=0.4\linewidth]{TOC}}
%\end{center}
%\bigskip
\end{abstract}

% Title
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Equilibrium geometry of ethylene}
%\label{sec:ethylene}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Equilibrium geometry of ethylene, in atomic units, obtained from Ref. \onlinecite{Loos_2018}:
\begin{singlespace}
\begin{verbatim}
C    0.00000000       1.26026583        0.00000000 
C    0.00000000      -1.26026583        0.00000000 
H    0.00000000       2.32345976        1.74287672 
H    0.00000000      -2.32345976        1.74287672 
H    0.00000000       2.32345976       -1.74287672 
H    0.00000000      -2.32345976       -1.74287672 
\end{verbatim}
\end{singlespace}

Equilibrium geometry of vinyl, in atomic units, obtained from Ref. \onlinecite{Loos_2020}:
\begin{singlespace}
\begin{verbatim}
C    0.00000000       1.16769663       -0.04303146
C    0.00000000      -1.29945364        0.15810072
H    0.00000000       2.38429609        1.59801822
H    0.00000000       2.08759130       -1.87998309
H    0.00000000      -2.90307925       -1.08814513
\end{verbatim}
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{\ce{Computational details}}
%\label{sec:comp_details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

By fitting the computed potential energy curves with a Morse potential, we obtained harmonic vibrational frequencies and equilibrium geometries.
The following intervals have been considered for the fitting:
\begin{itemize}
\item \ce{HF}: \SI{0.8}{\angstrom} to \SI{1.3}{\angstrom}
\item \ce{F2}: \SI{1.25}{\angstrom} to \SI{1.65}{\angstrom}
\item ethylene: \SI{2.2}{\bohr} to \SI{2.9}{\bohr}
\item \ce{N2}: \SI{0.95}{\angstrom} to \SI{1.3}{\angstrom}
\item \ce{H4}: \SI{1.45}{\bohr} to \SI{1.95}{\bohr}
\item \ce{H8}: \SI{1.6}{\bohr} to \SI{2.05}{\bohr}
\item \ce{OH}: \SI{0.85}{\angstrom} to \SI{1.2}{\angstrom}
\item \ce{CN}: \SI{1.0}{\angstrom} to \SI{1.4}{\angstrom}
\item vinyl: \SI{2.35}{\bohr} to \SI{2.7}{\bohr}
\item \ce{H7}: \SI{1.6}{\bohr} to \SI{2.05}{\bohr}
\end{itemize}

The non-parallelity error and distance error were computed for the following intervals:
\begin{itemize}
\item \ce{HF}: \SI{0.5}{\angstrom} to \SI{6}{\angstrom}
\item \ce{F2}: \SI{0.95}{\angstrom} to \SI{8}{\angstrom}
\item ethylene: \SI{1.5}{\bohr} to \SI{16}{\bohr}
\item \ce{N2}: \SI{0.7}{\angstrom} to \SI{4}{\angstrom}
\item \ce{H4}: \SI{1}{\bohr} to \SI{10}{\bohr}
\item \ce{H8}: \SI{1}{\bohr} to \SI{10}{\bohr}
\item \ce{OH}: \SI{0.6}{\angstrom} to \SI{8}{\angstrom}
\item \ce{CN}: \SI{0.7}{\angstrom} to \SI{4}{\angstrom}
\item vinyl: \SI{1.7}{\bohr} to \SI{9}{\bohr}
\item \ce{H7}: \SI{1}{\bohr} to \SI{10}{\bohr}
\end{itemize}

\clearpage

%%% FIG S1 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pes}
\caption{
Potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
according to hCI (green), excitation-based CI (red) and sCI (blue) models.
}
\label{fig:plot_pes}
\end{figure*}
%%% %%% %%%

%%% FIG S2 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pes_rpt2}
\caption{
Potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
according to hCI (green), excitation-based CI (red) and sCI (blue) models
corrected with second-order Epstein-Nesbet perturbative correction.
}
\label{fig:plot_pes_rpt2}
\end{figure*}
%%% %%% %%%

%%% FIG S3 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pes_closed}
\caption{
Potential energy curves for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
according to hCI (green), excitation-based CI (red) and sCI (blue) models.
Results first reported in Ref. \onlinecite{Kossoski_2022}.
}
\label{fig:plot_pes_closed}
\end{figure*}
%%% %%% %%%

%%% FIG S4 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_pes_closed_rpt2}
\caption{
Potential energy curves for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
according to hCI (green), excitation-based CI (red) and sCI (blue) models
corrected with second-order Epstein-Nesbet perturbative correction.
}
\label{fig:plot_pes_closed_rpt2}
\end{figure*}
%%% %%% %%%

%%% FIG S5 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_all_rpt2}
\caption{
Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{OH}, vinyl, and \ce{H7},
as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models
corrected with second-order Epstein-Nesbet perturbative correction.
For \ce{CN}, the perturbative correction produces discontinuous potential energy curves due to the change of character of the unperturbed ground state, as explained in the main text.
}
\label{fig:plot_all_rpt2}
\end{figure*}
%%% %%% %%%

%%% FIG S6 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_all_closed}
\caption{
Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models.
Results first reported in Ref. \onlinecite{Kossoski_2022}.
}
\label{fig:plot_all_closed}
\end{figure*}
%%% %%% %%%

%%% FIG S7 %%%
\begin{figure*}%[h!]
\includegraphics[width=1.0\linewidth]{plot_all_closed_rpt2}
\caption{
Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models,
corrected with second-order Epstein-Nesbet perturbative correction.
}
\label{fig:plot_all_closed_rpt2}
\end{figure*}
%%% %%% %%%

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{manuscript}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2023-07-12 N2_cc-pvdz/pt2/compute_NPE.pl

@@ -10,38 +10,38 @@ my $xcut = $ARGV[1];
 my(@files,@index,$detfile,$outfile);
 
 if($case==1){
-@files = qw( pes_rohf.dat pes_CIo1.dat pes_CIo1.5.dat pes_CIo2.dat pes_CIo2.5.dat pes_CIo3.dat pes_CIo3.5.dat pes_CIo4.dat);
+@files = qw( pes_rhf.dat pes_CIo1.dat pes_CIo1.5.dat pes_CIo2.dat pes_CIo2.5.dat pes_CIo3.dat pes_CIo3.5.dat pes_CIo4.dat);
 @index = qw( 0 1 1.5 2 2.5 3 3.5 4);
 $detfile = "det_CIo.dat";
 $outfile = "stat_CIo.dat";
 }
 elsif($case==2){
-@files = qw( pes_rohf.dat pes_ooCIo1.dat pes_ooCIo1.5.dat pes_ooCIo2.dat pes_ooCIo2.5.dat pes_ooCIo3.dat pes_ooCIo3.5.dat pes_ooCIo4.dat);
+@files = qw( pes_rhf.dat pes_ooCIo1.dat pes_ooCIo1.5.dat pes_ooCIo2.dat pes_ooCIo2.5.dat pes_ooCIo3.dat pes_ooCIo3.5.dat pes_ooCIo4.dat);
 @index = qw( 0 1 1.5 2 2.5 3 3.5 4);
 $detfile = "det_CIo.dat";
 $outfile = "stat_ooCIo.dat";
 }
 elsif($case==3){
-@files = qw( pes_rohf.dat pes_CIS.dat pes_CISD.dat pes_CISDT.dat pes_CISDTQ.dat);
+@files = qw( pes_rhf.dat pes_rhf.dat pes_CISD.dat pes_CISDT.dat pes_CISDTQ.dat);
 @index = qw( 0 1 2 3 4);
 $detfile = "det_CI.dat";
 $outfile = "stat_CI.dat";
 }
 elsif($case==4){
-@files = qw( pes_rohf.dat pes_ooCIS.dat pes_ooCISD.dat pes_ooCISDT.dat pes_ooCISDTQ.dat);
+@files = qw( pes_rhf.dat pes_ooCIS.dat pes_ooCISD.dat pes_ooCISDT.dat pes_ooCISDTQ.dat);
 @index = qw( 0 1 2 3 4);
 $detfile = "det_CI.dat";
[…]

### 2023-07-12 OH_cc-pvdz/pt2/compute_NPE.pl

@@ -114,7 +114,7 @@ close(Fi);
 
         if($x==$xcut){
           my $npe = $maxy - $miny;
-          my $closeness = abs($maxy) + abs($miny);
+          my $closeness = $maxy + $miny;
           print Fo "$index[$j] $index2[$j] $det[$j]  $npe  $closeness  $maxy  $miny \n";
 	}
 	#     }

### 2023-07-12 ethylene_cc-pvdz/rpt2/compute_NPE.pl

@@ -114,7 +114,7 @@ close(Fi);
 
         if($x==$xcut){
           my $npe = $maxy - $miny;
-          my $closeness = $maxy + $miny;
+          my $closeness = abs($maxy) + abs($miny);
           print Fo "$index[$j] $index2[$j] $det[$j]  $npe  $closeness  $maxy  $miny \n";
 	}
 	#     }

### 2023-07-12 vinyl_cc-pvdz/compute_NPE.pl

@@ -114,7 +114,7 @@ close(Fi);
 
         if($x==$xcut){
           my $npe = $maxy - $miny;
-          my $closeness = $maxy + $miny;
+	  my $closeness = abs($maxy) + abs($miny);
           print Fo "$index[$j] $index2[$j] $det[$j]  $npe  $closeness  $maxy  $miny \n";
 	}
 	#     }

### 2023-07-12 CN_cc-pvdz/fit_eq.sh
added and updated gnu and sh

@@ -156,7 +156,7 @@ if [[ $(grep 'a25 ' a0.dat) ]]; then grep '3   ' det_CIo.dat | cut -c 4- >> det.
 paste det.dat a.dat D.dat > det_aD.dat
 
 grep 'FCI'      det_aD.dat | sed 's/FCI/1 /g'     >  det_aD_FCI.dat
-grep 'FCI'      det_aD.dat | sed 's/FCI/1.0E9 /g' >> det_aD_FCI.dat
+grep 'FCI'      det_aD.dat | sed 's/FCI/1.0E10 /g' >> det_aD_FCI.dat
 
 grep 'ROHF'     det_aD.dat | sed 's/ROHF/0 /g'     >  det_aD_CI.dat
 grep 'hfCIS '   det_aD.dat | sed 's/hfCIS /1 /g'   >> det_aD_CI.dat
@@ -256,7 +256,7 @@ grep 'xe25 ' xe0.dat | sed 's/xe25/ooCIo3   /g' >> xe.dat
 paste det.dat xe.dat D.dat > det_xe.dat
 
 grep 'FCI' det_xe.dat | sed 's/FCI/1 /g'   >  det_xe_FCI.dat
-grep 'FCI' det_xe.dat | sed 's/FCI/1E9 /g' >> det_xe_FCI.dat
+grep 'FCI' det_xe.dat | sed 's/FCI/1E10 /g' >> det_xe_FCI.dat
 
 grep 'ROHF'     det_xe.dat | sed 's/ROHF/0 /g'     >  det_xe_CI.dat
 grep 'hfCIS '   det_xe.dat | sed 's/hfCIS /1 /g'   >> det_xe_CI.dat

### 2023-07-12 CN_cc-pvdz/run_stat.sh

@@ -2,7 +2,7 @@
 
 norbitals=28
 nelectrons=13
-xcut=5.5
+xcut=4.0
 
 if [ -f "input" ]; then
   rm "input"

### 2023-07-12 N2_cc-pvdz/pt2/run_stat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+norbitals=26
+nelectrons=10
+xcut=4.0
+
+if [ -f "input" ]; then
+  rm "input"
+fi
+echo "$nelectrons $norbitals" > input
+../../determinants/determinants.x < input
+rm "input"
+
+for i in {1..6}
+do
+./compute_NPE.pl $i $xcut 
+done

### 2023-07-12 OH_cc-pvdz/pt2/fit_eq.sh

@@ -1,0 +1,298 @@
+#!/bin/bash
+
+#######################################################################
+
+if [ -f "fit.log" ]; then
+  rm "fit.log"
+fi
+
+gnuplot fit_eq.gnu
+
+#######################################################################
+grep ^a fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/a0 /FCI      /g' |
+        sed 's/a1 /ROHF     /g' |
+        sed 's/a2 /hfCIS    /g' |
+        sed 's/a3 /hfCISD   /g' |
+        sed 's/a4 /hfCISDT  /g' |
+        sed 's/a5 /hfCISDTQ /g' |
+        sed 's/a6 /hfCIs1   /g' |
+        sed 's/a7 /hfCIs3   /g' |
+        sed 's/a8 /hfCIs5   /g' |
+        sed 's/a9 /hfCIo1   /g' |
+        sed 's/a10/hfCIo1.5 /g' |
+        sed 's/a11/hfCIo2   /g' |
+        sed 's/a12/hfCIo2.5 /g' |
+        sed 's/a13/hfCIo3   /g' |
+        sed 's/a14/hfCIo3.5 /g' |
+        sed 's/a15/hfCIo4   /g' |
+        sed 's/a16/ooCIS    /g' |
+        sed 's/a17/ooCISD   /g' |
+        sed 's/a18/ooCISDT  /g' |
+        sed 's/a19/ooCIs1   /g' |
+        sed 's/a20/ooCIs3   /g' |
+        sed 's/a21/ooCIo1   /g' |
+        sed 's/a22/ooCIo1.5 /g' |
+        sed 's/a23/ooCIo2   /g' |
+        sed 's/a24/ooCIo2.5 /g' |
+        sed 's/a25/ooCIo3   /g' > a.dat
+
+grep ^D fit.log | grep = | sed 's/=//g' | cut -c -30 |
+        sed 's/D0 /FCI      /g' |
+        sed 's/D1 /ROHF     /g' |
+        sed 's/D2 /hfCIS    /g' |
[…]

### 2023-07-12 OH_cc-pvdz/pt2/plot_pes.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+gnuplot plot_pes.gnu
+epspdf plot_pes.eps
+okular plot_pes.pdf

### 2023-07-12 vinyl_cc-pvdz/run_stat.sh

@@ -2,7 +2,7 @@
 
 norbitals=43
 nelectrons=15
-xcut=12.0
+xcut=9.0
 
 if [ -f "input" ]; then
   rm "input"

### 2023-07-12 Manuscript/manuscript.tex
saving work

@@
 \begin{figure*}%[h!]
-\includegraphics[width=0.3\linewidth]{table_hCI_closed_shell}
-\hspace{0.02\linewidth}
-\includegraphics[width=0.3\linewidth]{table_hCI_one_open_shell}
-\hspace{0.02\linewidth}
-\includegraphics[width=0.3\linewidth]{table_hCI_two_open_shell}
-\caption{Partitioning of the Hilbert space according to the excitation degree $e$ with respect to a given reference determinant and the seniority number $s$.
-The color tones represent the determinants that are included at a given hCI model,
-for a closed-shell reference (left), an open-shell reference with one unpaired electron (center) and an open-shell reference with two unpaired electrons (right).}
+\includegraphics[width=0.05\linewidth]{det_closed.eps}
+\includegraphics[width=0.25\linewidth]{table_hCI_closed_shell}
+\hspace{0.03\linewidth}
+\includegraphics[width=0.05\linewidth]{det_1open.eps}
+\includegraphics[width=0.25\linewidth]{table_hCI_one_open_shell}
+\hspace{0.03\linewidth}
+\includegraphics[width=0.05\linewidth]{det_2open.eps}
+\includegraphics[width=0.25\linewidth]{table_hCI_two_open_shell}
+\caption{Partitioning of the Hilbert space according to the seniority number $s$ and the excitation degree $e$ with respect to a given reference determinant (shown by its side),
+for a closed-shell reference (left), an open-shell reference with one unpaired electron (center) and an open-shell reference with two unpaired electrons (right).
+The color tones represent the determinants that are included at a given hCI model.}
 \label{fig:hCI}
@@
 Actually, the hCI series presents the hCI0.5 level, accounting for the single excitations from and into the singly-occupied orbital only.
+For HF orbitals, however, these excitations do not connect with the reference, and hCI0.5 thus provides the same energy as restricted open-shell HF.
 More generally, since there are more types of orbitals for the open-shell case (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell case (doubly-occpupied or unoccupied),
@@
 These calculations employed the ground-state restricted open-shell HF orbitals and the frozen core approximation.
-%which display a variable number of bond breaking.
 The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020},
-%also reproduced in the \SupInf,
-and the PECs were computed along the \ce{C=C} double bond breaking coordinate, with the remaining internal coordinates kept frozen.
+and is also reproduced in the {\SupInf}.
+Their PECs were computed along the \ce{C=C} double bond breaking coordinate, with the remaining internal coordinates kept frozen.
 For \ce{H7}, we considered equally spaced and linearly arranged hydrogen atoms, and the PECs were computed along the symmetric dissociation coordinate.
+
 The results were analyzed along the same lines of our report on hCI for closed-shell systems. \cite{Kossoski_2022}
@@
 we evaluated the convergence of the non-parallelity error (NPE), the distance error, the harmonic vibrational frequencies, and the equilibrium bond lengths, as functions of $\Ndet$.
-The NPE (distance error) of a given level of theory is defined as the maximum minus (plus) the minimum differences between its corresponding PEC and the FCI one, for a given range of coordinates.
+The NPE of a given level of theory is defined as the maximum minus the minimum differences between its corresponding PEC and the FCI one, for a given range of coordinates.
+Here we redefine the previous definition of the distance error \cite{Kossoski_2022} to accommodate for the fact the approximate PEC might appear below the FCI when perturbative corrections are employed.
+The previous definition could provide negative values in this case, whereas this metric should provide the overall extent of proximity between two PECs.
+Here, we define the distance error based on the signed differences between two PECs, as the absolute value of their maximum difference plus the absolute value of their minimum difference, 
+evaluated at given coordinate interval.
 Details about how the harmonic vibrational frequencies and equilibrium geometries were obtained from the calculated PECs,
-and the ranges defining the NPE and distance errors can be found in the \SupInf.
+along with the ranges defining the NPE and distance errors can be found in the \SupInf.
 
@@
 
-% redefine the distance errors because of negative values?
 % vinyl missing point at 5.2 for s3i+pt2 and at 2.0 for s5+pt2
@@
 The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf},
-along with the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies, plotted as a function of $\Ndet$ for the different CI models.
-The present results for open-shell systems are discussed in detail in the following,
-and display similar trends to the ones previously reported for closed-shell systems. \cite{Kossoski_2022}
-hCI typically improves the convergence of the different properties when compared to excitation-based CI.
-Concerning the NPEs, such improvement is clear for \ce{OH} and \ce{CN} (involving a single bond breaking), less so for vinyl (double bond breaking),
+whereas Fig.~\ref{fig:plot_all} contains the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies plotted as a function of $\Ndet$ for the different CI models.
+The present results for open-shell systems display similar trends to the ones previously reported for closed-shell systems. \cite{Kossoski_2022}
+The main finding is that hCI typically improves the convergence of the different properties when compared to excitation-based CI.
+Concerning the NPEs, such improvement is clear for \ce{OH} and \ce{CN} (involving a single bond breaking), becoming less so for vinyl (double bond breaking),
 whereas for \ce{H7} (multiple bond breaking), hCI and excitation-based CI perform similarly.
-This behavior had also been observed for dissociation of closed-shell systems. \cite{Kossoski_2022}
+This dependence on the order of bond breaking had also been observed for closed-shell systems. \cite{Kossoski_2022}
 The same trend is observed for the distance error, although the improvement from excitation-based CI to hCI is less marked than for the NPEs.
-hCI further leads to a faster convergence of the equilibrium geometry of \ce{CN}, while comparable results are found for the geometries of the other systems.
-Meanwhile, a clear improvement from excitation-based CI to hCI is observed for the vibrational frequencies, except for \ce{H7}, where no difference is found.
+hCI further leads to a faster convergence of the equilibrium geometry of \ce{CN}, but does not improve upon excitation-based CI for the geometries of the other systems.
+Finally, hCI systematically outperforms excitation-based CI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+
+%%% FIG 3 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_all}
+\caption{
+Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{OH}, \ce{CN}, \ce{H7}, and vinyl,
+as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models.}
+\label{fig:plot_all}
+\end{figure*}
+%%% %%% %%%
+
+sCI models deliver a poor convergence of all observables.
+The single exception is the NPE for \ce{OH} obtained from sCI0, smaller than the NPEs obtained from CISD and hCI1.5 or hCI2, methods having a comparable computational cost in this case.
+Going to larger basis sets or larger systems, the computational burden of sCI models would increase considerably more than hCI or excitation-based CI models, 
+due to its exponential scaling, rendering this route even less atractive.
+Analogous conclusions had been drawn for sCI applied to closed-shell systems, \cite{Kossoski_2022}
+and here we show that the same holds for radicals.
 
@@
 and the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) addressed in our first report on hCI. \cite{Kossoski_2022}
-The full set of PECSs is shown in the {\SupInf}.
+The corresponding PECSs are shown in the {\SupInf}.
 At stretched geometries, the perturbative correction can lead to kinks and large jumps in the PECs.
@@
 Around the equilibrium geometry, the perturbative correction produced well-behaved PECs at all levels of CI and for all systems, 
-except for \ce{CN}, which is related to crossing of states around equilibrium.
+except for \ce{CN}.
+In this case, there are several close-lying excited states around the equilibrium geometry. \cite{}
+Since the unperturbed PEC corresponds to the lowest-lying solution, it often presents kinks where the ground state changes its character.
+Because of that, the perturbative correction produces discontinuous PECs.
 
-We found that the EN2 perturbative correction generally reduces the errors of all observables considered here, for a given level of excitation-based CI and hCI routes.
-As such, it accelerates the rate of convergence as a function of the excitation degree or hierarchy parameter.
-Importantly, the perturbative corrected hCI route still outperforms the excitation-based counterpart, just as for the uncorrected comparison.
+%%% FIG 4 %%%
+\begin{figure}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_oh_rpt2}
+\includegraphics[width=1.0\linewidth]{plot_f2_rpt2}
+\caption{
+Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies, for \ce{OH} (top four pannels) and \ce{F2} (bottom four pannels),
+as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+with (light tones) and without (dark tones) the EN2 perturbative correction.
+}
+\label{fig:plot_rpt2}
+\end{figure}
+%%% %%% %%%
+
+To illustrate the effect of the EN2 perturbative correction, we show the results for \ce{OH} and \ce{F2} in Fig.\ref{fig:plot_rpt2}.
+The full set of results can be found in the \SupInf.
+We found that the EN2 correction generally reduces the errors of all observables considered here, for a given level of CI.
+As such, it accelerates the rate of convergence as a function of the number of determinants in the unperturbed calculation.
+Importantly, the perturbative corrected hCI route still outperforms the excitation-based counterpart.
 Furthermore, the reduction of the errors is overall more monotonic than without the perturbative correction,
-which is outlined in Fig.\ref{} for the particular case of \ce{F2}.
-For the sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation-based and hCI alternatives.
+clearly seen in the case of \ce{F2}, shown in Fig.\ref{fig:plot_rpt2}.
+In some cases, the EN2 correction does not improve the convergence of the sCI route.
+This can be seen for the equilibrium geometry and vibrational frequency of \ce{F2}.
+
+%sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation-based and hCI alternatives.
 

### 2023-07-18 Manuscript/manuscript.tex
saving work

@@
 hCI can also be interpreted as the CI model that provides increasingly more dissimilar determinants from the given reference determinant.
-This dissimilarity is represented by the hierarchy $h$, which accounts for differences in orbital occupation (through the excitation degree $e$)
+The level of dissimilarity is represented by the hierarchy $h$, which accounts for differences in orbital occupation (through the excitation degree $e$)
 and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
@@
 Similarly, for excitation-based CI, $\Ndet$ scales as $N^{2e}$, where $e$ is the excitation degree.
+The computational scaling along the two routes is shown in Tab.~\ref{tab:scaling}.
+
+\begin{table}[ht!]
+\caption{Computational scaling along the hierarchy-based and excitation-base CI routes, with and without the EN2 perturbative correction.}
+\label{tab:scaling}
+\begin{ruledtabular}
+\begin{tabular}{ll|ll|l}
+ \mc{2}{c|}{hierarchy-based}   &  \mc{2}{c|}{excitation-based} &  scaling \\
+\hline
+hCI1   &    -       & CIS    &   -       & $\order*{N^4}$ \\
+hCI1.5 &    -       &  -     &   -       & $\order*{N^5}$ \\
+\hline
+hCI2   & hCI1+EN2   & CISD   & CIS+EN2   & $\order*{N^6}$ \\
+hCI2.5 & hCI1.5+EN2 &  -     &   -       & $\order*{N^7}$ \\
+\hline
+hCI3   & hCI2+EN2   & CISDT  & CISD+EN2  & $\order*{N^8}$ \\
+hCI3.5 & hCI2.5+EN2 &   -    &   -       & $\order*{N^9}$ \\
+\hline
+hCI4   & hCI3+EN2   & CISDTQ & CISDT+EN2 & $\order*{N^{10}}$ \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
 
@@
 which clearly outnumber the fewer classes associated with a closed-shell reference.
-One could adopt even higher-spin references, further displacing the hCI sequence to the right.
+One could adopt even higher-spin references, further displacing the hCI sequence to the right in the seniority-excitation map.
 In this case, hCI0 would include all the lower spin determinants obtained by partially or totally pairing the unpaired electrons.
 Notice that the number of such determinants do not depend on the size of the system (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
-hCI can be built in a similar way on top of a high-spin reference for odd number of electrons.
+Similarly, hCI can be built on top of a high-spin reference for odd number of electrons.
 
@@
 though with a small prefactor stemming from the EN2 calculation.
+The computational scaling associated with the CI+EN2 calculations is also presented in Tab.~\ref{tab:scaling}.
 
@@
 evaluated at given coordinate interval.
+In the discussion, we say equilibrium properties when referring to both the equilibrium geometry and the vibrational frequency.
 Details about how the harmonic vibrational frequencies and equilibrium geometries were obtained from the calculated PECs,
@@
 
-% vinyl missing point at 5.2 for s3i+pt2 and at 2.0 for s5+pt2
-% CN correct dissociating part for h4 and h3.5
-% H8 missing points for Q+pt2
-% H8 missing 9.0 for T+pt2
+% vinyl eq.geom e4 s5 h3.5 rpt2
+% vinyl 5.0 h3+rpt2
+% update pecs of ethylene
 
-The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf},
-whereas Fig.~\ref{fig:plot_all} contains the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies plotted as a function of $\Ndet$ for the different CI models.
-The present results for open-shell systems display similar trends to the ones previously reported for closed-shell systems. \cite{Kossoski_2022}
-The main finding is that hCI typically improves the convergence of the different properties when compared to excitation-based CI.
-Concerning the NPEs, such improvement is clear for \ce{OH} and \ce{CN} (involving a single bond breaking), becoming less so for vinyl (double bond breaking),
-whereas for \ce{H7} (multiple bond breaking), hCI and excitation-based CI perform similarly.
-This dependence on the order of bond breaking had also been observed for closed-shell systems. \cite{Kossoski_2022}
-The same trend is observed for the distance error, although the improvement from excitation-based CI to hCI is less marked than for the NPEs.
-hCI further leads to a faster convergence of the equilibrium geometry of \ce{CN}, but does not improve upon excitation-based CI for the geometries of the other systems.
-Finally, hCI systematically outperforms excitation-based CI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf}.
+From these PECs, we obtain the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies, which are plotted as a function of $\Ndet$ in Fig.~\ref{fig:plot_all}.
+The corresponding results for the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) are shown in Fig.~\ref{fig:plot_all_closed}.
+When we first introduced hCI, \cite{Kossoski_2022} we surveyed the same molecules, though only for the bare CI models (without a perturbative correction).
+These previous results are reproduced in Fig.~\ref{fig:plot_all_closed} together with the present results, which include such correction.
+First we will focus on the results for the bare CI models (hCI, excitation-based CI and sCI) for the open-shell systems, represented by the dark tones,
+leaving the discussion about the perturbative correction (light tones) for later.
 
@@
 \caption{
-Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{OH}, \ce{CN}, \ce{H7}, and vinyl,
-as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models.}
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{OH}, \ce{CN}, \ce{H7}, and vinyl,
+as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+with (light tones) and without (dark tones) the EN2 perturbative correction.
+The dashed lines represent the FCI results.}
 \label{fig:plot_all}
@@
 
+We found that hCI typically improves the convergence with respect to $\Ndet$ when compared to excitation-based CI, for the four properties surveyed here.
+The present finding for open-shell systems therefore parallels the previously reported superiority of hCI for closed-shell systems. \cite{Kossoski_2022}
+Concerning the NPEs, the advantage of hCI is pronounced for \ce{OH} (involving a single bond breaking), 
+becoming less so, though still clear, for vinyl (double bond breaking) and \ce{CN} (triple bond breaking),
+whereas for \ce{H7} (multiple bond breaking), it still shows a small advantage in comparison to excitation-based CI.
+This dependence on the order of bond breaking had also been observed for closed-shell systems \cite{Kossoski_2022} (as can be seen in Fig.~\ref{fig:plot_all_closed}).
+We also found overall smaller distance errors with hCI than with excitation-based CI, although the improvement is less noticeable in this case.
+hCI further leads to a somewhat faster convergence of the equilibrium geometries for the open-shell systems, most noticeably for \ce{CN}.
+Finally, hCI systematically outperforms excitation-based CI for the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+
+In great contrast to hCI (and excitation-based CI),
 sCI models deliver a poor convergence of all observables.
-The single exception is the NPE for \ce{OH} obtained from sCI0, smaller than the NPEs obtained from CISD and hCI1.5 or hCI2, methods having a comparable computational cost in this case.
-Going to larger basis sets or larger systems, the computational burden of sCI models would increase considerably more than hCI or excitation-based CI models, 
-due to its exponential scaling, rendering this route even less atractive.
-Analogous conclusions had been drawn for sCI applied to closed-shell systems, \cite{Kossoski_2022}
-and here we show that the same holds for radicals.
+The single exception is the NPE for \ce{OH} obtained from sCI0, smaller than the NPEs obtained from CISD and hCI1.5 or hCI2, methods having a comparable computational cost in this particular case.
+Going to larger basis sets or bigger systems, the computational burden of sCI models would increase considerably more than hCI or excitation-based CI models, 
+due to its formal exponential scaling.
+In their own, we conclude that the sCI route is unatractive to open-shell systems,
+in line with similar findings for closed-shell systems \cite{Kossoski_2022} (also clear from Fig.~\ref{fig:plot_all_closed}).
 
@@
 
-We further assesssed the effect of the EN2 perturbative correction for both the open-shell systems studied here 
-and the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) addressed in our first report on hCI. \cite{Kossoski_2022}
-The corresponding PECSs are shown in the {\SupInf}.
-At stretched geometries, the perturbative correction can lead to kinks and large jumps in the PECs.
-Such problems were encountered at some levels of CI for \ce{CN}, vinyl, \ce{H8}, and \ce{H7}, and to a less extent, for \ce{N2} (only for hCI1).
-For the remaining five systems (\ce{HF}, \ce{F2}, ethylene, \ce{H4}, and \ce{OH}),
-the EN2 correction produced smooth PECs at dissociation.
-Around the equilibrium geometry, the perturbative correction produced well-behaved PECs at all levels of CI and for all systems, 
-except for \ce{CN}.
-In this case, there are several close-lying excited states around the equilibrium geometry. \cite{}
-Since the unperturbed PEC corresponds to the lowest-lying solution, it often presents kinks where the ground state changes its character.
-Because of that, the perturbative correction produces discontinuous PECs.
+The results for the CI models corrected by the EN2 energy are represented by the light tones in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
+%The corresponding PECSs are shown in the {\SupInf}.
+We excluded cases where there are important discontinuities in the PECs.
+These discontinuities reflect crossings between states of different symmetries in the unperturbed calculation,
+appearing as kinks in the computed ground-state PEC.
+Due to the abrupt change of character, the EN2 correction is not uniform and gives rise to the observed discontinuities.
+This happens for the open-shell systems at dissociation (except for \ce{OH}), for which we do not present NPE and distance errors.
+For \ce{OH} and the six closed-shell systems, the EN2 correction produced smooth PECs at dissociation.
+Around the equilibrium geometry, the EN2 correction also leads to well-behaved PECs, for all CI models and for all systems, except for \ce{CN}
+(thus no equilibrium properties are presented for this radical).
+This is simply due to its several close-lying excited states. \cite{}
+If we were to follow the PEC of a given symmetry (and not the lowest-lying state as we did here), there would be no discontinuity stemming from the EN2 correction.
 
 %%% FIG 4 %%%
-\begin{figure}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_oh_rpt2}
-\includegraphics[width=1.0\linewidth]{plot_f2_rpt2}
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_all_closed}
 \caption{
-Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies, for \ce{OH} (top four pannels) and \ce{F2} (bottom four pannels),
-as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, 
+as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
 with (light tones) and without (dark tones) the EN2 perturbative correction.
-}
-\label{fig:plot_rpt2}
-\end{figure}
+The dashed lines represent the FCI results.}
+\label{fig:plot_all_closed}
+\end{figure*}
 %%% %%% %%%
 
-To illustrate the effect of the EN2 perturbative correction, we show the results for \ce{OH} and \ce{F2} in Fig.\ref{fig:plot_rpt2}.
-The full set of results can be found in the \SupInf.
-We found that the EN2 correction generally reduces the errors of all observables considered here, for a given level of CI.
-As such, it accelerates the rate of convergence as a function of the number of determinants in the unperturbed calculation.
-Importantly, the perturbative corrected hCI route still outperforms the excitation-based counterpart.
-Furthermore, the reduction of the errors is overall more monotonic than without the perturbative correction,
-clearly seen in the case of \ce{F2}, shown in Fig.\ref{fig:plot_rpt2}.
-In some cases, the EN2 correction does not improve the convergence of the sCI route.
-This can be seen for the equilibrium geometry and vibrational frequency of \ce{F2}.
-
-%sCI route, however, the EN2 correction does not bring a clear improvement to render it competitive in relation to the excitation-based and hCI alternatives.
-
-% rPT2 vs PT2 there are important differences
+We found that the EN2 correction significantly reduces the errors with respect to the FCI results.
+This is observed for all the CI models, systems, and observables considered here.
+In the particular case of \ce{OH}, the improvement is massive, as even the lower levels of CI+EN2 deliver very close results to FCI.
+The effect of the EN2 correction is most noticeable for the distance error and equilibrium properties, and less so for the NPE.
+In few cases, however, it does not improve the bare CI results, suggesting substantial cancelation of errors in the latter.
+This is seen at the hCI1 and hCI1.5 models and their EN2 corrected couterparts, for the NPE of \ce{HF}, \ce{F2} and \ce{N2}, and to some extent for the equilibrium properties of \ce{F2}.
+These are exceptions though, as the correction virtually always improve the accuracy of the computed observables.
 
@@
 
+Including the EN2 correction thus accelerates the rate of convergence as a function of $\Ndet$, when compared to the unperturbed calculations.
+In fact, the improvement is such that a given CI+EN2 model often provides more accurate results than the higher-order bare CI model sharing the same computational scaling.
+For instance, hCI1+EN2 and hCI2 display the same $\order*{N^6}$ computational scaling, and the former outperforms the latter in many cases.
+In Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}, hCI1+EN2 correlates with the second light green mark, and hCI2 with the fourth dark green mark.
+The former model provides smaller NPEs and distance errors for \ce{OH}, \ce{HF}, \ce{F2} (all single bond breaking),
+the opposite being true for multiple bond breaking.
+The advantage of the EN2 correction is more striking for the equilibrium properties,
+where hCI1+EN2 is more accurate than hCI2 for \ce{OH}, vinyl, ethylene, and \ce{N2},
+whereas both models perform similarly for \ce{HF}, \ce{F2}, \ce{H4}, \ce{H7}, and \ce{H8}.
+
+A similar trend on the effect of the EN2 correction also holds for excitation-based CI.
+For that, we compare CISD+EN2 and CISDT, both sharing a $\order*{N^8}$ scaling.
+CISD+EN2 correlates with the third dark red mark and CISDT with the fourth light red mark in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
+The errors for the four observables are systematically smaller with the former model,
+except for the NPE and distance error of \ce{H8} and the equilibrium properties of \ce{H7}, where they are comparable.
+%\ce{OH}, \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4} (small)
+
+Importantly, the hCI+EN2 route outperforms its excitation-based counterpart, thus preserving the advantage displayed without the perturbative correction.
+Furthermore, the convergence of the observables is overall more monotonic with the EN2 correction.
+In this sense, it regularizes oscillations seen in the unperturbed case.
+This is clearly seen in the equilibrium properties of vinyl, \ce{H7}, and \ce{F2},
+and in the NPE of \ce{OH}, \ce{HF}, \ce{F2}, and ethylene.
+
+The EN2 correction usually ameliorates the performance of the sCI route, though much less observed for hCI and excitation-based CI.
+This reflects the poorer reference provided by sCI, from which perturbation theory has a hard time to recover from.
+In a handful of cases, however, the EN2 correction does not improve the bare sCI results,
+for both NPEs (\ce{OH}, \ce{N2}, \ce{F2}, and ethylene)
+and equilibrium properties (vinyl, \ce{HF}, \ce{F2}, and ethylene).
+
+% rPT2 vs PT2 there are important differences
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 with respect to reference theoretical values for the excitation energies.
-The statistical errors are presented in Tab.~\ref{tab:1} and Tab.~\ref{tab:2},
+The statistical errors are presented in Tab.~\ref{tab:stat1} and Tab.~\ref{tab:stat2},
 for singly excited states of closed-shell and open-shell systems, respectively.
-For the lower order models, the calculations were performed for 46 (closed-shell) and 20 (open-shell) excited states.
-In turn, a subset of 14 (closed-shell) and 5 (open-shell) excited states was considered for the higher order models, given their more intensive computational cost.
+For the lower order models, the calculations were performed for 48 (closed-shell) and 19 (open-shell) excited states.
+In turn, a subset of 16 (closed-shell) and 6 (open-shell) excited states was considered for the higher order models, given their more intensive computational cost.
 Even though these subsets are too small for meaningful absolute statistics, they can still show the main trends on the higher order CI models.
@@
 
-We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Tab.~\ref{tab:1}.
+We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Tab.~\ref{tab:stat1}.
 %
@@
 %we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
-%For that, we performed calculations for a subset of 14 singly-excited states of small closed-shell molecular systems (indicated in the {\SupInf}).
+%For that, we performed calculations for a subset of 16 singly-excited states of small closed-shell molecular systems (indicated in the {\SupInf}).
 %(\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
-%with corresponding statistical errors shown in Tab.~\ref{tab:1}.
+%with corresponding statistical errors shown in Tab.~\ref{tab:stat1}.
 %Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order models.
@@
 }
-\label{tab:1}
+\label{tab:stat1}
 \begin{ruledtabular}
@@
 \hline
-CIS               & 46  & +0.04 & 0.61 & 0.60 \\
-CISD              & 14  & +4.09 & 4.09 & 4.18 \\
-CISDT             & 14  & +0.12 & 0.17 & 0.18 \\
-CISDTQ            & 14  & +0.15 & 0.15 & 0.17 \\
+CIS               & 48  & +0.04 & 0.61 & 0.60 \\
+CISD              & 16  & +4.09 & 4.09 & 4.18 \\
+CISDT             & 16  & +0.12 & 0.17 & 0.18 \\
+CISDTQ            & 16  & +0.15 & 0.15 & 0.17 \\
 %
 \hline
-hCI1              & 46  & +0.97 & 1.07 & 1.20 \\
-hCI1.5            & 46  & +1.91 & 1.91 & 2.00 \\
-hCI2              & 14  & +2.99 & 3.53 & 3.61 \\
-hCI2.5            & 14  & +1.95 & 1.95 & 2.06 \\
-hCI3              & 14  & +0.19 & 0.19 & 0.21 \\
-hCI3.5            & 14  & +0.11 & 0.11 & 0.13 \\
+hCI1              & 48  & +0.97 & 1.07 & 1.20 \\
+hCI1.5            & 48  & +1.91 & 1.91 & 2.00 \\
+hCI2              & 16  & +2.99 & 3.53 & 3.61 \\
+hCI2.5            & 16  & +1.95 & 1.95 & 2.06 \\
+hCI3              & 16  & +0.19 & 0.19 & 0.21 \\
+hCI3.5            & 16  & +0.11 & 0.11 & 0.13 \\
 %
 \hline
-$\Delta$CSF       & 46  & -0.73 & 0.80 & 0.94 \\
-$\Delta$CISD      & 46  & -0.13 & 0.17 & 0.22 \\
-$\Delta$CISDT     & 14  & -0.20 & 0.20 & 0.22 \\
-$\Delta$CISDTQ    & 14  & -0.02 & 0.02 & 0.02 \\
+$\Delta$CSF       & 48  & -0.73 & 0.80 & 0.94 \\
+$\Delta$CISD      & 48  & -0.13 & 0.17 & 0.22 \\
+$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 \\
+$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 \\
 %
 \hline
-$\Delta$hCI1      & 46  & -1.36 & 1.40 & 1.50 \\
-$\Delta$hCI1.5    & 46  & -2.91 & 2.91 & 3.10 \\
-$\Delta$hCI2      & 46  & -0.19 & 0.20 & 0.25 \\
-$\Delta$hCI2.5    & 14  & -0.27 & 0.27 & 0.30 \\
-$\Delta$hCI3      & 14  & -0.22 & 0.22 & 0.24 \\
-$\Delta$hCI3.5    & 14  & -0.08 & 0.08 & 0.09 \\
+$\Delta$hCI1      & 48  & -1.36 & 1.40 & 1.50 \\
+$\Delta$hCI1.5    & 48  & -2.91 & 2.91 & 3.10 \\
+$\Delta$hCI2      & 48  & -0.19 & 0.20 & 0.25 \\
+$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 \\
+$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 \\
+$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 \\
 %
 \hline
-sCI2/sCI2         & 46  & +1.26 & 1.34 & 1.49 \\
-sCI2/sCI0         & 46  & -0.41 & 0.57 & 0.89 \\
-$\Delta$sCI2/sCI2 & 46  & +0.69 & 0.79 & 0.90 \\
-$\Delta$sCI2/sCI0 & 46  & -0.99 & 0.99 & 1.15 \\
+sCI2/sCI2         & 48  & +1.26 & 1.34 & 1.49 \\
+sCI2/sCI0         & 48  & -0.41 & 0.57 & 0.89 \\
+$\Delta$sCI2/sCI2 & 48  & +0.69 & 0.79 & 0.90 \\
+$\Delta$sCI2/sCI0 & 48  & -0.99 & 0.99 & 1.15 \\
 %
@@
 We now shift to excitations for open-shell systems.
-The statistical errors are shown in Tab.~\ref{tab:2}.
+The statistical errors are shown in Tab.~\ref{tab:stat2}.
 %in the following comparison to the results for closed-shell excitations,
@@
 $\Delta$hCI2.5 is more accurate than $\Delta$hCI2, producing a MAE of \SI{0.06}{\eV}.
-(Accounting only for the 5 states considered for $\Delta$hCI2.5, $\Delta$hCI2 would have a MAE of \SI{0.09}{\eV}.)
+(Accounting only for the 6 states considered for $\Delta$hCI2.5, $\Delta$hCI2 would have a MAE of \SI{0.09}{\eV}.)
 The statistics at the next orders ($h=3$ and $h=3.5$) is limited, 
@@
 }
-\label{tab:2}
+\label{tab:stat2}
 \begin{ruledtabular}
@@
 \hline
-CIS               & 20 & +0.33 & 0.36 & 0.50 \\
-CISD              & 5  & +1.98 & 1.98 & 2.16 \\
-CISDT             & 5  & +0.02 & 0.03 & 0.05 \\
-CISDTQ            & 5  & +0.03 & 0.03 & 0.05 \\
+CIS               & 19 & +0.33 & 0.36 & 0.50 \\
+CISD              & 6  & +1.98 & 1.98 & 2.16 \\
+CISDT             & 6  & +0.02 & 0.03 & 0.05 \\
+CISDTQ            & 6  & +0.03 & 0.03 & 0.05 \\
 %
 \hline
-hCI1              & 20 & +0.70 & 0.77 & 0.96 \\
-hCI1.5            & 20 & +1.11 & 1.11 & 1.30 \\
-hCI2              & 5  & +0.69 & 0.69 & 0.85 \\
-hCI2.5            & 5  & +0.01 & 0.03 & 0.04 \\
-hCI3              & 5  & -0.00 & 0.02 & 0.03 \\
-hCI3.5            & 5  & -0.00 & 0.01 & 0.02 \\
-%hCI4              & 5  &       &      &      \\
+hCI1              & 19 & +0.70 & 0.77 & 0.96 \\
+hCI1.5            & 19 & +1.11 & 1.11 & 1.30 \\
+hCI2              & 6  & +0.69 & 0.69 & 0.85 \\
+hCI2.5            & 6  & +0.01 & 0.03 & 0.04 \\
+hCI3              & 6  & -0.00 & 0.02 & 0.03 \\
+hCI3.5            & 6  & -0.00 & 0.01 & 0.02 \\
+%hCI4              & 6  &       &      &      \\
 %
 \hline
-$\Delta$CSF       & 20 & -0.04 & 0.41 & 0.58 \\
-$\Delta$CISD      & 20 & +0.01 & 0.12 & 0.21 \\
-$\Delta$CISDT     & 5  & -0.05 & 0.05 & 0.09 \\
-$\Delta$CISDTQ    & 5  & -0.02 & 0.02 & 0.02 \\
+$\Delta$CSF       & 19 & -0.04 & 0.41 & 0.58 \\
+$\Delta$CISD      & 19 & +0.01 & 0.12 & 0.21 \\
+$\Delta$CISDT     & 6  & -0.05 & 0.05 & 0.09 \\
+$\Delta$CISDTQ    & 6  & -0.02 & 0.02 & 0.02 \\
 %
 \hline
-$\Delta$hCI1      & 20 & -0.19 & 0.34 & 0.48 \\
-$\Delta$hCI1.5    & 20 & -0.10 & 0.46 & 0.70 \\
-$\Delta$hCI2      & 20 & -0.01 & 0.11 & 0.16 \\
-$\Delta$hCI2.5    & 5  & -0.06 & 0.06 & 0.12 \\
-$\Delta$hCI3      & 5  & -0.03 & 0.03 & 0.06 \\
-$\Delta$hCI3.5    & 5  & -0.02 & 0.02 & 0.02 \\
-%$\Delta$hCI4      & 5  &       &      &      \\
+$\Delta$hCI1      & 19 & -0.19 & 0.34 & 0.48 \\
+$\Delta$hCI1.5    & 19 & -0.10 & 0.46 & 0.70 \\
+$\Delta$hCI2      & 19 & -0.01 & 0.11 & 0.16 \\
+$\Delta$hCI2.5    & 6  & -0.06 & 0.06 & 0.12 \\
+$\Delta$hCI3      & 6  & -0.03 & 0.03 & 0.06 \\
+$\Delta$hCI3.5    & 6  & -0.02 & 0.02 & 0.02 \\
+%$\Delta$hCI4      & 6  &       &      &      \\
 %
 \hline
-sCI1              & 20 & +0.24 & 0.39 & 0.61 \\
-$\Delta$sCI1      & 20 & -0.11 & 0.27 & 0.38 \\
+sCI1              & 19 & +0.24 & 0.39 & 0.61 \\
+$\Delta$sCI1      & 19 & -0.11 & 0.27 & 0.38 \\
 \end{tabular}
@@
 %\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
-%for the Set of 14 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+%for the Set of 16 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
 %}
-%\label{tab:3}
+%\label{tab:stat3}
 %\begin{ruledtabular}
@@
 
-We refer back to Tab.~\ref{tab:1} to discuss the performance of seniority-based CI for excited states of closed-shell systems.
+We refer back to Tab.~\ref{tab:stat1} to discuss the performance of seniority-based CI for excited states of closed-shell systems.
 The first model, sCI2/sCI2, systematically overestimates the excitation energies, and presents a MAEs of \SI{1.34}{\eV}.
@@
 %This improvement, however, is not as expressive as observed for excitataion-based and hierarchy-based methods.
-%Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:1}.
+%Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:stat1}.
 %The large errors, even following the state-specific approach, combined with the formal $\order*{e^N}$ scaling of these methods, render seniority-based CI unsuitable for calculations of excitation energies.
@@
 The situation for open-shell systems is quite different, in close parallel to the previous discussion of hCI.
-As shown in Tab.~\ref{tab:2}, sCI provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
+As shown in Tab.~\ref{tab:stat2}, sCI provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
 The reason should not be surprising at this point:

### 2023-07-18 Manuscript/sup.tex
saving work

@@
 
-%%% FIG S2 %%%
-\begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_pes_rpt2}
-\caption{
-Potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
-according to hCI (green), excitation-based CI (red) and sCI (blue) models
-corrected with second-order Epstein-Nesbet perturbative correction.
-}
-\label{fig:plot_pes_rpt2}
-\end{figure*}
-%%% %%% %%%
-
 %%% FIG S3 %%%
@@
 Potential energy curves for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
-according to hCI (green), excitation-based CI (red) and sCI (blue) models.
+according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+with (light tones) and without (dark tones) the EN2 perturbative correction.
 Results first reported in Ref. \onlinecite{Kossoski_2022}.
@@
 \label{fig:plot_pes_closed}
-\end{figure*}
-%%% %%% %%%
-
-%%% FIG S4 %%%
-\begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_pes_closed_rpt2}
-\caption{
-Potential energy curves for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
-according to hCI (green), excitation-based CI (red) and sCI (blue) models
-corrected with second-order Epstein-Nesbet perturbative correction.
-}
-\label{fig:plot_pes_closed_rpt2}
-\end{figure*}
-%%% %%% %%%
-
-%%% FIG S5 %%%
-\begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_all_rpt2}
-\caption{
-Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{OH}, vinyl, and \ce{H7},
-as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models
-corrected with second-order Epstein-Nesbet perturbative correction.
-For \ce{CN}, the perturbative correction produces discontinuous potential energy curves due to the change of character of the unperturbed ground state, as explained in the main text.
-}
-\label{fig:plot_all_rpt2}
-\end{figure*}
-%%% %%% %%%
-
-%%% FIG S6 %%%
-\begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_all_closed}
-\caption{
-Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
-as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models.
-Results first reported in Ref. \onlinecite{Kossoski_2022}.
-}
-\label{fig:plot_all_closed}
-\end{figure*}
-%%% %%% %%%
-
-%%% FIG S7 %%%
-\begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_all_closed_rpt2}
-\caption{
-Non-parallelity errors, distance errors, equilibrium distances, and vibrational frequencies (or force constants), for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
-as a function of the number of determinants, according to hCI (green), excitation-based CI (red) and sCI (blue) models,
-corrected with second-order Epstein-Nesbet perturbative correction.
-}
-\label{fig:plot_all_closed_rpt2}
 \end{figure*}

### 2023-07-18 H4_cc-pvdz/compute_NPE.pl

@@ -104,6 +104,10 @@ close(Fi);
             my @line =  split( /\s+/, $_ );
             $m++;
             $x = $line[0];
+            if($x ne $x0[$m]) {
+                    print "Not the same coordinates for file $files[$j]! \n";
+                    exit;
+            }
             $y = $line[1];
             $dy    = $y - $y0[$m];
             $maxy = max($maxy,$dy);

### 2023-07-18 OH_cc-pvdz/rpt2/compute_NPE.pl

@@ -104,6 +104,10 @@ close(Fi);
             my @line =  split( /\s+/, $_ );
             $m++;
             $x = $line[0];
+	    if($x ne $x0[$m]) {
+		    print "Not the same coordinates for file $files[$j]! \n";
+		    exit;
+	    }
             $y = $line[1];
             $dy    = $y - $y0[$m];
             $maxy = max($maxy,$dy);

### 2023-07-18 vinyl_cc-pvdz/rpt2/compute_NPE.pl

@@ -104,6 +104,10 @@ close(Fi);
             my @line =  split( /\s+/, $_ );
             $m++;
             $x = $line[0];
+    	    if($x ne $x0[$m]) {
+     		    print "Not the same coordinates for file $files[$j]! \n";
+    		    exit;
+	    }
             $y = $line[1];
             $dy    = $y - $y0[$m];
             $maxy = max($maxy,$dy);
@@ -114,7 +118,7 @@ close(Fi);
 
         if($x==$xcut){
           my $npe = $maxy - $miny;
-          my $closeness = abs($maxy) + abs($miny);
+	  my $closeness = abs($maxy) + abs($miny);
           print Fo "$index[$j] $index2[$j] $det[$j]  $npe  $closeness  $maxy  $miny \n";
 	}
 	#     }

### 2023-07-21 Manuscript/manuscript.tex
saving work

@@
 %excitations of increasingly higher order give rise to a systematically improvable sequence of CI methods:
-The second goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction, \cite{Garniron_2019}
+The second goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction (both standard and renormalized), \cite{Garniron_2019}
 impacts different properties of ground-state closed-shell systems and radicals, for excitation-based, sCI and hCI methods.
@@
 {\QP} \cite{Garniron_2019} was also employed to perform all the CI calculations presented here.
-In a given calculation, the energies are considered to be converged when the (largest) renormalized EN2 correction computed in the truncated Hilbert space 
+In a given calculation, the energies are considered to be converged when the (largest) EN2 correction computed in the truncated Hilbert space 
 lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
@@
 The ground- and excited-state CI energies are obtained with the Davidson iterative algorithm. \cite{Davidson_1975}
-For a given approximate CI model, we futher evaluated the bare and the renormalized EN2 perturbative correction. \cite{Garniron_2019} 
+For a given approximate CI model, we futher evaluated the standard and the renormalized EN2 perturbative correction. \cite{Garniron_2019} 
 This calculation involves a single loop over the determinants left outside the truncated (internal) space but connected to it via at most double excitations.
@@
 From these PECs, we obtain the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies, which are plotted as a function of $\Ndet$ in Fig.~\ref{fig:plot_all}.
+Each point in the figure denotes one CI model.
+For instance, the hCI route is represented by the dark green line, with the first point being HF, the second being hCI1, the third being hCI1.5, and so on.
 The corresponding results for the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) are shown in Fig.~\ref{fig:plot_all_closed}.
@@
 These previous results are reproduced in Fig.~\ref{fig:plot_all_closed} together with the present results, which include such correction.
-First we will focus on the results for the bare CI models (hCI, excitation-based CI and sCI) for the open-shell systems, represented by the dark tones,
-leaving the discussion about the perturbative correction (light tones) for later.
+First we will focus on the results for the bare CI models (hCI, excitation-based CI and sCI) for the open-shell systems (represented by the dark tones in the figures),
+leaving the discussion about the perturbative correction (light tones) on both open-shell and closed-shell systems for later.
 
@@
 as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
-with (light tones) and without (dark tones) the EN2 perturbative correction.
+with (light tones) and without (dark tones) the standard EN2 perturbative correction.
 The dashed lines represent the FCI results.}
@@
 
+%%% FIG 4 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_all_closed}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, 
+as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+with (light tones) and without (dark tones) the standard EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_all_closed}
+\end{figure*}
+%%% %%% %%%
+
 We found that hCI typically improves the convergence with respect to $\Ndet$ when compared to excitation-based CI, for the four properties surveyed here.
 The present finding for open-shell systems therefore parallels the previously reported superiority of hCI for closed-shell systems. \cite{Kossoski_2022}
-Concerning the NPEs, the advantage of hCI is pronounced for \ce{OH} (involving a single bond breaking), 
+The smaller NPEs obtained with hCI are linked to the better description of the PECs at dissociation.
+This is due to a larger fraction of static correlation being recovered,
+stemming from the classes of determinants appearing lower in the seniority-excitation map (see Fig.~\ref{fig:hCI}),
+namely, determinants high in excitation degree $e$ and low in seniority $s$, absent in excitation-based CI of the same order.
+
+Still concerning the NPEs, the advantage of hCI is pronounced for \ce{OH} (involving a single bond breaking), 
 becoming less so, though still clear, for vinyl (double bond breaking) and \ce{CN} (triple bond breaking),
-whereas for \ce{H7} (multiple bond breaking), it still shows a small advantage in comparison to excitation-based CI.
-This dependence on the order of bond breaking had also been observed for closed-shell systems \cite{Kossoski_2022} (as can be seen in Fig.~\ref{fig:plot_all_closed}).
-We also found overall smaller distance errors with hCI than with excitation-based CI, although the improvement is less noticeable in this case.
+whereas for \ce{H7} (multiple bond breaking), it still displays a small advantage in comparison to excitation-based CI.
+This dependence on the order of bond breaking had also been observed for closed-shell systems \cite{Kossoski_2022} (as can be seen in Fig.~\ref{fig:plot_all_closed}),
+and would be expected, as describing dissociation becomes increasingly more challenging as the bond order increases.
+%This trend can be explained by the growing amount of static correlation as more bonds are broken.
+Even though hCI manages to recover more static correlation than excitation-based CI,
+the result is more striking for single bond breaking, whereas multiple bond breaking (like \ce{H7} and \ce{H8}) inevitably requires higher order excited determinants.
+
+We also found overall smaller distance errors with hCI than with excitation-based CI, although the improvement is not as remarkable as observed for the NPEs.
 hCI further leads to a somewhat faster convergence of the equilibrium geometries for the open-shell systems, most noticeably for \ce{CN}.
-Finally, hCI systematically outperforms excitation-based CI for the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+Finally, hCI systematically outperforms excitation-based CI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+Given the overall superiority of hCI for the equilibrium properties, it also manages to account for more dynamic correlation than excitation-based CI,
+for both open-shell and closed-shell systems.
 
@@
 
-The results for the CI models corrected by the EN2 energy are represented by the light tones in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
+% EN2 correction
+
+The results for the CI models corrected by the standard EN2 energy are represented by the light tones in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
 %The corresponding PECSs are shown in the {\SupInf}.
@@
 This is simply due to its several close-lying excited states. \cite{}
-If we were to follow the PEC of a given symmetry (and not the lowest-lying state as we did here), there would be no discontinuity stemming from the EN2 correction.
+If we were to follow the PEC of a given symmetry (and not the lowest-lying state as we did here), there would be no discontinuity stemming from the EN2 contribution.
 
-%%% FIG 4 %%%
-\begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_all_closed}
-\caption{
-Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, 
-as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
-with (light tones) and without (dark tones) the EN2 perturbative correction.
-The dashed lines represent the FCI results.}
-\label{fig:plot_all_closed}
-\end{figure*}
-%%% %%% %%%
-
-We found that the EN2 correction significantly reduces the errors with respect to the FCI results.
+We found that the EN2 perturbative correction significantly reduces the errors of the bare CI calculations.
 This is observed for all the CI models, systems, and observables considered here.
-In the particular case of \ce{OH}, the improvement is massive, as even the lower levels of CI+EN2 deliver very close results to FCI.
-The effect of the EN2 correction is most noticeable for the distance error and equilibrium properties, and less so for the NPE.
-In few cases, however, it does not improve the bare CI results, suggesting substantial cancelation of errors in the latter.
-This is seen at the hCI1 and hCI1.5 models and their EN2 corrected couterparts, for the NPE of \ce{HF}, \ce{F2} and \ce{N2}, and to some extent for the equilibrium properties of \ce{F2}.
-These are exceptions though, as the correction virtually always improve the accuracy of the computed observables.
+For \ce{OH}, in particular, the improvement is massive, as even the lower levels of CI+EN2 deliver very close results to FCI.
+Importantly, the hCI+EN2 route outperforms its excitation-based counterpart, thus preserving the advantage observed without the perturbative correction.
+Similarly, the correction typically maintains the comparable performance of the two routes when this is the case according to the unperturbed CI calculations.
 
@@
 
-Including the EN2 correction thus accelerates the rate of convergence as a function of $\Ndet$, when compared to the unperturbed calculations.
-In fact, the improvement is such that a given CI+EN2 model often provides more accurate results than the higher-order bare CI model sharing the same computational scaling.
+%Including the EN2 correction thus accelerates the rate of convergence as a function of $\Ndet$, when compared to the unperturbed calculations.
+%In fact, 
+The improvement brought by the perturbative correction is such that a given CI+EN2 model often provides more accurate results than the higher-order bare CI model sharing the same computational scaling.
 For instance, hCI1+EN2 and hCI2 display the same $\order*{N^6}$ computational scaling, and the former outperforms the latter in many cases.
@@
 where hCI1+EN2 is more accurate than hCI2 for \ce{OH}, vinyl, ethylene, and \ce{N2},
-whereas both models perform similarly for \ce{HF}, \ce{F2}, \ce{H4}, \ce{H7}, and \ce{H8}.
+whereas both models perform similarly for the remaining systems.
+It is worth recalling that the prefactor associated with the EN2 calculation is smaller than that associated with the higher-order CI calculation,
+which in practice makes hCI1+EN2 cheaper than hCI2.
+In the above examples, the accuracy of hCI1+EN2 is actually better or comparable to that of hCI2.5 or CISDT, considerably more expensive models.
+%\ce{HF}, \ce{F2}, \ce{H4}, \ce{H7}, and \ce{H8}.
 
-A similar trend on the effect of the EN2 correction also holds for excitation-based CI.
+A similar impact on the effect of the EN2 correction also holds for excitation-based CI.
 For that, we compare CISD+EN2 and CISDT, both sharing a $\order*{N^8}$ scaling.
@@
 except for the NPE and distance error of \ce{H8} and the equilibrium properties of \ce{H7}, where they are comparable.
+
 %\ce{OH}, \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4} (small)
 
-Importantly, the hCI+EN2 route outperforms its excitation-based counterpart, thus preserving the advantage displayed without the perturbative correction.
-Furthermore, the convergence of the observables is overall more monotonic with the EN2 correction.
-In this sense, it regularizes oscillations seen in the unperturbed case.
+The EN2 correction usually ameliorates the performance of the sCI route, though to a lesser extent than observed for hCI and excitation-based CI.
+This reflects the poorer reference provided by sCI, from which perturbation theory struggles to recover from.
+In a handful of cases, the EN2 correction does not improve the bare sCI results,
+for both NPEs (\ce{OH}, \ce{N2}, \ce{F2}, and ethylene)
+and equilibrium properties (vinyl, \ce{HF}, \ce{F2}, and ethylene).
+Overall, the sCI+EN2 route falls behind its alternatives.
+
+In few cases, however, it does not improve the bare CI results, suggesting important cancellation of errors in the latter.
+This is seen when employing the hCI1 and hCI1.5 models and their EN2 corrected couterparts to compute the NPE of \ce{HF}, \ce{F2} and \ce{N2}, and to some extent the equilibrium properties of \ce{F2}.
+These are exceptions though, as the EN2 correction practically always improve the accuracy of the computed observables.
+It further leads to overall more monotonic convergence of the observables.
+In this sense, it regularizes oscillations seen in the unperturbed case, probably related to cancellation of errors at the lower orders.
 This is clearly seen in the equilibrium properties of vinyl, \ce{H7}, and \ce{F2},
@@
 
-The EN2 correction usually ameliorates the performance of the sCI route, though much less observed for hCI and excitation-based CI.
-This reflects the poorer reference provided by sCI, from which perturbation theory has a hard time to recover from.
-In a handful of cases, however, the EN2 correction does not improve the bare sCI results,
-for both NPEs (\ce{OH}, \ce{N2}, \ce{F2}, and ethylene)
-and equilibrium properties (vinyl, \ce{HF}, \ce{F2}, and ethylene).
-
-% rPT2 vs PT2 there are important differences
+Instead of the usual EN2 correction, discussed so far and shown in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}, 
+one can compute the renormalized EN2 correction. \cite{Garniron_2019}
+Analogous figures showing the convergence of observables with the usual and with the renormalized forms are shown in the {\SupInf}.
+There are often significant differences between the two forms at the lower orders of CI, which in turn become negligeable at higher orders.
+More often than not, the usual correction performs better.
+The difference is expressive for ethylene, \ce{N2}, \ce{H7}, and \ce{H8}, and to a lesser extent for vinyl.
+For \ce{H4} the results are mixed.
+The renormalized correction is slightly more accurate for \ce{OH}, \ce{HF}, and \ce{F2}.
+These results suggest that renormalizing the EN2 energy may only be helpful for single bond breaking, though by a small amount,
+whereas it worsens the results for multiple bond breaking.
+Overall, the usual EN2 correction should be prefered when employed in combination with the approximate CI models considered here.
 
@@
 
-We now shift to excitations for open-shell systems.
+We now shift to the discussion of excitations for open-shell systems.
 The statistical errors are shown in Tab.~\ref{tab:stat2}.
-%in the following comparison to the results for closed-shell excitations,
-%In comparison to the previous discussion,
 The key difference on the calculations for open-shell excitations is that the same type of reference determinant were employed for ground- and excited-state calculations,

### 2023-08-06 Manuscript/manuscript.tex
saving work

@@
 {\bf Abstract:}
+Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number, 
+which showed to efficiently recover dynamic and static correlation for closed-shell molecular systems [J. Phys. Chem. Lett. 2022, 13, 19, 4342].
+Here, we generalize hCI for arbitrary reference determinants, allowing calculations for radicals and for excited states in a state-specific approach.
+We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI)
+by evaluating how different properties of ground-state radicals converge to the full CI limit.
+hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
+Employing the second-order Epstein-Nesbet perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
+We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
+Our results underline that both the choice of reference and orbitals drive the fine balance between correlation of ground and excited states.
+State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.
+In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
 %\bigskip
@@
 
-The most well-known CI route is excitation-based.
+The most well-known CI route is based on the excitation degree.
 Starting from a reference determinant, typically the Hartree-Fock (HF) determinant,
-one generates all determinants that are connected to it by excitating at most $e$ electrons.
-The excitation degree $e$ thus defines the order of the approximate CI model:
+one generates all determinants that are connected to it by exciting at most $e$ electrons.
+The excitation degree $e$ thus defines the order of the approximate excitation-based CI (eCI) model:
 CI with single excitations (CIS); CI with single and double excitations (CISD), CI with single, double and triple excitations (CISDT), and so on.
+The eCI route rather quickly captures dynamic (weak) correlation, though it struggles with the description of static (strong) correlation.
+
 A different CI route is based on the seniority number $s$ (the number of unpaired electrons in a given determinant).
 In seniority-based CI (sCI) \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}, there is no reference determinant, and one accounts for all determinants having seniority equal or less than $s$.
-For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually refered to as doubly-occupied CI (DOCI),
+In contrast to eCI, sCI recovers static correlation more efficiently and works well to describe molecular dissociation, \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
+at the expense of a poorer account of dynamic correlation and a higher computational cost.
+For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually referred to as doubly-occupied CI (DOCI),
 which is followed by the higher-order models sCI2, sCI4, and so on.
 For an odd number of electrons, the seniority route follows along odd numbers of $s$: sCI1, sCI3, and so on.
-%Limited studies for radicals and excited states
 
-%sCI works well to describe molecular dissociation \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
-%For small systems, FCI is still attainable and one can thus gauge the performance of approximate CI models.
 We have recently introduced a third CI route, hierarchy CI (hCI), \cite{Kossoski_2022}
-where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$ according to $h = (e+s/2)/2$.
-This definition ensures that all types of determinants whose number share the same scaling with system size are included at the same hierarchy $h$.
+where the Hilbert space is partitioned according to a hierarchy parameter $h$ that combines the excitation degree $e$ and the seniority number $s$ defined as $h = (e+s/2)/2$.
+This definition ensures that all classes of determinants whose number share the same scaling with system size are included at the same hierarchy $h$.
 For different properties and closed-shell molecular systems, 
-hCI was found to display an overall faster convergence to the FCI results than the traditional excitation-based CI. \cite{Kossoski_2022}
-However, hCI has only been defined for a closed-shell reference, thus systems with an even number of electrons.
-Here, our first goal is to generalize hCI for open-shell systems and to assess the performance of this class of methods by performing calculations for various properties of several radicals.
-%A generalization of the method for open-shell systems has not yet been proposed.
-%excitations of increasingly higher order give rise to a systematically improvable sequence of CI methods:
-The second goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction (both standard and renormalized), \cite{Garniron_2019}
-impacts different properties of ground-state closed-shell systems and radicals, for excitation-based, sCI and hCI methods.
+hCI was found to display an overall faster convergence toward the FCI results than the traditional eCI route. \cite{Kossoski_2022}
+In this sense, it was able to recover both static and dynamic correlation more effectively than either eCI or sCI.
+However, hCI has only been defined for a closed-shell reference, thus being limited to the ground state of systems with an even number of electrons.
 
-%Whereas the first CI root represents the electronic ground state, higher-lying roots give access to excited states.
-%Similarly, the favourable results of hCI when compared to excitation-based CI for ground states raises the question on how well it can describe excitation energies.
-%The performance of excitation-based CI for excited states is well stablished, \cite{} but remains an open question for sCI methods,
-Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022} our third goal is to define and assess the accuracy of hCI models for excited states.
+There are two possible approaches to target excited states with CI methods.
+One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix, which we refer to as ground-state based approach.
+Instead, one may optimize the orbitals for the excited state of interest (described with an appropriate reference), followed by a separate CI calculation,
+in a so-called state-specific approach ($\Delta$CI).
+There has been a recent surge in the development of state-specific methods, covering 
+single-reference and multiconfigurational self-consistent field,
+\cite{Ziegler_1977,Burton_2021,Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Hanscam_2022,Kossoski_2023,Marie_2023}
+% references for SCF
+%\cite{Ziegler_1977,Burton_2021}
+% multiconfigurational SCF, \cite{Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Kossoski_2023,Marie_2023}
+density functional theory, 
+\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022}
+and coupled cluster (CC)
+\cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023}
+methods.
+In particular, by employing a minimal configuration state function (CSF) reference, 
+we have recently shown that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues. \cite{Kossoski_2023}
+
+Here, our first goal is to generalize hCI for an arbitrary type of reference, thus extending its applicability from ground-state closed-shells \cite{Kossoski_2022}
+to radicals and state-specific excited state calculations.
+This is done in Sec.~\ref{sec:hCI}.
+The computational details about the implementation and the specific calculations performed here are presented in Sec.~\ref{sec:compdet}.
+In Sec.~\ref{sec:res_A}, we assess the performance of hCI, eCI and sCI models by calculating various properties of four ground-state radicals, which comprises our second goal.
+Our third goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction (both standard and renormalized), \cite{Garniron_2019}
+impacts various properties of ground-state closed-shell systems and radicals, for hCI, eCI, and sCI models.
+This part is presented in Sec.~\ref{sec:res_B}.
+Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022} 
+and by the promising set of excitation-based $\Delta$CI, \cite{Kossoski_2023} here we explore hCI models for excited states, following both ground-state based and state-specific approaches.
+In thise sense, our fourth goal is to assess the accuracy of hCI models for excited states, which is the subject of Sec.~\ref{sec:res_C}.
 Furthermore, to the best of our knowledge, sCI models have not yet been directly used to target excited states,
@@
 \cite{Boguslawski_2016b,Boguslawski_2016c,Boguslawski_2019,Nowak_2019,Kossoski_2021,Marie_2021,Tecmer_2022,Rishi_2023,Nowak_2023} 
-Our fourth goal is therefore to introduce and gauge sCI models for excited states.
-There are two possible approaches to target excited states with CI.
-One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix, which we refer to as ground-state based approach.
-Instead, one may optimize the orbitals for the excited state of interest (described with an appropriate reference), followed by a separate CI calculation,
-in a so-called state-specific approach ($\Delta$CI).
-There has been a recent surge in the development of state-specific methods, covering 
-single-reference and multiconfigurational self-consistent field,
-\cite{Ziegler_1977,Burton_2021,Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Hanscam_2022,Kossoski_2023,Marie_2023}
-% add references for SCF
-%\cite{Ziegler_1977,Burton_2021}
-%multiconfigurational SCF, \cite{Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Kossoski_2023,Marie_2023}
-density functional theory, 
-\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022}
-and coupled-cluster (CC)
-\cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023}
-methods.
-In particular, by employing a minimal configuration state function (CSF) reference, 
-we have recently shown that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues. \cite{Kossoski_2023}
-Based on this promising set of results for excitation-based CI, here we explore both ground-state based and state-specific approaches for sCI and hCI models.
+Our fifth and last goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state based and state-specific sCI models for excited states.
+Sec.~\ref{sec:conclusion} closes the present contribution with the main conclusions and perspectives.
 
@@
 
-We introduced hCI \cite{Kossoski_2022} as a truncation of the Hilbert space based on a two-dimensional map of determinants 
-based on their seniority $s$ and their excitation degree $e$ with respect to a reference determinant (see Fig.~\ref{fig:hCI}).
-It is built such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ enter at the same order of truncation,
-which means moving diagonaly in this seniority-excitation map.
+We introduced hCI \cite{Kossoski_2022} as a particular truncation of the Hilbert space,
+represented as a two-dimensional map of determinants built from their seniority $s$ and their excitation degree $e$ with respect to a reference determinant (see Fig.~\ref{fig:hCI}).
+hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
+which means moving diagonally in the seniority-excitation map (denoted by the color tones in Fig.~\ref{fig:hCI}).
 
@@
 
-hCI was initially defined for a closed-shell reference. \cite{Kossoski_2022}
-Here we generalize hCI for an arbitrary type of reference Slater determinant and for systems with an odd number of electrons.
-With respect to a given reference determinant, we define the hierarchy $h$ of a target determinant as
+hCI was initially introduced for a closed-shell reference. \cite{Kossoski_2022}
+Here, we generalize it for an arbitrary Slater determinant reference, including systems with an odd number of electrons.
+With respect to a given reference determinant, we define the hierarchy $h$ of a candidate determinant to be included in the truncated hCI model as
 \begin{equation}
@@
 \end{equation}
-where $s$ and $s_0$ denote the seniority of target and reference determinants, respectively, and $e$ represents the excitation degree that connects the two determinants.
-The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with an even (odd) numbers of electrons, whereas $e$ is also an integer.
+where $s$ and $s_0$ denote the seniority of the candidate and reference determinants, respectively, and $e$ represents the excitation degree that connects them.
+The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with an even (odd) numbers of electrons. The excitation degree $e$ is also an integer.
 Therefore, $h$ can assume integer or half-integer values, for any type of reference determinant.
-For a given hCI model defined by $h$, we include all the determinants having a hierarchy less than or equal to $h$ with respect to any determinant in the reference.
+For a given hCI model defined by $h$, we include all the candidate determinants having a hierarchy less than or equal to $h$ with respect to any determinant in the reference.
+This definition allow us to build multireference hCI models as well.
 Notice that Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant, when $s_0 = 0$.
+hCI can be viewed as the CI model that provides increasingly more dissimilar determinants from the given reference determinant.
+The level of dissimilarity is represented by the hierarchy $h$, which accounts for differences in orbital occupation (through the excitation degree $e$)
+and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
 
-The above definition ensures that all types of determinants with a given scaling enter at the same hierarchy $h$.
+The definition in Eq.~\eqref{eq:h} guarantees that all classes of determinants sharing the same scaling of $\Ndet$ with $N$ enter at the same hierarchy $h$.
 % computational scaling
 We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
-For these three types of reference, Fig.~\ref{fig:determinants} shows the classes of determinants included up to hCI1.
-hCI can also be interpreted as the CI model that provides increasingly more dissimilar determinants from the given reference determinant.
-The level of dissimilarity is represented by the hierarchy $h$, which accounts for differences in orbital occupation (through the excitation degree $e$)
-and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
+These references are suitable for describing the ground state of a closed-shell system, of an open-shell system, and a singly-excited state of a closed-shell, respectively.
+For these three types of reference, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
 
@@
 closed-shell (top left), open-shell with one unpaired electron (top right) and open-shell with two unpaired electrons (bottom).
-Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied and unoccupied).
-}
+Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied and unoccupied).}
 \label{fig:determinants}
@@
 
-The first order for a closed-shell reference determinant is hCI1 (see top left of Fig.~\ref{fig:determinants}), which accounts for all single excitations (as CIS) 
-plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital).
-For both types of determinants, their number scales as $N^2$, and are thus taken into account at the same hierarchy of hCI ($h=1$ in this case).
-An odd number of excitations from a closed-shell reference always leave unpaired electrons, thus the empty blocks at $s=0$ and $e$ odd.
-hCI1.5 includes to the hCI1 determinants the set of double excitations where two electrons are unpaired.
-At the next integer order, hCI2 accounts for all types of determinants where $\Ndet$ scale as $N^4$ or less.
-That means all single and double excitations (as CISD), plus the subset of triple excitations that leave only two unpaired electrons, 
+The first order for a closed-shell reference determinant is hCI1 (see left of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS) 
+plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital),
+as shown in the top left of Fig.~\ref{fig:determinants}.
+For both classes of determinants, their number scales as $N^2$, and are thus taken into account at the same hierarchy of hCI ($h=1$ in this case).
+An odd number of excitations from a closed-shell reference always leave unpaired electrons, hence the empty blocks at $s=0$ for odd $e$.
+hCI1.5 augments the set of hCI1 determinants with the set of double excitations where two electrons are unpaired.
+At the next integer order, hCI2 incorporates all classes of determinants where $\Ndet$ scale as $N^4$.
+In total, it accounts for all single and double excitations (as CISD), plus the subset of triple excitations that leave only two unpaired electrons, 
 plus the subset of quadruple excitations where no electrons are unpaired.
 More generally, in hCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2h}$.
-Similarly, for excitation-based CI, $\Ndet$ scales as $N^{2e}$, where $e$ is the excitation degree.
+Likewise, in eCI, $\Ndet$ scales as $N^{2e}$.
 The computational scaling along the two routes is shown in Tab.~\ref{tab:scaling}.
@@
 \begin{table}[ht!]
-\caption{Computational scaling along the hierarchy-based and excitation-base CI routes, with and without the EN2 perturbative correction.}
+\caption{Computational scaling in terms of the number of basis functions $N$ for the hierarchy-based and excitation-base CI routes, with and without the EN2 perturbative correction.}
 \label{tab:scaling}
@@
 The simplest reference is an open-shell determinant with one unpaired electron (see center pannel of Fig.~\ref{fig:hCI} and top right of Fig.~\ref{fig:determinants}).
-In contrast to the closed-shell reference case, here there are no empty blocks in the excitation-seniority map.
-Actually, the hCI series presents the hCI0.5 level, accounting for the single excitations from and into the singly-occupied orbital only.
-For HF orbitals, however, these excitations do not connect with the reference, and hCI0.5 thus provides the same energy as restricted open-shell HF.
-More generally, since there are more types of orbitals for the open-shell case (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell case (doubly-occpupied or unoccupied),
-there is correspondingly more classes of determinants at a given level of hCI.
+In contrast to the closed-shell case, here there are no empty blocks in the excitation-seniority map.
+Actually, the hCI series displays the hCI0.5 level, accounting only for the single excitations from and into the singly-occupied orbital.
+For HF orbitals, however, these excitations do not connect with the reference, and hCI0.5 provides the same energy as restricted open-shell HF.
+More generally, since there are more types of orbitals for the open-shell reference (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell reference (doubly-occupied or unoccupied),
+there are correspondingly more classes of determinants at a given level of hCI.
 This can be appreciated from the comparison of hCI1 for closed- and open-shell references, shown in Fig.~\ref{fig:determinants}.
@@
 Finally, one can employ a high-spin reference determinant to define the hCI models.
-The case of two unpaired electrons is shown in the right pannel of Fig.~\ref{fig:hCI}.
-The key difference with respect to the closed-shell case is that the hCI sequence is displaced one block to the right in the seniority-excitation map,
+The case of two unpaired electrons is shown in the right pannel of Fig.~\ref{fig:hCI},
+which would be appropriate for singly-excited states of closed-shells.
+The key difference in the hCI sequence, with respect to the closed-shell case, lies in the displacement by one block to the right in the seniority-excitation map,
 reflecting the shift from the $s_0=0$ to the $s_0=2$ reference.
-In the two previous cases, the hCI0 level only includes the reference determinant.
+In the two cases discussed, the hCI0 level only included the reference determinant.
 For a $s=2$ reference, however, hCI0 not only accounts for the reference determinant,
-but also the two closed-shell determinants produced by the single excitation that pairs the two unpaired electrons.
-The lower order classes of determinants can be seen in the bottom of Fig.~\ref{fig:determinants},
-which clearly outnumber the fewer classes associated with a closed-shell reference.
+but also for the two closed-shell determinants produced by the single excitation that pairs the two unpaired electrons.
+The hCI1 classes of determinants for the $s_0=2$ reference can be seen in the bottom of Fig.~\ref{fig:determinants},
+which clearly outnumber the fewer classes associated with the $s_0=0$ and $s_0=1$ references.
 One could adopt even higher-spin references, further displacing the hCI sequence to the right in the seniority-excitation map.
@@
 Notice that the number of such determinants do not depend on the size of the system (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
-Similarly, hCI can be built on top of a high-spin reference for odd number of electrons.
+In a similar fashion, hCI can be built on top of a high-spin reference for odd number of electrons.
 
@@
 By allowing only for the determinants that are connected with the reference determinant(s) up to a given maximum hierarchy $h$,
-the CIPSI algorithm is restricted to the truncated Hilbert space defined by the reference determinant(s) and the value of $h$.
-{\QP} \cite{Garniron_2019} was also employed to perform all the CI calculations presented here.
+the CIPSI algorithm is restricted to the truncated Hilbert space specified by the reference determinant(s) and the value of $h$.
+{\QP} \cite{Garniron_2019} was also employed to perform all the present eCI, sCI and FCI calculations.
 In a given calculation, the energies are considered to be converged when the (largest) EN2 correction computed in the truncated Hilbert space 
 lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
-%by spanning the most important regions of the truncated Hilbert space
 This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space,
-while delievering fairly converged absolute and excitation energies.
+while delivering fairly converged energies.
 The ground- and excited-state CI energies are obtained with the Davidson iterative algorithm. \cite{Davidson_1975}
-For a given approximate CI model, we futher evaluated the standard and the renormalized EN2 perturbative correction. \cite{Garniron_2019} 
+For a given approximate CI model, we further evaluated the standard and the renormalized EN2 perturbative correction. \cite{Garniron_2019} 
 This calculation involves a single loop over the determinants left outside the truncated (internal) space but connected to it via at most double excitations.
 Looping over these external doubly excited determinants has a computational scaling equal to $\Ndet$,
-thus $\order*{N^{2e+4}}$ for excitation-based CI and $\order*{N^{2h+4}}$ for hCI, where $e$ or $h$ define the internal CI space.
+thus $\order*{N^{2e+4}}$ for eCI and $\order*{N^{2h+4}}$ for hCI, where $e$ or $h$ define the internal CI space.
 For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^8}$,
@@
 
-To gauge the performance of hCI against excitation-based and seniority-based CI for radical systems,
-we have calculated the ground-state potential energy curves (PECs) for the dissociation of four radical systems:
+To gauge the performance of hCI, eCI and sCI for radical systems,
+we have calculated the ground-state potential energy curves (PECs) for the dissociation of four radicals:
 \ce{OH}, \ce{CN}, vinyl (\ce{C2H3}), and \ce{H7}.
-These calculations employed the ground-state restricted open-shell HF orbitals and the frozen core approximation.
-The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020},
-and is also reproduced in the {\SupInf}.
+The CI calculations employed the ground-state restricted open-shell HF orbitals, described with the cc-pVDZ basis set, and within the frozen core approximation.
+For such small systems and basis set, FCI is attainable and provides the reference results for gauging the approximate CI models.
+The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020}, and is also reproduced in the {\SupInf}.
 Their PECs were computed along the \ce{C=C} double bond breaking coordinate, with the remaining internal coordinates kept frozen.
@@
 
-The results were analyzed along the same lines of our report on hCI for closed-shell systems. \cite{Kossoski_2022}
+The results were analyzed along the same lines of our previous report on hCI for closed-shell systems. \cite{Kossoski_2022}
 Namely, for the different CI models considered here, 
 we evaluated the convergence of the non-parallelity error (NPE), the distance error, the harmonic vibrational frequencies, and the equilibrium bond lengths, as functions of $\Ndet$.
-The NPE of a given level of theory is defined as the maximum minus the minimum differences between its corresponding PEC and the FCI one, for a given range of coordinates.
-Here we redefine the previous definition of the distance error \cite{Kossoski_2022} to accommodate for the fact the approximate PEC might appear below the FCI when perturbative corrections are employed.
-The previous definition could provide negative values in this case, whereas this metric should provide the overall extent of proximity between two PECs.
-Here, we define the distance error based on the signed differences between two PECs, as the absolute value of their maximum difference plus the absolute value of their minimum difference, 
+The NPE of a given level of theory is defined as the maximum minus the minimum differences between its corresponding PEC and the FCI PEC, for a given range of coordinates.
+Here, we redefine the previous definition of the distance error \cite{Kossoski_2022} to accommodate for the fact the approximate PEC might appear below the FCI one when perturbative corrections are employed.
+In such cases and with the previous definition, undesired negative values could be attained.
+The distance error is redefined based on the signed differences between two PECs, as the absolute value of their maximum difference plus the absolute value of their minimum difference, 
 evaluated at given coordinate interval.
-In the discussion, we say equilibrium properties when referring to both the equilibrium geometry and the vibrational frequency.
-Details about how the harmonic vibrational frequencies and equilibrium geometries were obtained from the calculated PECs,
+This new definition measures how close are two PECs, remaining always non-negative.
+From here on, we employ equilibrium properties when referring to both the equilibrium geometry and the harmonic vibrational frequency.
+Details about how the equilibrium properties were obtained from the calculated PECs,
 along with the ranges defining the NPE and distance errors can be found in the \SupInf.
@@
 Core orbitals were frozen.
-For the radicals, we further impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies accounting for a set of appropriate spin-flipped determinants.
-In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS, \cite{Maurice_1996} and similarly for the other hCI and excitation-based CI models.
+For the radicals, we further impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies in accounting for a set of appropriate spin-flipped determinants.
+In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS, \cite{Maurice_1996} and equivalently for the other hCI, eCI and sCI models.
 We performed calculations following both the standard ground-state based CI route and the state-specific CI route. \cite{Kossoski_2023}
 For the latter, we employed the state-specific orbitals obtained in Ref.~\onlinecite{Kossoski_2023}.
-Notice that, in contrast to excitation-based CI, the energies obtained with sCI and hCI models are not invariant under rotations within the occupied and virtual subspaces.
-The computed excitation energies were compared against the reference values provided in the QUEST database. \cite{Veril_2021}
-%For completeness, we also reproduce the geometries in the \SupInf.
+Notice that, in contrast to eCI, the energies obtained with sCI and hCI models are not invariant under rotations within the occupied and virtual subspaces.
+The computed excitation energies were benchmarked against the reference values of FCI quality provided in the QUEST database. \cite{Veril_2021}
 
-The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as reference determinant for the hCI ground state calculations.
-For the state-specific hCI calculations, we employed a minimal CSF reference, \cite{Kossoski_2023}
-meaning, for the states considered here, a single open-shell determinant for the doublet excited states and a single CSF for the excited states from closed-shells.
-As for the seniority-based CI models, 
+The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as the reference determinant for the hCI and eCI ground state calculations 
+(left and middle pannels of Fig.~\ref{fig:hCI})
+For the state-specific hCI and eCI calculations, we employed a minimal CSF reference: \cite{Kossoski_2023}
+a single open-shell determinant for the doublet excited states and a single open-shell CSF for the excited states from closed-shell systems
+(middle and right pannels of Fig.~\ref{fig:hCI}, respectively).
+
+As for the sCI models, 
 we considered sCI1 for both ground and (state-specific) excited state calculations for the open-shell systems (odd number of electrons).
 For the closed-shell systems (even number of electrons), we explored two models.
-In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal seniority-based CI model for computing singly-excited states of closed-shell systems.
+In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal sCI calculation for computing singly-excited states of closed-shell systems.
 Further including the $s=2$ sector for the ground state calculations gives rise to the sCI2/sCI2 model.
-CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it imply a ground-state based approach.
-% comments on the different references
+From here on, CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it correspond to a ground-state based approach.
 
@@
 
-% vinyl eq.geom e4 s5 h3.5 rpt2
-% vinyl 5.0 h3+rpt2
-% update pecs of ethylene
-
 The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf}.
@@
 These previous results are reproduced in Fig.~\ref{fig:plot_all_closed} together with the present results, which include such correction.
-First we will focus on the results for the bare CI models (hCI, excitation-based CI and sCI) for the open-shell systems (represented by the dark tones in the figures),
-leaving the discussion about the perturbative correction (light tones) on both open-shell and closed-shell systems for later.
+First we will discuss the results for the bare CI models (hCI, eCI and sCI) for the open-shell systems (represented by the dark tones in the figures),
+leaving the discussion about the perturbative correction (light tones) for both open-shell and closed-shell systems to the next subsection.
 
@@
 \caption{
-Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{OH}, \ce{CN}, \ce{H7}, and vinyl,
-as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
 with (light tones) and without (dark tones) the standard EN2 perturbative correction.
@@
 Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, 
-as a function of the number of determinants ($\Ndet$), according to hCI (green), excitation-based CI (red) and sCI (blue) models,
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
 with (light tones) and without (dark tones) the standard EN2 perturbative correction.
@@
 
-We found that hCI typically improves the convergence with respect to $\Ndet$ when compared to excitation-based CI, for the four properties surveyed here.
+We found that hCI typically improves the convergence with respect to $\Ndet$ when compared to eCI, for the four properties surveyed here.
 The present finding for open-shell systems therefore parallels the previously reported superiority of hCI for closed-shell systems. \cite{Kossoski_2022}
-The smaller NPEs obtained with hCI are linked to the better description of the PECs at dissociation.
+
+The smaller NPEs obtained with hCI are related to the better description of the PECs at dissociation.
 This is due to a larger fraction of static correlation being recovered,
 stemming from the classes of determinants appearing lower in the seniority-excitation map (see Fig.~\ref{fig:hCI}),
-namely, determinants high in excitation degree $e$ and low in seniority $s$, absent in excitation-based CI of the same order.
-
-Still concerning the NPEs, the advantage of hCI is pronounced for \ce{OH} (involving a single bond breaking), 
-becoming less so, though still clear, for vinyl (double bond breaking) and \ce{CN} (triple bond breaking),
-whereas for \ce{H7} (multiple bond breaking), it still displays a small advantage in comparison to excitation-based CI.
+namely, determinants high in excitation degree $e$ and low in seniority $s$, absent in eCI of the same order.
+The smaller NPEs attained with hCI are clear for \ce{OH} (involving a single bond breaking), 
+vinyl (double bond breaking) and \ce{CN} (triple bond breaking),
+whereas for \ce{H7} (multiple bond breaking), hCI and eCI are comparable.
 This dependence on the order of bond breaking had also been observed for closed-shell systems \cite{Kossoski_2022} (as can be seen in Fig.~\ref{fig:plot_all_closed}),
 and would be expected, as describing dissociation becomes increasingly more challenging as the bond order increases.
-%This trend can be explained by the growing amount of static correlation as more bonds are broken.
-Even though hCI manages to recover more static correlation than excitation-based CI,
-the result is more striking for single bond breaking, whereas multiple bond breaking (like \ce{H7} and \ce{H8}) inevitably requires higher order excited determinants.
+Even though hCI manages to recover more static correlation than eCI,
+the advantage of the former is more striking for single bond breaking, whereas multiple bond breaking (like \ce{H7} and \ce{H8}) inevitably requires higher order excited determinants.
 
-We also found overall smaller distance errors with hCI than with excitation-based CI, although the improvement is not as remarkable as observed for the NPEs.
-hCI further leads to a somewhat faster convergence of the equilibrium geometries for the open-shell systems, most noticeably for \ce{CN}.
-Finally, hCI systematically outperforms excitation-based CI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
-Given the overall superiority of hCI for the equilibrium properties, it also manages to account for more dynamic correlation than excitation-based CI,
+The distance errors obtained with hCI and eCI are overall comparable for open-shells, whereas the former presents slightly better convergence for the closed-shell systems.
+In comparison to eCI, hCI leads to either comparable or somewhat faster (for \ce{CN}) convergence of the equilibrium geometries for the open-shell systems.
+Finally, hCI systematically outperforms eCI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+Given the overall slight superiority of hCI for the equilibrium properties, it also manages to account for more dynamic correlation than eCI,
 for both open-shell and closed-shell systems.
 
-In great contrast to hCI (and excitation-based CI),
-sCI models deliver a poor convergence of all observables.
-The single exception is the NPE for \ce{OH} obtained from sCI0, smaller than the NPEs obtained from CISD and hCI1.5 or hCI2, methods having a comparable computational cost in this particular case.
-Going to larger basis sets or bigger systems, the computational burden of sCI models would increase considerably more than hCI or excitation-based CI models, 
+In great contrast to hCI and eCI, sCI models deliver a poor convergence of all observables.
+The single exception is the NPE for \ce{OH} obtained from sCI0, in-between the NPEs obtained from hCI2 and CISDT, methods having a comparable computational cost in this particular case.
+Going to larger basis sets or bigger systems, the computational burden of sCI models would increase considerably more than hCI or eCI models, 
 due to its formal exponential scaling.
-In their own, we conclude that the sCI route is unatractive to open-shell systems,
-in line with similar findings for closed-shell systems \cite{Kossoski_2022} (also clear from Fig.~\ref{fig:plot_all_closed}).
+As far as configuration interaction methods are concerned, we conclude that the sCI route is unattractive to open-shell systems,
+in line with similar findings for closed-shell systems \cite{Kossoski_2022} (also see Fig.~\ref{fig:plot_all_closed}).
 
-%orbital optimization?
-
-% EN2 correction
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{hCI plus EN2 correction}
+\label{sec:res_B}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 The results for the CI models corrected by the standard EN2 energy are represented by the light tones in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
-%The corresponding PECSs are shown in the {\SupInf}.
-We excluded cases where there are important discontinuities in the PECs.
-These discontinuities reflect crossings between states of different symmetries in the unperturbed calculation,
+We excluded instances where the PECs present important discontinuities.
+These discontinuities reflect crossings between states of different symmetries in the unperturbed CI calculation,
 appearing as kinks in the computed ground-state PEC.
 Due to the abrupt change of character, the EN2 correction is not uniform and gives rise to the observed discontinuities.
-This happens for the open-shell systems at dissociation (except for \ce{OH}), for which we do not present NPE and distance errors.
+This happens for the open-shell systems at dissociation (except for \ce{OH}), for which we do not present NPEs and distance errors.
 For \ce{OH} and the six closed-shell systems, the EN2 correction produced smooth PECs at dissociation.
@@
 (thus no equilibrium properties are presented for this radical).
-This is simply due to its several close-lying excited states. \cite{}
-If we were to follow the PEC of a given symmetry (and not the lowest-lying state as we did here), there would be no discontinuity stemming from the EN2 contribution.
+This is simply due to its several close-lying excited states.
+If we were to follow the PEC of a given symmetry (and not the lowest-lying state as we did here), there would be no discontinuities stemming from the EN2 contribution.
 
@@
 For \ce{OH}, in particular, the improvement is massive, as even the lower levels of CI+EN2 deliver very close results to FCI.
-Importantly, the hCI+EN2 route outperforms its excitation-based counterpart, thus preserving the advantage observed without the perturbative correction.
+Importantly, the hCI+EN2 route outperforms its eCI+EN2 counterpart, thus preserving the advantage observed without the perturbative correction.
 Similarly, the correction typically maintains the comparable performance of the two routes when this is the case according to the unperturbed CI calculations.
 
-%xe
-%\ce{OH} massive improvement
-%\ce{CN} \ce{H4} \ce{N2} \ce{F2} \ce{HF} big improvement
-%\ce{H7} \ce{H8} vinyl ethylene some improvement
-
-%freq
-%\ce{OH} massive improvement
-%\ce{CN} \ce{F2} \ce{N2} \ce{H4} \ce{H7} big improvement
-%\ce{HF} \ce{H8} vinyl ethylene some improvement
-
-%Including the EN2 correction thus accelerates the rate of convergence as a function of $\Ndet$, when compared to the unperturbed calculations.
-%In fact, 
 The improvement brought by the perturbative correction is such that a given CI+EN2 model often provides more accurate results than the higher-order bare CI model sharing the same computational scaling.
@@
 It is worth recalling that the prefactor associated with the EN2 calculation is smaller than that associated with the higher-order CI calculation,
-which in practice makes hCI1+EN2 cheaper than hCI2.
-In the above examples, the accuracy of hCI1+EN2 is actually better or comparable to that of hCI2.5 or CISDT, considerably more expensive models.
-%\ce{HF}, \ce{F2}, \ce{H4}, \ce{H7}, and \ce{H8}.
+which, in the above example, makes hCI1+EN2 cheaper than hCI2.
+In many cases, the accuracy of hCI1+EN2 is actually better or comparable to that of hCI2.5 or CISDT, considerably more expensive models.
 
-A similar impact on the effect of the EN2 correction also holds for excitation-based CI.
+A similar impact on the effect of the EN2 correction also holds for eCI.
 For that, we compare CISD+EN2 and CISDT, both sharing a $\order*{N^8}$ scaling.
@@
 
-%\ce{OH}, \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4} (small)
-
-The EN2 correction usually ameliorates the performance of the sCI route, though to a lesser extent than observed for hCI and excitation-based CI.
+The EN2 correction usually ameliorates the performance of the sCI route, though to a lesser extent than observed for hCI and eCI.
 This reflects the poorer reference provided by sCI, from which perturbation theory struggles to recover from.
@@
 and equilibrium properties (vinyl, \ce{HF}, \ce{F2}, and ethylene).
-Overall, the sCI+EN2 route falls behind its alternatives.
+Overall, the sCI+EN2 route is too expensive for the attained accuracy.
 
-In few cases, however, it does not improve the bare CI results, suggesting important cancellation of errors in the latter.
-This is seen when employing the hCI1 and hCI1.5 models and their EN2 corrected couterparts to compute the NPE of \ce{HF}, \ce{F2} and \ce{N2}, and to some extent the equilibrium properties of \ce{F2}.
+In very few cases, the EN2 correction does not improve the bare CI results, suggesting important cancellation of errors in the latter.
+This is seen when employing the hCI1 and hCI1+EN2 models to compute the NPE of \ce{OH} and \ce{F2}, and to some extent the equilibrium properties of \ce{F2}.
 These are exceptions though, as the EN2 correction practically always improve the accuracy of the computed observables.
@@
 one can compute the renormalized EN2 correction. \cite{Garniron_2019}
-Analogous figures showing the convergence of observables with the usual and with the renormalized forms are shown in the {\SupInf}.
-There are often significant differences between the two forms at the lower orders of CI, which in turn become negligeable at higher orders.
+Analogous results showing the convergence of observables for both usual and renormalized EN2 corrections are shown in the {\SupInf}.
+Significant differences can be encountered at the lower orders of CI, becoming negligible at higher orders.
 More often than not, the usual correction performs better.
-The difference is expressive for ethylene, \ce{N2}, \ce{H7}, and \ce{H8}, and to a lesser extent for vinyl.
-For \ce{H4} the results are mixed.
+The difference is noticeable for ethylene, \ce{N2}, \ce{H7}, and \ce{H8}, and to a lesser extent for vinyl,
+whereas for \ce{H4} the results are mixed.
 The renormalized correction is slightly more accurate for \ce{OH}, \ce{HF}, and \ce{F2}.
@@
 whereas it worsens the results for multiple bond breaking.
-Overall, the usual EN2 correction should be prefered when employed in combination with the approximate CI models considered here.
+Overall, the usual EN2 correction should probably be favored when employed in combination with the approximate CI models.
 
@@
 \subsection{hCI for excited states}
-\label{sec:res_B}
+\label{sec:res_C}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% intro
-
 For each CI model considered here, we evaluated the mean signed error (MSE), mean absolute error (MAE), and root-mean square error (RMSE) 
-with respect to reference theoretical values for the excitation energies.
-The statistical errors are presented in Tab.~\ref{tab:stat1} and Tab.~\ref{tab:stat2},
-for singly excited states of closed-shell and open-shell systems, respectively.
+with respect to the reference theoretical values for the excitation energies.
 For the lower order models, the calculations were performed for 48 (closed-shell) and 19 (open-shell) excited states.
-In turn, a subset of 16 (closed-shell) and 6 (open-shell) excited states was considered for the higher order models, given their more intensive computational cost.
-Even though these subsets are too small for meaningful absolute statistics, they can still show the main trends on the higher order CI models.
-%A detailed comparison of excitation-based CI along the ground-state and the state-specific routes can be found elsewhere. \cite{Kossoski_2023}
-%The focus of the present discussion lies on the comparison between hCI, sCI, and excitation-based CI, for both routes,
-
-% ground-state based hCI
+In turn, subsets of 16 (closed-shell) and 6 (open-shell) excited states were considered for the higher order CI models, given the more intensive computational cost.
+Even though these subsets are too small for meaningful absolute statistics, they should be enough to reveal main trends.
+A detailed comparison of eCI based on ground-state and state-specific approaches can be found elsewhere. \cite{Kossoski_2023}
+The focus of the present discussion lies on the comparison between hCI and eCI, and sCI routes, 
+and on a similar comparison between ground-state and state-specific approaches for the hCI and sCI routes.
 
 We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Tab.~\ref{tab:stat1}.
-%
 The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.07}{\eV}.
 This model shares the same computational scaling with CIS, but includes the paired double excitations, 
-which clearly do not improve the descent CIS results (MAE of \SI{0.61}{\eV}).
-Moving one rank up, to hCI1.5, the MAE increases to \SI{1.91}{\eV}, and reaches a maxium of \SI{3.53}{\eV} at the hCI2 level.
-Here we notice that hCI2 delivers somewhat smaller errors than CISD (MAE of \SI{4.09}{\eV}), even though both are way too large.
+which clearly worsen the decent CIS results (MAE of \SI{0.61}{\eV}).
+Moving one rank up, to hCI1.5, the MAE increases to \SI{1.91}{\eV}, and then reaches a maximum of \SI{3.53}{\eV} at the hCI2 level.
+We notice that hCI2 delivers somewhat smaller errors than CISD (MAE of \SI{4.09}{\eV}), even though both are way too large.
 From that point on, the next hCI models generate progressively better results.
-Despite the improvement with respect to hCI2, hCI2.5 is stil as inaccurate (MAE of \SI{1.95}{\eV}) as hCI1.5.
+Despite the improvement with respect to hCI2, hCI2.5 is still as inaccurate (MAE of \SI{1.95}{\eV}) as hCI1.5.
 At the hCI3 level, significantly smaller errors are finally achieved (MAE of \SI{0.19}{\eV}).
-This is quite close to that obtained with the excitation-based model of same order, CISDT (\SI{0.17}{\eV}).
+This is close to that obtained with the eCI model of same order, CISDT (\SI{0.17}{\eV}).
 Even smaller errors are produced at the hCI3.5 level (MAE of \SI{0.11}{\eV}), but at a considerable computational cost.
 
-% first origin of the unbalance hCI
-
-%There are two key elements that explain the main trends on the observed statistical errors. The first concerns the set of molecular orbitals.
-The errors of the hCI models first increase as one augments the hierarchy parameter $h$, reaching a maximum at hCI2, and then decrease towards higher orders.
-This behavior paralles what is found with excitation-based CI, where CISD is far less accurate than CIS, which in turn is less accurate than CISDT and CISDTQ.
-This can be understood based on the role of the excited determinants for ground and excited states.
-The excited determinants of the low-order hCI models (hCI1 and hCI1.5) already account for some correlation of the ground state,
-but represent mostly orbital relaxation of the excited state, which thus remains less correlated.
-This favored description of the ground state explains the overestimated excitation energies.
-The effect becomes exagerated at the hCI2 (and CISD) models, since it captures most of the ground state correlation through the unpaired double excitations,
-but still little correlation for the excited state.
-Higher orders models are needed to recover most of the excited state correlation, and at this point the errors on the excitation energies start to decrease.
-%(though always overestimating the reference values).
-The same argument also explains why hCI systematically overestimates the excitation energies, just as for excitation-based CI.
-
-% state-specific hCI and second origin of the unbalance
-
-The trend discussed above stems from the use of ground-state HF orbitals.
-Employing state-specific orbitals would be expected to suppress the bias toward the ground state and lead to improved results, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
-However, for the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, though by underestimating the excitation energies.
-Besides the set of orbitals, the types of determinants included at each order play an equally important role and explains our observation.
-For a given hierarchy parameter $h$, different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF).
-The case of $\Delta$hCI1 can be understood by comparing the hCI1 determinants for these two references, shown in Fig.~\ref{fig:determinants}.
-%For the ground state calculation, it accounts for two classes of excited determinants, generated by single excitation and by paired double excitation from the closed-shell reference.
-%For the singly-excited state, four types of determinants are generated. On top of the single and paired double excitations from the doubly-occupied orbitals, 
-There is clearly more diversity in the classes of determinants employed for the excited-state calculation than for the ground-state one, reflecting the different references employed.
-Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the excited state than for ground state.
-%and are thus not suitable for calculations of excitation energies for closed-shell systems.
-Although hCI is constructed to account for all classes of determinants whose number share the same computational scaling,
-at the lower orders this procedure does not lead to a balanced description of correlation, at least for excitations of closed-shell systems.
-As will be discussed later for the excitations of open-shell systems, this inbalance probably arises from the distinct reference determinants employed for the closed-shell excitations.
-%
-Another important point to realize is that a state-specific approach is not necessarily more accurate than a ground-state based approach.
-Both the orbitals and the classes of determinants of a given CI model control the fine balance of ground- and excited-state correlation effects.
-
-%Ramping up to $\Delta$hCI2, this inbalance is reduced to a great extent.
-%even though the excitation energies are still underestimated.
-%Even though this unbalance is still observed for $\Delta$hCI2, it is much less serious than observed for the lower-order state-specific hCI methods.
-
-% state-specific vs. ground-state hCI
-
-The state-specific advantage only shows up at the $\Delta$hCI2 level, which is substantially more accurate than hCI2 (MAEs of \SI{0.20}{\eV} and \SI{3.53}{\eV}).
-This improvement parallels our previous finding on the excitation-based CI models of the same order, CISD and $\Delta$CISD. \cite{Kossoski_2023}
-Even though hCI2 is somewhat more accurate than CISD (both have large errors), their state-specific versions share a comparable performance.
-Actually, $\Delta$hCI2 is slightly less accurate than $\Delta$CISD (MAEs of \SI{0.20}{\eV} and \SI{0.17}{\eV}).
-Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.19}{\eV} against \SI{-0.13}{\eV}).
-Albeit small, these differences are statistically significant, and $\Delta$CISD would still be preferable to $\Delta$hCI2 by some margin.
-
-%This means that the additional determinants present in the former model recover a larger fraction of the correlation energy for excited state than for the ground state, thus creating a bias toward underestimated excitation energies.
-Ramping up to $\Delta$hCI2.5 increases the errors (MAE of \SI{0.27}{\eV}),
-which probably reflects another set of unbalanced state-specific determinants that enters at this stage.
-(Despite the differing number of states considered, the observed variation on the accuracy of $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant.)
-Even though $\Delta$hCI2.5 is much more accurate than hCI2.5, $\Delta$hCI2 or $\Delta$CISD remain cheaper and more accurate.
-
-The situation improves at $\Delta$hCI3 (MAE of \SI{0.21}{\eV}),
-while $\Delta$hCI3.5 generates considerably more accurate results (MAE of \SI{0.08}{\eV}).
-There is no gain in going from hCI3 to $\Delta$hCI3 and from hCI3.5 to $\Delta$hCI3.5,
-similarly to what had been found for CISDT and $\Delta$CISDT. \cite{Kossoski_2023}
-%which also parallels the comparison between CISDT and $\Delta$CISDT. \cite{Kossoski_2023}
-We notice, however, that the state-specific route presents negative MSEs, in contrast to the positives values obtained with the ground-state based route.
-Furthermore, excitation-based CI and hCI perform similarly at this order (the former perhaps slightly better),
-as also observed for $\Delta$CISD and $\Delta$hCI2.
-
-%As discussed above, adopting the state-specific approach at lower-order hCI models ($\Delta$hCI1 and $\Delta$hCI1.5)
-%actually worsens the results with respect to the ground-state approach (hCI1 and hCI1.5).
-
-%we further investigated whether higher-order versions could be competitive against the traditional excitation-based counterparts.
-%For that, we performed calculations for a subset of 16 singly-excited states of small closed-shell molecular systems (indicated in the {\SupInf}).
-%(\ce{H2O}, \ce{BF}, \ce{HCF}, \ce{HCl}, and \ce{N2}),
-%with corresponding statistical errors shown in Tab.~\ref{tab:stat1}.
-%Even though this is a small set, it should be enough to obtain the main trends on the accuracy of higher-order models.
-
-%We found that the errors of both ground-state based hCI and state-specific $\Delta$hCI to systematically decrease when going toward higher orders of $h$, which would be expected.
-%Unfortunately, the pace of decrease is too slow to render these approaches very useful.
-%In line with what has been found for excitation-based CI \cite{Kossoski_2023},
+% HCF cisdtq ground
+% sil cisdtq sA2 and sB2
 
@@
 
+As discussed above, the errors on excitation energies obtained with the hCI models first increase as one augments the hierarchy parameter $h$, reaching a maximum at hCI2, and then decrease towards higher orders.
+This behavior parallels what is well stablished for the excitation energies computed with eCI, where CISD is far worse than CIS, which in turn is inferior to CISDT, whereas CISDTQ is considerably more accurate.
+This can be understood based on the role of the excited determinants for ground and excited states.
+The excited determinants of the low-order hCI models (hCI1 and hCI1.5) already account for some correlation for the ground state,
+but represent mostly orbital relaxation of the excited state, which thus remains less correlated.
+This favored description of the ground state explains the overestimated excitation energies at these orders.
+The effect becomes exaggerated at the hCI2 (and CISD) models, since it captures most of the ground state correlation through the unpaired double excitations,
+which in turn just start to describe correlation for the excited state.
+Higher orders models, starting at hCI3 and CISDT, are needed to recover a large fraction of the excited state correlation, and at this point the errors on the excitation energies decline progressively.
+Because the excited determinants are accessed from the ground-state determinant and due to ground-state HF orbitals, the description of the ground state is always remain favoured.
+This explains why hCI systematically overestimates the excitation energies, just as eCI does.
+
+Employing state-specific orbitals would be expected to suppress the bias toward the ground state and lead to improved results, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
+However, for the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, this time by underestimating the excitation energies.
+Besides the set of orbitals, the classes of determinants included at each order play an equally important role and explains our observation.
+Still considering the excitations of the closed-shell systems,
+different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF),
+for a given hierarchy parameter $h$.
+The case of $\Delta$hCI1 can be understood by comparing the hCI1 determinants for these two references, shown in Fig.~\ref{fig:determinants}.
+There is far more diversity in the classes of determinants employed for the excited-state calculation than for the ground-state one, a consequence of the different references.
+Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the excited state than for ground state.
+This is further supported by the more accurate results obtained for excitations of open-shell systems (discussed in detail later), where the same type of reference is employed.
+Although hCI is constructed to account for all classes of determinants whose number share the same computational scaling,
+at the lower orders this procedure does not lead to a balanced description of correlation, at least for excitations of closed-shell systems and the present choice of minimal references.
+An important point to realize from the present results is that a state-specific approach is not necessarily more accurate than a ground-state based approach.
+Both the orbitals and the classes of determinants of a given CI model control the fine balance of ground- and excited-state correlation effects.
+
+At the $\Delta$hCI2 level, the unbalance associated with the state-specific classes of determinants is reduced to a great extent.
+At this level, the state-specific advantage clearly manifests, with $\Delta$hCI2 having a MAE of \SI{0.20}{\eV}, substantially smaller than the MAE of \SI{3.53}{\eV} obtained from hCI2.
+This comparison between ground-state based and state-specific approaches is analogous to our previous finding on the eCI models of the same order, CISD and $\Delta$CISD. \cite{Kossoski_2023}
+Even though hCI2 is somewhat more accurate than CISD (both have large errors), their state-specific versions share a comparable performance.
+Actually, $\Delta$hCI2 is slightly less accurate than $\Delta$CISD, with MAEs of \SI{0.20}{\eV} and \SI{0.17}{\eV}.
+Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.19}{\eV} against \SI{-0.13}{\eV}).
+Albeit small, these differences are statistically significant, and $\Delta$CISD would still be preferable to $\Delta$hCI2 by some margin.
+The slightly worse performance of $\Delta$hCI2 probably stems from a residual unbalance associated with the state-specific classes of determinants.
+
+Moving to $\Delta$hCI2.5 increases the errors (MAE of \SI{0.27}{\eV}) when compared to $\Delta$hCI2 (MAE of \SI{0.20}{\eV}).
+This could be due to another set of unbalanced state-specific determinants that enters at this stage, likewise to what was observed between $\Delta$hCI1 and $\Delta$hCI1.5.
+Despite the differing number of states considered, the observed variation on the accuracy of $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant,
+which is confirmed by calculating the statistical errors for the same subset of excited states.
+Even though $\Delta$hCI2.5 is much more accurate than hCI2.5, $\Delta$hCI2 or $\Delta$CISD remain cheaper and more accurate.
+
+The situation improves at $\Delta$hCI3 (MAE of \SI{0.22}{\eV}), though it still remains less accurate than $\Delta$hCI2,
+only surpassed at the $\Delta$hCI3.5 level (MAE of \SI{0.08}{\eV}).
+There is no gain in going from hCI3 (MAE of \SI{0.19}{\eV}) to $\Delta$hCI3 and from hCI3.5 (MAE of \SI{0.11}{\eV}) to $\Delta$hCI3.5,
+likewise to what had been found for CISDT and $\Delta$CISDT. \cite{Kossoski_2023}
+We notice, however, that the state-specific route presents negative MSEs, in contrast to the positives values obtained with the ground-state based route, for both hCI and eCI.
+Furthermore, eCI and hCI present comparable performances at this order, with an arguable preference of the former,
+as also discussed above for $\Delta$CISD and $\Delta$hCI2.
+
 We now shift to the discussion of excitations for open-shell systems.
 The statistical errors are shown in Tab.~\ref{tab:stat2}.
-The key difference on the calculations for open-shell excitations is that the same type of reference determinant were employed for ground- and excited-state calculations,
-whereas the closed-shell excitations relied on different types of reference determinants.
-This accounts for the more accurate results for open-shell excitations and explains most of the differences with respect to excitations from closed-shell systems, outlined in detail below.
-
-%We first point out the similarities and later the key differences with respect to the closed-shell discussion.
-
-When using ground-state orbitals, we found the performance initially degrates and later improves as the order increases, for both excitation-based CI and hCI.
-This is similar to what was observed for the closed-shell excitations, 
-and can be traced back to a bias towards the ground state due to the choice of orbitals.
-%more ground-state correlation being recovered at the lower orders since one starts from ground-state orbitals.
-An important difference, however, is that the maximum error appears at a lower order (hCI1.5) and is significantly smaller (MAE of \SI{1.11}{\eV}) for open-shells
-than for closed-shells (at hCI2 with a MAE of \SI{3.53}{\eV}).
-This reflects the choice of reference, since both ground and excited states of the open shells considered here are qualitatively described with the same type of determinant.
-%This reflects the fact that ground and excited states of the open shells considered here are qualitatively described with the same type of determinant.
-For the same reason, the MSEs are systematically smaller for the open-shell excitations.
-
-Once state-specific orbitals are employed, there is no longer a bias towards the ground state related to the orbitals.
-In a big contrast to the closed-shell case, here the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.34}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.46}{\eV}), 
-become more accurate than their ground-state based counterparts, hCI1 (MAE of \SI{0.77}{\eV}) and hCI1.5 (MAE of \SI{1.11}{\eV}).
-Again, thanks to the same type of reference in ground- and excited-state calculations.
-The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:determinants}, in comparison to the more unbalanced case of excitation from closed-shells.
-We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5).
-%and that $\Delta$hCI1 is slightly more accurate than $\Delta$hCI1.5.
-
-This residual bias disappears at the next order, $\Delta$hCI2, whereas for the closed-shell systems even $\Delta$hCI3.5 still understimated excitation energies. 
-The same holds for excitation-based CI.
-For open-shells, $\Delta$hCI2 is arguably slightly more accurate (MAE of \SI{0.11}{\eV}) than $\Delta$CISD (MAE of \SI{0.12}{\eV}), also based on the larger differences on the RMSEs.
-In further contrast to the closed-shell case,
-$\Delta$hCI2.5 is more accurate than $\Delta$hCI2, producing a MAE of \SI{0.06}{\eV}.
-(Accounting only for the 6 states considered for $\Delta$hCI2.5, $\Delta$hCI2 would have a MAE of \SI{0.09}{\eV}.)
-The statistics at the next orders ($h=3$ and $h=3.5$) is limited, 
-but suggests a possible advantage of the hCI models against the excitation-based ones,
-and that the ground-state routes perform slightly better than the state-specific ones.
+The key difference on the calculations for open-shell excitations is that the same type of reference was employed for ground- and excited-state calculations, namely,
+a single open-shell determinant, shown in the center panel of Fig.~\ref{fig:hCI}.
+In contrast, the excited states from the closed-shell systems relied on different classes of reference determinants.
+This accounts for the more accurate results observed for open-shell excitations and explains most of the differences with respect to excitations from closed-shell systems, as discussed in detail in the following.
 
@@
 CIS               & 19 & +0.33 & 0.36 & 0.50 \\
-CISD              & 6  & +1.98 & 1.98 & 2.16 \\
-CISDT             & 6  & +0.02 & 0.03 & 0.05 \\
-CISDTQ            & 6  & +0.03 & 0.03 & 0.05 \\
+CISD              & 6  & +2.97 & 2.97 & 3.19 \\
+CISDT             & 6  & +0.06 & 0.07 & 0.09 \\
+CISDTQ            & 6  & +0.08 & 0.08 & 0.11 \\
 %
@@
 hCI1.5            & 19 & +1.11 & 1.11 & 1.30 \\
-hCI2              & 6  & +0.69 & 0.69 & 0.85 \\
-hCI2.5            & 6  & +0.01 & 0.03 & 0.04 \\
-hCI3              & 6  & -0.00 & 0.02 & 0.03 \\
-hCI3.5            & 6  & -0.00 & 0.01 & 0.02 \\
-%hCI4              & 6  &       &      &      \\
+hCI2              & 6  & +1.32 & 1.32 & 1.60 \\
+hCI2.5            & 6  & +0.03 & 0.05 & 0.06 \\
+hCI3              & 6  & +0.00 & 0.04 & 0.05 \\
+hCI3.5            & 6  & +0.01 & 0.03 & 0.05 \\
 %
@@
 $\Delta$CISD      & 19 & +0.01 & 0.12 & 0.21 \\
-$\Delta$CISDT     & 6  & -0.05 & 0.05 & 0.09 \\
+$\Delta$CISDT     & 6  & -0.07 & 0.07 & 0.10 \\
 $\Delta$CISDTQ    & 6  & -0.02 & 0.02 & 0.02 \\
@@
 $\Delta$hCI2      & 19 & -0.01 & 0.11 & 0.16 \\
-$\Delta$hCI2.5    & 6  & -0.06 & 0.06 & 0.12 \\
-$\Delta$hCI3      & 6  & -0.03 & 0.03 & 0.06 \\
-$\Delta$hCI3.5    & 6  & -0.02 & 0.02 & 0.02 \\
-%$\Delta$hCI4      & 6  &       &      &      \\
+$\Delta$hCI2.5    & 6  & -0.09 & 0.09 & 0.14 \\
+$\Delta$hCI3      & 6  & -0.06 & 0.06 & 0.08 \\
+$\Delta$hCI3.5    & 6  & -0.02 & 0.02 & 0.03 \\
 %
@@
 
+When using ground-state orbitals, we found the performance initially degrades and later improves as the order increases, for both eCI and hCI.
+This is similar to what was observed for the closed-shell excitations, 
+and can be traced back to a biased description of the ground state related to the choice of ground-state orbitals.
+An important difference, however, is that the maximum error, also at hCI2, is significantly smaller (MAE of \SI{1.32}{\eV}) for open-shells
+than for closed-shells (MAE of \SI{3.53}{\eV}).
+This clearly reflects the choice of reference, as both ground and excited states of the open shells considered here are qualitatively described with the same type of determinant.
+For the same reason, the MSEs are systematically smaller for the open-shell excitations (though still positive because of the ground-state orbitals).
+
+Once state-specific orbitals are employed, there is no remaining bias towards the ground state.
+In a stark contrast to the closed-shell case, the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.34}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.46}{\eV}), 
+become more accurate than their ground-state based counterparts, hCI1 (MAE of \SI{0.77}{\eV}) and hCI1.5 (MAE of \SI{1.11}{\eV}).
+Once again, this is thanks to the same type of reference in ground- and excited-state calculations.
+The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:determinants}, in comparison to the more unbalanced case of excitation from closed-shells.
+We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5),
+and that $\Delta$hCI1.5 is slightly less accurate than $\Delta$hCI1.
+
+This residual bias virtually disappears at the next order, $\Delta$hCI2 (MSE of \SI{-0.01}{\eV}), 
+whereas for the closed-shell systems even $\Delta$hCI3.5 significantly underestimated (MSE of \SI{-0.08}{\eV}) the excitation energies. 
+The same comparison between open- and closed-shells holds for eCI models.
+For the open-shell transitions, $\Delta$hCI2 is slightly more accurate (MAE of \SI{0.11}{\eV} and RMSE of \SI{0.16}{\eV}) than $\Delta$CISD (MAE of \SI{0.12}{\eV} and RMSE of \SI{0.21}{\eV}).
+In further contrast to the closed-shell case,
+$\Delta$hCI2.5 is as accurate as $\Delta$hCI2, producing a MAE of \SI{0.09}{\eV}.
+(Accounting only for the 6 states considered for $\Delta$hCI2.5, $\Delta$hCI2 would also have a MAE of \SI{0.09}{\eV}.)
+Likewise, higher order hCI models become progressively more accurate.
+Even though the statistics become more limited,
+they suggest both a small advantage of the hCI models against eCI,
+and that the ground-state routes perform slightly better than the state-specific ones.
+
 Methods that make use of the concept of seniority to generate excited determinants 
 are not invariant under orbital rotations within the occupied and virtual subspaces. \cite{Bytautas_2011,Limacher_2014,Stein_2014}
-This applies for the hCI and sCI models considered here.
-Here we employed HF ground-state orbitals or excited-state orbitals optimized for a minimal reference. \cite{Kossoski_2023}
-Further exploiting the rotations within occupied and virtual subspaces could render more suitable orbitals for hCI and sCI calculations.
-Alternatively, one could fully optimize the orbitals at a given CI model.
-
-%\begin{table}[ht!]
-%\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values,
-%for the Set of 16 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
-%}
-%\label{tab:stat3}
-%\begin{ruledtabular}
-%\begin{tabular}{lddd}
-%method            &     \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
-%\hline
-%CISD              & +4.09 & 4.09 & 4.18  \\
-%CISDT             & +0.09 & 0.18 & 0.20  \\
-%CISDTQ            & +0.14 & 0.14 & 0.16  \\
-%$\Delta$CISD      & -0.13 & 0.17 & 0.21  \\
-%$\Delta$CISDT     & -0.19 & 0.19 & 0.23  \\
-%$\Delta$CISDTQ    & -0.02 & 0.02 & 0.02  \\
-%\hline
-%hCI1              & +1.62 & 1.62 & 1.65 \\
-%hCI1.5            & +2.10 & 2.10 & 2.11 \\
-%hCI2              & +2.89 & 3.00 & 3.28  \\
-%hCI2.5            & +1.93 & 1.93 & 2.05  \\
-%hCI3              & +0.18 & 0.18 & 0.19  \\
-%$\Delta$hCI1      & -1.37 & 1.50 & 1.66 \\
-%$\Delta$hCI1.5    & -2.62 & 2.62 & 2.72 \\
-%$\Delta$hCI2      & -0.22 & 0.22 & 0.27  \\
-%$\Delta$hCI2.5    & -0.27 & 0.27 & 0.30  \\
-%$\Delta$hCI3      & -0.21 & 0.21 & 0.24  \\
-%$\Delta$hCI3.5    & -0.08 & 0.08 & 0.09  \\
-%\end{tabular}
-%\end{ruledtabular}
-%\end{table}
+This applies for the hCI and sCI models addressed here,
+which employed HF ground-state orbitals or excited-state orbitals optimized for a minimal reference.
+Further exploiting the rotations within occupied and virtual subspaces (by using localization techniques) could render more suitable orbitals for hCI and sCI calculations.
+Alternatively, one could fully optimize the orbitals at a given level of CI. \cite{Kossoski_2022}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Seniority CI for excited states}
-\label{sec:res_C}
+\subsection{sCI for excited states}
+\label{sec:res_D}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%This provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models.
-
-We refer back to Tab.~\ref{tab:stat1} to discuss the performance of seniority-based CI for excited states of closed-shell systems.
-The first model, sCI2/sCI2, systematically overestimates the excitation energies, and presents a MAEs of \SI{1.34}{\eV}.
+We refer back to Tab.~\ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
+The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.34}{\eV}.
 The same two factors discussed above for hCI explain such large errors.
@@
 Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAE to \SI{0.79}{\eV}, though still overestimating the excitation energies.
-Second, there is an unbalanced description of correlation for ground and excited states.
-%While the ground state reference (which has $s=0$) can access determinants with unpaired electrons (in the $s=2$ sector)
-While sCI2/sCI2 accounts for an additional seniority sector ($s=2$) for the ground state reference (which has $s=0$), 
-no further electron unpaired is allowed for the excited state, whose reference already lives in the $s=2$ seniority sector.
-In other words, unpaired excitations are allowed for the ground state but not for the excited state, relative to their respective reference determinants, which creates a bias towards the former.
-A possible solution to this unbalance is to restrict the determinants to the maximum seniority number of the reference.
+Second, there is an unbalanced description of correlation for ground and excited states, associated with the classes of determinants.
+While sCI2/sCI2 accounts for an additional $s=2$ sector for the ground state (which can be qualitatively described in the $s=0$ sector),
+no addition pair of electrons is allowed to become unpaired in the excited state calculation, even though the state is qualitatively described by a determinant that is contained in the $s=2$ sector.
+In other words, unpaired excitations are allowed to correlate the ground state, but not the excited state, given their respective closed-shell and open-shell characters, 
+thus creating a bias towards the former.
+
+A possible solution to this unbalance is to restrict the determinants to a maximum seniority number of $s=0$ for the ground state and $s=2$ for the excited state calculation.
 This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.57}{\eV}, compared to \SI{1.34}{\eV} of sCI2/sCI2.
@@
 Ultimately, we did not find a combination of sCI model and orbitals that produced reasonable excitation energies for closed-shell systems.
-%sCI2 delivers quite poor results (MAE of \SI{1.06}{eV}), whereas $\Delta$sCI2 reduces the statistical errors (MAE of \SI{0.60}{eV}).
-%This improvement, however, is not as expressive as observed for excitataion-based and hierarchy-based methods.
-%Compare, for instance, with the cases of hCI2.5 and $\Delta$hCI2.5 shown in Tab.~\ref{tab:stat1}.
-%The large errors, even following the state-specific approach, combined with the formal $\order*{e^N}$ scaling of these methods, render seniority-based CI unsuitable for calculations of excitation energies.
-pCCD-CI for excited states \cite{Nowak_2023}
 
-The situation for open-shell systems is quite different, in close parallel to the previous discussion of hCI.
-As shown in Tab.~\ref{tab:stat2}, sCI provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
-The reason should not be surprising at this point:
-both ground and excited states of open-shells share the same type of reference (a single $s=1$ open-shell determinant).
-% check for s=3 reference
-Moreover, the state-specific approach is clearly superior.
-sCI1 presents a MAE of \SI{0.39}{\eV} and overestimates the excitation energy (MSE of \SI{+0.24}{\eV}), reflecting the bias toward the ground state due to ground-state orbitals.
+The situation for open-shell systems is quite different, in close analogy to the previous discussion regarding hCI.
+As shown in Tab.~\ref{tab:stat2}, sCI1 provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
+The reason should not be surprising at this point.
+Both ground and excited states of open-shells can be qualitatively described by the same type of reference (a single $s=1$ open-shell determinant).
+Moreover, the state-specific approach is superior.
+sCI1 presents a MAE of \SI{0.39}{\eV} and overestimates the excitation energies (MSE of \SI{+0.24}{\eV}), in view of the bias introduced by the ground-state orbitals.
 With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.27}{\eV} and the MSE to \SI{-0.11}{\eV}, which is smaller in absolute value than found for sCI1.
@@
 Even though $\Delta$sCI1 is less accurate and has less favourable computational scaling than CI models like $\Delta$CISD and $\Delta$hCI2,
-its descent errors are encouraging for another reason.
+its decent errors are encouraging for another reason:
 the development of polynomial scaling CC methods based on the concept of seniority.
-For closed-shell systems, sCI0 energies can be very well reproduced with pair coupled cluster doubles (pCCD), \cite{Kossoski_2021} a method that has mean-field cost.
-Similarly, a formulation of pCCD to open-shell systems might share a analogous connection to low-order seniority-based CI models like sCI1.
+For closed-shell systems, DOCI (here referred to as sCI0) energies can be very well reproduced with state-specific pair coupled cluster doubles (pCCD), a method that has mean-field cost,
+for both ground \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967} and excited \cite{Kossoski_2021,Marie_2021} states.
+Likewise, a formulation of pCCD to open-shell systems might share a analogous connection to low-order sCI models like sCI1.
 If that is the case, then a state-specific approach of such pCCD formulation adapted to open-shells may approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.27}{\eV}) at a mean-field cost.
-This method could then provide a better starting point to recover the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.41}{\eV}).
-To the best of our knoweledge, there has been a single proposed extension of pCCD to open-shell systems, \cite{Boguslawski_2021}
-based on the ionization-potential equation-of-motion (EOM) formalism.
+This method could then provide an improved starting point to recover the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.41}{\eV}).
+To the best of our knowledge, there has been a single proposed extension of pCCD to open-shell systems, \cite{Boguslawski_2021}
+based on the ionization-potential equation-of-motion CC (EOM-CC) formalism.
 
@@
 \section{Conclusion}
-\label{sec:ccl}
+\label{sec:conclusion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%three CI routes (excitation, hierarchy, and seniority-based).
-%Here we address the knowledge gaps outlined above by setting four goals.
-%Extending our previous findings for closed-shell \cite{Kossoski_2022} to open-shell systems,
-%We further explored Epstein-Nesbet second-order perturbative corrections to the various approximate CI methods.
-%The second is to assess the performance of this method for ground state radicals, similarly to what has been done for closed-shell systems. \cite{Kossoski_2022}
+Here, we have generalized hCI \cite{Kossoski_2022} for an arbitrary reference determinant, thus extending its applicability to radical species and state-specific excited state calculations.
 
-%The third goal is to assess the improvement from perturbative corrections to ground-state radicals and closed-shell systems, for both excitation-based and hierarchy-based CI methods.
-%For hCI and excitation-based CI, the perturbative correction tends to accelerate and stabilize the convergence of different properties toward the FCI results.
-%EN2 perturbative corrections lead to faster and more monotonic convergence of different properties
-%We have considered both closed and open shell systems, considering ground-state properties and excitation energies, for a diverse set of molecular systems.
-%Finally, we introduce and gauge seniority-based and hierarchy-based CI methods for excited states, along both the standard ground-state based route and the state-specific route. \cite{Kossoski_2023}
+By surveying the dissociation of four radicals,
+we found that the hCI route outperforms or matches eCI, for both weakly correlated (equilibrium properties) and strongly correlating (dissociation) regimes.
+These and previous \cite{Kossoski_2022} findings demonstrate the ability of hCI models to recover weak and strong correlation for both open- and closed-shell systems.
+Meanwhile, sCI leads to far less accurate results in comparison to hCI or eCI, for a given computational cost.
+For closed- and open-shell systems, the EN2 perturbative correction
+substantially accelerates and stabilizes the convergence of hCI and eCI routes (while keeping the advantage of the former),
+though it ameliorates the sCI route to a lesser extent.
+The standard EN2 correction typically produced more accurate results than its renormalized form, except perhaps for describing single bond breaking.
+Overall, lower order models like hCI1+EN2 were found to be fairly accurate given its low computational cost.
+In the future, it may be worth combining orbital optimization at a lower level of hCI \cite{Kossoski_2022} with the correction provided by perturbation theory.
 
-Here we have generalized hCI \cite{Kossoski_2022} for the case of an open-shell reference determinant, thus extending its applicability to radical species and excited states.
-In the first part,
-by surveying the dissociation of several open-shell systems, 
-we found that the hCI route generally outperforms excitation-based CI, for both weakly correlated (equilibrium geometries) and strongly correlating (dissociation) regimes.
-Meanwhile, seniority-based CI leads to far less accurate results in comparison to hCI or excitation-based CI, for a given computational cost.
-These findings are in line with what had been found for closed-shell systems. \cite{Kossoski_2022}
-We further evaluated, for closed- and open-shell systems, the impact of EN2 perturbative corrections.
-They lead to substantial improvements for hCI and excitation-based CI routes (while keeping the advantage of the former), 
-but do not help much with the deficiencies of the sCI route.
+We further gauged the performance of hCI to describe excited states of closed-shell and open-shell systems, based on either HF ground-state orbitals or state-specific orbitals.
+For excitations of closed-shells, $\Delta$hCI2 and $\Delta$hCI3 are comparable to their excitation-based counterparts, $\Delta$CISD and $\Delta$CISDT,
+whereas $\Delta$hCI1, $\Delta$hCI1.5 and $\Delta$hCI2.5 models are inaccurate bearing in mind their computational cost.
+The poor performance at lower orders is ascribed to the different minimal references employed,
+a single $s=0$ determinant for the ground state and a single $s=2$ CSF for the excited state,
+which introduces a strong bias on the classes of determinants accessed in the hCI calculations.
+$\Delta$hCI performs significantly better for the doublet transitions, when compared to the results for closed-shells.
+In this case, the reference of both ground and excited states comprises the same type of single $s=1$ determinant.
+The advantage of using state-specific orbitals over ground-state ones depends on the choice of reference and the order in which the CI is truncated.
+When similar references are adopted (as for the doublet transitions), such advantage is already evident at $\Delta$hCI1.
+In contrast, for unlike references (as for the closed-shell excitations), the state-specific approach only becomes advantageous at somewhat higher orders ($\Delta$hCI2).
+The present findings highlight the challenge in describing on an equal footing states with qualitatively different characters, in particular ground and singly-excited states of closed-shell systems.
 
-In the second part,
-we gauged the performance of hCI to describe excited states of closed-shell and open-shell systems, based on either HF ground-state orbitals or state-specific orbitals.
-We discussed the 
-%
-For lower orders of hCI (hCI1 and hCI1.5), the performance is poor,
-with ground-state (state-specific) orbitals leading to largely overestimated (underestimated) excitation energies.
-The situation improves at the state-specific $\Delta$hCI2 order, even though its accuracy is somewhat worse than $\Delta$CISD, both models sharing the same computational scaling. 
-Unfortunately, $\Delta$hCI2.5 does not lead to more accurate results, whereas all the CI models at next next order (CISDT, $\Delta$CISDT, hCI3, and $\Delta$hCI3) typically show comparable performances.
-Overall, hCI models for excited states are either innacurate given their computational cost ($h=1$, $h=1.5$ and $h=2.5$) or comparable to excitation-based CI at the same order ($h=2$ and $h=3$).
-%it is fair to conclude that 
-The present findings highlight the challenge in describing on an equal footing states with qualitatively different characters (like ground and singly-excited states of closed-shell systems).
-%, and thus should not be useful in practice.
+We put forward the interesting perspective of developing hierarchy-based CC (hCC) methods (hCC1, hCC2, and so on),
+along with derived EOM-CC formulations to target excited states (EOM-hCC1, EOM-hCC2, and so on).
+EOM-CC with single and double excitations (EOM-CCSD) \cite{Rowe_1968,Monkhorst_1977,Koch_1990,Stanton_1993}
+is very accurate for describing singly-excited states, \cite{Loos_2018,Loos_2020,Loos_2020a} albeit it slightly overestimates their excitation energies.
+In contrast, CISD provides too big excitation energies, \cite{Koch_1990,Kossoski_2023}
+even though both CISD and EOM-CCSD methods span the same excited determinants and rely on HF ground-state orbitals.
+Here, we found that hCI2 is more accurate than CISD, despite their large absolute errors. In particular, the overestimated energies are less exaggerated in the former.
+We ponder whether this improvement would be transferable to EOM-hCC2.
+Along the same line, it would be interesting to develop and gauge the performance of cheaper methods like EOM-hCC1.
+Importantly, it is not obvious what the computational scaling of such methods would be.
+Alternatively, one can envision state-specific hCC methods to target excited states.
 
-Here we put forward the interesting perspective of developing hierarchy-based CC (hCC) methods for excited states.
-For instance, EOM-CC with single and double excitations (EOM-CCSD) \cite{} is very accurate for excitation energies of singly-excited states, \cite{Loos_2018,Loos_2020,Loos_2020a},
-whereas CISD provides too big excitation energies, 
-even though both methods use ground-state orbitals and span the same excited determinants.
-Here we found that hCI2 is more accurate than CISD to some extent. In particular, the overestimated energies are slightly less exaggerated in the former.
-We wonder whether this improvement would be transferable to an analogous EOM formulation built on top of hCC truncated at $h=2$, EOM-hCC2.
-Importantly, it is not obvious what the computational scaling of such a method would be.
-Along the same line, it would be interesting to develop and test the performance of cheaper orders like EOM-hCC1.
-Alternatively, one could develop hCC methods to target excited states along the state-specific route.
-
-%This general observation can be understood from the fact that 
-%The same type of reference provides a fair starting point for the subsequent sCI calculations, similarly to what was as discussed above for hCI models applied to excitations of open-shell systems.
-%thus significantly minimizing an unbalanced account of correlation, encountered for excitations from closed-shells.
-
-Finally, in the third part of this contribution we employed different sCI models to compute excitation energies of closed-shell and open-shell systems.
-In spite of their formal exponential computational scaling, the sCI models did not produce attractive results compared with more competitive CI alternatives.
-The relevance of these calculations is related to the connection between seniority-based CI and pCCD methods.
-We raised the question whether there is an extension of pCCD to open-shell systems that can provide very similar energies to sCI1 (but with a polynomial scaling),
-which is already established (pCCD and sCI0) for closed-shell systems. \cite{Kossoski_2021}
-In this sense, the present results obtained with sCI1 for open-shell excitations are appealing and encourage the development of such a generalized pCCD.
+Finally, we employed different sCI models to compute excitation energies of closed-shell and open-shell systems.
+Despite the four different models employed for the closed-shell excitations, the results are overall disappointing.
+For the open-shell transitions, the outcome is more encouraging.
+The relevance of these sCI models lies in its possible connection with CC methods.
+While this is already established for closed-shell systems (pCCD and DOCI deliver very similar energies), \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967,Kossoski_2021,Marie_2021}
+here we raised the question of whether there exists a polynomial scaling extension of pCCD to open-shell systems that matches sCI1 in terms of computed energies.
+In this sense, the present results obtained with $\Delta$sCI1 for open-shell excitations are appealing and encourage the development of a generalized pCCD method.
 
@@
 
+Equilibrium geometries of ethylene and vinyl; details about the fitting of potential energy curves;
+potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
+computed according to FCI and the various hCI, eCI, and sCI models considered here;
+and convergence of the NPE, distance errors, and equilibrium properties of
+open-shell (\ce{OH}, \ce{CN}, vinyl, and \ce{H7})
+and closed-shell (\ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}) systems,
+as a function of the number of determinants, according to the various CI models considered here while
+corrected with the standard and the renormalized EN2 perturbative energy.
+
+For the full set of 48 excited states for closed-shell molecules and 19 excited states for open-shell radicals,
+total energies and excitation energies obtained with 
+the various hCI, eCI, and sCI models considered here, in both ground-state based and state-specific approches;
+number of determinants in the reference; saddle point order associated with the $\Delta$CSF solutions;
+and the reference excitation energies and corresponding methods.
+For a subset of 16 excited states for closed-shell molecules and 6 excited states for open-shell radicals,
+additional total energies and excitation energies obtained with higher-order hCI and eCI models.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-08-06 Manuscript/sup.tex
saving work

@@
 \newcommand{\QP}{\textsc{quantum package}}
+
+\newcommand{\Ndet}{N_\text{det}}
 
@@
 Potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
-according to hCI (green), excitation-based CI (red) and sCI (blue) models.
-}
+according to FCI (black), hCI (green) (dashed lines for half-integer $h$), eCI (red), and sCI (blue) models, with the cc-pVDZ basis set and frozen-core approximation.}
 \label{fig:plot_pes}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S2 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_OH}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{OH},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_oh}
 \end{figure*}
@@
 \begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_pes_closed}
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_CN}
 \caption{
-Potential energy curves for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
-according to hCI (green), excitation-based CI (red) and sCI (blue) models,
-with (light tones) and without (dark tones) the EN2 perturbative correction.
-Results first reported in Ref. \onlinecite{Kossoski_2022}.
-}
-\label{fig:plot_pes_closed}
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{CN},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_cn}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S4 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_vinyl}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for vinyl,
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_vinyl}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S5 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H7}
+\caption{
+\label{fig:plot_pt2_rpt2_h7}
+Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H7},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S6 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_HF}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{HF},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_hf}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S7 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_F2}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{F2},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_f2}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S8 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_ethylene}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for ethylene,
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_ethylene}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S9 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_N2}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{N2},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_n2}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S10 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H4}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H4},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_h4}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S11 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H8}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H8},
+as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_h8}
 \end{figure*}

### 2023-08-11 Manuscript/manuscript.tex
map
corrections

@@
 for a closed-shell reference (left), an open-shell reference with one unpaired electron (center) and an open-shell reference with two unpaired electrons (right).
-The color tones represent the determinants that are included at a given hCI model.}
+The color tones represent the determinants that are included at a given hCI model. hCI0 reduces to HF in the former two cases, where $s_0 = 0$ and $s_0 = 1$.}
 \label{fig:hCI}
@@
 
-The first order for a closed-shell reference determinant is hCI1 (see left of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS) 
+For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the restricted HF determinant.
+The first non-trivial order is hCI1 (see left of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS)
 plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital),
@@
 For an odd number of electrons, $s$ assumes odd values starting from $s=1$.
-The simplest reference is an open-shell determinant with one unpaired electron (see center pannel of Fig.~\ref{fig:hCI} and top right of Fig.~\ref{fig:determinants}).
-In contrast to the closed-shell case, here there are no empty blocks in the excitation-seniority map.
+The simplest reference is an open-shell determinant with one unpaired electron ($s_0=1$), shown in the center pannel of Fig.~\ref{fig:hCI} and in the top right of Fig.~\ref{fig:determinants}.
+hCI0 also reduces to the restricted open-shell HF determinant.
+In contrast to the closed-shell case, here there are no empty blocks in the seniority-excitation map.
 Actually, the hCI series displays the hCI0.5 level, accounting only for the single excitations from and into the singly-occupied orbital.
@@
 
-Finally, one can employ a high-spin reference determinant to define the hCI models.
-The case of two unpaired electrons is shown in the right pannel of Fig.~\ref{fig:hCI},
-which would be appropriate for singly-excited states of closed-shells.
+Finally, one can employ a reference determinant having more unpaired electrons to define the hCI models.
+The case of two unpaired electrons ($s_0=2$) is shown in the right pannel of Fig.~\ref{fig:hCI},
+which would be appropriate for describing singly-excited states of closed-shells.
 The key difference in the hCI sequence, with respect to the closed-shell case, lies in the displacement by one block to the right in the seniority-excitation map,
 reflecting the shift from the $s_0=0$ to the $s_0=2$ reference.
-In the two cases discussed, the hCI0 level only included the reference determinant.
-For a $s=2$ reference, however, hCI0 not only accounts for the reference determinant,
+In addition, in the two previous cases, the hCI0 level only included the reference determinant.
+whereas for $s_0=2$, hCI0 not only accounts for the reference determinant,
 but also for the two closed-shell determinants produced by the single excitation that pairs the two unpaired electrons.
 The hCI1 classes of determinants for the $s_0=2$ reference can be seen in the bottom of Fig.~\ref{fig:determinants},
-which clearly outnumber the fewer classes associated with the $s_0=0$ and $s_0=1$ references.
-One could adopt even higher-spin references, further displacing the hCI sequence to the right in the seniority-excitation map.
+which clearly outnumber the fewer classes associated with the $s_0=0$ and $s_0=1$ cases.
+One could adopt references with even larger $s_0$, further displacing the hCI sequence to the right in the seniority-excitation map.
 In this case, hCI0 would include all the lower spin determinants obtained by partially or totally pairing the unpaired electrons.
 Notice that the number of such determinants do not depend on the size of the system (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
-In a similar fashion, hCI can be built on top of a high-spin reference for odd number of electrons.
+In a similar fashion, hCI can be built on top of a larger $s_0$ reference for odd number of electrons.
 
@@
 \begin{acknowledgements}
-This work was performed using HPC resources from CALMIP (Toulouse) under allocation 2021-18005.
+This work was performed using HPC resources from CALMIP (Toulouse) under allocation 2023-18005.
 This project has received funding from the European Research Council (ERC) under the European Union's Horizon 2020 research and innovation programme (Grant agreement No.~863481).

### 2023-08-14 Manuscript/manuscript.tex
s2 and high-spin

@@
 
+For a reference containing unpaired electrons, an approximate CI model will generally produce spin-contaminated states.
+One can impose the solutions to have well-defined spin by including additional determinants generated via higher-order hierarchy, which account for the missing spin-flip configurations.
+The spin-contamination problem and its solution are therefore equivalent to that encountered in eCI. \cite{Maurice_1996}
+Here, we have employed this procedure and considered pure spin states.
+For the $s_0 = 2$ reference, one could alternatively employ a high-spin triplet determinant (two unpaired spin-up electrons), 
+rather than the low-spin determinant shown in Fig.~\ref{fig:ref}.
+Regardless of the choice, hCI (as eCI) produces the same energies for the triplet states, provided that a spin eigenstate is imposed for the low-spin reference.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Core orbitals were frozen.
-For the radicals, we further impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies in accounting for a set of appropriate spin-flipped determinants.
-In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS, \cite{Maurice_1996} and equivalently for the other hCI, eCI and sCI models.
+We impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies in accounting for a set of appropriate spin-flipped determinants stemming from higher-order excitation or hierarchy.
+In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS \cite{Maurice_1996} for instance, and equivalently for the other hCI, eCI and sCI models.
+Notice that spin-contaminated solutions would have different energies than the spin eigenstates considered here.
+
 We performed calculations following both the standard ground-state based CI route and the state-specific CI route. \cite{Kossoski_2023}
@@
 The computed excitation energies were benchmarked against the reference values of FCI quality provided in the QUEST database. \cite{Veril_2021}
-
 The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as the reference determinant for the hCI and eCI ground state calculations
-(left and middle pannels of Fig.~\ref{fig:hCI})
+(left and middle pannels of Fig.~\ref{fig:hCI}).
 For the state-specific hCI and eCI calculations, we employed a minimal CSF reference: \cite{Kossoski_2023}

### 2023-08-14 Manuscript/manuscript.tex
correction
fix ref

@@
 For the $s_0 = 2$ reference, one could alternatively employ a high-spin triplet determinant (two unpaired spin-up electrons), 
-rather than the low-spin determinant shown in Fig.~\ref{fig:ref}.
-Regardless of the choice, hCI (as eCI) produces the same energies for the triplet states, provided that a spin eigenstate is imposed for the low-spin reference.
+rather than the low-spin determinant shown in Fig.~\ref{fig:hCI}.
+Regardless of the choice, hCI (as eCI) produces the same energies for the triplet states, provided that a spin eigenstate is imposed.
 

### 2023-08-17 Manuscript/manuscript.tex
added appendix

@@
 
+\newcommand{\Na}{N_{\alpha}}
+\newcommand{\Nb}{N_{\beta}}
+\newcommand{\Nd}{N^{d}}
+\newcommand{\Nv}{N^{v}}
+\newcommand{\Nas}{N^s_{\alpha}}
+\newcommand{\Nbs}{N^s_{\beta}}
+
 \newcommand{\EHF}{E_\text{HF}}
-\newcommand{\EDOCI}{E_\text{DOCI}}
 \newcommand{\EFCI}{E_\text{FCI}}
@@
 In a similar fashion, hCI can be built on top of a larger $s_0$ reference for odd number of electrons.
+The deduction for the number of determinants in a given hCI model, for the $s_0=0$ closed-shell, $s_0=1$ open-shell, as well as for an arbitrary reference determinant,
+can be found in Sec.~\ref{sec:appendix}.
 
@@
 
+\appendix
+
+\section{Number of determinants}
+\label{sec:appendix}
+
+What is the number of determinants in a given hCI model, defined by the hierarchy $h$ [see Eq.~\eqref{eq:h}] and the reference determinant?
+Here, we address this question by first working out the simpler case of a closed-shell reference.
+Then, we move to a slightly more complicated case of an open-shell reference with a single unpaired electron.
+Finally, we deduce the general case.
+The reference determinant of each case is shown in Fig.~\ref{fig:appendix}.
+
+%%% FIG 5 %%%
+\begin{figure}[h!]
+\includegraphics[width=1.0\linewidth]{determinants_appendix}
+\caption{
+$s_0=0$ closed-shell determinant with $N$ electrons (left),
+$s_0=1$ open-shell determinant with $\Na$ $\alpha$-electrons and $\Nb=\Na-1$ $\beta$-electrons (center),
+and general $s_0=\Nas+\Nbs$ determinant, with $\Nd$ doubly-occupied orbitals, $\Nas$ singly-occupied $\alpha$-electrons and $\Nbs$ singly-occupied $\beta$-electrons (right).
+In the three cases, there are $K$ spatial orbitals.
+The number of determinants generated from these three reference determinants, from left to right, which have excitation degree $e$ and seniority $s$
+are given in Eqs.~\eqref{eq:det_1}, ~\eqref{eq:det_2}, and ~\eqref{eq:det_3}.
+}
+\label{fig:appendix}
+\end{figure}
+%%% %%% %%%
+
+The gist of the deduction is as follows.
+We start from the known number of determinants based only on the excitation degree $e$.
+Then, we systematically decompose each term into specific contributions,
+based on whether the excitation increases or reduces the seniority.
+For the decomposition, we make use of the Vandermonde's identity,
+\begin{equation}
+	\binom{n}{k} = \sum_{j}^{k} \binom{m}{j} \binom{n-m}{k-j}.
+\end{equation}
+Next, an incremental seniority can be assigned to each type of contribution.
+Finally, the final answer can be obtained by summing only the contributions with the desired seniority $s$.
+
+We start with the simpler case of a closed-shell determinant with $N$ electrons and $K$ spatial orbitals (left of Fig.~\ref{fig:appendix}).
+From this reference determinant, the number of excited determinants generated by exciting exactly $e$ electrons, is given by \cite{SzaboBook}
+\begin{equation}
+\sum_{p=0}^{e} \binom{N/2}{p} \binom{K-N/2}{p} \binom{N/2}{e-p} \binom{K-N/2}{e-p},
+\label{eq:e_closed}
+\end{equation}
+where the sum expresses the different combinations for exciting $\alpha$- or $\beta$-electrons,
+the first binomial term accounts for the number of possibilities for exciting $p$ $\alpha$-electrons from the $N/2$ orbitals,
+the second term represents the number of ways of placing these $p$ electrons into $K-N/2$ orbitals,
+and similarly for the latter two terms and the $e-p$ remaining $\beta$-electrons.
+
+To account for the seniority of the excited determinants, one must disentangle the excitations based on how they change the seniority number,
+while keeping track of the previously excited electrons.
+Starting from the closed-shell determinant, $p$ $\alpha$-electrons are excited, which increases the seniority by $2p$
+(a factor of $p$ due to the unpaired $\alpha$-electrons just excited,
+and another factor of $p$ due to the unpaired $\beta$-electrons left behind).
+The first two binomial terms in Eq.~\ref{eq:e_closed} are left untouched since they are always accompanied by the same change in seniority.
+Next, the $\beta$-electrons can be excited from the same orbitals from which the $\alpha$-electrons were excited (decreasing the seniority by one)
+or instead from an orbital that remained doubly-occupied (increasing the seniority by one).
+The two possibilities are expressed by decomposing the third binomial term in Eq.~\ref{eq:e_closed} as a sum over the two corresponding binomials,
+\begin{equation}
+\binom{N/2}{e-p} = \sum_{q=0}^{e-p} \binom{p}{q} \binom{N/2-p}{e-p-q}.
+\end{equation}
+The first counts the number of possibilities for exciting $q$ $\beta$-electrons from one of $p$ orbitals for which an $\alpha$-electron was already excited.
+By removing the unpaired $\beta$-electron left behind, the seniority thus decreases by a factor of $q$.
+The second term accounts for the complementary excitations, where the $(e-p-q)$ $\beta$-electrons are chosen out of the $N/2-p$ orbitals that 
+remained doubly-occupied after exciting the $p$ $\alpha$-electrons.
+Therefore, this term increases the seniority by $(e-p-q)$.
+We proceed similarly for the fourth binomial term of Eq.~\ref{eq:e_closed}, by decomposing it into the sum of two other binomials,
+where $r$ $\beta$-electrons pair with the previously $\alpha$-electrons (thus reducing the seniority) and $(e-p-r)$ do not (which increases the seniority).
+By collecting the seniority changes from each term: $(2p)$ from the unmodified binomials, $(-q)$ and $(e-p-q)$ from the third binomial, $(-r)$ and $(e-p-r)$ from the fourth one,
+the seniority of the excited determinant is given by $s = 2 (e-q-r)$.
+Finally, by combining the seniority specific binomials and imposing the desired seniority via a Kronecker delta,
+the number of determinants with a given excitation degree $e$ and seniority $s$ produced from a closed-shell reference ($s_0=0$) is given as
+\begin{equation}
+\begin{split}
+	&\sum_{p=0}^{e} \binom{N/2}{p} \binom{K-N/2}{p}
+	\sum_{q=0}^{e-p} \binom{p}{q} \binom{N/2-p}{e-p-q}  \\
+	&\sum_{r=0}^{e-p} \binom{p}{r} \binom{K-N/2-p}{e-p-r}
+ 	\delta_{s,2(e-q-r)}.
+\label{eq:det_1}
+\end{split}
+\end{equation}
+To obtain the final number of determinants for a given hierarchy $h$, we simply sum over the allowed combinations of $e$ and $s$ according to Eq.~\eqref{eq:h}.
+
+Moving to the case where the number of $\alpha$-electrons ($\Na$) and $\beta$-electrons ($\Nb$) differ by one ($\Na=\Nb+1$), which is illustrated in the center of Fig.~\ref{fig:appendix}.
+The number of excited determinants generated exclusively by exciting $e$ electrons %$\Ndet(e)$, 
+is analogous to Eq.~\eqref{eq:e_closed}, being given by
+\begin{equation}
+\sum_{p=0}^{e} 
+\binom{\Na}{p} \binom{K-\Na}{p} \binom{\Nb}{e-p} \binom{K-\Nb}{e-p}.
+\label{eq:e_open}
+\end{equation}
+
+The deduction proceeds similarly to the closed-shell case. 
+The main differences are the following.
+The first binomial term of Eq.~\ref{eq:e_open} is decomposed into the sum of $t$ $\alpha$-electrons excited from the singly-occupied orbital and $(p-t)$ that are excited from the $\Nb$ doubly-occupied ones.
+The second binomial is left untouched, as before.
+The third one is also decomposed as explained for the closed-shell case, with the difference that the $q$ $\beta$-electrons are excited from $(p-t)$ orbitals (instead of $p$) 
+from which $\alpha$-electrons were excited. 
+Similarly, in the fourth binomial of Eq.~\ref{eq:e_open}, $r$ $\beta$-electrons are placed into one of the $(1+p-t)$ orbitals (rather than $p$ in the closed-shell case) which contain an unpaired $\alpha$-electron.
+Collecting the individual contributions and setting the targeted seniority,
+the number of determinants generated from a $s_0=1$ reference determinant, with excitation degree $e$ and seniority $s$ is given by
+\begin{equation}
+\begin{split}
+	& \sum_{p=0}^{e} 
+	\sum_{t=0}^{p} \binom{1}{t} \binom{\Nb}{p-t} 
+	\binom{K-\Na}{p} 
+	\sum_{q=0}^{e-p} \binom{p-t}{q} \binom{\Nb-p+t}{e-p-q} \\
+	& \sum_{r=0}^{e-p} \binom{1+p-t}{r} \binom{K-\Nb-1-p+t}{e-p-r}
+ 	\delta_{s,1+2(e-q-r-t)}.
+\label{eq:det_2}
+\end{split}
+\end{equation}
+
+In the general case, an arbitrary reference determinant is defining by three parameters, the number of doubly-occupied orbitals ($\Nd$),
+the number of singly-occupied $\alpha$-electrons ($\Nas = \Na - \Nd$),
+and the number of singly-occupied $\beta$-electrons ($\Nbs = \Nb - \Nd$).
+(Alternatively, one could employ the number of electrons $N$, $\Nd$, and the spin quantum number $S_z = (\Na-\Nb)/2$.)
+For $K$ spatial orbitals, the number of virtual orbitals is $\Nv = K - \Nd - \Nas - \Nbs$.
+
+The deduction is somewhat more involved, but follows along the same lines as the two previously discussed cases.
+The starting point is also given by Eq.~\ref{eq:e_open}, with each binomial being written as a sum of two or more binomials that distinguish the change in seniority.
+Here, we just state how each term is decomposed, where the details can be checked by inspection of the final result presented below [Eq.~\eqref{eq:det_3}].
+The first binomial of Eq.~\ref{eq:e_open} is decomposed into the contributions of $\alpha$-electrons excited from the $\Nd$ doubly- and from the $\Nas$ singly-occupied orbitals.
+The second binomial is decomposed into the $\alpha$-electrons being excited to the $\Nv$ virtual orbitals and to the $\Nbs$ singly-occupied orbitals.
+The third one undergoes two decompositions, and is thus expressed as a double sum over three binomials.
+The $\beta$-electrons are distinguished between the $\Nbs$ singly-occupied orbitals, a subset of $\Nd$ doubly-occupied orbitals for which $\alpha$-electrons were excited,
+and the complementary subset for which no $\alpha$-electrons were excited.
+Finally, the fourth binomial of Eq.~\ref{eq:e_open} counts the number of possibilities to place the excited $\beta$-electrons.
+It is first decomposed into two terms, based on whether they are promoted to the $\Nv$ virtual orbitals or to the $\Nas$ singly-occupied orbitals.
+Each term is further decomposed into two terms, accounting for the subset of $\Nv$ orbitals which now contain $\alpha$-electrons,
+and the subset of $\Nas$ orbitals from which $\alpha$-electrons were removed,
+thus leading to three sums over four binomials.
+The final result for the number of determinants generated from an arbitrary reference determinant, constrained to have excitation degree $e$ and seniority $s$ is given by
+\begin{equation}
+\begin{split}
+	& \sum_{p=0}^{e} 
+	\sum_{t=0}^{p} \binom{\Nas}{t} \binom{\Nd}{p-t} 
+	\sum_{w=0}^{p} \binom{\Nv}{w} \binom{\Nbs}{p-w} \\
+	& \sum_{u=0}^{e-p} \binom{\Nbs}{u} 
+	\sum_{v=0}^{e-p-u} \binom{p-t}{v} \binom{\Nd-p+t}{e-p-u-v} \\
+	& \sum_{r=0}^{e-p} \sum_{m=0}^{r} \binom{w}{m} \binom{\Nv-w}{r-m} 
+	\sum_{n=0}^{e-p-r} \binom{t}{n} \binom{\Nas-t}{e-p-r-n} \\
+	& \delta_{s,\Nas+\Nbs+2(e+w-p-t-u-v-m-n)}.
+\label{eq:det_3}
+\end{split}
+\end{equation}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-08-18 Manuscript/manuscript.tex
abstract

@@
 \begin{abstract}
-{\bf Abstract:}
-Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number,
-which showed to efficiently recover dynamic and static correlation for closed-shell molecular systems [J. Phys. Chem. Lett. 2022, 13, 19, 4342].
-Here, we generalize hCI for arbitrary reference determinants, allowing calculations for radicals and for excited states in a state-specific approach.
-We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI)
-by evaluating how different properties of ground-state radicals converge to the full CI limit.
-hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
+Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number, which showed to efficiently recover both dynamic and static correlations for closed-shell molecular systems [\href{https://doi.org/10.1021/acs.jpclett.2c00730}{\textit{J.~Phys.~Chem.~Lett.}~\textbf{2022}, \textit{13}, 4342}].
+Here, we generalize hCI for an arbitrary reference determinant, allowing calculations for radicals and for excited states in a state-specific way.
+We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI) by evaluating how different ground-state properties of radicals converge to the full CI limit.
+We find that hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
 Employing the second-order Epstein-Nesbet perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
 We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
-Our results underline that both the choice of reference and orbitals drive the fine balance between correlation of ground and excited states.
+Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance \titou{between correlation of ground and excited states}.
 State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.

### 2023-08-18 Manuscript/manuscript.tex
Tab. -> Table

@@
 Likewise, in eCI, $\Ndet$ scales as $N^{2e}$.
-The computational scaling along the two routes is shown in Tab.~\ref{tab:scaling}.
+The computational scaling along the two routes is shown in Table~\ref{tab:scaling}.
 
@@
 though with a small prefactor stemming from the EN2 calculation.
-The computational scaling associated with the CI+EN2 calculations is also presented in Tab.~\ref{tab:scaling}.
+The computational scaling associated with the CI+EN2 calculations is also presented in Table~\ref{tab:scaling}.
 
@@
 
-We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Tab.~\ref{tab:stat1}.
+We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Table~\ref{tab:stat1}.
 The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.07}{\eV}.
@@
 We now shift to the discussion of excitations for open-shell systems.
-The statistical errors are shown in Tab.~\ref{tab:stat2}.
+The statistical errors are shown in Table~\ref{tab:stat2}.
 The key difference on the calculations for open-shell excitations is that the same type of reference was employed for ground- and excited-state calculations, namely,
@@
 
-We refer back to Tab.~\ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
+We refer back to Table~\ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
 The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.34}{\eV}.
@@
 The situation for open-shell systems is quite different, in close analogy to the previous discussion regarding hCI.
-As shown in Tab.~\ref{tab:stat2}, sCI1 provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
+As shown in Table~\ref{tab:stat2}, sCI1 provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
 The reason should not be surprising at this point.

### 2023-08-18 Manuscript/manuscript.tex
intro

@@
 
-The most well-known CI route is based on the excitation degree.
+The most well-known CI route is based on the excitation degree $e$.
 Starting from a reference determinant, typically the Hartree-Fock (HF) determinant,
 one generates all determinants that are connected to it by exciting at most $e$ electrons.
-The excitation degree $e$ thus defines the order of the approximate excitation-based CI (eCI) model:
-CI with single excitations (CIS); CI with single and double excitations (CISD), CI with single, double and triple excitations (CISDT), and so on.
+The excitation degree thus defines the order of the approximate excitation-based CI (eCI) model:
+CI with single excitations (CIS), CI with single and double excitations (CISD), CI with single, double, and triple excitations (CISDT), and so on.
 The eCI route rather quickly captures dynamic (weak) correlation, though it struggles with the description of static (strong) correlation.
@@
 A different CI route is based on the seniority number $s$ (the number of unpaired electrons in a given determinant).
-In seniority-based CI (sCI) \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}, there is no reference determinant, and one accounts for all determinants having seniority equal or less than $s$.
+In seniority-based CI (sCI), \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967} there is no reference determinant and one accounts for all determinants having seniority equal or less than $s$.
 In contrast to eCI, sCI recovers static correlation more efficiently and works well to describe molecular dissociation, \cite{Bytautas_2015,Alcoba_2014,Alcoba_2014b}
@@
 For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually referred to as doubly-occupied CI (DOCI),
-which is followed by the higher-order models sCI2, sCI4, and so on.
-For an odd number of electrons, the seniority route follows along odd numbers of $s$: sCI1, sCI3, and so on.
+which is followed by the higher-order models: sCI2, sCI4, and so on.
+For odd numbers of electrons, the seniority route follows along odd numbers of $s$: sCI1, sCI3, and so on.
 
@@
 hCI was found to display an overall faster convergence toward the FCI results than the traditional eCI route. \cite{Kossoski_2022}
-In this sense, it was able to recover both static and dynamic correlation more effectively than either eCI or sCI.
+In this sense, it was able to recover both static and dynamic correlations more effectively than either eCI or sCI.
 However, hCI has only been defined for a closed-shell reference, thus being limited to the ground state of systems with an even number of electrons.
@@
 There are two possible approaches to target excited states with CI methods.
-One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix, which we refer to as ground-state based approach.
+One can employ the ground-state HF orbitals and obtain excitation energies from the higher-lying eigenvalues of the CI matrix, which we refer to as the ground-state-based approach.
 Instead, one may optimize the orbitals for the excited state of interest (described with an appropriate reference), followed by a separate CI calculation,
@@
 % multiconfigurational SCF, \cite{Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Kossoski_2023,Marie_2023}
-density functional theory,
-\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022}
+density-functional theory,
+\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022,Schmerwitz_2023}
 and coupled cluster (CC)
@@
 In particular, by employing a minimal configuration state function (CSF) reference,
-we have recently shown that excitation-based $\Delta$CI models deliver far more accurate excitation energies than the ground-state based analogues. \cite{Kossoski_2023}
+we have recently shown that excitation-based $\Delta$CI models deliver far more accurate excitation energies than their ground-state-based analogs. \cite{Kossoski_2023}
 
 Here, our first goal is to generalize hCI for an arbitrary type of reference, thus extending its applicability from ground-state closed-shells \cite{Kossoski_2022}
-to radicals and state-specific excited state calculations.
+to radicals and state-specific excited-state calculations.
 This is done in Sec.~\ref{sec:hCI}.
 The computational details about the implementation and the specific calculations performed here are presented in Sec.~\ref{sec:compdet}.
-In Sec.~\ref{sec:res_A}, we assess the performance of hCI, eCI and sCI models by calculating various properties of four ground-state radicals, which comprises our second goal.
+In Sec.~\ref{sec:res_A}, we assess the performance of hCI, eCI, and sCI models by calculating various properties of four ground-state radicals, which comprises our second goal.
 Our third goal is to evaluate how perturbation theory, more precisely the second-order Epstein-Nesbet (EN2) perturbative correction (both standard and renormalized), \cite{Garniron_2019}
@@
 Inspired by the results of hCI for ground-state closed-shell systems, \cite{Kossoski_2022}
-and by the promising set of excitation-based $\Delta$CI, \cite{Kossoski_2023} here we explore hCI models for excited states, following both ground-state based and state-specific approaches.
-In thise sense, our fourth goal is to assess the accuracy of hCI models for excited states, which is the subject of Sec.~\ref{sec:res_C}.
+and by the promising set of excitation-based $\Delta$CI, \cite{Kossoski_2023} here we explore hCI models for excited states, following both the ground-state-based and state-specific approaches.
+In this sense, our fourth goal is to assess the accuracy of hCI models for excited states, which is the subject of Sec.~\ref{sec:res_C}.
 Furthermore, to the best of our knowledge, sCI models have not yet been directly used to target excited states,
-despite the growing number of methods that exploit the concept of seniority number,
+despite the growing number of methods that exploit the concept of seniority,
 for both ground
@@
 \cite{Boguslawski_2016b,Boguslawski_2016c,Boguslawski_2019,Nowak_2019,Kossoski_2021,Marie_2021,Tecmer_2022,Rishi_2023,Nowak_2023}
-Our fifth and last goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state based and state-specific sCI models for excited states.
-Sec.~\ref{sec:conclusion} closes the present contribution with the main conclusions and perspectives.
+Our fifth and last goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state-based and state-specific sCI models for excited states.
+Section \ref{sec:conclusion} closes the present contribution with the main conclusions and perspectives.
 

### 2023-08-18 Manuscript/manuscript.tex
conclusion

@@
 
-Here, we have generalized hCI \cite{Kossoski_2022} for an arbitrary reference determinant, thus extending its applicability to radical species and state-specific excited state calculations.
+Here, we have generalized hCI \cite{Kossoski_2022} for an arbitrary reference determinant, thus extending its applicability to radical species and state-specific excited-state calculations.
 
 By surveying the dissociation of four radicals,
-we found that the hCI route outperforms or matches eCI, for both weakly correlated (equilibrium properties) and strongly correlating (dissociation) regimes.
-These and previous \cite{Kossoski_2022} findings demonstrate the ability of hCI models to recover weak and strong correlation for both open- and closed-shell systems.
+we found that the hCI route outperforms or matches eCI, for both weakly correlated (equilibrium properties) and strongly correlated (dissociation) regimes.
+These and previous \cite{Kossoski_2022} findings demonstrate the ability of hCI models to recover weak and strong correlations for both open- and closed-shell systems.
 Meanwhile, sCI leads to far less accurate results in comparison to hCI or eCI, for a given computational cost.
 For closed- and open-shell systems, the EN2 perturbative correction
-substantially accelerates and stabilizes the convergence of hCI and eCI routes (while keeping the advantage of the former),
+substantially accelerates and stabilizes the convergence of hCI and eCI (while keeping the advantage of the former),
 though it ameliorates the sCI route to a lesser extent.
-The standard EN2 correction typically produced more accurate results than its renormalized form, except perhaps for describing single bond breaking.
-Overall, lower order models like hCI1+EN2 were found to be fairly accurate given its low computational cost.
+The standard EN2 correction typically produced more accurate results than its renormalized form, except perhaps for describing single-bond breaking.
+Overall, lower-order models (like hCI1+EN2) were found to be fairly accurate given their low computational cost.
 In the future, it may be worth combining orbital optimization at a lower level of hCI \cite{Kossoski_2022} with the correction provided by perturbation theory.
 
-We further gauged the performance of hCI to describe excited states of closed-shell and open-shell systems, based on either HF ground-state orbitals or state-specific orbitals.
-For excitations of closed-shells, $\Delta$hCI2 and $\Delta$hCI3 are comparable to their excitation-based counterparts, $\Delta$CISD and $\Delta$CISDT,
-whereas $\Delta$hCI1, $\Delta$hCI1.5 and $\Delta$hCI2.5 models are inaccurate bearing in mind their computational cost.
+We further gauged the performance of hCI to describe excited states of closed- and open-shell systems, based on either HF ground-state orbitals or state-specific orbitals.
+For excitations of closed-shell systems, $\Delta$hCI2 and $\Delta$hCI3 are comparable to their excitation-based counterparts, $\Delta$CISD and $\Delta$CISDT,
+whereas $\Delta$hCI1, $\Delta$hCI1.5, and $\Delta$hCI2.5 are inaccurate bearing in mind their computational cost.
 The poor performance at lower orders is ascribed to the different minimal references employed,
-a single $s=0$ determinant for the ground state and a single $s=2$ CSF for the excited state,
+a single seniority-zero determinant for the ground state and a single seniority-two CSF for the excited state,
 which introduces a strong bias on the classes of determinants accessed in the hCI calculations.
-$\Delta$hCI performs significantly better for the doublet transitions, when compared to the results for closed-shells.
-In this case, the reference of both ground and excited states comprises the same type of single $s=1$ determinant.
+$\Delta$hCI performs significantly better for the doublet transitions when compared to the results for closed shells.
+In this case, the reference of both ground and excited states comprises the same type of single seniority-one determinant.
 The advantage of using state-specific orbitals over ground-state ones depends on the choice of reference and the order in which the CI is truncated.
@@
 In contrast, for unlike references (as for the closed-shell excitations), the state-specific approach only becomes advantageous at somewhat higher orders ($\Delta$hCI2).
-The present findings highlight the challenge in describing on an equal footing states with qualitatively different characters, in particular ground and singly-excited states of closed-shell systems.
+The present findings highlight the challenge in describing, on an equal footing, states with qualitatively different characters, in particular ground and singly-excited states of closed-shell systems.
 
@@
 is very accurate for describing singly-excited states, \cite{Loos_2018,Loos_2020,Loos_2020a} albeit it slightly overestimates their excitation energies.
-In contrast, CISD provides too big excitation energies, \cite{Koch_1990,Kossoski_2023}
+In contrast, CISD provides too large excitation energies, \cite{Koch_1990,Kossoski_2023}
 even though both CISD and EOM-CCSD methods span the same excited determinants and rely on HF ground-state orbitals.
@@
 
-Finally, we employed different sCI models to compute excitation energies of closed-shell and open-shell systems.
+Finally, we employed different sCI models to compute excitation energies of closed- and open-shell systems.
 Despite the four different models employed for the closed-shell excitations, the results are overall disappointing.
 For the open-shell transitions, the outcome is more encouraging.
-The relevance of these sCI models lies in its possible connection with CC methods.
+The relevance of these sCI models lies in their possible connection with CC methods.
 While this is already established for closed-shell systems (pCCD and DOCI deliver very similar energies), \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967,Kossoski_2021,Marie_2021}

### 2023-08-20 Manuscript/manuscript.tex
hCI section

@@
 We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
-Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance \titou{between correlation of ground and excited states}.
+Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance between correlation of ground and excited states.
 State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.
@@
 The most well-known CI route is based on the excitation degree $e$.
-Starting from a reference determinant, typically the Hartree-Fock (HF) determinant,
-one generates all determinants that are connected to it by exciting at most $e$ electrons.
+Starting from a reference determinant, typically the Hartree-Fock (HF) determinant, one generates all connected determinants by exciting at most $e$ electrons.
 The excitation degree thus defines the order of the approximate excitation-based CI (eCI) model:
@@
 
-hCI was initially introduced for a closed-shell reference. \cite{Kossoski_2022}
+hCI was initially introduced for a closed-shell reference (left panel of Fig.~\ref{fig:hCI}). \cite{Kossoski_2022}
 Here, we generalize it for an arbitrary Slater determinant reference, including systems with an odd number of electrons.
@@
 where $s$ and $s_0$ denote the seniority of the candidate and reference determinants, respectively, and $e$ represents the excitation degree that connects them.
-The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with an even (odd) numbers of electrons. The excitation degree $e$ is also an integer.
-Therefore, $h$ can assume integer or half-integer values, for any type of reference determinant.
+The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with even (odd) numbers of electrons. The excitation degree $e$ is also an integer.
+Therefore, $h$ assumes integer or half-integer values, for any type of reference determinant.
 For a given hCI model defined by $h$, we include all the candidate determinants having a hierarchy less than or equal to $h$ with respect to any determinant in the reference.
-This definition allow us to build multireference hCI models as well.
-Notice that Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant, when $s_0 = 0$.
-hCI can be viewed as the CI model that provides increasingly more dissimilar determinants from the given reference determinant.
+This definition allows us to build multireference hCI models as well.
+Notice that Eq.~\eqref{eq:h} simplifies to the previous definition \cite{Kossoski_2022} for the case of a closed-shell reference determinant where $s_0 = 0$.
+hCI can be viewed as a CI model that includes increasingly more dissimilar determinants from a given reference determinant.
 The level of dissimilarity is represented by the hierarchy $h$, which accounts for differences in orbital occupation (through the excitation degree $e$)
-and differences in the number of unpaired electrons (through the term $(s-s_0)/2$).
+and differences in the number of unpaired electrons [through the term $(s-s_0)/2$].
 
-The definition in Eq.~\eqref{eq:h} guarantees that all classes of determinants sharing the same scaling of $\Ndet$ with $N$ enter at the same hierarchy $h$.
-% computational scaling
-We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference, and for open-shell references with one or two unpaired electrons.
-These references are suitable for describing the ground state of a closed-shell system, of an open-shell system, and a singly-excited state of a closed-shell, respectively.
-For these three types of reference, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
+\titou{The definition in Eq.~\eqref{eq:h} guarantees that all classes of determinants sharing the same scaling of $\Ndet$ with $N$ enter at the same hierarchy $h$.}
+We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference ($s_0 = 0$), and for open-shell references with one ($s_0 = 1$) or two ($s_0 = 2$) unpaired electrons.
+\titou{These references are suitable for describing ground states of a closed-shell system, of open-shell systems with doublet ground states, and (singlet or triplet) singly-excited states of closed-shell systems, respectively.}
+For these three types of references, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
 
@@
 \includegraphics[width=1.0\linewidth]{determinants}
-\caption{Classes of determinants generated in hCI0, hCI0.5 and hCI1, for three different reference determinants:
-closed-shell (top left), open-shell with one unpaired electron (top right) and open-shell with two unpaired electrons (bottom).
-Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied and unoccupied).}
+\caption{Classes of determinants generated in hCI0, hCI0.5, and hCI1, for three different reference determinants:
+closed-shell (top left), open-shell with one unpaired electron (top right), and open-shell with two unpaired electrons (bottom).
+\titou{Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied, and unoccupied).}}
 \label{fig:determinants}
@@
 
-For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the restricted HF determinant.
-The first non-trivial order is hCI1 (see left of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS)
-plus all paired doubled excitations (two electrons promoted from the same occupied orbital into the same virtual orbital),
-as shown in the top left of Fig.~\ref{fig:determinants}.
+For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the \titou{restricted HF} determinant.
+The first non-trivial order is hCI1 (see left panel of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS)
+plus all paired double excitations (two electrons promoted from the same occupied orbital into the same virtual orbital),
+as shown in the top left panel of Fig.~\ref{fig:determinants}.
 For both classes of determinants, their number scales as $N^2$, and are thus taken into account at the same hierarchy of hCI ($h=1$ in this case).
@@
 hCI1.5 augments the set of hCI1 determinants with the set of double excitations where two electrons are unpaired.
-At the next integer order, hCI2 incorporates all classes of determinants where $\Ndet$ scale as $N^4$.
+At the next integer order, hCI2 incorporates all classes of determinants where $\Ndet$ scales as $N^4$.
 In total, it accounts for all single and double excitations (as CISD), plus the subset of triple excitations that leave only two unpaired electrons,
@@
 Likewise, in eCI, $\Ndet$ scales as $N^{2e}$.
-The computational scaling along the two routes is shown in Table~\ref{tab:scaling}.
+\titou{The computational scaling} along the two routes is shown in Table~\ref{tab:scaling}.
 
 \begin{table}[ht!]
-\caption{Computational scaling in terms of the number of basis functions $N$ for the hierarchy-based and excitation-base CI routes, with and without the EN2 perturbative correction.}
+\caption{\titou{Computational scaling} in terms of the number of basis functions $N$ for the hierarchy- and excitation-based CI routes, with and without the EN2 perturbative correction.}
 \label{tab:scaling}
@@
 For an odd number of electrons, $s$ assumes odd values starting from $s=1$.
-The simplest reference is an open-shell determinant with one unpaired electron ($s_0=1$), shown in the center pannel of Fig.~\ref{fig:hCI} and in the top right of Fig.~\ref{fig:determinants}.
-hCI0 also reduces to the restricted open-shell HF determinant.
+The simplest reference is an open-shell determinant with one unpaired electron ($s_0=1$), shown in the center panel of Fig.~\ref{fig:hCI} and in the top right panel of Fig.~\ref{fig:determinants}.
+hCI0 also reduces to the \titou{restricted open-shell HF} determinant.
 In contrast to the closed-shell case, here there are no empty blocks in the seniority-excitation map.
 Actually, the hCI series displays the hCI0.5 level, accounting only for the single excitations from and into the singly-occupied orbital.
-For HF orbitals, however, these excitations do not connect with the reference, and hCI0.5 provides the same energy as restricted open-shell HF.
+For HF orbitals, however, these excitations do not connect with the reference. Thus, hCI0.5 provides the same energy as restricted open-shell HF.
 More generally, since there are more types of orbitals for the open-shell reference (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell reference (doubly-occupied or unoccupied),
@@
 
-Finally, one can employ a reference determinant having more unpaired electrons to define the hCI models.
-The case of two unpaired electrons ($s_0=2$) is shown in the right pannel of Fig.~\ref{fig:hCI},
-which would be appropriate for describing singly-excited states of closed-shells.
+Finally, one can employ a reference determinant having more unpaired electrons to define hCI models.
+The case of two unpaired electrons ($s_0=2$) is shown in the right panel of Fig.~\ref{fig:hCI},
+which would be appropriate for describing \titou{singlet and triplet} singly-excited states of closed-shells.
 The key difference in the hCI sequence, with respect to the closed-shell case, lies in the displacement by one block to the right in the seniority-excitation map,
-reflecting the shift from the $s_0=0$ to the $s_0=2$ reference.
-In addition, in the two previous cases, the hCI0 level only included the reference determinant.
-whereas for $s_0=2$, hCI0 not only accounts for the reference determinant,
+reflecting the shift from the seniority-zero to the seniority-two reference.
+In addition, in the two previous cases, the hCI0 level only included the reference determinant
+whereas, for $s_0=2$, hCI0 does not only account for the reference determinant
 but also for the two closed-shell determinants produced by the single excitation that pairs the two unpaired electrons.
 The hCI1 classes of determinants for the $s_0=2$ reference can be seen in the bottom of Fig.~\ref{fig:determinants},
-which clearly outnumber the fewer classes associated with the $s_0=0$ and $s_0=1$ cases.
+which clearly outnumbers the fewer classes associated with the $s_0=0$ and $s_0=1$ cases.
 One could adopt references with even larger $s_0$, further displacing the hCI sequence to the right in the seniority-excitation map.
 In this case, hCI0 would include all the lower spin determinants obtained by partially or totally pairing the unpaired electrons.
-Notice that the number of such determinants do not depend on the size of the system (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
-In a similar fashion, hCI can be built on top of a larger $s_0$ reference for odd number of electrons.
-The deduction for the number of determinants in a given hCI model, for the $s_0=0$ closed-shell, $s_0=1$ open-shell, as well as for an arbitrary reference determinant,
-can be found in Sec.~\ref{sec:appendix}.
+Notice that the number of such determinants does not depend on the system size (scaling as $N^0$), and are thus included at the $h=0$ level, in line with the spirit of hCI.
+In a similar fashion, hCI can be built on top of a larger $s_0$ reference for odd numbers of electrons.
+The deduction for the number of determinants in a given hCI model, for the $s_0=0$ closed-shell, $s_0=1$ open-shell, as well as for an arbitrary reference determinant, can be found in Appendix \ref{app:appendix}.
 
-For a reference containing unpaired electrons, an approximate CI model will generally produce spin-contaminated states.
-One can impose the solutions to have well-defined spin by including additional determinants generated via higher-order hierarchy, which account for the missing spin-flip configurations.
+For a reference containing unpaired electrons, an approximate CI model generally produces spin-contaminated states.
+One can impose the solutions to have a well-defined spin by including additional determinants generated via higher-order hierarchies, which account for the missing spin-flip configurations.
 The spin-contamination problem and its solution are therefore equivalent to that encountered in eCI. \cite{Maurice_1996}
@@
 By allowing only for the determinants that are connected with the reference determinant(s) up to a given maximum hierarchy $h$,
-the CIPSI algorithm is restricted to the truncated Hilbert space specified by the reference determinant(s) and the value of $h$.
-{\QP} \cite{Garniron_2019} was also employed to perform all the present eCI, sCI and FCI calculations.
+the CIPSI algorithm is restricted to the truncated Hilbert space specified by the reference determinant(s), and the value of $h$.
+{\QP} \cite{Garniron_2019} was also employed to perform all the present eCI, sCI, and FCI calculations.
 In a given calculation, the energies are considered to be converged when the (largest) EN2 correction computed in the truncated Hilbert space
 lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
-This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space,
+This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space
 while delivering fairly converged energies.
@@
 The CI calculations employed the ground-state restricted open-shell HF orbitals, described with the cc-pVDZ basis set, and within the frozen core approximation.
-For such small systems and basis set, FCI is attainable and provides the reference results for gauging the approximate CI models.
+For such small systems and basis sets, FCI is attainable and provides the reference results for gauging the approximate CI models.
 The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020}, and is also reproduced in the {\SupInf}.
 Their PECs were computed along the \ce{C=C} double bond breaking coordinate, with the remaining internal coordinates kept frozen.
-For \ce{H7}, we considered equally spaced and linearly arranged hydrogen atoms, and the PECs were computed along the symmetric dissociation coordinate.
+For \ce{H7}, we considered equally spaced and linearly arranged hydrogen atoms and the PECs were computed along the symmetric dissociation coordinate.
 
-The results were analyzed along the same lines of our previous report on hCI for closed-shell systems. \cite{Kossoski_2022}
+The results were analyzed along the same lines as our previous report on hCI for closed-shell systems. \cite{Kossoski_2022}
 Namely, for the different CI models considered here,
@@
 The distance error is redefined based on the signed differences between two PECs, as the absolute value of their maximum difference plus the absolute value of their minimum difference,
-evaluated at given coordinate interval.
+evaluated at a given coordinate interval.
 This new definition measures how close are two PECs, remaining always non-negative.
@@
 We impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies in accounting for a set of appropriate spin-flipped determinants stemming from higher-order excitation or hierarchy.
-In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS \cite{Maurice_1996} for instance, and equivalently for the other hCI, eCI and sCI models.
+In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS \cite{Maurice_1996} for instance, and equivalently for the other hCI, eCI, and sCI models.
 Notice that spin-contaminated solutions would have different energies than the spin eigenstates considered here.
 
-We performed calculations following both the standard ground-state based CI route and the state-specific CI route. \cite{Kossoski_2023}
+We performed calculations following both the standard ground-state-based CI route and the state-specific CI route. \cite{Kossoski_2023}
 For the latter, we employed the state-specific orbitals obtained in Ref.~\onlinecite{Kossoski_2023}.
@@
 Further including the $s=2$ sector for the ground state calculations gives rise to the sCI2/sCI2 model.
-From here on, CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it correspond to a ground-state based approach.
+From here on, CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it correspond to a ground-state-based approach.
 
@@
 \section{Number of determinants}
-\label{sec:appendix}
+\label{app:appendix}
 

### 2023-08-20 Manuscript/manuscript.tex
hCI section
comp det

@@
 \titou{The computational scaling} along the two routes is shown in Table~\ref{tab:scaling}.
+\PFL{Should we mention something about the scaling of seniority-based CI here?}
 
@@
 One can impose the solutions to have a well-defined spin by including additional determinants generated via higher-order hierarchies, which account for the missing spin-flip configurations.
+\cite{Chilkuri_2021}
 The spin-contamination problem and its solution are therefore equivalent to that encountered in eCI. \cite{Maurice_1996}
@@
 For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^8}$,
-though with a small prefactor stemming from the EN2 calculation.
+though with a small prefactor stemming from the EN2 calculation that employs a very efficient semistochastic algorithm. \cite{Garniron_2018}
 The computational scaling associated with the CI+EN2 calculations is also presented in Table~\ref{tab:scaling}.
 
-To gauge the performance of hCI, eCI and sCI for radical systems,
-we have calculated the ground-state potential energy curves (PECs) for the dissociation of four radicals:
+To gauge the performance of hCI, eCI, and sCI for radicals, we have calculated the ground-state potential energy curves (PECs) for the dissociation of four radicals:
 \ce{OH}, \ce{CN}, vinyl (\ce{C2H3}), and \ce{H7}.
-The CI calculations employed the ground-state restricted open-shell HF orbitals, described with the cc-pVDZ basis set, and within the frozen core approximation.
+The CI calculations employed the ground-state restricted open-shell HF orbitals, described with the cc-pVDZ basis set, and within the frozen-core approximation.
 For such small systems and basis sets, FCI is attainable and provides the reference results for gauging the approximate CI models.
-The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020}, and is also reproduced in the {\SupInf}.
+The equilibrium geometry of vinyl was taken from Ref.~\onlinecite{Loos_2020} and is also reproduced in the {\SupInf}.
 Their PECs were computed along the \ce{C=C} double bond breaking coordinate, with the remaining internal coordinates kept frozen.
@@
 we evaluated the convergence of the non-parallelity error (NPE), the distance error, the harmonic vibrational frequencies, and the equilibrium bond lengths, as functions of $\Ndet$.
-The NPE of a given level of theory is defined as the maximum minus the minimum differences between its corresponding PEC and the FCI PEC, for a given range of coordinates.
+The NPE of a given level of theory is defined as the maximum minus the minimum energy differences between its corresponding PEC and the FCI PEC, for a given range of coordinates.
 Here, we redefine the previous definition of the distance error \cite{Kossoski_2022} to accommodate for the fact the approximate PEC might appear below the FCI one when perturbative corrections are employed.
@@
 We employed the aug-cc-pVDZ basis set for systems having up to three non-hydrogen atoms and the 6-31+G(d) basis set for the larger ones.
-Core orbitals were frozen.
-We impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies in accounting for a set of appropriate spin-flipped determinants stemming from higher-order excitation or hierarchy.
+Core orbitals were frozen systematically.
+We impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies accounting for a set of appropriate spin-flipped determinants stemming from higher-order excitations or hierarchies.
+\cite{Chilkuri_2021}
 In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS \cite{Maurice_1996} for instance, and equivalently for the other hCI, eCI, and sCI models.
@@
 The computed excitation energies were benchmarked against the reference values of FCI quality provided in the QUEST database. \cite{Veril_2021}
-The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as the reference determinant for the hCI and eCI ground state calculations
-(left and middle pannels of Fig.~\ref{fig:hCI}).
+The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as the reference determinant for the hCI and eCI ground-state calculations
+(left and middle panels of Fig.~\ref{fig:hCI}).
 For the state-specific hCI and eCI calculations, we employed a minimal CSF reference: \cite{Kossoski_2023}
 a single open-shell determinant for the doublet excited states and a single open-shell CSF for the excited states from closed-shell systems
-(middle and right pannels of Fig.~\ref{fig:hCI}, respectively).
+(middle and right panels of Fig.~\ref{fig:hCI}, respectively).
 
 As for the sCI models,
-we considered sCI1 for both ground and (state-specific) excited state calculations for the open-shell systems (odd number of electrons).
-For the closed-shell systems (even number of electrons), we explored two models.
+we considered sCI1 for both ground and (state-specific) excited-state calculations for the open-shell systems (odd numbers of electrons).
+For the closed-shell systems (\titou{even numbers of electrons}), we explored two models.
 In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal sCI calculation for computing singly-excited states of closed-shell systems.
-Further including the $s=2$ sector for the ground state calculations gives rise to the sCI2/sCI2 model.
+Further including the seniority-two sector for the ground-state calculations defines the sCI2/sCI2 model.
 From here on, CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it correspond to a ground-state-based approach.

### 2023-08-20 Manuscript/manuscript.tex
radicals

@@
 The excitation degree thus defines the order of the approximate excitation-based CI (eCI) model:
-CI with single excitations (CIS), CI with single and double excitations (CISD), CI with single, double, and triple excitations (CISDT), and so on.
+CI with single excitations (CIS), CI with single and double excitations (CISD), CI with single, double, and triple excitations (CISDT), etc.
 The eCI route rather quickly captures dynamic (weak) correlation, though it struggles with the description of static (strong) correlation.
@@
 For systems with an even number of electrons, the first approximate model is defined by $s=0$ (sCI0), usually referred to as doubly-occupied CI (DOCI),
-which is followed by the higher-order models: sCI2, sCI4, and so on.
+which is followed by the higher-order models: sCI2, sCI4, \ldots.
 For odd numbers of electrons, the seniority route follows along odd numbers of $s$: sCI1, sCI3, and so on.
@@
 The full set of PECs for the open-shell systems (\ce{OH}, \ce{CN}, \ce{H7}, and vinyl) are presented in the {\SupInf}.
-From these PECs, we obtain the NPEs, distance errors, equilibrium bond lengths and harmonic vibrational frequencies, which are plotted as a function of $\Ndet$ in Fig.~\ref{fig:plot_all}.
+From these PECs, we obtain the NPEs, distance errors, equilibrium bond lengths, and harmonic vibrational frequencies, which are plotted as functions of $\Ndet$ in Fig.~\ref{fig:plot_all}.
 Each point in the figure denotes one CI model.
-For instance, the hCI route is represented by the dark green line, with the first point being HF, the second being hCI1, the third being hCI1.5, and so on.
+For instance, the hCI route is represented by the dark green line, with the first point being HF \titou{(or hCI0)}, the second being hCI1, the third being hCI1.5, etc.
 The corresponding results for the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) are shown in Fig.~\ref{fig:plot_all_closed}.
@@
 These previous results are reproduced in Fig.~\ref{fig:plot_all_closed} together with the present results, which include such correction.
-First we will discuss the results for the bare CI models (hCI, eCI and sCI) for the open-shell systems (represented by the dark tones in the figures),
+First, we discuss the results for the bare CI models (hCI, eCI, and sCI) for the open-shell systems (represented by the dark tones in the figures),
 leaving the discussion about the perturbative correction (light tones) for both open-shell and closed-shell systems to the next subsection.
@@
 Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
-as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red), and sCI (blue) models,
 with (light tones) and without (dark tones) the standard EN2 perturbative correction.
@@
 \caption{
-Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8},
-as a function of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red) and sCI (blue) models,
 with (light tones) and without (dark tones) the standard EN2 perturbative correction.
@@
 
-We found that hCI typically improves the convergence with respect to $\Ndet$ when compared to eCI, for the four properties surveyed here.
-The present finding for open-shell systems therefore parallels the previously reported superiority of hCI for closed-shell systems. \cite{Kossoski_2022}
+We find that hCI typically improves the convergence with respect to $\Ndet$ when compared to eCI, for the four properties surveyed here.
+The present finding for open-shell systems, therefore, parallels the previously reported superiority of hCI for closed-shell systems. \cite{Kossoski_2022}
 
@@
 The smaller NPEs attained with hCI are clear for \ce{OH} (involving a single bond breaking),
-vinyl (double bond breaking) and \ce{CN} (triple bond breaking),
+vinyl (double bond breaking), and \ce{CN} (triple bond breaking),
 whereas for \ce{H7} (multiple bond breaking), hCI and eCI are comparable.
@@
 Even though hCI manages to recover more static correlation than eCI,
-the advantage of the former is more striking for single bond breaking, whereas multiple bond breaking (like \ce{H7} and \ce{H8}) inevitably requires higher order excited determinants.
+the advantage of the former is more striking for single bond breaking, whereas multiple bond breaking (like \ce{H7} and \ce{H8}) inevitably requires higher-order excited determinants.
 
@@
 In great contrast to hCI and eCI, sCI models deliver a poor convergence of all observables.
-The single exception is the NPE for \ce{OH} obtained from sCI0, in-between the NPEs obtained from hCI2 and CISDT, methods having a comparable computational cost in this particular case.
+The single exception is the NPE for \ce{OH} obtained from sCI0, in between the NPEs obtained from hCI2 and CISDT, methods having a comparable computational cost in this particular case.
 Going to larger basis sets or bigger systems, the computational burden of sCI models would increase considerably more than hCI or eCI models,
@@
 in line with similar findings for closed-shell systems \cite{Kossoski_2022} (also see Fig.~\ref{fig:plot_all_closed}).
+\PFL{Shall we comment on orbital optimization here?}
 
@@
 
-We put forward the interesting perspective of developing hierarchy-based CC (hCC) methods (hCC1, hCC2, and so on),
-along with derived EOM-CC formulations to target excited states (EOM-hCC1, EOM-hCC2, and so on).
+We put forward the interesting perspective of developing hierarchy-based CC (hCC) methods (hCC1, hCC2, \ldots),
+along with derived EOM-CC formulations to target excited states (EOM-hCC1, EOM-hCC2, \ldots).
 EOM-CC with single and double excitations (EOM-CCSD) \cite{Rowe_1968,Monkhorst_1977,Koch_1990,Stanton_1993}

### 2023-08-20 Manuscript/manuscript.tex
EN2

@@
 
-The results for the CI models corrected by the standard EN2 energy are represented by the light tones in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
+The results for the CI models corrected by the standard EN2 energy are represented by the light tones in Figs.~\ref{fig:plot_all} and \ref{fig:plot_all_closed}.
 We excluded instances where the PECs present important discontinuities.
 These discontinuities reflect crossings between states of different symmetries in the unperturbed CI calculation,
-appearing as kinks in the computed ground-state PEC.
+appearing as kinks in the computed ground-state PEC \titou{(intruder states)}.
 Due to the abrupt change of character, the EN2 correction is not uniform and gives rise to the observed discontinuities.
@@
 For \ce{OH} and the six closed-shell systems, the EN2 correction produced smooth PECs at dissociation.
-Around the equilibrium geometry, the EN2 correction also leads to well-behaved PECs, for all CI models and for all systems, except for \ce{CN}
+Around the equilibrium geometry, the EN2 correction also leads to well-behaved PECs, for all CI models, and for all systems, except for \ce{CN}
 (thus no equilibrium properties are presented for this radical).
@@
 For instance, hCI1+EN2 and hCI2 display the same $\order*{N^6}$ computational scaling, and the former outperforms the latter in many cases.
-In Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}, hCI1+EN2 correlates with the second light green mark, and hCI2 with the fourth dark green mark.
-The former model provides smaller NPEs and distance errors for \ce{OH}, \ce{HF}, \ce{F2} (all single bond breaking),
+In Figs.~\ref{fig:plot_all} and \ref{fig:plot_all_closed}, hCI1+EN2 correlates with the second light green mark, and hCI2 with the fourth dark green mark.
+The former model provides smaller NPEs and distance errors for \ce{OH}, \ce{HF}, \ce{F2} (all single-bond breakings),
 the opposite being true for multiple bond breaking.
@@
 whereas both models perform similarly for the remaining systems.
-It is worth recalling that the prefactor associated with the EN2 calculation is smaller than that associated with the higher-order CI calculation,
+It is worth recalling that the prefactor associated with the EN2 calculation is smaller than the one associated with the higher-order CI calculation,
 which, in the above example, makes hCI1+EN2 cheaper than hCI2.
@@
 CISD+EN2 correlates with the third dark red mark and CISDT with the fourth light red mark in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed}.
-The errors for the four observables are systematically smaller with the former model,
+The errors for the four observables are systematically smaller than the former model,
 except for the NPE and distance error of \ce{H8} and the equilibrium properties of \ce{H7}, where they are comparable.
@@
 This is seen when employing the hCI1 and hCI1+EN2 models to compute the NPE of \ce{OH} and \ce{F2}, and to some extent the equilibrium properties of \ce{F2}.
-These are exceptions though, as the EN2 correction practically always improve the accuracy of the computed observables.
-It further leads to overall more monotonic convergence of the observables.
-In this sense, it regularizes oscillations seen in the unperturbed case, probably related to cancellation of errors at the lower orders.
+These are exceptions though, as the EN2 correction practically always improves the accuracy of the computed observables.
+It further leads to an overall more monotonic convergence of the observables.
+In this sense, it regularizes oscillations seen in the unperturbed case, probably related to the cancellation of errors at the lower orders.
 This is clearly seen in the equilibrium properties of vinyl, \ce{H7}, and \ce{F2},
@@
 
-Instead of the usual EN2 correction, discussed so far and shown in Figs.~\ref{fig:plot_all} and ~\ref{fig:plot_all_closed},
+Instead of the usual EN2 correction, discussed so far and shown in Figs.~\ref{fig:plot_all} and \ref{fig:plot_all_closed},
 one can compute the renormalized EN2 correction. \cite{Garniron_2019}

### 2023-08-20 Manuscript/manuscript.tex
hCI for excited states

@@
 Likewise, in eCI, $\Ndet$ scales as $N^{2e}$.
-\titou{The computational scaling} along the two routes is shown in Table~\ref{tab:scaling}.
+\titou{The computational scaling} along the two routes is shown in Table \ref{tab:scaling}.
 \PFL{Should we mention something about the scaling of seniority-based CI here?}
@@
 \begin{table}[ht!]
-\caption{\titou{Computational scaling} in terms of the number of basis functions $N$ for the hierarchy- and excitation-based CI routes, with and without the EN2 perturbative correction.}
+\caption{\titou{Computational Scaling} in Terms of the Number of Basis Functions $N$ for the Hierarchy- and Excitation-Based CI Routes, With and Without the EN2 Perturbative Correction.}
 \label{tab:scaling}
@@
 though with a small prefactor stemming from the EN2 calculation that employs a very efficient semistochastic algorithm. \cite{Garniron_2018}
-The computational scaling associated with the CI+EN2 calculations is also presented in Table~\ref{tab:scaling}.
+The computational scaling associated with the CI+EN2 calculations is also presented in Table \ref{tab:scaling}.
 
@@
 
-For each CI model considered here, we evaluated the mean signed error (MSE), mean absolute error (MAE), and root-mean square error (RMSE)
+For each CI model considered here, we evaluate the mean signed error (MSE), mean absolute error (MAE), and root-mean-square error (RMSE)
 with respect to the reference theoretical values for the excitation energies.
-For the lower order models, the calculations were performed for 48 (closed-shell) and 19 (open-shell) excited states.
-In turn, subsets of 16 (closed-shell) and 6 (open-shell) excited states were considered for the higher order CI models, given the more intensive computational cost.
+For the lower-order models, the calculations were performed for 48 (closed-shell) and 19 (open-shell) excited states.
+In turn, subsets of 16 (closed-shell) and 6 (open-shell) excited states were considered for the higher-order CI models, given the more intensive computational cost.
 Even though these subsets are too small for meaningful absolute statistics, they should be enough to reveal main trends.
 A detailed comparison of eCI based on ground-state and state-specific approaches can be found elsewhere. \cite{Kossoski_2023}
-The focus of the present discussion lies on the comparison between hCI and eCI, and sCI routes,
-and on a similar comparison between ground-state and state-specific approaches for the hCI and sCI routes.
+The focus of the present discussion lies on the comparison between hCI and eCI, and sCI,
+and on a similar comparison between ground-state and state-specific approaches for hCI and sCI.
 
-We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Table~\ref{tab:stat1}.
+We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Table \ref{tab:stat1}.
 The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.07}{\eV}.
@@
 At the hCI3 level, significantly smaller errors are finally achieved (MAE of \SI{0.19}{\eV}).
-This is close to that obtained with the eCI model of same order, CISDT (\SI{0.17}{\eV}).
+This is close to that obtained with the eCI model of the same order, CISDT (\SI{0.17}{\eV}).
 Even smaller errors are produced at the hCI3.5 level (MAE of \SI{0.11}{\eV}), but at a considerable computational cost.
@@
 As discussed above, the errors on excitation energies obtained with the hCI models first increase as one augments the hierarchy parameter $h$, reaching a maximum at hCI2, and then decrease towards higher orders.
-This behavior parallels what is well stablished for the excitation energies computed with eCI, where CISD is far worse than CIS, which in turn is inferior to CISDT, whereas CISDTQ is considerably more accurate.
+This behavior parallels what is well established for the excitation energies computed with eCI, where CISD is far worse than CIS, which in turn is inferior to CISDT, whereas CISDTQ is considerably more accurate.
 This can be understood based on the role of the excited determinants for ground and excited states.
@@
 This favored description of the ground state explains the overestimated excitation energies at these orders.
-The effect becomes exaggerated at the hCI2 (and CISD) models, since it captures most of the ground state correlation through the unpaired double excitations,
+The effect becomes exaggerated at the hCI2 (and CISD) level since it captures most of the ground state correlation through the unpaired double excitations,
 which in turn just start to describe correlation for the excited state.
-Higher orders models, starting at hCI3 and CISDT, are needed to recover a large fraction of the excited state correlation, and at this point the errors on the excitation energies decline progressively.
-Because the excited determinants are accessed from the ground-state determinant and due to ground-state HF orbitals, the description of the ground state is always remain favoured.
+Higher-order models, starting at hCI3 and CISDT, are needed to recover a large fraction of the excited-state correlation, and at this point, the errors on the excitation energies decline progressively.
+Because the excited determinants are accessed from the ground-state determinant and due to ground-state HF orbitals, the description of the ground state is always favored.
 This explains why hCI systematically overestimates the excitation energies, just as eCI does.
@@
 Employing state-specific orbitals would be expected to suppress the bias toward the ground state and lead to improved results, as observed, for instance, when going from CISD to $\Delta$CISD \cite{Kossoski_2023}.
-However, for the lower orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state based approach, this time by underestimating the excitation energies.
+However, for the low orders of state-specific hCI ($\Delta$hCI1 and $\Delta$hCI1.5), we actually found larger errors than with the ground-state-based approach, this time by underestimating the excitation energies.
 Besides the set of orbitals, the classes of determinants included at each order play an equally important role and explains our observation.
 Still considering the excitations of the closed-shell systems,
-different classes of determinants are accessed from the ground state reference (a closed-shell determinant) and from the excited state reference (a single open-shell CSF),
+different classes of determinants are accessed from the ground-state reference (a closed-shell determinant) and from the excited-state reference (a single open-shell CSF),
 for a given hierarchy parameter $h$.
@@
 There is far more diversity in the classes of determinants employed for the excited-state calculation than for the ground-state one, a consequence of the different references.
-Because of that, lower orders of hCI may be expected to capture a larger fraction of correlation for the excited state than for ground state.
+Because of that, low orders of hCI may be expected to capture a larger fraction of correlation for the excited state than for ground state.
 This is further supported by the more accurate results obtained for excitations of open-shell systems (discussed in detail later), where the same type of reference is employed.
 Although hCI is constructed to account for all classes of determinants whose number share the same computational scaling,
-at the lower orders this procedure does not lead to a balanced description of correlation, at least for excitations of closed-shell systems and the present choice of minimal references.
-An important point to realize from the present results is that a state-specific approach is not necessarily more accurate than a ground-state based approach.
+at the lower orders, this procedure does not lead to a balanced description of correlation, at least for excitations of closed-shell systems and the present choice of minimal references.
+An important point to realize from the present results is that a state-specific approach is not necessarily more accurate than a ground-state-based approach.
 Both the orbitals and the classes of determinants of a given CI model control the fine balance of ground- and excited-state correlation effects.
@@
 At this level, the state-specific advantage clearly manifests, with $\Delta$hCI2 having a MAE of \SI{0.20}{\eV}, substantially smaller than the MAE of \SI{3.53}{\eV} obtained from hCI2.
-This comparison between ground-state based and state-specific approaches is analogous to our previous finding on the eCI models of the same order, CISD and $\Delta$CISD. \cite{Kossoski_2023}
+This comparison between ground-state-based and state-specific approaches is analogous to our previous finding on the eCI models of the same order, CISD, and $\Delta$CISD. \cite{Kossoski_2023}
 Even though hCI2 is somewhat more accurate than CISD (both have large errors), their state-specific versions share a comparable performance.
@@
 This could be due to another set of unbalanced state-specific determinants that enters at this stage, likewise to what was observed between $\Delta$hCI1 and $\Delta$hCI1.5.
-Despite the differing number of states considered, the observed variation on the accuracy of $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant,
+Despite the differing number of states considered, the observed variation in the accuracy of $\Delta$hCI2 and $\Delta$hCI2.5 is statistically significant,
 which is confirmed by calculating the statistical errors for the same subset of excited states.
@@
 likewise to what had been found for CISDT and $\Delta$CISDT. \cite{Kossoski_2023}
-We notice, however, that the state-specific route presents negative MSEs, in contrast to the positives values obtained with the ground-state based route, for both hCI and eCI.
-Furthermore, eCI and hCI present comparable performances at this order, with an arguable preference of the former,
+We notice, however, that the state-specific route presents negative MSEs, in contrast to the positive values obtained with the ground-state-based route, for both hCI and eCI.
+Furthermore, eCI and hCI present comparable performances at this order, with an arguable preference for the former,
 as also discussed above for $\Delta$CISD and $\Delta$hCI2.
@@
 We now shift to the discussion of excitations for open-shell systems.
-The statistical errors are shown in Table~\ref{tab:stat2}.
+The statistical errors are shown in Table \ref{tab:stat2}.
 The key difference on the calculations for open-shell excitations is that the same type of reference was employed for ground- and excited-state calculations, namely,
@@
 When using ground-state orbitals, we found the performance initially degrades and later improves as the order increases, for both eCI and hCI.
-This is similar to what was observed for the closed-shell excitations,
+This is similar to what was observed for the closed-shell excitations
 and can be traced back to a biased description of the ground state related to the choice of ground-state orbitals.
@@
 
-Once state-specific orbitals are employed, there is no remaining bias towards the ground state.
-In a stark contrast to the closed-shell case, the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.34}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.46}{\eV}),
-become more accurate than their ground-state based counterparts, hCI1 (MAE of \SI{0.77}{\eV}) and hCI1.5 (MAE of \SI{1.11}{\eV}).
+Once state-specific orbitals are employed, there is no remaining bias toward the ground state.
+In stark contrast to the closed-shell case, the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.34}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.46}{\eV}),
+become more accurate than their ground-state-based counterparts, hCI1 (MAE of \SI{0.77}{\eV}) and hCI1.5 (MAE of \SI{1.11}{\eV}).
 Once again, this is thanks to the same type of reference in ground- and excited-state calculations.
-The case of $\Delta$hCI1 for open-shell excitation is depicted in Fig.~\ref{fig:determinants}, in comparison to the more unbalanced case of excitation from closed-shells.
+The case of $\Delta$hCI1 for open-shell excitations is depicted in Fig.~\ref{fig:determinants}, in comparison to the more unbalanced case of excitations from closed-shells.
 We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5),
@@
 (Accounting only for the 6 states considered for $\Delta$hCI2.5, $\Delta$hCI2 would also have a MAE of \SI{0.09}{\eV}.)
-Likewise, higher order hCI models become progressively more accurate.
+Likewise, higher-order hCI models become progressively more accurate.
 Even though the statistics become more limited,
-they suggest both a small advantage of the hCI models against eCI,
+they suggest both a small advantage of the hCI models against eCI
 and that the ground-state routes perform slightly better than the state-specific ones.
@@
 are not invariant under orbital rotations within the occupied and virtual subspaces. \cite{Bytautas_2011,Limacher_2014,Stein_2014}
-This applies for the hCI and sCI models addressed here,
+This applies to the hCI and sCI models addressed here,
 which employed HF ground-state orbitals or excited-state orbitals optimized for a minimal reference.
@@
 
-We refer back to Table~\ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
+We refer back to Table \ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
 The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.34}{\eV}.
@@
 The situation for open-shell systems is quite different, in close analogy to the previous discussion regarding hCI.
-As shown in Table~\ref{tab:stat2}, sCI1 provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
+As shown in Table \ref{tab:stat2}, sCI1 provides considerably more accurate excitation energies for open-shell than for closed-shell systems.
 The reason should not be surprising at this point.

### 2023-08-20 Manuscript/manuscript.tex
sCI for excited states

@@
 The same two factors discussed above for hCI explain such large errors.
-First, the ground state description is favored due to the use of ground-state HF orbitals.
+First, the ground-state description is favored due to the use of ground-state HF orbitals.
 Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAE to \SI{0.79}{\eV}, though still overestimating the excitation energies.
-Second, there is an unbalanced description of correlation for ground and excited states, associated with the classes of determinants.
+Second, there is an unbalanced description of the correlation for ground and excited states, associated with the classes of determinants.
 While sCI2/sCI2 accounts for an additional $s=2$ sector for the ground state (which can be qualitatively described in the $s=0$ sector),
-no addition pair of electrons is allowed to become unpaired in the excited state calculation, even though the state is qualitatively described by a determinant that is contained in the $s=2$ sector.
+no additional pairs of electrons are allowed to become unpaired in the excited state calculation, even though the state is qualitatively described by a determinant that is contained in the $s=2$ sector.
 In other words, unpaired excitations are allowed to correlate the ground state, but not the excited state, given their respective closed-shell and open-shell characters,
@@
 However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{0.99}{\eV}.
-Clearly, the $s=2$ sector captures more correlation for the excited states than the $s=0$ does for the ground state.
-Ultimately, we did not find a combination of sCI model and orbitals that produced reasonable excitation energies for closed-shell systems.
+Clearly, the seniority-two sector captures more correlation for the excited states than the seniority-zero does for the ground state.
+Ultimately, we did not find a combination of sCI models and orbitals that produced reasonable excitation energies for closed-shell systems.
 
@@
 
-Even though $\Delta$sCI1 is less accurate and has less favourable computational scaling than CI models like $\Delta$CISD and $\Delta$hCI2,
+Even though $\Delta$sCI1 is less accurate and has less favorable computational scaling than CI models like $\Delta$CISD and $\Delta$hCI2,
 its decent errors are encouraging for another reason:
@@
 for both ground \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967} and excited \cite{Kossoski_2021,Marie_2021} states.
-Likewise, a formulation of pCCD to open-shell systems might share a analogous connection to low-order sCI models like sCI1.
+Likewise, a formulation of pCCD to open-shell systems might share an analogous connection to low-order sCI models like sCI1.
 If that is the case, then a state-specific approach of such pCCD formulation adapted to open-shells may approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.27}{\eV}) at a mean-field cost.
@@
 To the best of our knowledge, there has been a single proposed extension of pCCD to open-shell systems, \cite{Boguslawski_2021}
-based on the ionization-potential equation-of-motion CC (EOM-CC) formalism.
+based on the ionization-potential equation-of-motion CC (EOM-CC) formalism. \cite{Stanton_1994,Muneaki_2006,Bomble_2005}
 

### 2023-08-20 Manuscript/manuscript.tex
appendix

@@
 
-\newcommand{\Na}{N_{\alpha}}
-\newcommand{\Nb}{N_{\beta}}
-\newcommand{\Nd}{N^{d}}
-\newcommand{\Nv}{N^{v}}
-\newcommand{\Nas}{N^s_{\alpha}}
-\newcommand{\Nbs}{N^s_{\beta}}
+\newcommand{\Na}{N^{\uparrow}}
+\newcommand{\Nb}{N^{\downarrow}}
+\newcommand{\Nd}{N_\text{d}}
+\newcommand{\Nv}{N_\text{v}}
+\newcommand{\Nas}{N_\text{s}^{\uparrow}}
+\newcommand{\Nbs}{N_\text{s}^{\downarrow}}
 
@@
 total energies and excitation energies obtained with
-the various hCI, eCI, and sCI models considered here, in both ground-state based and state-specific approches;
+the various hCI, eCI, and sCI models considered here, in both ground-state-based and state-specific approaches;
 number of determinants in the reference; saddle point order associated with the $\Delta$CSF solutions;
 and the reference excitation energies and corresponding methods.
-For a subset of 16 excited states for closed-shell molecules and 6 excited states for open-shell radicals,
+For a subset of 16 excited states for closed-shell molecules and 6 excited states for radicals,
 additional total energies and excitation energies obtained with higher-order hCI and eCI models.
@@
 Finally, we deduce the general case.
-The reference determinant of each case is shown in Fig.~\ref{fig:appendix}.
+The reference determinant for each case is shown in Fig.~\ref{fig:appendix}.
 
@@
 \caption{
-$s_0=0$ closed-shell determinant with $N$ electrons (left),
-$s_0=1$ open-shell determinant with $\Na$ $\alpha$-electrons and $\Nb=\Na-1$ $\beta$-electrons (center),
-and general $s_0=\Nas+\Nbs$ determinant, with $\Nd$ doubly-occupied orbitals, $\Nas$ singly-occupied $\alpha$-electrons and $\Nbs$ singly-occupied $\beta$-electrons (right).
+Closed-shell determinant ($s_0=0$) with $N$ electrons (left),
+open-shell determinant ($s_0=1$) with $\Na$ spin-up electrons and $\Nb=\Na-1$ spin-down electrons (center),
+and general determinant ($s_0=\Nas+\Nbs$), with $\Nd$ doubly-occupied orbitals, $\Nas$ singly-occupied spin-up electrons and $\Nbs$ singly-occupied spin-down electrons (right).
 In the three cases, there are $K$ spatial orbitals.
 The number of determinants generated from these three reference determinants, from left to right, which have excitation degree $e$ and seniority $s$
-are given in Eqs.~\eqref{eq:det_1}, ~\eqref{eq:det_2}, and ~\eqref{eq:det_3}.
+are given in Eqs.~\eqref{eq:det_1}, \eqref{eq:det_2}, and \eqref{eq:det_3}.
 }
@@
 based on whether the excitation increases or reduces the seniority.
-For the decomposition, we make use of the Vandermonde's identity,
+For the decomposition, we make use of Vandermonde's identity,
 \begin{equation}
@@
 
-We start with the simpler case of a closed-shell determinant with $N$ electrons and $K$ spatial orbitals (left of Fig.~\ref{fig:appendix}).
-From this reference determinant, the number of excited determinants generated by exciting exactly $e$ electrons, is given by \cite{SzaboBook}
+We start with the simpler case of a closed-shell determinant with $N$ electrons and $K$ spatial orbitals (left panel of Fig.~\ref{fig:appendix}).
+From this reference determinant, the number of excited determinants generated by exciting $e$ electrons, is given by \cite{SzaboBook}
 \begin{equation}
@@
 \end{equation}
-where the sum expresses the different combinations for exciting $\alpha$- or $\beta$-electrons,
-the first binomial term accounts for the number of possibilities for exciting $p$ $\alpha$-electrons from the $N/2$ orbitals,
+where the sum expresses the different combinations for exciting spin-up or spin-down electrons,
+the first binomial term accounts for the number of possibilities for exciting $p$ spin-up electrons from the $N/2$ orbitals,
 the second term represents the number of ways of placing these $p$ electrons into $K-N/2$ orbitals,
-and similarly for the latter two terms and the $e-p$ remaining $\beta$-electrons.
+and similarly for the latter two terms and the $e-p$ remaining spin-down electrons.
 
@@
 while keeping track of the previously excited electrons.
-Starting from the closed-shell determinant, $p$ $\alpha$-electrons are excited, which increases the seniority by $2p$
-(a factor of $p$ due to the unpaired $\alpha$-electrons just excited,
-and another factor of $p$ due to the unpaired $\beta$-electrons left behind).
-The first two binomial terms in Eq.~\ref{eq:e_closed} are left untouched since they are always accompanied by the same change in seniority.
-Next, the $\beta$-electrons can be excited from the same orbitals from which the $\alpha$-electrons were excited (decreasing the seniority by one)
+Starting from the closed-shell determinant, $p$ spin-up electrons are excited, which increases the seniority by $2p$
+(a factor of $p$ due to the unpaired spin-up electrons just excited,
+and another factor of $p$ due to the unpaired spin-down electrons left behind).
+The first two binomial terms in Eq.~\eqref{eq:e_closed} are left untouched since they are always accompanied by the same change in seniority.
+Next, the spin-down electrons can be excited from the same orbitals from which the spin-up electrons were excited (decreasing the seniority by one)
 or instead from an orbital that remained doubly-occupied (increasing the seniority by one).
-The two possibilities are expressed by decomposing the third binomial term in Eq.~\ref{eq:e_closed} as a sum over the two corresponding binomials,
+The two possibilities are expressed by decomposing the third binomial term in Eq.~\eqref{eq:e_closed} as a sum over the two corresponding binomials,
 \begin{equation}
@@
 \end{equation}
-The first counts the number of possibilities for exciting $q$ $\beta$-electrons from one of $p$ orbitals for which an $\alpha$-electron was already excited.
-By removing the unpaired $\beta$-electron left behind, the seniority thus decreases by a factor of $q$.
-The second term accounts for the complementary excitations, where the $(e-p-q)$ $\beta$-electrons are chosen out of the $N/2-p$ orbitals that 
-remained doubly-occupied after exciting the $p$ $\alpha$-electrons.
+The first binomial counts the number of possibilities for exciting $q$ spin-down electrons from one of $p$ orbitals for which a spin-up electron was already excited.
+By removing the unpaired spin-down electron left behind, the seniority thus decreases by a factor of $q$.
+The second term accounts for the complementary excitations, where the $(e-p-q)$ spin-down electrons are chosen out of the $N/2-p$ orbitals that 
+remained doubly-occupied after exciting the $p$ spin-up electrons.
 Therefore, this term increases the seniority by $(e-p-q)$.
-We proceed similarly for the fourth binomial term of Eq.~\ref{eq:e_closed}, by decomposing it into the sum of two other binomials,
-where $r$ $\beta$-electrons pair with the previously $\alpha$-electrons (thus reducing the seniority) and $(e-p-r)$ do not (which increases the seniority).
+We proceed similarly for the fourth binomial term of Eq.~\eqref{eq:e_closed}, by decomposing it into the sum of two other binomials,
+where $r$ spin-down electron pairs with the previously spin-up electrons (thus reducing the seniority) and $(e-p-r)$ do not (which increases the seniority).
 By collecting the seniority changes from each term: $(2p)$ from the unmodified binomials, $(-q)$ and $(e-p-q)$ from the third binomial, $(-r)$ and $(e-p-r)$ from the fourth one,
@@
 Finally, by combining the seniority specific binomials and imposing the desired seniority via a Kronecker delta,
-the number of determinants with a given excitation degree $e$ and seniority $s$ produced from a closed-shell reference ($s_0=0$) is given as
+the number of determinants with a given excitation degree $e$ and seniority $s$ produced from a closed-shell reference ($s_0=0$) is given by
 \begin{equation}
@@
 
-Moving to the case where the number of $\alpha$-electrons ($\Na$) and $\beta$-electrons ($\Nb$) differ by one ($\Na=\Nb+1$), which is illustrated in the center of Fig.~\ref{fig:appendix}.
+Moving to the case where the number of spin-up electrons ($\Na$) and spin-down electrons ($\Nb$) differ by one ($\Na=\Nb+1$), which is illustrated in the center of Fig.~\ref{fig:appendix}.
 The number of excited determinants generated exclusively by exciting $e$ electrons %$\Ndet(e)$, 
@@
 The main differences are the following.
-The first binomial term of Eq.~\ref{eq:e_open} is decomposed into the sum of $t$ $\alpha$-electrons excited from the singly-occupied orbital and $(p-t)$ that are excited from the $\Nb$ doubly-occupied ones.
+The first binomial term of Eq.~\eqref{eq:e_open} is decomposed into the sum of $t$ spin-up electrons excited from the singly-occupied orbital and $(p-t)$ that are excited from the $\Nb$ doubly-occupied ones.
 The second binomial is left untouched, as before.
-The third one is also decomposed as explained for the closed-shell case, with the difference that the $q$ $\beta$-electrons are excited from $(p-t)$ orbitals (instead of $p$) 
-from which $\alpha$-electrons were excited. 
-Similarly, in the fourth binomial of Eq.~\ref{eq:e_open}, $r$ $\beta$-electrons are placed into one of the $(1+p-t)$ orbitals (rather than $p$ in the closed-shell case) which contain an unpaired $\alpha$-electron.
+The third one is also decomposed as explained for the closed-shell case, with the difference that the $q$ spin-down electrons are excited from $(p-t)$ orbitals (instead of $p$) 
+from which spin-up electrons were excited. 
+Similarly, in the fourth binomial of Eq.~\eqref{eq:e_open}, $r$ spin-down electrons are placed into one of the $(1+p-t)$ orbitals (rather than $p$ in the closed-shell case) which contain an unpaired spin-up electron.
 Collecting the individual contributions and setting the targeted seniority,
@@
 
-In the general case, an arbitrary reference determinant is defining by three parameters, the number of doubly-occupied orbitals ($\Nd$),
-the number of singly-occupied $\alpha$-electrons ($\Nas = \Na - \Nd$),
-and the number of singly-occupied $\beta$-electrons ($\Nbs = \Nb - \Nd$).
+In the general case, an arbitrary reference determinant is defined by three parameters, the number of doubly-occupied orbitals ($\Nd$),
+the number of singly-occupied spin-up electrons ($\Nas = \Na - \Nd$),
+and the number of singly-occupied spin-down electrons ($\Nbs = \Nb - \Nd$).
 (Alternatively, one could employ the number of electrons $N$, $\Nd$, and the spin quantum number $S_z = (\Na-\Nb)/2$.)
@@
 
-The deduction is somewhat more involved, but follows along the same lines as the two previously discussed cases.
-The starting point is also given by Eq.~\ref{eq:e_open}, with each binomial being written as a sum of two or more binomials that distinguish the change in seniority.
-Here, we just state how each term is decomposed, where the details can be checked by inspection of the final result presented below [Eq.~\eqref{eq:det_3}].
-The first binomial of Eq.~\ref{eq:e_open} is decomposed into the contributions of $\alpha$-electrons excited from the $\Nd$ doubly- and from the $\Nas$ singly-occupied orbitals.
-The second binomial is decomposed into the $\alpha$-electrons being excited to the $\Nv$ virtual orbitals and to the $\Nbs$ singly-occupied orbitals.
-The third one undergoes two decompositions, and is thus expressed as a double sum over three binomials.
-The $\beta$-electrons are distinguished between the $\Nbs$ singly-occupied orbitals, a subset of $\Nd$ doubly-occupied orbitals for which $\alpha$-electrons were excited,
-and the complementary subset for which no $\alpha$-electrons were excited.
-Finally, the fourth binomial of Eq.~\ref{eq:e_open} counts the number of possibilities to place the excited $\beta$-electrons.
+The deduction is somewhat more involved but follows along the same lines as the two previously discussed cases.
+The starting point is also given by Eq.~\eqref{eq:e_open}, with each binomial being written as a sum of two or more binomials that distinguish the change in seniority.
+Here, we just state how each term is decomposed, where the details can be checked by inspection of the final result presented below [see Eq.~\eqref{eq:det_3}].
+The first binomial of Eq.~\eqref{eq:e_open} is decomposed into the contributions of spin-up electrons excited from the $\Nd$ doubly- and from the $\Nas$ singly-occupied orbitals.
+The second binomial is decomposed into the spin-up electrons being excited to the $\Nv$ virtual orbitals and to the $\Nbs$ singly-occupied orbitals.
+The third one undergoes two decompositions and is thus expressed as a double sum over three binomials.
+The spin-down electrons are distinguished between the $\Nbs$ singly-occupied orbitals, a subset of $\Nd$ doubly-occupied orbitals for which spin-up electrons were excited,
+and the complementary subset for which no spin-up electrons were excited.
+Finally, the fourth binomial of Eq.~\eqref{eq:e_open} counts the number of possibilities to place the excited spin-down electrons.
 It is first decomposed into two terms, based on whether they are promoted to the $\Nv$ virtual orbitals or to the $\Nas$ singly-occupied orbitals.
-Each term is further decomposed into two terms, accounting for the subset of $\Nv$ orbitals which now contain $\alpha$-electrons,
-and the subset of $\Nas$ orbitals from which $\alpha$-electrons were removed,
+Each term is further decomposed into two terms, accounting for the subset of $\Nv$ orbitals which now contain spin-up electrons,
+and the subset of $\Nas$ orbitals from which spin-up electrons were removed,
 thus leading to three sums over four binomials.

### 2023-08-20 Manuscript/manuscript.tex
sup

@@
 and closed-shell (\ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}) systems,
-as a function of the number of determinants, according to the various CI models considered here while
+as functions of the number of determinants, according to the various CI models considered here while
 corrected with the standard and the renormalized EN2 perturbative energy.

### 2023-08-20 Manuscript/sup.tex
sup

--- 2023-07-18 Manuscript/sup.tex
@@
 \begin{figure*}%[h!]
-\includegraphics[width=1.0\linewidth]{plot_pes_closed}
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_CN}
 \caption{
-Potential energy curves for \ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8},
-according to hCI (green), excitation-based CI (red) and sCI (blue) models,
-with (light tones) and without (dark tones) the EN2 perturbative correction.
-Results first reported in Ref. \onlinecite{Kossoski_2022}.
-}
-\label{fig:plot_pes_closed}
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{CN},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_cn}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S4 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_vinyl}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for vinyl,
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_vinyl}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S5 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H7}
+\caption{
+\label{fig:plot_pt2_rpt2_h7}
+Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H7},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S6 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_HF}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{HF},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_hf}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S7 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_F2}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{F2},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_f2}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S8 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_ethylene}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for ethylene,
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_ethylene}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S9 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_N2}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency, for \ce{N2},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_n2}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S10 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H4}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H4},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_h4}
+\end{figure*}
+%%% %%% %%%
+
+%%% FIG S11 %%%
+\begin{figure*}%[h!]
+\includegraphics[width=1.0\linewidth]{plot_pt2_rpt2_H8}
+\caption{
+Non-parallelity error (NPE), distance error, equilibrium distance, and force constant, for \ce{H8},
+as functions of the number of determinants ($\Ndet$), according to hCI (green), eCI (red) and sCI (blue) models,
+with the standard (full lines with circles) and renormalized (dashed lines with triangles) EN2 perturbative correction.
+The dashed lines represent the FCI results.}
+\label{fig:plot_pt2_rpt2_h8}
 \end{figure*}

### 2023-08-22 Manuscript/manuscript.tex
corrections

@@
 \usepackage[normalem]{ulem}
-\newcommand{\fk}[1]{\textcolor{blue}{#1}}
+\newcommand{\fk}[1]{\textcolor{orange}{#1}}
 \newcommand{\titou}[1]{\textcolor{red}{#1}}
@@
 
-\title{Seniority and hierarchy configuration interaction for radicals and excited states}
+\title{Seniority and hierarchy configuration interaction for radicals and \\ excited states}
 
@@
 \maketitle
+
+% check refs to Alcoba, check report from JPCL
+% reference determinants for other types of states. Diradicals in particular
+% computational scaling and scaling of number of determinants
 
@@
 We introduced hCI \cite{Kossoski_2022} as a particular truncation of the Hilbert space,
-represented as a two-dimensional map of determinants built from their seniority $s$ and their excitation degree $e$ with respect to a reference determinant (see Fig.~\ref{fig:hCI}).
+viewed as a two-dimensional map of determinants built from their seniority $s$ and their excitation degree $e$ with respect to a reference determinant, as shown in Fig.~\ref{fig:hCI}. 
+\fk{By defining orbital subspaces according to the occupancy of the reference determinant (doubly-occupied, singly-occupied, or unoccupied),
+we refer to the class of an excited determinant as the combination of number of electrons and seniority of each orbital subspace.}
 hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
 which means moving diagonally in the seniority-excitation map (denoted by the color tones in Fig.~\ref{fig:hCI}).
+\fk{In comparison, eCI spans the map horizontally (top to bottom), whereas sCI does it vertically (left to right).}
 
@@
 for a closed-shell reference (left), an open-shell reference with one unpaired electron (center) and an open-shell reference with two unpaired electrons (right).
-The color tones represent the determinants that are included at a given hCI model. hCI0 reduces to HF in the former two cases, where $s_0 = 0$ and $s_0 = 1$.}
+The color tones represent the determinants that are included at a given hCI model.
+hCI0 reduces to \fk{the reference determinant (usually the HF solution)} in the former two cases, where $s_0 = 0$ and $s_0 = 1$.}
 \label{fig:hCI}
@@
 where $s$ and $s_0$ denote the seniority of the candidate and reference determinants, respectively, and $e$ represents the excitation degree that connects them.
+\fk{The definition in Eq.~\eqref{eq:h} guarantees the sought after relation between the classes of determinants and their scaling.
+Namely, all classes whose number of determinants $\Ndet$ share the same scaling with $N$ enter at the hierarchy $h$.
+Had we employed the absolute value of $s-s_0$ or discarded $s_0$ in the definition of $h$, this property would not hold.}
+%\titou{The definition in Eq.~\eqref{eq:h} guarantees that all classes of determinants sharing the same scaling of $\Ndet$ with $N$ enter at the same hierarchy $h$.}
 The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with even (odd) numbers of electrons. The excitation degree $e$ is also an integer.
@@
 
-\titou{The definition in Eq.~\eqref{eq:h} guarantees that all classes of determinants sharing the same scaling of $\Ndet$ with $N$ enter at the same hierarchy $h$.}
 We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference ($s_0 = 0$), and for open-shell references with one ($s_0 = 1$) or two ($s_0 = 2$) unpaired electrons.
-\titou{These references are suitable for describing ground states of a closed-shell system, of open-shell systems with doublet ground states, and (singlet or triplet) singly-excited states of closed-shell systems, respectively.}
-For these three types of references, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
+%\titou{These references are suitable for describing ground states of a closed-shell system, of open-shell systems with doublet ground states, and (singlet or triplet) singly-excited states of closed-shell systems, respectively.}
+\fk{For these three types of references, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
+The closed-shell reference determinant is the natural choice for describing the ground state of closed-shell systems, as well as their doubly-excited states.
+\cite{Barca_2018,Lee_2019,Hait_2020,Kossoski_2021,Marie_2021,Kossoski_2023}
+Meanwhile, the determinant having one unpaired electron is suitable for open-shell systems with doublet ground and excited states.
+Singlet and triplet singly-excited states of closed-shell systems as well as diradicals \cite{Krylov_2006,Horbatenko_2021,Monino_2022} would require
+the reference determinant with two unpaired electrons.}
 
@@
 closed-shell (top left), open-shell with one unpaired electron (top right), and open-shell with two unpaired electrons (bottom).
-\titou{Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied, and unoccupied).}}
+%\titou{Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied, and unoccupied).}
+\fk{By dividing the orbitals of the reference determinant into doubly-occupied, singly-occupied, and unoccupied subspaces,
+the class of an excited determinant is defined by the number of electrons and seniority of each orbital subspace.}}
 \label{fig:determinants}
@@
 
-For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the \titou{restricted HF} determinant.
+\fk{In eCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2e}$. \cite{SzaboBook}
+Likewise, in hCI, $\Ndet$ scales as $N^{2h}$.
+The actual computational cost scales as $\order*{N^{2e+2}}$ for eCI, and similarly for hCI, scaling as $\order*{N^{2h+2}}$.
+Both the scaling of $\Ndet$ and the computational scaling along the two routes are shown in Table \ref{tab:scaling}.}
+%\titou{The computational scaling} along the two routes is shown in Table \ref{tab:scaling}.
+%\PFL{Should we mention something about the scaling of seniority-based CI here?}
+\fk{As for sCI models, the scaling of $\Ndet$ with respect to $N$ is exponential at all orders. \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}
+This is because excitations of all degrees are included, even at lower orders, sCI0 and sCI1.}
+
+\begin{table}[ht!]
+\caption{\fk{Scaling of $\Ndet$ and Computational Scaling (Comput.)} in Terms of the Number of Basis Functions $N$ for the Hierarchy- and Excitation-Based CI Routes, With and Without the EN2 Perturbative Correction.}
+\label{tab:scaling}
+\begin{ruledtabular}
+\begin{tabular}{ll|ll|cc}
+\mc{2}{c|}{Hierarchy-based}   &  \mc{2}{c|}{Excitation-based} &  $\Ndet$ & Comput. \\
+\hline
+hCI1   &            & CIS    &           & $\order*{N^2}$ & $\order*{N^4}$ \\
+hCI1.5 &            &        &           & $\order*{N^3}$ & $\order*{N^5}$ \\
+\hline
+hCI2   & hCI1+EN2   & CISD   & CIS+EN2   & $\order*{N^4}$ & $\order*{N^6}$ \\
+hCI2.5 & hCI1.5+EN2 &        &           & $\order*{N^5}$ & $\order*{N^7}$ \\
+\hline
+hCI3   & hCI2+EN2   & CISDT  & CISD+EN2  & $\order*{N^6}$ & $\order*{N^8}$ \\
+hCI3.5 & hCI2.5+EN2 &        &           & $\order*{N^7}$ & $\order*{N^9}$ \\
+\hline
+hCI4   & hCI3+EN2   & CISDTQ & CISDT+EN2 & $\order*{N^8}$ & $\order*{N^{10}}$ \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the \fk{reference determinant, usually chosen as the HF one}.
 The first non-trivial order is hCI1 (see left panel of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS)
@@
 plus the subset of quadruple excitations where no electrons are unpaired.
-More generally, in hCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2h}$.
-Likewise, in eCI, $\Ndet$ scales as $N^{2e}$.
-\titou{The computational scaling} along the two routes is shown in Table \ref{tab:scaling}.
-\PFL{Should we mention something about the scaling of seniority-based CI here?}
-
-\begin{table}[ht!]
-\caption{\titou{Computational Scaling} in Terms of the Number of Basis Functions $N$ for the Hierarchy- and Excitation-Based CI Routes, With and Without the EN2 Perturbative Correction.}
-\label{tab:scaling}
-\begin{ruledtabular}
-\begin{tabular}{ll|ll|l}
- \mc{2}{c|}{hierarchy-based}   &  \mc{2}{c|}{excitation-based} &  scaling \\
-\hline
-hCI1   &    -       & CIS    &   -       & $\order*{N^4}$ \\
-hCI1.5 &    -       &  -     &   -       & $\order*{N^5}$ \\
-\hline
-hCI2   & hCI1+EN2   & CISD   & CIS+EN2   & $\order*{N^6}$ \\
-hCI2.5 & hCI1.5+EN2 &  -     &   -       & $\order*{N^7}$ \\
-\hline
-hCI3   & hCI2+EN2   & CISDT  & CISD+EN2  & $\order*{N^8}$ \\
-hCI3.5 & hCI2.5+EN2 &   -    &   -       & $\order*{N^9}$ \\
-\hline
-hCI4   & hCI3+EN2   & CISDTQ & CISDT+EN2 & $\order*{N^{10}}$ \\
-\end{tabular}
-\end{ruledtabular}
-\end{table}
+\fk{}
 
@@
 The simplest reference is an open-shell determinant with one unpaired electron ($s_0=1$), shown in the center panel of Fig.~\ref{fig:hCI} and in the top right panel of Fig.~\ref{fig:determinants}.
-hCI0 also reduces to the \titou{restricted open-shell HF} determinant.
+hCI0 also reduces to the \fk{reference determinant (the restricted open-shell HF solution being a natural choice)}.
 In contrast to the closed-shell case, here there are no empty blocks in the seniority-excitation map.
 Actually, the hCI series displays the hCI0.5 level, accounting only for the single excitations from and into the singly-occupied orbital.
-For HF orbitals, however, these excitations do not connect with the reference. Thus, hCI0.5 provides the same energy as restricted open-shell HF.
+\fk{When HF orbitals are employed, these excitations do not connect with the reference, and in this case hCI0.5 provides the same energy as restricted open-shell HF.}
 More generally, since there are more types of orbitals for the open-shell reference (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell reference (doubly-occupied or unoccupied),
@@
 Finally, one can employ a reference determinant having more unpaired electrons to define hCI models.
-The case of two unpaired electrons ($s_0=2$) is shown in the right panel of Fig.~\ref{fig:hCI},
-which would be appropriate for describing \titou{singlet and triplet} singly-excited states of closed-shells.
+The case of two unpaired electrons ($s_0=2$) is shown in the right panel of Fig.~\ref{fig:hCI}.
 The key difference in the hCI sequence, with respect to the closed-shell case, lies in the displacement by one block to the right in the seniority-excitation map,
@@
 which clearly outnumbers the fewer classes associated with the $s_0=0$ and $s_0=1$ cases.
+
 One could adopt references with even larger $s_0$, further displacing the hCI sequence to the right in the seniority-excitation map.
@@
 The ground- and excited-state CI energies are obtained with the Davidson iterative algorithm. \cite{Davidson_1975}
+
 For a given approximate CI model, we further evaluated the standard and the renormalized EN2 perturbative correction. \cite{Garniron_2019}
@@
 Notice that, in contrast to eCI, the energies obtained with sCI and hCI models are not invariant under rotations within the occupied and virtual subspaces.
-The computed excitation energies were benchmarked against the reference values of FCI quality provided in the QUEST database. \cite{Veril_2021}
 The restricted HF solution (restricted open-shell HF for the open-shell systems) was taken as the reference determinant for the hCI and eCI ground-state calculations
@@
 (middle and right panels of Fig.~\ref{fig:hCI}, respectively).
+The computed excitation energies were benchmarked against the reference values of FCI quality provided in the QUEST database. \cite{Veril_2021}
 
 As for the sCI models,
-we considered sCI1 for both ground and (state-specific) excited-state calculations for the open-shell systems (odd numbers of electrons).
-For the closed-shell systems (\titou{even numbers of electrons}), we explored two models.
+we considered sCI1 for both ground and (state-specific) excited-state calculations for the \fk{doublet} open-shell systems.
+For the closed-shell systems \fk{and their excited states}, we explored two models.
 In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal sCI calculation for computing singly-excited states of closed-shell systems.
 Further including the seniority-two sector for the ground-state calculations defines the sCI2/sCI2 model.
+
 From here on, CI models carrying the $\Delta$ symbol denote a state-specific approach, whereas those without it correspond to a ground-state-based approach.
@@
 Each point in the figure denotes one CI model.
-For instance, the hCI route is represented by the dark green line, with the first point being HF \titou{(or hCI0)}, the second being hCI1, the third being hCI1.5, etc.
+For instance, the hCI route is represented by the dark green line, with the first point being \fk{hCI0 (which corresponds to HF in the present cases)}, the second being hCI1, the third being hCI1.5, etc.
 The corresponding results for the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) are shown in Fig.~\ref{fig:plot_all_closed}.
@@
 in line with similar findings for closed-shell systems \cite{Kossoski_2022} (also see Fig.~\ref{fig:plot_all_closed}).
-\PFL{Shall we comment on orbital optimization here?}
+
+%\PFL{Shall we comment on orbital optimization here?}
+\fk{We recall that ground-state restricted open-shell HF orbitals were employed in all calculations for the ground-state radicals.
+In contrast to eCI, methods that rely on the seniority to generate excited determinants (like the hCI and sCI models addressed here)
+are not invariant under orbital rotations within the occupied and virtual subspaces. \cite{Bytautas_2011,Limacher_2014,Stein_2014}
+One could exploit the rotations within each subspace (by means of orbital localization for example) to hopefully render more suitable orbitals for hCI and sCI calculations.
+In our first contribution on hCI to closed-shell systems, \cite{Kossoski_2022} we went one step further
+and variationally optimized the orbitals at the correlated CI level, thus allowing rotations between occupied and virtual subspaces.
+Except for the lower-order CI models, the cost and complications of orbital optimization outweight the mixed improvement on the computed properties. \cite{Kossoski_2022}
+For the closed-shell systems (surveyed both here and in the previous study \cite{Kossoski_2022}),
+the results become significantly more accurate with the EN2 perturbative correction rather than by variationally optimizing the orbitals.}
 
@@
 
-Methods that make use of the concept of seniority to generate excited determinants
-are not invariant under orbital rotations within the occupied and virtual subspaces. \cite{Bytautas_2011,Limacher_2014,Stein_2014}
-This applies to the hCI and sCI models addressed here,
-which employed HF ground-state orbitals or excited-state orbitals optimized for a minimal reference.
-Further exploiting the rotations within occupied and virtual subspaces (by using localization techniques) could render more suitable orbitals for hCI and sCI calculations.
-Alternatively, one could fully optimize the orbitals at a given level of CI. \cite{Kossoski_2022}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\fk{As pointed out before, hCI and sCI models are not invariant under rotations within the subspaces of occupied and virtual orbitals.
+For the state-specific calculations, such rotations for the excited-state orbitals represent additional degrees of freedom to those of the ground-state orbitals. \cite{Kossoski_2022}
+As discussed in the end of Sec.~\ref{sec:res_A}, these degrees of freedom could be exploited,
+by localizing occupied and virtual subspaces or by variationally optimizing all orbitals at a correlated level, for instance.
+We did not pursue these ideas here, and their impact on the excitation energies computed with hCI and sCI remains an open question.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Overall, lower-order models (like hCI1+EN2) were found to be fairly accurate given their low computational cost.
+\fk{At a given CI level, the perturbative correction is significantly more effective in recovering the missing correlation energy than variationally optimizing the orbitals. \cite{Kossoski_2022}}
 In the future, it may be worth combining orbital optimization at a lower level of hCI \cite{Kossoski_2022} with the correction provided by perturbation theory.
@@
 The present findings highlight the challenge in describing, on an equal footing, states with qualitatively different characters, in particular ground and singly-excited states of closed-shell systems.
+\fk{In this regard, rotations within occupied and virtual orbitals represent yet another factor worth exploring in the future of hCI models.}
 

### 2023-08-23 Manuscript/manuscript.tex
validating fabris corrections

@@
 for both ground
-\cite{Limacher_2013,Limacher_2014,Tecmer_2014,Boguslawski_2014a,Boguslawski_2015,Boguslawski_2014b,Boguslawski_2014c,Johnson_2020,Henderson_2014,Stein_2014,Henderson_2015,Chen_2015,Bytautas_2018,Marie_2021,Boguslawski_2021,Tecmer_2022,Mamache_2023}
+\cite{Limacher_2013,Limacher_2014,Tecmer_2014,Boguslawski_2014a,Boguslawski_2015,Boguslawski_2014b,Boguslawski_2014c,Johnson_2020,Henderson_2014,Stein_2014,Henderson_2015,Chen_2015,Bytautas_2018,Marie_2021,Boguslawski_2021,Tecmer_2022,Mamache_2023,Fecteau_2022}
 and excited states.
-\cite{Boguslawski_2016b,Boguslawski_2016c,Boguslawski_2019,Nowak_2019,Kossoski_2021,Marie_2021,Tecmer_2022,Rishi_2023,Nowak_2023}
+\cite{Boguslawski_2016b,Boguslawski_2016c,Boguslawski_2019,Nowak_2019,Kossoski_2021,Marie_2021,Tecmer_2022,Rishi_2023,Nowak_2023,Fecteau_2022}
 Our fifth and last goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state-based and state-specific sCI models for excited states.
@@
 viewed as a two-dimensional map of determinants built from their seniority $s$ and their excitation degree $e$ with respect to a reference determinant, as shown in Fig.~\ref{fig:hCI}. 
-\fk{By defining orbital subspaces according to the occupancy of the reference determinant (doubly-occupied, singly-occupied, or unoccupied),
-we refer to the class of an excited determinant as the combination of number of electrons and seniority of each orbital subspace.}
+By defining orbital subspaces according to the occupancy in the reference determinant (doubly-occupied, singly-occupied, or unoccupied),
+we refer to the class of an excited determinant as the combination of the number of electrons and seniority \titou{in} each orbital subspace.
 hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
 which means moving diagonally in the seniority-excitation map (denoted by the color tones in Fig.~\ref{fig:hCI}).
-\fk{In comparison, eCI spans the map horizontally (top to bottom), whereas sCI does it vertically (left to right).}
+In comparison, eCI spans the map horizontally (top to bottom), whereas sCI does it vertically (left to right).
 
@@
 The color tones represent the determinants that are included at a given hCI model.
-hCI0 reduces to \fk{the reference determinant (usually the HF solution)} in the former two cases, where $s_0 = 0$ and $s_0 = 1$.}
+hCI0 reduces to the reference determinant (usually the HF solution) in the former two cases, where $s_0 = 0$ and $s_0 = 1$.}
 \label{fig:hCI}
@@
 where $s$ and $s_0$ denote the seniority of the candidate and reference determinants, respectively, and $e$ represents the excitation degree that connects them.
-\fk{The definition in Eq.~\eqref{eq:h} guarantees the sought after relation between the classes of determinants and their scaling.
+The definition in Eq.~\eqref{eq:h} guarantees the sought-after relation between the classes of determinants and their scaling.
 Namely, all classes whose number of determinants $\Ndet$ share the same scaling with $N$ enter at the hierarchy $h$.
-Had we employed the absolute value of $s-s_0$ or discarded $s_0$ in the definition of $h$, this property would not hold.}
-%\titou{The definition in Eq.~\eqref{eq:h} guarantees that all classes of determinants sharing the same scaling of $\Ndet$ with $N$ enter at the same hierarchy $h$.}
+Had we employed the absolute value of $s-s_0$ or discarded $s_0$ in the definition of $h$, this property would not hold.
 The term $(s-s_0)/2$ is always an integer, with $s$ and $s_0$ being even (odd) for systems with even (odd) numbers of electrons. The excitation degree $e$ is also an integer.
@@
 We show in Fig.~\ref{fig:hCI} how the seniority-excitation map of determinants is partitioned for a closed-shell reference ($s_0 = 0$), and for open-shell references with one ($s_0 = 1$) or two ($s_0 = 2$) unpaired electrons.
-%\titou{These references are suitable for describing ground states of a closed-shell system, of open-shell systems with doublet ground states, and (singlet or triplet) singly-excited states of closed-shell systems, respectively.}
-\fk{For these three types of references, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
-The closed-shell reference determinant is the natural choice for describing the ground state of closed-shell systems, as well as their doubly-excited states.
+For these three types of references, the classes of determinants included up to hCI1 are presented in Fig.~\ref{fig:determinants}.
+The closed-shell reference determinant is the natural choice for describing the ground state of a closed-shell system, as well as their doubly-excited states.
 \cite{Barca_2018,Lee_2019,Hait_2020,Kossoski_2021,Marie_2021,Kossoski_2023}
-Meanwhile, the determinant having one unpaired electron is suitable for open-shell systems with doublet ground and excited states.
+Meanwhile, a determinant having one unpaired electron is suitable for an open-shell system with doublet ground and excited states.
 Singlet and triplet singly-excited states of closed-shell systems as well as diradicals \cite{Krylov_2006,Horbatenko_2021,Monino_2022} would require
-the reference determinant with two unpaired electrons.}
+a reference determinant with two unpaired electrons.
 
@@
 closed-shell (top left), open-shell with one unpaired electron (top right), and open-shell with two unpaired electrons (bottom).
-%\titou{Each class is determined by the specific seniority and occupation of the orbital blocks of the reference (doubly-occupied, singly-occupied, and unoccupied).}
-\fk{By dividing the orbitals of the reference determinant into doubly-occupied, singly-occupied, and unoccupied subspaces,
-the class of an excited determinant is defined by the number of electrons and seniority of each orbital subspace.}}
+By dividing the orbitals of the reference determinant into doubly-occupied, singly-occupied, and unoccupied subspaces,
+the class of an excited determinant is defined by the number of electrons and seniority \titou{in} each orbital subspace.}
 \label{fig:determinants}
@@
 
-\fk{In eCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2e}$. \cite{SzaboBook}
+In eCI, the number of determinants $\Ndet$ scales polynomially with the number of basis functions $N$ as $N^{2e}$. \cite{SzaboBook}
 Likewise, in hCI, $\Ndet$ scales as $N^{2h}$.
 The actual computational cost scales as $\order*{N^{2e+2}}$ for eCI, and similarly for hCI, scaling as $\order*{N^{2h+2}}$.
-Both the scaling of $\Ndet$ and the computational scaling along the two routes are shown in Table \ref{tab:scaling}.}
-%\titou{The computational scaling} along the two routes is shown in Table \ref{tab:scaling}.
-%\PFL{Should we mention something about the scaling of seniority-based CI here?}
-\fk{As for sCI models, the scaling of $\Ndet$ with respect to $N$ is exponential at all orders. \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}
-This is because excitations of all degrees are included, even at lower orders, sCI0 and sCI1.}
+Both the scaling of $\Ndet$ and the computational scaling along the two routes are shown in Table \ref{tab:scaling}.
+As for sCI models, the scaling of $\Ndet$ with respect to $N$ is exponential at all orders. \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967}
+This is because excitations of all degrees are included, even at lower orders, such as sCI0 and sCI1.
 
 \begin{table}[ht!]
-\caption{\fk{Scaling of $\Ndet$ and Computational Scaling (Comput.)} in Terms of the Number of Basis Functions $N$ for the Hierarchy- and Excitation-Based CI Routes, With and Without the EN2 Perturbative Correction.}
+\caption{Scaling of $\Ndet$ and Computational Scaling (Cost) in Terms of the Number of Basis Functions $N$ for the Hierarchy- and Excitation-Based CI Routes, With and Without the EN2 Perturbative Correction.}
 \label{tab:scaling}
@@
 \begin{tabular}{ll|ll|cc}
-\mc{2}{c|}{Hierarchy-based}   &  \mc{2}{c|}{Excitation-based} &  $\Ndet$ & Comput. \\
+\mc{2}{c|}{Hierarchy-based}   &  \mc{2}{c|}{Excitation-based} &  $\Ndet$ & Cost \\
 \hline
@@
 
-For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the \fk{reference determinant, usually chosen as the HF one}.
+For a closed-shell reference determinant ($s_0=0$), hCI0 reduces to the reference determinant, usually chosen as the HF one.
 The first non-trivial order is hCI1 (see left panel of Fig.~\ref{fig:hCI}), which accounts for all single excitations (as CIS)
@@
 The simplest reference is an open-shell determinant with one unpaired electron ($s_0=1$), shown in the center panel of Fig.~\ref{fig:hCI} and in the top right panel of Fig.~\ref{fig:determinants}.
-hCI0 also reduces to the \fk{reference determinant (the restricted open-shell HF solution being a natural choice)}.
+hCI0 also reduces to the reference determinant (the restricted open-shell HF solution being a natural choice).
 In contrast to the closed-shell case, here there are no empty blocks in the seniority-excitation map.
 Actually, the hCI series displays the hCI0.5 level, accounting only for the single excitations from and into the singly-occupied orbital.
-\fk{When HF orbitals are employed, these excitations do not connect with the reference, and in this case hCI0.5 provides the same energy as restricted open-shell HF.}
-More generally, since there are more types of orbitals for the open-shell reference (doubly-occupied, singly-occupied and unoccupied) than for the closed-shell reference (doubly-occupied or unoccupied),
+When HF orbitals are employed, these excitations do not connect with the reference, and in this case, hCI0.5 provides the same energy as restricted open-shell HF.
+More generally, since there are more types of orbitals for the open-shell reference (doubly-occupied, singly-occupied, and unoccupied) than for the closed-shell reference (doubly-occupied or unoccupied),
 there are correspondingly more classes of determinants at a given level of hCI.
@@
 In a given calculation, the energies are considered to be converged when the (largest) EN2 correction computed in the truncated Hilbert space
-lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2018}
+lies below \SI{0.01}{\milli\hartree}. \cite{Garniron_2017}
 This selected CI procedure requires considerably fewer determinants than the total number of determinants in the truncated Hilbert space
@@
 For example, hCI2 and CISD present a $\order*{N^6}$ computational scaling, whereas hCI2+EN2 and CISD+EN2 scale as $\order*{N^8}$,
-though with a small prefactor stemming from the EN2 calculation that employs a very efficient semistochastic algorithm. \cite{Garniron_2018}
+though with a small prefactor stemming from the EN2 calculation that employs a very efficient semistochastic algorithm. \cite{Garniron_2017}
 The computational scaling associated with the CI+EN2 calculations is also presented in Table \ref{tab:scaling}.
@@
 As for the sCI models,
-we considered sCI1 for both ground and (state-specific) excited-state calculations for the \fk{doublet} open-shell systems.
-For the closed-shell systems \fk{and their excited states}, we explored two models.
+we considered sCI1 for both ground and (state-specific) excited-state calculations for the doublet open-shell systems.
+For the closed-shell systems and their excited states, we explored two models.
 In the sCI2/sCI0 model, the ground state was described with sCI0 and the excited state with sCI2, which is the minimal sCI calculation for computing singly-excited states of closed-shell systems.
@@
 Each point in the figure denotes one CI model.
-For instance, the hCI route is represented by the dark green line, with the first point being \fk{hCI0 (which corresponds to HF in the present cases)}, the second being hCI1, the third being hCI1.5, etc.
+For instance, the hCI route is represented by the dark green line, with the first point being hCI0 (which corresponds to HF in the present cases), the second being hCI1, the third being hCI1.5, etc.
 The corresponding results for the closed-shell systems (\ce{HF}, \ce{F2}, \ce{N2}, ethylene, \ce{H4}, and \ce{H8}) are shown in Fig.~\ref{fig:plot_all_closed}.
@@
 
-%\PFL{Shall we comment on orbital optimization here?}
-\fk{We recall that ground-state restricted open-shell HF orbitals were employed in all calculations for the ground-state radicals.
+We recall that ground-state restricted open-shell HF orbitals were employed in all calculations for the ground-state radicals.
 In contrast to eCI, methods that rely on the seniority to generate excited determinants (like the hCI and sCI models addressed here)
@@
 and variationally optimized the orbitals at the correlated CI level, thus allowing rotations between occupied and virtual subspaces.
-Except for the lower-order CI models, the cost and complications of orbital optimization outweight the mixed improvement on the computed properties. \cite{Kossoski_2022}
+Except for the lower-order CI models, the cost and complications of orbital optimization outweigh the mixed improvement in the computed properties. \cite{Kossoski_2022}
 For the closed-shell systems (surveyed both here and in the previous study \cite{Kossoski_2022}),
-the results become significantly more accurate with the EN2 perturbative correction rather than by variationally optimizing the orbitals.}
+the results become significantly more accurate with the EN2 perturbative correction rather than by variationally optimizing the orbitals.
 
@@
 
-\fk{As pointed out before, hCI and sCI models are not invariant under rotations within the subspaces of occupied and virtual orbitals.
+As pointed out before, hCI and sCI models are not invariant under rotations within the subspaces of occupied and virtual orbitals.
 For the state-specific calculations, such rotations for the excited-state orbitals represent additional degrees of freedom to those of the ground-state orbitals. \cite{Kossoski_2022}
@@
 by localizing occupied and virtual subspaces or by variationally optimizing all orbitals at a correlated level, for instance.
-We did not pursue these ideas here, and their impact on the excitation energies computed with hCI and sCI remains an open question.}
+We did not pursue these ideas here, and their impact on the excitation energies computed with hCI and sCI remains an open question.
 
@@
 Overall, lower-order models (like hCI1+EN2) were found to be fairly accurate given their low computational cost.
-\fk{At a given CI level, the perturbative correction is significantly more effective in recovering the missing correlation energy than variationally optimizing the orbitals. \cite{Kossoski_2022}}
+At a given CI level, the perturbative correction is significantly more effective in recovering the missing correlation energy than variationally optimizing the orbitals. \cite{Kossoski_2022}
 In the future, it may be worth combining orbital optimization at a lower level of hCI \cite{Kossoski_2022} with the correction provided by perturbation theory.
@@
 The present findings highlight the challenge in describing, on an equal footing, states with qualitatively different characters, in particular ground and singly-excited states of closed-shell systems.
-\fk{In this regard, rotations within occupied and virtual orbitals represent yet another factor worth exploring in the future of hCI models.}
+In this regard, rotations within occupied and virtual orbitals represent yet another factor worth exploring in the future developments of hCI models.
 

### 2023-08-23 Manuscript/manuscript.tex
ok with appendix

@@
 open-shell determinant ($s_0=1$) with $\Na$ spin-up electrons and $\Nb=\Na-1$ spin-down electrons (center),
-and general determinant ($s_0=\Nas+\Nbs$), with $\Nd$ doubly-occupied orbitals, $\Nas$ singly-occupied spin-up electrons and $\Nbs$ singly-occupied spin-down electrons (right).
+and general determinant ($s_0=\Nas+\Nbs$), with $\Nd$ doubly-occupied orbitals, $\Nas$ singly-occupied spin-up electrons, and $\Nbs$ singly-occupied spin-down electrons (right).
 In the three cases, there are $K$ spatial orbitals.
@@
 based on whether the excitation increases or reduces the seniority.
-For the decomposition, we make use of Vandermonde's identity,
+For the decomposition, we make use of Vandermonde's identity:
 \begin{equation}
@@
 \end{equation}
-where the sum expresses the different combinations for exciting spin-up or spin-down electrons,
-the first binomial term accounts for the number of possibilities for exciting $p$ spin-up electrons from the $N/2$ orbitals,
+where the sum expresses the different combinations for exciting spin-up or spin-down electrons.
+The first binomial term accounts for the number of possibilities for exciting $p$ spin-up electrons from the $N/2$ orbitals,
 the second term represents the number of ways of placing these $p$ electrons into $K-N/2$ orbitals,
@@
 Next, the spin-down electrons can be excited from the same orbitals from which the spin-up electrons were excited (decreasing the seniority by one)
-or instead from an orbital that remained doubly-occupied (increasing the seniority by one).
-The two possibilities are expressed by decomposing the third binomial term in Eq.~\eqref{eq:e_closed} as a sum over the two corresponding binomials,
+or instead from an orbital that remained doubly occupied (increasing the seniority by one).
+The two possibilities are expressed by decomposing the third binomial term in Eq.~\eqref{eq:e_closed} as a sum over the two corresponding binomials, such as
 \begin{equation}
@@
 the number of determinants with a given excitation degree $e$ and seniority $s$ produced from a closed-shell reference ($s_0=0$) is given by
-\begin{equation}
-\begin{split}
-	&\sum_{p=0}^{e} \binom{N/2}{p} \binom{K-N/2}{p}
-	\sum_{q=0}^{e-p} \binom{p}{q} \binom{N/2-p}{e-p-q}  \\
-	&\sum_{r=0}^{e-p} \binom{p}{r} \binom{K-N/2-p}{e-p-r}
+\begin{multline}
+	\sum_{p=0}^{e} \binom{N/2}{p} \binom{K-N/2}{p}
+	\sum_{q=0}^{e-p} \binom{p}{q} \binom{N/2-p}{e-p-q} 
+	\\ \times
+	\sum_{r=0}^{e-p} \binom{p}{r} \binom{K-N/2-p}{e-p-r}
  	\delta_{s,2(e-q-r)}.
 \label{eq:det_1}
-\end{split}
-\end{equation}
+\end{multline}
 To obtain the final number of determinants for a given hierarchy $h$, we simply sum over the allowed combinations of $e$ and $s$ according to Eq.~\eqref{eq:h}.
 
-Moving to the case where the number of spin-up electrons ($\Na$) and spin-down electrons ($\Nb$) differ by one ($\Na=\Nb+1$), which is illustrated in the center of Fig.~\ref{fig:appendix}.
-The number of excited determinants generated exclusively by exciting $e$ electrons %$\Ndet(e)$, 
+Moving to the case where the number of spin-up electrons ($\Na$) and spin-down electrons ($\Nb$) differ by one ($\Na=\Nb+1$), which is illustrated in the center panel of Fig.~\ref{fig:appendix},
+the number of excited determinants generated exclusively by exciting $e$ electrons %$\Ndet(e)$, 
 is analogous to Eq.~\eqref{eq:e_closed}, being given by
@@
 the number of determinants generated from a $s_0=1$ reference determinant, with excitation degree $e$ and seniority $s$ is given by
-\begin{equation}
-\begin{split}
-	& \sum_{p=0}^{e} 
+\begin{multline}
+	\sum_{p=0}^{e} 
 	\sum_{t=0}^{p} \binom{1}{t} \binom{\Nb}{p-t} 
 	\binom{K-\Na}{p} 
-	\sum_{q=0}^{e-p} \binom{p-t}{q} \binom{\Nb-p+t}{e-p-q} \\
-	& \sum_{r=0}^{e-p} \binom{1+p-t}{r} \binom{K-\Nb-1-p+t}{e-p-r}
+	\sum_{q=0}^{e-p} \binom{p-t}{q} \binom{\Nb-p+t}{e-p-q} 
+	\\ \times
+	\sum_{r=0}^{e-p} \binom{1+p-t}{r} \binom{K-\Nb-1-p+t}{e-p-r}
  	\delta_{s,1+2(e-q-r-t)}.
 \label{eq:det_2}
-\end{split}
-\end{equation}
+\end{multline}
 
@@
 The final result for the number of determinants generated from an arbitrary reference determinant, constrained to have excitation degree $e$ and seniority $s$ is given by
-\begin{equation}
-\begin{split}
-	& \sum_{p=0}^{e} 
+\begin{widetext}
+\begin{multline}
+	\sum_{p=0}^{e} 
 	\sum_{t=0}^{p} \binom{\Nas}{t} \binom{\Nd}{p-t} 
-	\sum_{w=0}^{p} \binom{\Nv}{w} \binom{\Nbs}{p-w} \\
-	& \sum_{u=0}^{e-p} \binom{\Nbs}{u} 
-	\sum_{v=0}^{e-p-u} \binom{p-t}{v} \binom{\Nd-p+t}{e-p-u-v} \\
-	& \sum_{r=0}^{e-p} \sum_{m=0}^{r} \binom{w}{m} \binom{\Nv-w}{r-m} 
-	\sum_{n=0}^{e-p-r} \binom{t}{n} \binom{\Nas-t}{e-p-r-n} \\
-	& \delta_{s,\Nas+\Nbs+2(e+w-p-t-u-v-m-n)}.
+	\sum_{w=0}^{p} \binom{\Nv}{w} \binom{\Nbs}{p-w} 
+	\sum_{u=0}^{e-p} \binom{\Nbs}{u} 
+	\sum_{v=0}^{e-p-u} \binom{p-t}{v} \binom{\Nd-p+t}{e-p-u-v} 
+	\sum_{r=0}^{e-p} \sum_{m=0}^{r} \binom{w}{m} \binom{\Nv-w}{r-m} 
+	\\ \times
+	\sum_{n=0}^{e-p-r} \binom{t}{n} \binom{\Nas-t}{e-p-r-n} 
+	\delta_{s,\Nas+\Nbs+2(e+w-p-t-u-v-m-n)}.
 \label{eq:det_3}
-\end{split}
-\end{equation}
+\end{multline}
+\end{widetext}
 

### 2023-08-27 Manuscript/manuscript.tex
discussion on spin-contamination

@@
 By defining orbital subspaces according to the occupancy in the reference determinant (doubly-occupied, singly-occupied, or unoccupied),
-we refer to the class of an excited determinant as the combination of the number of electrons and seniority \titou{in} each orbital subspace.
+we refer to the class of an excited determinant as the combination of the number of electrons and seniority in each orbital subspace.
 hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
@@
 By dividing the orbitals of the reference determinant into doubly-occupied, singly-occupied, and unoccupied subspaces,
-the class of an excited determinant is defined by the number of electrons and seniority \titou{in} each orbital subspace.}
+the class of an excited determinant is defined by the number of electrons and seniority in each orbital subspace.}
 \label{fig:determinants}
@@
 These discontinuities reflect crossings between states of different symmetries in the unperturbed CI calculation,
-appearing as kinks in the computed ground-state PEC \titou{(intruder states)}.
+appearing as kinks in the computed ground-state PEC.
 Due to the abrupt change of character, the EN2 correction is not uniform and gives rise to the observed discontinuities.

### 2023-08-27 Manuscript/sup.tex
discussion on spin-contamination

@@
 Potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
-according to FCI (black), hCI (green) (dashed lines for half-integer $h$), eCI (red), and sCI (blue) models, with the cc-pVDZ basis set and frozen-core approximation.}
+according to restricted open-shell HF (gray), FCI (black), hCI (green) (dashed lines for half-integer $h$), eCI (red), and sCI (blue) models, with the cc-pVDZ basis set and frozen-core approximation.}
 \label{fig:plot_pes}

### 2023-08-27 Manuscript/manuscript.tex
now it comes

@@
 
-% check refs to Alcoba, check report from JPCL
-% reference determinants for other types of states. Diradicals in particular
-% computational scaling and scaling of number of determinants
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \cite{Ziegler_1977,Burton_2021,Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Hanscam_2022,Kossoski_2023,Marie_2023}
-% references for SCF
-%\cite{Ziegler_1977,Burton_2021}
-% multiconfigurational SCF, \cite{Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Kossoski_2023,Marie_2023}
 density-functional theory,
@@
 plus the subset of quadruple excitations where no electrons are unpaired.
-\fk{}
 
@@
 The spin-contamination problem and its solution are therefore equivalent to that encountered in eCI. \cite{Maurice_1996}
-Here, we have employed this procedure and considered pure spin states.
+\fk{Here, we have employed this procedure and considered pure spin states.
+For the calculation of excitation energies, however, we have also assessed the effect of not imposing this condition.}
+
 For the $s_0 = 2$ reference, one could alternatively employ a high-spin triplet determinant (two unpaired spin-up electrons),
@@
 
-For each CI model considered here, we evaluate the mean signed error (MSE), mean absolute error (MAE), and root-mean-square error (RMSE)
+For each CI model considered here, we evaluate the mean signed error (MSE), mean absolute error (MAE), root-mean-square error (RMSE), and standard deviation of the errors (SDE).
 with respect to the reference theoretical values for the excitation energies.
-For the lower-order models, the calculations were performed for 48 (closed-shell) and 19 (open-shell) excited states.
+For the lower-order models, the calculations were performed for 50 (closed-shell) and 19 (open-shell) excited states.
 In turn, subsets of 16 (closed-shell) and 6 (open-shell) excited states were considered for the higher-order CI models, given the more intensive computational cost.
@@
 We start the discussion with the excitations from closed-shell systems, with corresponding statistical errors shown in Table \ref{tab:stat1}.
-The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.07}{\eV}.
+The first level of hCI, hCI1, produces poor excitation energies, with a MAE of \SI{1.16}{\eV}.
 This model shares the same computational scaling with CIS, but includes the paired double excitations,
 which clearly worsen the decent CIS results (MAE of \SI{0.61}{\eV}).
-Moving one rank up, to hCI1.5, the MAE increases to \SI{1.91}{\eV}, and then reaches a maximum of \SI{3.53}{\eV} at the hCI2 level.
+Moving one rank up, to hCI1.5, the MAE increases to \SI{1.95}{\eV}, and then reaches a maximum of \SI{3.53}{\eV} at the hCI2 level.
 We notice that hCI2 delivers somewhat smaller errors than CISD (MAE of \SI{4.09}{\eV}), even though both are way too large.
@@
 
-% HCF cisdtq ground
-% sil cisdtq sA2 and sB2
-
 \begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
 Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
@@
 \begin{ruledtabular}
-\begin{tabular}{ldddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
-\hline
-CIS               & 48  & +0.04 & 0.61 & 0.60 \\
-CISD              & 16  & +4.09 & 4.09 & 4.18 \\
-CISDT             & 16  & +0.12 & 0.17 & 0.18 \\
-CISDTQ            & 16  & +0.15 & 0.15 & 0.17 \\
+\begin{tabular}{lddddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
+\hline
+CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77 \\
+CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84 \\
+CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14 \\
+CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08 \\
 %
 \hline
-hCI1              & 48  & +0.97 & 1.07 & 1.20 \\
-hCI1.5            & 48  & +1.91 & 1.91 & 2.00 \\
-hCI2              & 16  & +2.99 & 3.53 & 3.61 \\
-hCI2.5            & 16  & +1.95 & 1.95 & 2.06 \\
-hCI3              & 16  & +0.19 & 0.19 & 0.21 \\
-hCI3.5            & 16  & +0.11 & 0.11 & 0.13 \\
+hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89 \\
+hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59 \\
+hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76 \\
+hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66 \\
+hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08 \\
+hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07 \\
 %
 \hline
-$\Delta$CSF       & 48  & -0.73 & 0.80 & 0.94 \\
-$\Delta$CISD      & 48  & -0.13 & 0.17 & 0.22 \\
-$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 \\
-$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 \\
+$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58 \\
+$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18 \\
+$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11 \\
+$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02 \\
 %
 \hline
-$\Delta$hCI1      & 48  & -1.36 & 1.40 & 1.50 \\
-$\Delta$hCI1.5    & 48  & -2.91 & 2.91 & 3.10 \\
-$\Delta$hCI2      & 48  & -0.19 & 0.20 & 0.25 \\
-$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 \\
-$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 \\
-$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 \\
+$\Delta$hCI1      & 50  & -1.29 & 1.47 & 1.63 & 1.00 \\
+$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15 \\
+$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16 \\
+$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13 \\
+$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10 \\
+$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05 \\
 %
 \hline
-sCI2/sCI2         & 48  & +1.26 & 1.34 & 1.49 \\
-sCI2/sCI0         & 48  & -0.41 & 0.57 & 0.89 \\
-$\Delta$sCI2/sCI2 & 48  & +0.69 & 0.79 & 0.90 \\
-$\Delta$sCI2/sCI0 & 48  & -0.99 & 0.99 & 1.15 \\
+sCI2/sCI2         & 50  & +1.22 & 1.29 & 1.45 & 0.78 \\
+sCI2/sCI0         & 50  & -0.47 & 0.62 & 0.92 & 0.80 \\
+$\Delta$sCI2/sCI2 & 50  & +0.65 & 0.78 & 0.89 & 0.60 \\
+$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62 \\
 %
@@
 Actually, $\Delta$hCI2 is slightly less accurate than $\Delta$CISD, with MAEs of \SI{0.20}{\eV} and \SI{0.17}{\eV}.
-Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.19}{\eV} against \SI{-0.13}{\eV}).
+Moreover, $\Delta$hCI2 has a more negative MSE than $\Delta$CISD (\SI{-0.18}{\eV} against \SI{-0.12}{\eV}).
 Albeit small, these differences are statistically significant, and $\Delta$CISD would still be preferable to $\Delta$hCI2 by some margin.
@@
 \begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), and Root-Mean Square Error (RMSE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
 Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
@@
 \begin{ruledtabular}
-\begin{tabular}{ldddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} \\
-\hline
-CIS               & 19 & +0.33 & 0.36 & 0.50 \\
-CISD              & 6  & +2.97 & 2.97 & 3.19 \\
-CISDT             & 6  & +0.06 & 0.07 & 0.09 \\
-CISDTQ            & 6  & +0.08 & 0.08 & 0.11 \\
+\begin{tabular}{lddddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
+\hline
+CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50 \\
+CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17 \\
+CISDT             &  6 & +0.06 & 0.07 & 0.09 & 0.06 \\
+CISDTQ            &  6 & +0.08 & 0.08 & 0.11 & 0.08 \\
 %
 \hline
-hCI1              & 19 & +0.70 & 0.77 & 0.96 \\
-hCI1.5            & 19 & +1.11 & 1.11 & 1.30 \\
-hCI2              & 6  & +1.32 & 1.32 & 1.60 \\
-hCI2.5            & 6  & +0.03 & 0.05 & 0.06 \\
-hCI3              & 6  & +0.00 & 0.04 & 0.05 \\
-hCI3.5            & 6  & +0.01 & 0.03 & 0.05 \\
+hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64 \\
+hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68 \\
+hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90 \\
+hCI2.5            &  6 & +0.03 & 0.05 & 0.06 & 0.05 \\
+hCI3              &  6 & +0.00 & 0.04 & 0.05 & 0.05 \\
+hCI3.5            &  6 & +0.01 & 0.03 & 0.05 & 0.04 \\
 %
 \hline
-$\Delta$CSF       & 19 & -0.04 & 0.41 & 0.58 \\
-$\Delta$CISD      & 19 & +0.01 & 0.12 & 0.21 \\
-$\Delta$CISDT     & 6  & -0.07 & 0.07 & 0.10 \\
-$\Delta$CISDTQ    & 6  & -0.02 & 0.02 & 0.02 \\
+$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59 \\
+$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21 \\
+$\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08 \\
+$\Delta$CISDTQ    &  6 & -0.02 & 0.02 & 0.02 & 0.02 \\
 %
 \hline
-$\Delta$hCI1      & 19 & -0.19 & 0.34 & 0.48 \\
-$\Delta$hCI1.5    & 19 & -0.10 & 0.46 & 0.70 \\
-$\Delta$hCI2      & 19 & -0.01 & 0.11 & 0.16 \\
-$\Delta$hCI2.5    & 6  & -0.09 & 0.09 & 0.14 \\
-$\Delta$hCI3      & 6  & -0.06 & 0.06 & 0.08 \\
-$\Delta$hCI3.5    & 6  & -0.02 & 0.02 & 0.03 \\
+$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44 \\
+$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69 \\
+$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16 \\
+$\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10 \\
+$\Delta$hCI3      &  6 & -0.06 & 0.06 & 0.08 & 0.05 \\
+$\Delta$hCI3.5    &  6 & -0.02 & 0.02 & 0.03 & 0.02 \\
 %
 \hline
-sCI1              & 19 & +0.24 & 0.39 & 0.61 \\
-$\Delta$sCI1      & 19 & -0.11 & 0.27 & 0.38 \\
+sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53 \\
+$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37 \\
 \end{tabular}
@@
 Once state-specific orbitals are employed, there is no remaining bias toward the ground state.
-In stark contrast to the closed-shell case, the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.34}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.46}{\eV}),
-become more accurate than their ground-state-based counterparts, hCI1 (MAE of \SI{0.77}{\eV}) and hCI1.5 (MAE of \SI{1.11}{\eV}).
+In stark contrast to the closed-shell case, the lower orders of hCI, $\Delta$hCI1 (MAE of \SI{0.35}{\eV}) and $\Delta$hCI1.5 (MAE of \SI{0.45}{\eV}),
+become more accurate than their ground-state-based counterparts, hCI1 (MAE of \SI{0.73}{\eV}) and hCI1.5 (MAE of \SI{1.07}{\eV}).
 Once again, this is thanks to the same type of reference in ground- and excited-state calculations.
 The case of $\Delta$hCI1 for open-shell excitations is depicted in Fig.~\ref{fig:determinants}, in comparison to the more unbalanced case of excitations from closed-shells.
-We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.19}{\eV} with $\Delta$hCI1 and \SI{-0.10}{\eV} with $\Delta$hCI1.5),
+We notice, however, an apparent residual bias from the somewhat negative MSEs (\SI{-0.21}{\eV} with $\Delta$hCI1 and \SI{-0.14}{\eV} with $\Delta$hCI1.5),
 and that $\Delta$hCI1.5 is slightly less accurate than $\Delta$hCI1.
 
-This residual bias virtually disappears at the next order, $\Delta$hCI2 (MSE of \SI{-0.01}{\eV}),
+This residual bias virtually disappears at the next order, $\Delta$hCI2 (MSE of \SI{-0.02}{\eV}),
 whereas for the closed-shell systems even $\Delta$hCI3.5 significantly underestimated (MSE of \SI{-0.08}{\eV}) the excitation energies.
 The same comparison between open- and closed-shells holds for eCI models.
-For the open-shell transitions, $\Delta$hCI2 is slightly more accurate (MAE of \SI{0.11}{\eV} and RMSE of \SI{0.16}{\eV}) than $\Delta$CISD (MAE of \SI{0.12}{\eV} and RMSE of \SI{0.21}{\eV}).
+For the open-shell transitions, $\Delta$hCI2 is slightly more accurate (MAE of \SI{0.10}{\eV} and RMSE of \SI{0.16}{\eV}) than $\Delta$CISD (MAE of \SI{0.12}{\eV} and RMSE of \SI{0.21}{\eV}).
 In further contrast to the closed-shell case,
@@
 
+\fk{For the state-specific models, we have further assessed the impact of not imposing spin eigenstates (the results and statistics can be found in the {\SupInf}).
+This amounts to not including the appropriate spin-flip determinants lying above the hierarchy or excitation degree of choice.
+The effect is minimal for both excitations of both closed- and open-shell systems, in the $\Delta$hCI1.5, $\Delta$hCI2, and $\Delta$CISD models.
+The mean difference on the individual excitation energies lies between \SI{0.01}{\eV} and \SI{0.02}{\eV} for the latter two and one order of magnitude less for $\Delta$hCI1.5.
+Similarly small effects are seen in the global statistics.
+In turn, $\Delta$hCI1 is more affected, which would be expected given it is a low order model.
+Average individual excitations of closed- and open-shell systems vary by \SI{0.26}{\eV} and \SI{0.06}{\eV}, respectively, always in the sense of improving the energies in the former case,
+though not enough to cause a considerable reduction of the large absolute errors.
+Overall, not constraining the CI solutions to have well-defined spin brings practical complications (as more roots have to be calculated)
+and does not improve the computed excitation energies obtained with the more competitive $\Delta$CISD and $\Delta$hCI2 models.
+The sCI models, discussed in the next subsection, are naturally spin eigenstates, and there is no need to enforce that.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We refer back to Table \ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
-The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.34}{\eV}.
+The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.29}{\eV}.
 The same two factors discussed above for hCI explain such large errors.
 First, the ground-state description is favored due to the use of ground-state HF orbitals.
-Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAE to \SI{0.79}{\eV}, though still overestimating the excitation energies.
+Indeed, with state-specific orbitals, the $\Delta$sCI2/sCI2 model reduces the MAE to \SI{0.78}{\eV}, though still overestimating the excitation energies.
 Second, there is an unbalanced description of the correlation for ground and excited states, associated with the classes of determinants.
@@
 A possible solution to this unbalance is to restrict the determinants to a maximum seniority number of $s=0$ for the ground state and $s=2$ for the excited state calculation.
-This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.57}{\eV}, compared to \SI{1.34}{\eV} of sCI2/sCI2.
-However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{0.99}{\eV}.
+This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.62}{\eV}, compared to \SI{1.29}{\eV} of sCI2/sCI2.
+However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{1.04}{\eV}.
 Clearly, the seniority-two sector captures more correlation for the excited states than the seniority-zero does for the ground state.
@@
 Moreover, the state-specific approach is superior.
-sCI1 presents a MAE of \SI{0.39}{\eV} and overestimates the excitation energies (MSE of \SI{+0.24}{\eV}), in view of the bias introduced by the ground-state orbitals.
-With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.27}{\eV} and the MSE to \SI{-0.11}{\eV}, which is smaller in absolute value than found for sCI1.
+sCI1 presents a MAE of \SI{0.35}{\eV} and overestimates the excitation energies (MSE of \SI{+0.20}{\eV}), in view of the bias introduced by the ground-state orbitals.
+With state-specific orbitals, $\Delta$sCI1 reduces the MAE to \SI{0.28}{\eV} and the MSE to \SI{-0.12}{\eV}, which is smaller in absolute value than found for sCI1.
 
@@
 Likewise, a formulation of pCCD to open-shell systems might share an analogous connection to low-order sCI models like sCI1.
-If that is the case, then a state-specific approach of such pCCD formulation adapted to open-shells may approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.27}{\eV}) at a mean-field cost.
-This method could then provide an improved starting point to recover the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.41}{\eV}).
+If that is the case, then a state-specific approach of such pCCD formulation adapted to open-shells may approach the accuracy of $\Delta$sCI1 (MAE of \SI{0.28}{\eV}) at a mean-field cost.
+This method could then provide an improved starting point to recover the remaining weak correlation than if starting from the (also mean-field though less accurate) $\Delta$CSF model (MAE of \SI{0.43}{\eV}).
 To the best of our knowledge, there has been a single proposed extension of pCCD to open-shell systems, \cite{Boguslawski_2021}
@@
 
-For the full set of 48 excited states for closed-shell molecules and 19 excited states for open-shell radicals,
+For the full set of 50 excited states for closed-shell molecules and 19 excited states for open-shell radicals,
 total energies and excitation energies obtained with
@@
 and the reference excitation energies and corresponding methods.
+\fk{For a subset of 27 excited states for closed-shell molecules and 15 excited states for radicals,
+additional total energies and excitation energies obtained with lower-order state-specific hCI and eCI models without enforcing spin pure states.}
 For a subset of 16 excited states for closed-shell molecules and 6 excited states for radicals,
@@
 Moving to the case where the number of spin-up electrons ($\Na$) and spin-down electrons ($\Nb$) differ by one ($\Na=\Nb+1$), which is illustrated in the center panel of Fig.~\ref{fig:appendix},
-the number of excited determinants generated exclusively by exciting $e$ electrons %$\Ndet(e)$, 
+the number of excited determinants generated exclusively by exciting $e$ electrons,
 is analogous to Eq.~\eqref{eq:e_closed}, being given by

### 2023-08-28 Manuscript/manuscript.tex
TOC
validating fabris changes

@@
 In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
-%\bigskip
-%\begin{center}
-%        \boxed{\includegraphics[keepaspectratio,width=2in]{TOC}}
-%\end{center}
-%\bigskip
+\bigskip
+\begin{center}
+        \boxed{\includegraphics[keepaspectratio,width=3.25in]{TOC}}
+\end{center}
+\bigskip
 \end{abstract}
@@
 The spin-contamination problem and its solution are therefore equivalent to that encountered in eCI. \cite{Maurice_1996}
-\fk{Here, we have employed this procedure and considered pure spin states.
-For the calculation of excitation energies, however, we have also assessed the effect of not imposing this condition.}
+Here, we have employed this procedure and considered pure spin states.
+For the calculation of excitation energies, however, we have also assessed the effect of not imposing this condition.
 
@@
 
-\fk{For the state-specific models, we have further assessed the impact of not imposing spin eigenstates (the results and statistics can be found in the {\SupInf}).
+For the state-specific models, we have further assessed the impact of not imposing spin eigenstates (the results and statistics can be found in the {\SupInf}).
 This amounts to not including the appropriate spin-flip determinants lying above the hierarchy or excitation degree of choice.
-The effect is minimal for both excitations of both closed- and open-shell systems, in the $\Delta$hCI1.5, $\Delta$hCI2, and $\Delta$CISD models.
+The effect is minimal for the excitation energies of both closed- and open-shell systems, in the $\Delta$hCI1.5, $\Delta$hCI2, and $\Delta$CISD models.
 The mean difference on the individual excitation energies lies between \SI{0.01}{\eV} and \SI{0.02}{\eV} for the latter two and one order of magnitude less for $\Delta$hCI1.5.
-Similarly small effects are seen in the global statistics.
+Similarly, small effects are seen in the global statistics.
 In turn, $\Delta$hCI1 is more affected, which would be expected given it is a low order model.
@@
 and does not improve the computed excitation energies obtained with the more competitive $\Delta$CISD and $\Delta$hCI2 models.
-The sCI models, discussed in the next subsection, are naturally spin eigenstates, and there is no need to enforce that.}
+The sCI models, discussed in the next subsection, are naturally spin eigenstates, and there is no need to enforce that.
 
@@
 and the reference excitation energies and corresponding methods.
-\fk{For a subset of 27 excited states for closed-shell molecules and 15 excited states for radicals,
-additional total energies and excitation energies obtained with lower-order state-specific hCI and eCI models without enforcing spin pure states.}
+For a subset of 27 excited states for closed-shell molecules and 15 excited states for radicals,
+additional total energies and excitation energies obtained with low-order state-specific hCI and eCI models without enforcing spin pure states.
 For a subset of 16 excited states for closed-shell molecules and 6 excited states for radicals,

### 2023-08-28 Manuscript/manuscript.tex
fixing and adding references

@@
 \usepackage[version=4]{mhchem}
-%\usepackage{natbib}
+\usepackage{natbib}
 \bibliographystyle{achemso}
@@
 single-reference and multiconfigurational self-consistent field,
-\cite{Ziegler_1977,Burton_2021,Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Burton_2022,Hanscam_2022,Kossoski_2023,Marie_2023}
+\cite{Ziegler_1977,Burton_2021,Shea_2018,Tran_2019,Tran_2020,Hardikar_2020,Shea_2020,Burton_2021,Burton_2022,Hanscam_2022,Kossoski_2023,Marie_2023,Tran_2023}
 density-functional theory,
-\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Hardikar_2020,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022,Schmerwitz_2023}
+\cite{Filatov_1999,Kowalczyk_2011,Kowalczyk_2013,Gilbert_2008,Barca_2018,Hait_2020,Hait_2021,Zhao_2020,Levi_2020,Carter-Fenk_2020,Toffoli_2022,Schmerwitz_2022,Schmerwitz_2023}
+perturbation theory, \cite{Clune_2020,Zhao_2020b,Clune_2023}
+quantum Monte Carlo, \cite{Scemama_2018a,Scemama_2018b,Dash_2018,Dash_2019,Dash_2021,Cuzzocrea_2022,Shepard_2022,Otis_2020,Otis_2023}
 and coupled cluster (CC)
-\cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023}
+\cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023,Tuckman_2023}
 methods.

### 2023-08-28 Manuscript/Cover_Letter/Cover_Letter.tex
Cover Letter

\documentclass[10pt]{letter}
\usepackage{UPS_letterhead,color,mhchem,mathpazo,ragged2e}
\newcommand{\alert}[1]{\textcolor{red}{#1}}

\usepackage[
        colorlinks=true,
    citecolor=blue,
    breaklinks=true
        ]{hyperref}
\urlstyle{same}

\begin{document}

\begin{letter}%
{To the Editors of the The Journal of Chemical Theory and Computation}

\opening{Dear Editors,}

\justifying
Please find enclosed our manuscript entitled \textit{``Seniority and hierarchy configuration interaction for radicals and excited states''}, 
which we would like you to consider as a Regular Article to the \textit{Journal of Chemical Theory and Computation}.
This contribution has never been submitted in total nor in parts to any other journal, and has been seen and approved by all authors.

Describing static and dynamic correlation on an equal footing is an ever-present goal in the development of novel electronic structure methods.
With this motivation in mind, we have recently proposed hierarchy configuration interaction (hCI)
[\href{https://doi.org/10.1021/acs.jpclett.2c00730}{\textit{J.~Phys.~Chem.~Lett.}~\textbf{2022}, \textit{13}, 4342}],
a new route that presented superior performance to the traditional excitation-based and the more recent seniority-based alternatives.
Despite the promising initial results, hCI had only been defined for ground-state closed-shell systems.

In the present contribution, we report a general formulation of hCI, significantly expanding its domain of applicability.
By tackling challenging and diverse chemical problems, including the dissociation of radical species and the excited states of closed- and open-shell systems,
hCI is extensively gauged against other configuration interaction (CI) alternatives.
The present findings establish hCI as an appealing method for recovering static and dynamic correlation.
They further encourage the development of hierarchy-based coupled-cluster methods, along with corresponding equation-of-motion formulations for excited states.

This contribution also concerns seniority-based CI for excited states.
Despite the growing interest seniority-based methods have attracted over the past decade,
here we demonstrate its poor performance in the calculation of excitation energies for closed-shell systems.
The results are more encouraging for excitations of open-shell systems,
which is relevant for the pursuit of seniority-based coupled cluster methods.

We suggest Paul Johnson, Gianluca Levi, Hannes Jonsson, Stijn de Baerdemacker, Eric Neuscamman, Katharina Boguslawski, and Peter Knowles as potential referees.
We look forward to hearing from you soon.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2023-08-28 Manuscript/Cover_Letter/Cover_Letter.tex
response letter

@@
 
-We suggest Paul Johnson, Gianluca Levi, Hannes Jonsson, Stijn de Baerdemacker, Eric Neuscamman, Katharina Boguslawski, and Peter Knowles as potential referees.
+We suggest Paul Johnson, Gianluca Levi, Hannes Jonsson, Stijn de Baerdemacker, Eric Neuscamman, and Katharina Boguslawski as potential referees.
 We look forward to hearing from you soon.

### 2023-08-28 Manuscript/manuscript.tex
more refs

@@
 This definition ensures that all classes of determinants whose number share the same scaling with system size are included at the same hierarchy $h$.
+\fk{This key feature distinguishes hCI from previous schemes combining excitation and seniority. \cite{Alcoba_2014,Raemdonck_2015,Alcoba_2018}
+By allowing for higher-order excitations of paired electrons (as explained in detail below), hCI is reminiscent of perfect pairing models for closed-shell systems.
+\cite{Hurley_1953,Cullen_1996,VanVoorhis_2000,Parkhill_2009,Parkhill_2010,Lehtola_2016,Lehtola_2018}}
 For different properties and closed-shell molecular systems,

### 2023-08-28 Manuscript/manuscript.tex
capital letters
OK

@@
 
-\title{Seniority and hierarchy configuration interaction for radicals and \\ excited states}
+\title{Seniority and Hierarchy Configuration Interaction for Radicals and \\ Excited States}
 
@@
 This definition ensures that all classes of determinants whose number share the same scaling with system size are included at the same hierarchy $h$.
-\fk{This key feature distinguishes hCI from previous schemes combining excitation and seniority. \cite{Alcoba_2014,Raemdonck_2015,Alcoba_2018}
+This key feature distinguishes hCI from previous schemes combining excitation and seniority. \cite{Alcoba_2014,Raemdonck_2015,Alcoba_2018}
 By allowing for higher-order excitations of paired electrons (as explained in detail below), hCI is reminiscent of perfect pairing models for closed-shell systems.
-\cite{Hurley_1953,Cullen_1996,VanVoorhis_2000,Parkhill_2009,Parkhill_2010,Lehtola_2016,Lehtola_2018}}
+\cite{Hurley_1953,Cullen_1996,VanVoorhis_2000,Parkhill_2009,Parkhill_2010,Lehtola_2016,Lehtola_2018}
 For different properties and closed-shell molecular systems,

### 2023-08-28 Manuscript/Cover_Letter/Cover_Letter.tex
more refs
capital letters

@@
 \justifying
-Please find enclosed our manuscript entitled \textit{``Seniority and hierarchy configuration interaction for radicals and excited states''}, 
+Please find enclosed our manuscript entitled \textit{``Seniority and Hierarchy Configuration Interaction for Radicals and Excited States''}, 
 which we would like you to consider as a Regular Article to the \textit{Journal of Chemical Theory and Computation}.

### 2023-10-05 Manuscript/Response_Letter/Response_Letter.tex
start Response Letter

\documentclass[10pt]{letter}
\usepackage{UPS_letterhead,xcolor,mhchem,ragged2e,hyperref}
\usepackage{physics,siunitx}
\newcommand{\alert}[1]{\textcolor{red}{#1}}
\definecolor{darkgreen}{HTML}{009900}

\newcommand{\yann}{\textcolor{green}}

\begin{document}

\begin{letter}%
{To the Editors of the \textit{Journal of Chemical Theory and Computation},}

\opening{Dear Editors,}

\justifying
Please find attached a revised version of the manuscript entitled
\begin{quote}
	\textit{``Seniority and Hierarchy Configuration Interaction for Radicals and Excited States''}
\end{quote}

We thank the reviewers for their support to publication of the present manuscript.
Our detailed responses to their comments can be found below.
For convenience, changes of the revised manuscript are highlighted in red in a separate file.

We look forward to hearing from you.

\closing{Sincerely, the authors.}

\newpage

%%% REVIEWER 1 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#1}

{In this manuscript, Loos and coworkers report the extension of the hCI method (J. Phys. Chem. Lett. 2022, 13, 4342) towards calculating the open-shell and excited-state chemical species. 
This method can be seen as a partitioning of the FCI space that combines the excitation-based CI (eCI, which is a conventional truncated CI) and the seniority-based CI (sCI).
The authors report various benchmark data for radicals (\ce{OH}, \ce{CN}, vinyl, \ce{H7}) and excited states (the molecules at the geometries from the QUEST database).
The authors also introduced the second-order Epstein-Nesbet perturbative corrections (EN2) in the hCI calculations,
and the impact of such correction is described for the closed-shell systems as well.
Overall, this work reasonably demonstrates applications of novel electronic-structure methods for diverse systems. 
The paper is interesting, well-written, and straightforward to follow, and I think this paper deserves publication in J. Chem. Theory Comput. after minor improvements. 
I have only a couple of brief comments below:
}

\alert{
	aaa
}

{This work presents the partitioning of the Hilbert space. Although the authors present the number
of determinants in the appendix, can the authors also provide the number of the FCI determinants
so that readers can easily evaluate the efficiency of the xCI calculations?
}

\alert{
	aaa
}

{Figures 3 and 4: The data is shown with respect to the number of determinants, which makes the
reader to evaluate the different methods efficiently. However, for readability, I think it will also be
helpful to provide n in eCIn, sCIn, and hCIn in these figures.
}

\alert{
	aaa
}

{In sections IV-C and IV-D: The EN2 corrections are not discussed at all in the excited states, making
it unnatural compared to the presentation of the radical systems. While it might be challenging to
perform the additional calculations for all the systems, I think it will be helpful for some sample
systems to investigate the impact of EN2 correction in hCI and DhCI calculations (maybe in the
same fashion presented in Figures 3 and 4).
}

\alert{
	aaa
}

\clearpage

%%% REVIEWER 2 %%%
\noindent \textbf{\large Authors' answer to Reviewer \#2}

{The authors present a generalization of the hierarchy configuration interaction (hCI) method to an arbitrary reference state, which allows them to extend their previously developed hCI model to open-shell systems and general excited-state electronic structures. They test the performance of their proposed extension against various test sets containing (small) molecules for which reliable reference data is available. Their numerical analysis and discussion are in-depth, while their methodology (the theory part) is comprehensive. Their numerical study allows them to obtain reliable energies and spectroscopic properties for their more elaborate hCI models. This manuscript is of potential interest to the readership of the Journal of Chemical Theory and Computation. However, a few remarks need to be addressed before publication.
}

\alert{
aaa
}

{1) The sentence ``[...] whose number Ndet share the same scaling with the number of basis functions N [...]'' is rather difficult to understand. The authors need to either rephrase or elaborate more.
}

\alert{
aaa
}

{2) the authors need to include more information concerning the fitting procedure in the main manuscript for reproducibility reasons. In the SI, they state that they used a Morse function. Why did the authors use a Morse function and not, for instance, a generalized Morse function, etc... Furthermore, it is unclear how they fitted their vibrational frequencies.
}

\alert{
aaa
}

{3) Some of the PES displayed in the SI are not smooth and rather kinky. How does the lack of smoothness affect the fitting procedure and, hence, the vibrational frequencies?
}

\alert{
aaa
}

{4) The authors need to mention (or draw) the molecules of their excited state study. Not everyone might be familiar with the QUEST database. It would be insightful to mention the character of the excited states. I assume these are all singly excited states (no double excitation character?). Can the authors provide additional information regarding their nature (local excitation, charge transfer, Rydberg, etc.)? Specifically, what kind of excited states can be described well or less well with hCI methods?
}

\alert{
aaa
}

{5) For reasons of transparency, the authors should define MSE, MAE, RMSE, and SDE using equations, even if it might be obvious.
}

\alert{
aaa
}

{6) The authors keep mentioning ``reference theoretical values.'' Please mention what kind of reference values were used.
}

\alert{
aaa
}

{7) The authors look at several small-sized molecules for which FCI can still be determined. There is nothing wrong with that. However, I am missing numerical studies on slightly larger systems, which are known to be challenging in computational chemistry. I recommend adding some ``larger'' molecules, which feature open- and closed-shell ground and excited states. Examples are cyclobutadiene (its automatization problem) and o-, m-, and p-benzynes. These systems are notoriously difficult to study and extensively discussed in the literature. Including just those four molecules would add more value to the paper and allow a direct comparison to other state-of-the-art computational methods.
}

\alert{
aaa
}

\end{letter}
\end{document}

### 2023-10-17 Manuscript/manuscript.tex
sentence about repo
ready to submit
update table

@@
 \usepackage{natbib}
+\usepackage{comment}
 \bibliographystyle{achemso}
@@
 
+\begin{comment}
 \begin{table}[ht!]
@@
 \hline
-$\Delta$hCI1      & 50  & -1.29 & 1.47 & 1.63 & 1.00 \\
+$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67 \\
 $\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15 \\
@@
 \hline
-sCI2/sCI2         & 50  & +1.22 & 1.29 & 1.45 & 0.78 \\
-sCI2/sCI0         & 50  & -0.47 & 0.62 & 0.92 & 0.80 \\
-$\Delta$sCI2/sCI2 & 50  & +0.65 & 0.78 & 0.89 & 0.60 \\
+sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68 \\
+sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66 \\
+$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60 \\
 $\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62 \\
@@
 \end{table}
+\end{comment}
+
+\begin{table*}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
+Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+}
+\label{tab:stat1}
+\begin{ruledtabular}
+\begin{tabular}{ldddddlddddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
+\hline
+CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  &\fk{CIS+EN2}               & \fk{48} &       &      &      &      \\
+CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84  & & & & & & \\
+CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14  & & & & & & \\
+CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08  & & & & & & \\
+%
+\hline
+hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  &\fk{hCI1+EN2}              & \fk{50}  & \fk{+0}.\fk{65}  & \fk{0}.\fk{65} & \fk{0}.\fk{84} & \fk{0}.\fk{53} \\
+hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  &\fk{hCI1.5+EN2}            & \fk{50}  & \fk{+0}.\fk{62}  & \fk{0}.\fk{63} & \fk{0}.\fk{76} & \fk{0}.\fk{44} \\
+hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76  & & & & & & \\
+hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66  & & & & & & \\
+hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08  & & & & & & \\
+hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07  & & & & & & \\
+%
+\hline
+$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  &\fk{$\Delta$CSF+EN2}       & \fk{50}  &        &      &      &      \\
+$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
+$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11  & & & & & & \\
+$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
+%
+\hline
+$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  &\fk{$\Delta$hCI1+EN2}      & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{24} & \fk{0}.\fk{30} & \fk{0}.\fk{27} \\
+$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
+$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{00}  & \fk{0}.\fk{06} & \fk{0}.\fk{10} & \fk{0}.\fk{09} \\
+$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
+$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10  & & & & & & \\
+$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05  & & & & & & \\
+%
+\hline
+sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  &\fk{sCI2/sCI2+EN2}         & \fk{50}  & \fk{+0}.\fk{60}  & \fk{0}.\fk{60} & \fk{0}.\fk{79} & \fk{0}.\fk{51} \\
+sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  &\fk{sCI2/sCI0+EN2}         & \fk{50}  & \fk{+0}.\fk{51}  & \fk{0}.\fk{52} & \fk{0}.\fk{69} & \fk{0}.\fk{47} \\
+$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  &\fk{$\Delta$sCI2/sCI2+EN2} & \fk{50}  & \fk{+0}.\fk{21}  & \fk{0}.\fk{26} & \fk{0}.\fk{35} & \fk{0}.\fk{28} \\
+$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  &\fk{$\Delta$sCI2/sCI0+EN2} & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{23} & \fk{0}.\fk{27} & \fk{0}.\fk{24} \\
+%
+\end{tabular}
+\end{ruledtabular}
+\end{table*}
 
@@
 \end{table}
+\end{comment}
+
+\begin{table*}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
+Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
+}
+\label{tab:stat2}
+\begin{ruledtabular}
+\begin{tabular}{ldddddlddddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
+\hline
+CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50  & \fk{CIS+EN2}            & \fk{19} &       &      &      &      \\
+CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17  & & & & & & \\
+CISDT             &  6 & +0.06 & 0.07 & 0.09 & 0.06  & & & & & & \\
+CISDTQ            &  6 & +0.08 & 0.08 & 0.11 & 0.08  & & & & & & \\
+%
+\hline
+hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64  & \fk{hCI1+EN2}           & \fk{19} & \fk{+0}.\fk{27} & \fk{0}.\fk{28} & \fk{0}.\fk{36} & \fk{0}.\fk{24} \\
+hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68  & \fk{hCI1.5+EN2}         & \fk{19} & \fk{+0}.\fk{22} & \fk{0}.\fk{25} & \fk{0}.\fk{33} & \fk{0}.\fk{24} \\
+hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90  & & & & & & \\
+hCI2.5            &  6 & +0.03 & 0.05 & 0.06 & 0.05  & & & & & & \\
+hCI3              &  6 & +0.00 & 0.04 & 0.05 & 0.05  & & & & & & \\
+hCI3.5            &  6 & +0.01 & 0.03 & 0.05 & 0.04  & & & & & & \\
+%
+\hline
+$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59  & \fk{$\Delta$CSF+EN2}    & \fk{19} &       &      &      &      \\
+$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21  & \fk{$\Delta$CISD+EN2}   & \fk{19} & \fk{-0}.\fk{02} & \fk{0}.\fk{04} & \fk{0}.\fk{06} & \fk{0}.\fk{06} \\
+$\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08  & & & & & & \\
+$\Delta$CISDTQ    &  6 & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
+%
+\hline
+$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44  & \fk{$\Delta$hCI1+EN2}   & \fk{19} & \fk{+0}.\fk{02} & \fk{0}.\fk{17} & \fk{0}.\fk{29} & \fk{0}.\fk{29} \\
+$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69  & \fk{$\Delta$hCI1.5+EN2} & \fk{19} & \fk{+0}.\fk{00} & \fk{0}.\fk{10} & \fk{0}.\fk{15} & \fk{0}.\fk{15} \\
+$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16  & \fk{$\Delta$hCI2+EN2}   & \fk{19} & \fk{-0}.\fk{00} & \fk{0}.\fk{05} & \fk{0}.\fk{08} & \fk{0}.\fk{08} \\
+$\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10  & & & & & & \\
+$\Delta$hCI3      &  6 & -0.06 & 0.06 & 0.08 & 0.05  & & & & & & \\
+$\Delta$hCI3.5    &  6 & -0.02 & 0.02 & 0.03 & 0.02  & & & & & & \\
+%
+\hline
+sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53  & \fk{sCI1+EN2}           & \fk{19} & \fk{+0}.\fk{23} & \fk{0}.\fk{25} & \fk{0}.\fk{32} & \fk{0}.\fk{21} \\
+$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37  & \fk{$\Delta$sCI1+EN2}   & \fk{19} & \fk{+0}.\fk{06} & \fk{0}.\fk{12} & \fk{0}.\fk{20} & \fk{0}.\fk{19} \\
+\end{tabular}
+\end{ruledtabular}
+\end{table*}
 
@@
 We refer back to Table \ref{tab:stat1} to discuss the performance of sCI for excited states of closed-shell systems.
-The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.29}{\eV}.
+The first model, sCI2/sCI2, systematically overestimates the excitation energies, producing a MAE of \SI{1.35}{\eV}.
 The same two factors discussed above for hCI explain such large errors.
@@
 A possible solution to this unbalance is to restrict the determinants to a maximum seniority number of $s=0$ for the ground state and $s=2$ for the excited state calculation.
-This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.62}{\eV}, compared to \SI{1.29}{\eV} of sCI2/sCI2.
+This is precisely the sCI2/sCI0 model, which delivers a MAE of \SI{0.55}{\eV}, compared to \SI{1.35}{\eV} of sCI2/sCI2.
 However, when going to state-specific orbitals, $\Delta$sCI2/sCI0 systematically undershoots the excitation energies and provides a MAE of \SI{1.04}{\eV}.
@@
 We did not pursue these ideas here, and their impact on the excitation energies computed with hCI and sCI remains an open question.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{\fk{EN2 correction for excited states}}
+\label{sec:res_E}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{comment}
+
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
+Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+}
+\label{tab:stat1}
+\begin{ruledtabular}
+\begin{tabular}{lddddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
+\hline
+CIS+EN2               & 48 &       &      &      &      \\
+\hline
+hCI1+EN2              & 50  & +0.65  & 0.65 & 0.84 & 0.53 \\
+hCI1.5+EN2            & 50  & +0.62  & 0.63 & 0.76 & 0.44 \\
+%
+\hline
+$\Delta$CSF+EN2       & 50  &        &      &      &      \\
+$\Delta$CISD+EN2      & 50  & +0.02  & 0.06 & 0.09 & 0.09 \\
+%
+\hline
+$\Delta$hCI1+EN2      & 50  & +0.12  & 0.24 & 0.30 & 0.27 \\
+$\Delta$hCI1.5+EN2    & 50  & -0.01  & 0.13 & 0.19 & 0.19 \\
+$\Delta$hCI2+EN2      & 50  & +0.00  & 0.06 & 0.10 & 0.09 \\
+%
+\hline
+sCI2/sCI2+EN2         & 50  & +0.51  & 0.60 & 0.79 & 0.51 \\
+sCI2/sCI0+EN2         & 50  & +0.60  & 0.52 & 0.69 & 0.47 \\
+$\Delta$sCI2/sCI2+EN2 & 50  & +0.20  & 0.25 & 0.34 & 0.28 \\
+$\Delta$sCI2/sCI0+EN2 & 50  & +0.11  & 0.22 & 0.26 & 0.24 \\
+%
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+\begin{table}[ht!]
+\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
+Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
+}
+\label{tab:stat2}
+\begin{ruledtabular}
+\begin{tabular}{lddddd}
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
+\hline
+CIS+EN2            & 19 &       &      &      &      \\
+%
+\hline
+hCI1+EN2           & 19 & +0.27 & 0.28 & 0.36 & 0.24 \\
+hCI1.5+EN2         & 19 & +0.22 & 0.25 & 0.33 & 0.24 \\
+%
+\hline
+$\Delta$CSF+EN2    & 19 &       &      &      &      \\
+$\Delta$CISD+EN2   & 19 & -0.02 & 0.04 & 0.06 & 0.06 \\
+%
+\hline
+$\Delta$hCI1+EN2   & 19 & +0.02 & 0.17 & 0.29 & 0.29 \\
+$\Delta$hCI1.5+EN2 & 19 & +0.00 & 0.10 & 0.15 & 0.15 \\
+$\Delta$hCI2+EN2   & 19 & -0.00 & 0.05 & 0.08 & 0.08 \\
+%
+\hline
+sCI1+EN2           & 19 & +0.23 & 0.25 & 0.32 & 0.21 \\
+$\Delta$sCI1+EN2   & 19 & +0.06 & 0.12 & 0.20 & 0.19 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+\end{comment}
 
@@
 corrected with the standard and the renormalized EN2 perturbative energy.
+The data associated with the PECs and derived properties can be found at \url{https://github.com/kossoski/open_shell_hCI}.
 

### 2023-10-18 Manuscript/Response_Letter/Response_Letter.tex
saving work

@@
 \newcommand{\yann}{\textcolor{green}}
+\newcommand{\Ndet}{N_\text{det}}
 
@@
 \alert{
-	aaa
+We thank the reviwer for the supportive comments.
 }
@@
 \alert{
-	aaa
+The number of determinants in FCI and in each approximate CI model is now presented in Tables S1 and S2 of the Supporting Information.
 }
@@
 \alert{
-	aaa
+We have experimented including the values of $n$ to the figures, as suggested by the reviewer.
+However, we consider that they become somewhat clutered and less readable, and therefore we prefer to leave them unchanged.
+With the intention of helping the reader, we have been supplemented the following sentence to the captions of Figures 3 (4):
+``Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (sCI1, sCI3, sCI5) (blue).''
 }
@@
 \alert{
-	aaa
+We agree with the reviewer's suggestion that including the EN2 correction for the excited state calculations is a natural and important addition to the manuscript.
+We have performed additional calculations for the full set of states with the different CI models surveyed here (except for the more computationally demanding higher-order models).
+These new results were included in Tables II, III, and in the Supporting Information, while a new subsection (IV.E), starting on page x, is devoted to their discussion.
 }
@@
 \alert{
-aaa
+We thank the reviwer for the positive appreciation of our work.
 }
 
-{1) The sentence ``[...] whose number Ndet share the same scaling with the number of basis functions N [...]'' is rather difficult to understand. The authors need to either rephrase or elaborate more.
+{1) The sentence ``[...] whose number $\Ndet$ share the same scaling with the number of basis functions $N$ [...]'' is rather difficult to understand.
+The authors need to either rephrase or elaborate more.
 }
@@
 \alert{
-aaa
+This sentence has been rephrased and now reads, on page x,
+``The number of determinants $\Ndet$ within a given class scales polynomially with the number of basis functions $N$, with the exponent depending on the specific class.
+For instance, for the class of doubly-excited determinants with no unpaired electrons ($e=2$ and $s=0$), $\Ndet = \order*{N^2}$.
+hCI was defined such that, at a given hierarchy $h$, all classes of determinants presenting a scaling of $\Ndet = \order*{N^{2h}}$ or less are accounted for.''
 }
@@
 \alert{
-aaa
+The reviewer's questions are addressed on page x of the revised Supporting Information, reading
+``By fitting the computed potential energy curves with a Morse potential $V(r) = D [1 - e^{-a(r-r_e)}]^2 + C $,
+we obtained equilibrium geometries $r_e$, force constants $2Da^2$, and harmonic vibrational frequencies $\left( Da^2/2\pi^2\mu \right)^{1/2}$,
+where $\mu$ is the reduced mass.
+Since the fitting is restricted to the equilibrium region, employing more elaborate functions would be inconsequential, such that the adopted Morse potential is well justified.
+The fitting was performed with the Marquardt-Levenberg algorithm implemented in Gnuplot.''
+%Since the fitting is restricted to the equilibrium region (the ranges are given in the Supporting Information),
+%using a more elaborate functional form would only produce minimal changes to the obtained fitted values. A Morse potential is thus enough for our purposes.
+%Moreover, any residual bias from the fitting function would be inconsequential to the goal of comparing to the (equally fitted) FCI values.
+We prefer to keep these details in the Supporting Information rather than in the main text.
 }
@@
 \alert{
-aaa
+\ce{CN} is the only system for which some of the potential energy curves are not smooth around the equilibrium geometry.
+This is mentioned on page x of the revised manuscript, reading
+``For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
+and somewhat less reliable equilibrium properties in these specific cases.''
 }
@@
 
+\clearpage
+
 {5) For reasons of transparency, the authors should define MSE, MAE, RMSE, and SDE using equations, even if it might be obvious.
@@
 \alert{
-aaa
+The statistical measures are now defined on page x of the Supporting Information, which are refered to on page x of the main text, reading
+``For completeness, the defition of these statistical measures can be found in the Supporting Information.''
 }
@@
 \alert{
-aaa
+This information is provided in the Supporting Information, which is stated on page x of the revised manuscript, reading
+``For each excited state considered here, the reference value and the corresponding method to obtain it (of high-order coupled cluster or extrapolated FCI quality) can be found in the Supporting Information.''
 }

### 2023-10-18 Manuscript/manuscript.tex
saving work

@@
 \cite{Boguslawski_2016b,Boguslawski_2016c,Boguslawski_2019,Nowak_2019,Kossoski_2021,Marie_2021,Tecmer_2022,Rishi_2023,Nowak_2023,Fecteau_2022}
-Our fifth and last goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state-based and state-specific sCI models for excited states.
+Our fifth goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state-based and state-specific sCI models for excited states.
+\fk{Finally, in Sec.~\ref{sec:res_E} we assess how the excitation energies are impacted by the EN2 perturbative correction to hCI and sCI models, our sixth and last goal.}
 Section \ref{sec:conclusion} closes the present contribution with the main conclusions and perspectives.
@@
 we refer to the class of an excited determinant as the combination of the number of electrons and seniority in each orbital subspace.
-hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
-which means moving diagonally in the seniority-excitation map (denoted by the color tones in Fig.~\ref{fig:hCI}).
+%hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
+\fk{The number of determinants $\Ndet$ within a given class scales polynomially with the number of basis functions $N$, with the exponent depending on the specific class.
+For instance, for the class of doubly-excited determinants with no unpaired electrons ($e=2$ and $s=0$), $\Ndet = \order*{N^2}$.
+hCI was defined such that, at a given hierarchy $h$, all classes of determinants presenting a scaling of $\Ndet = \order*{N^{2h}}$ or less are accounted for.}
+This means moving diagonally in the seniority-excitation map (denoted by the color tones in Fig.~\ref{fig:hCI}).
 In comparison, eCI spans the map horizontally (top to bottom), whereas sCI does it vertically (left to right).
@@
 (middle and right panels of Fig.~\ref{fig:hCI}, respectively).
-The computed excitation energies were benchmarked against the reference values of FCI quality provided in the QUEST database. \cite{Veril_2021}
+The computed excitation energies were benchmarked against the reference theoretical values provided in the QUEST database. \cite{Veril_2021}
+\fk{For each excited state considered here, the reference value and the method used to obtain it (of high-order coupled cluster or extrapolated FCI quality) can be found in the {\SupInf}.}
 
@@
 First, we discuss the results for the bare CI models (hCI, eCI, and sCI) for the open-shell systems (represented by the dark tones in the figures),
-leaving the discussion about the perturbative correction (light tones) for both open-shell and closed-shell systems to the next subsection.
+leaving the discussion about the perturbative correction (light tones) for both open-shell and closed-shell systems to subsection Sec.~\ref{sec:res_B}.
 
@@
 Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
-as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red), and sCI (blue) models,
+as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red), and sCI (blue) routes,
 with (light tones) and without (dark tones) the standard EN2 perturbative correction.
+\fk{Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI1, sCI3, and sCI5 (blue).}
 The dashed lines represent the FCI results.}
@@
 \caption{
-Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red) and sCI (blue) models,
+Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red) and sCI (blue) routes,
 with (light tones) and without (dark tones) the standard EN2 perturbative correction.
+\fk{Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (blue).}
 The dashed lines represent the FCI results.}
@@
 Finally, hCI systematically outperforms eCI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
+\fk{For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
+and somewhat less reliable equilibrium properties in these specific cases.}
 Given the overall slight superiority of hCI for the equilibrium properties, it also manages to account for more dynamic correlation than eCI,
@@
 with respect to the reference theoretical values for the excitation energies.
+\fk{For completeness, the defition of these statistical measures can be found in the {\SupInf}.}
 For the lower-order models, the calculations were performed for 50 (closed-shell) and 19 (open-shell) excited states.
@@
 and on a similar comparison between ground-state and state-specific approaches for hCI and sCI.
+\fk{The bare (unperturbed) hCI and sCI models are discussed in this section and in Sec.~\ref{sec:res_D}, respectively,
+whereas their EN2 perturbed analogues are left for Sec.~\ref{sec:res_E}.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\fk{EN2 correction for excited states}}
+\subsection{\fk{hCI and sCI plus EN2 correction for excited states}}
 \label{sec:res_E}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\fk{
+The statistical measures for the CI models corrected with the renormalized EN2 correction are presented in 
+Table~\ref{tab:stat1} (closed-shell systems) and Table~\ref{tab:stat2} (open-shell systems),
+along with the results for the bare (unperturbed) CI models.
+We have not considered the EN2 correction for the higher-order models.
+}
+
+\fk{
+The EN2 correction decreases the four statistical metrics for both closed- and open-shell systems and across most CI models (the exceptions are discussed in the following).
+Importantly, for a given scaling (see Table~\ref{tab:scaling}), whether the bare CI model or the lower-order CI model plus EN2 correction produces the smaller errors depends on the truncation order,
+and also on the type of excitations (from closed- or open-shells).
+For instance, $\Delta$hCI2 and $\Delta$hCI1+EN2 have the same $\order*{N^6}$ computational cost
+but the former has smaller MAEs, specially for the open-shells (\SI{0.10}{\eV} compared to \SI{0.17}{\eV}), but also for the closed-shells (\SI{0.20}{\eV} compared to \SI{0.24}{\eV}).
+In turn, $\Delta$hCI1.5+EN2 is significantly more accurate than $\Delta$hCI2.5 for closed-shells, 
+with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, though for open-shells both models produce a MAE of \SI{0.10}{\eV}.
+The same is found at the next order.
+$\Delta$hCI2+EN2 is much more accurate than $\Delta$hCI3 for closed-shells, with MAEs of \SI{0.06}{\eV} and \SI{0.22}{\eV},
+whereas for the open-shell transitions, they deliver comparable MAEs of \SIrange{0.05}{0.06}{\eV}.
+%We recall that the small prefactor associated with the calculation of the EN2 correction generally makes it an attractive option.
+}
+
+\fk{EN2 with ground-state HF orbitals}
+
+\fk{hCI+EN2 vs eCI+EN2}
+
+\fk{sCI+EN2 vs sCI}
 
@@
 
-Equilibrium geometries of ethylene and vinyl; details about the fitting of potential energy curves;
+\fk{Definition of the statistical measures;}
+equilibrium geometries of ethylene and vinyl; details about the fitting of potential energy curves;
+\fk{number of determinants of each CI model;}
 potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},

### 2023-10-18 Manuscript/sup.tex
TOC
fixing and adding references
added data.xlsx
saving work

@@
 \usepackage[version=4]{mhchem}
-%\usepackage{natbib}
-%\bibliographystyle{achemso}
+\usepackage{natbib}
+\bibliographystyle{achemso}
 
 \maxdeadcycles=200
+
+\newcommand{\fk}[1]{\textcolor{orange}{#1}}
 
@@
 
-\title{Supporting Information for ``Seniority and hierarchy configuration interaction for radicals and excited states''}
+\title{Supporting Information for ``Seniority and Hierarchy Configuration Interaction for Radicals and Excited States''}
 
@@
 \maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\section{Statistical measures}
+%\label{sec:stat}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\fk{
+Given two sets of $N$ paired variables $x_i$ and $y_i$
+(in our case, the approximate and reference excitation energies),
+the corresponding mean signed error (MSE), mean absolute error (MAE), root-mean-square error (RMSE), and standard deviation of the errors (SDE) are defined as:
+}
+
+\begin{equation}
+	\fk{\text{MSE} = \frac{1}{N} \sum_{i=1}^{N} (x_i - y_i),}
+\end{equation}
+
+\begin{equation}
+	\fk{\text{MAE} = \frac{1}{N} \sum_{i=1}^{N} |x_i - y_i|,}
+\end{equation}
+
+\begin{equation}
+	\fk{\text{RMSE} = \sqrt{ \frac{1}{N} \sum_{i=1}^{N} (x_i - y_i)^2 },}
+\end{equation}
+
+\begin{equation}
+	\fk{\text{SDE} = \sqrt{ \frac{1}{N} \sum_{i=1}^{N} (x_i - \mu)^2 },}
+\end{equation}
+
+\fk{where}
+\begin{equation}
+	\fk{\mu = \frac{1}{N} \sum_{i=1}^{N} x_i.}
+\end{equation}
 
@@
 
+\clearpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-By fitting the computed potential energy curves with a Morse potential, we obtained harmonic vibrational frequencies and equilibrium geometries.
+\fk{
+By fitting the computed potential energy curves with a Morse potential $V(r) = D [1 - e^{-a(r-r_e)}]^2 + C $,
+we obtained equilibrium geometries $r_e$, force constants $2Da^2$, and harmonic vibrational frequencies $\left( Da^2/2\pi^2\mu \right)^{1/2}$,
+where $\mu$ is the reduced mass.
+Since the fitting is restricted to the equilibrium region, employing more elaborate functions would be inconsequential, such that the adopted Morse potential is well justified.
+The fitting was performed with the Marquardt-Levenberg algorithm implemented in Gnuplot.
+}
 The following intervals have been considered for the fitting:
@@
 
+\clearpage
+
 The non-parallelity error and distance error were computed for the following intervals:
@@
 \end{itemize}
+
+\clearpage
+
+\begin{table}[ht!]
+\caption{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
+with the cc-pVDZ basis set and frozen-core approximation.}
+\label{tab:ndet_open}
+\begin{ruledtabular}
+\begin{tabular}{lrrrr}
+method &     \ce{OH} &         \ce{CN} &                 vinyl &     \ce{H7} \\
+\hline
+RHF    &           1 &               1 &                     1 &           1 \\
+\hline
+CIS    &         131 &             280 &                   533 &         221 \\
+CISD   &       5 871 &          27 559 &               100 983 &      16 403 \\
+CISDT  &     110 031 &       1 196 384 &             8 620 123 &     491 695 \\
+CISDTQ &   1 035 501 &      27 468 434 &           397 093 838 &   7 015 800 \\
+\hline
+CIs1   &      58 140 &       8 288 280 &         1 160 068 104 &     209 440 \\
+CIs3   &   1 686 060 &     530 449 920 &       143 268 410 844 &   9 948 400 \\
+CIs5   &  12 267 540 &   9 233 143 920 &     4 974 952 064 004 & 107 338 000 \\
+\hline
+hCI1   &         299 &             658 &                 1 268 &         500 \\
+hCI1.5 &       3 491 &          11 809 &                35 813 &      11 009 \\
+hCI2   &      25 401 &         136 234 &               569 178 &      96 290 \\
+hCI2.5 &     162 265 &       1 417 234 &             8 899 178 &     832 850 \\
+hCI3   &     732 107 &      10 879 834 &           102 007 753 &   5 530 745 \\
+hCI3.5 &   2 706 807 &      71 632 834 &         1 011 139 788 &  23 829 890 \\
+hCI4   &   7 774 870 &     384 458 809 &         8 484 940 738 &  92 360 660 \\
+\hline
+FCI    &  45 070 128 & 446 075 229 600 & 4 672 770 853 882 482 & 342 696 200 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+\begin{table}[ht!]
+\caption{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{H4}, \ce{N2}, \ce{F2}, \ce{HF}, \ce{H8}, and ethylene,
+with the cc-pVDZ basis set and frozen-core approximation.}
+\label{tab:ndet_closed}
+\begin{ruledtabular}
+\begin{tabular}{lrrr}
+method & \ce{H4} &       \ce{N2} &         \ce{F2} \\
+\hline
+RHF    &       1 &             1 &               1 \\
+\hline
+CIS    &      73 &           211 &             267 \\
+CISD   &   1 675 &        15 436 &          25 138 \\
+CISDT  &  12 691 &       483 036 &       1 048 174 \\
+CISDTQ &  36 100 &     7 745 886 &      23 236 165 \\
+\hline
+CIs0   &     190 &        65 780 &         657 800 \\
+CIs2   &   7 030 &     6 972 680 &      88 145 200 \\
+CIs4   &  36 100 &   145 110 680 &   2 450 305 000 \\
+\hline
+hCI1   &     109 &           316 &             400 \\
+hCI1.5 &     757 &         2 836 &           3 592 \\
+hCI2   &   3 052 &        34 336 &          57 457 \\
+hCI2.5 &  17 740 &       311 536 &         531 469 \\
+hCI3   &  36 100 &     1 913 836 &       4 209 850 \\
+hCI3.5 &  36 100 &    11 489 836 &      28 628 650 \\
+hCI4   &  36 100 &    49 045 711 &     153 775 000 \\
+\hline
+FCI    &  36 100 & 4 327 008 400 & 432 700 840 000 \\
+\hline
+\hline
+& & & \\
+\hline
+\hline
+method &    \ce{HF} &       \ce{H8} &           ethylene \\
+\hline
+RHF    &          1 &             1 &                  1 \\
+\hline
+CIS    &        113 &           289 &                409 \\
+CISD   &      4 341 &        28 585 &             58 855 \\
+CISDT  &     68 405 &     1 174 345 &          3 731 535 \\
+CISDTQ &    531 595 &    23 805 835 &        124 764 480 \\
+\hline
+CIs0   &      3 060 &        91 390 &          3 838 380 \\
+CIs2   &    174 420 &    13 251 550 &        786 867 900 \\
+CIs4   &  1 845 180 &   358 705 750 &     33 086 835 600 \\
+\hline
+hCI1   &        169 &           433 &                613 \\
+hCI1.5 &      1 065 &         5 905 &              8 365 \\
+hCI2   &      9 255 &        62 605 &            134 590 \\
+hCI2.5 &     55 119 &       879 085 &          1 952 230 \\
+hCI3   &    227 109 &     5 659 525 &         17 306 800 \\
+hCI3.5 &    856 101 &    40 616 965 &        163 795 120 \\
+hCI4   &  2 216 460 &   213 974 380 &      1 112 401 240 \\
+\hline
+FCI    &  9 363 600 & 8 352 132 100 & 14 733 161 024 400 \\
+\end{tabular}
+\end{ruledtabular}
+\end{table}
 

### 2023-10-22 Manuscript/Response_Letter/Response_Letter.tex
saving work

@@
 These new results were included in Tables II, III, and in the Supporting Information, while a new subsection (IV.E), starting on page x, is devoted to their discussion.
+Additions to the Conclusion (page x) and the abstract have also been made.
 }
@@
 \alert{
-aaa
+The set of systems we considered is now presented in Table II, whreas on page x we describe the character of the excited states, reading
+``Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
+and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.''
+The details on each specific excited state can be found in The Supporting Information.
+New additions to text can be found on page x:
+``aa''
+The new subsection (IV.E) also discusses how hCI+EN2 models perform for different types of excited states.
 }

### 2023-10-22 Manuscript/manuscript.tex
saving work

@@
 We find that hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
-Employing the second-order Epstein-Nesbet perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
+Employing the second-order Epstein-Nesbet \fk{(EN2)} perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
 We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
 Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance between correlation of ground and excited states.
-State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.
+State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results,
+\fk{unless supplemented by the EN2 correction, which substantially improves their accuracy.}
 In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
@@
 
-The various CI models introduced here were further assessed based on calculated vertical excitation energies for 60 electronic states,
-from 18 closed-shell systems and 8 radicals, with geometries extracted from the QUEST database. \cite{Veril_2021}
+The various CI models introduced here were further assessed based on calculated vertical excitation energies for 69 electronic states,
+from 17 closed-shell systems and 13 radicals \fk{(shown in Table~\ref{tab:systems})}, with geometries extracted from the QUEST database. \cite{Veril_2021}
+\fk{Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
+and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.}
 The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.
+
+\begin{table}[ht!]
+\caption{\fk{Systems Considered in the Present Excited State Calculations.}}
+\label{tab:systems}
+{\color{orange}
+\begin{ruledtabular}
+\begin{tabular}{lll|lll}
+\mc{3}{c|}{Closed-shells}           & \mc{3}{c}{Open-shells} \\
+\hline
+glyoxal      & \ce{HCF}  & \ce{H2O} & \ce{BeH} & \ce{NH2} & \ce{HCO} \\
+acetaldehyde & \ce{HCCl} & \ce{H2S} & \ce{BeF} & \ce{PH2} & \ce{HOC} \\
+silylidene   & \ce{HPO}  & \ce{N2}  & \ce{BH2} & vinyl    & \ce{OH}  \\
+nitroxyl     & \ce{CF2}  & \ce{NH3} & \ce{CN}  & allyl    & \ce{CO+} \\
+ethylene     & \ce{BH}   & \ce{HCl} & \ce{CH3} &          & \\
+methanimine  & \ce{BF}   &          &          &          & \\
+\end{tabular}
+\end{ruledtabular}
+}
+\end{table}
+
 We employed the aug-cc-pVDZ basis set for systems having up to three non-hydrogen atoms and the 6-31+G(d) basis set for the larger ones.
@@
 
-\begin{comment}
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
-Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
-}
-\label{tab:stat1}
-\begin{ruledtabular}
-\begin{tabular}{lddddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
-\hline
-CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77 \\
-CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84 \\
-CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14 \\
-CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08 \\
-%
-\hline
-hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89 \\
-hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59 \\
-hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76 \\
-hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66 \\
-hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08 \\
-hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07 \\
-%
-\hline
-$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58 \\
-$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18 \\
-$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11 \\
-$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02 \\
-%
-\hline
-$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67 \\
-$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15 \\
-$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16 \\
-$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13 \\
-$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10 \\
-$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05 \\
-%
-\hline
-sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68 \\
-sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66 \\
-$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60 \\
-$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62 \\
-%
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-\end{comment}
-
 \begin{table*}[ht!]
@@
 $\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
-$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{00}  & \fk{0}.\fk{06} & \fk{0}.\fk{10} & \fk{0}.\fk{09} \\
+$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{01}  & \fk{0}.\fk{07} & \fk{0}.\fk{10} & \fk{0}.\fk{10} \\
 $\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
@@
 
+\fk{[Performance of the hCI models for the different types of transitions.]}
+
 We now shift to the discussion of excitations for open-shell systems.
@@
 This accounts for the more accurate results observed for open-shell excitations and explains most of the differences with respect to excitations from closed-shell systems, as discussed in detail in the following.
-
-\begin{comment}
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
-Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
-}
-\label{tab:stat2}
-\begin{ruledtabular}
-\begin{tabular}{lddddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
-\hline
-CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50 \\
-CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17 \\
-CISDT             &  6 & +0.06 & 0.07 & 0.09 & 0.06 \\
-CISDTQ            &  6 & +0.08 & 0.08 & 0.11 & 0.08 \\
-%
-\hline
-hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64 \\
-hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68 \\
-hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90 \\
-hCI2.5            &  6 & +0.03 & 0.05 & 0.06 & 0.05 \\
-hCI3              &  6 & +0.00 & 0.04 & 0.05 & 0.05 \\
-hCI3.5            &  6 & +0.01 & 0.03 & 0.05 & 0.04 \\
-%
-\hline
-$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59 \\
-$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21 \\
-$\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08 \\
-$\Delta$CISDTQ    &  6 & -0.02 & 0.02 & 0.02 & 0.02 \\
-%
-\hline
-$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44 \\
-$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69 \\
-$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16 \\
-$\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10 \\
-$\Delta$hCI3      &  6 & -0.06 & 0.06 & 0.08 & 0.05 \\
-$\Delta$hCI3.5    &  6 & -0.02 & 0.02 & 0.03 & 0.02 \\
-%
-\hline
-sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53 \\
-$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37 \\
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-\end{comment}
 
@@
 
+\fk{[Performance of the sCI models for the different types of transitions.]}
+
 The situation for open-shell systems is quite different, in close analogy to the previous discussion regarding hCI.
@@
 \fk{
-The statistical measures for the CI models corrected with the renormalized EN2 correction are presented in 
+The renormalized EN2 correction significantly reduces the errors across most CI models, for both closed- and open-shell systems.
+The statistical measures for the CI models corrected with the EN2 correction can be seen in
 Table~\ref{tab:stat1} (closed-shell systems) and Table~\ref{tab:stat2} (open-shell systems),
@@
 \fk{
-The EN2 correction decreases the four statistical metrics for both closed- and open-shell systems and across most CI models (the exceptions are discussed in the following).
-Importantly, for a given scaling (see Table~\ref{tab:scaling}), whether the bare CI model or the lower-order CI model plus EN2 correction produces the smaller errors depends on the truncation order,
+For a given scaling (see Table~\ref{tab:scaling}), whether the bare CI model or the lower-order CI model plus EN2 correction produces the smaller errors depends on the truncation order,
 and also on the type of excitations (from closed- or open-shells).
 For instance, $\Delta$hCI2 and $\Delta$hCI1+EN2 have the same $\order*{N^6}$ computational cost
-but the former has smaller MAEs, specially for the open-shells (\SI{0.10}{\eV} compared to \SI{0.17}{\eV}), but also for the closed-shells (\SI{0.20}{\eV} compared to \SI{0.24}{\eV}).
+but the former has smaller MAEs, for closed-shells (\SI{0.20}{\eV} compared to \SI{0.24}{\eV}), and more significantly for open-shells (\SI{0.10}{\eV} compared to \SI{0.17}{\eV}).
 In turn, $\Delta$hCI1.5+EN2 is significantly more accurate than $\Delta$hCI2.5 for closed-shells,
-with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, though for open-shells both models produce a MAE of \SI{0.10}{\eV}.
+with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, whreas for open-shells both models produce a MAE of \SI{0.10}{\eV}.
 The same is found at the next order.
@@
 whereas for the open-shell transitions, they deliver comparable MAEs of \SIrange{0.05}{0.06}{\eV}.
-%We recall that the small prefactor associated with the calculation of the EN2 correction generally makes it an attractive option.
+We recall that the calculation of the EN2 correction has a small associated prefactor, which makes the above $\Delta$hCI+EN2 models attractive alternatives to their higher-order unperturbed $\Delta$hCI counterparts.
 }
 
-\fk{EN2 with ground-state HF orbitals}
-
-\fk{hCI+EN2 vs eCI+EN2}
-
-\fk{sCI+EN2 vs sCI}
-
-\begin{comment}
-
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
-Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
+\fk{
+Among the different CI models, it is worth highlighting the huge improvement observed from $\Delta$hCI1.5 to $\Delta$hCI1.5+EN2.
+For the closed-shells, the MAE drops from \SI{2.80}{\eV} to \SI{0.13}{\eV}, whereas a reduction from \SI{0.45}{\eV} to \SI{0.10}{\eV} is seen for the open-shells.
+In both cases, the MSE is virtually zero.
+In addition, the individual errors for singlets, triplets, Rydberg, and valence states are comparable, lying in the range \SIrange{0.11}{0.15}{\eV}.
 }
-\label{tab:stat1}
-\begin{ruledtabular}
-\begin{tabular}{lddddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
-\hline
-CIS+EN2               & 48 &       &      &      &      \\
-\hline
-hCI1+EN2              & 50  & +0.65  & 0.65 & 0.84 & 0.53 \\
-hCI1.5+EN2            & 50  & +0.62  & 0.63 & 0.76 & 0.44 \\
-%
-\hline
-$\Delta$CSF+EN2       & 50  &        &      &      &      \\
-$\Delta$CISD+EN2      & 50  & +0.02  & 0.06 & 0.09 & 0.09 \\
-%
-\hline
-$\Delta$hCI1+EN2      & 50  & +0.12  & 0.24 & 0.30 & 0.27 \\
-$\Delta$hCI1.5+EN2    & 50  & -0.01  & 0.13 & 0.19 & 0.19 \\
-$\Delta$hCI2+EN2      & 50  & +0.00  & 0.06 & 0.10 & 0.09 \\
-%
-\hline
-sCI2/sCI2+EN2         & 50  & +0.51  & 0.60 & 0.79 & 0.51 \\
-sCI2/sCI0+EN2         & 50  & +0.60  & 0.52 & 0.69 & 0.47 \\
-$\Delta$sCI2/sCI2+EN2 & 50  & +0.20  & 0.25 & 0.34 & 0.28 \\
-$\Delta$sCI2/sCI0+EN2 & 50  & +0.11  & 0.22 & 0.26 & 0.24 \\
-%
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-
-\begin{table}[ht!]
-\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
-Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
+
+\fk{
+We found a similar impact of the EN2 correction for $\Delta$CISD and $\Delta$hCI2.
+As discussed above, $\Delta$CISD is somewhat more accurate than $\Delta$hCI2 for the closed-shells (MAEs of \SI{0.17}{\eV} and \SI{0.20}{\eV}),
+whereas the EN2 correction decreases their respective MAE to \SI{0.06}{\eV} and \SI{0.07}{\eV}.
+In turn, while $\Delta$CISD is slightly less accurate than $\Delta$hCI2 for the open-shells (MAEs of \SI{0.12}{\eV} and \SI{0.10}{\eV}),
+the EN2 corrected models show a MAE of \SI{0.04}{\eV} and \SI{0.05}{\eV}, respectively.
+Furthermore, the accuracy of $\Delta$hCI2+EN2 is comparable for singlets and triplets (MAE of \SIrange{0.06}{0.07}{\eV}),
+and somewhat superior for Rydberg states when compared to valence states (MAE of \SI{0.04}{\eV} and \SI{0.08}{\eV}).
+The same trend is found for $\Delta$CISD+EN2, which slightly outperforms $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above.
 }
-\label{tab:stat2}
-\begin{ruledtabular}
-\begin{tabular}{lddddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
-\hline
-CIS+EN2            & 19 &       &      &      &      \\
-%
-\hline
-hCI1+EN2           & 19 & +0.27 & 0.28 & 0.36 & 0.24 \\
-hCI1.5+EN2         & 19 & +0.22 & 0.25 & 0.33 & 0.24 \\
-%
-\hline
-$\Delta$CSF+EN2    & 19 &       &      &      &      \\
-$\Delta$CISD+EN2   & 19 & -0.02 & 0.04 & 0.06 & 0.06 \\
-%
-\hline
-$\Delta$hCI1+EN2   & 19 & +0.02 & 0.17 & 0.29 & 0.29 \\
-$\Delta$hCI1.5+EN2 & 19 & +0.00 & 0.10 & 0.15 & 0.15 \\
-$\Delta$hCI2+EN2   & 19 & -0.00 & 0.05 & 0.08 & 0.08 \\
-%
-\hline
-sCI1+EN2           & 19 & +0.23 & 0.25 & 0.32 & 0.21 \\
-$\Delta$sCI1+EN2   & 19 & +0.06 & 0.12 & 0.20 & 0.19 \\
-\end{tabular}
-\end{ruledtabular}
-\end{table}
-
-\end{comment}
+
+\fk{
+In another contrast to eCI and hCI routes, the EN2 correction has a more limited effect to the sCI route.
+There is little to no advance for the models based on ground-state HF orbitals.
+The outcome is more favourable with state-specific orbitals,
+where, for the closed-shells, $\Delta$sCI2/sCI2+EN2 and $\Delta$sCI2/sCI0+EN2 show comparable MAEs (\SI{0.26}{\eV} and \SI{0.23}{\eV}, respectively).
+For the open-shells, the MAE decrease from \SI{0.28}{\eV} ($\Delta$sCI1) to \SI{0.12}{\eV} ($\Delta$sCI1+EN2),
+the latter error being comparable to that of the much less expensive $\Delta$hCI1.5+EN2 model (\SI{0.10}{\eV}).
+Despite the improvement, the EN2 correction is not enough to render the sCI models competitive.
+}
 
@@
 
+\fk{The accuracy of the CI models is significantly enhanced thanks to the EN2 correction,
+the case of $\Delta$hCI1.5 for closed-shells being the most striking, with a MAE plummeting from \SI{2.80}{\eV} to \SI{0.13}{\eV}.
+Low-orders of state-specific CI supplemented with a perturbative correction can thus provide a cost-effective option for the computation of excitation energies.
+}
+
 We put forward the interesting perspective of developing hierarchy-based CC (hCC) methods (hCC1, hCC2, \ldots),
@@
 For the open-shell transitions, the outcome is more encouraging.
+\fk{Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than eCI or hCI options.}
 The relevance of these sCI models lies in their possible connection with CC methods.
@@
 number of determinants in the reference; saddle point order associated with the $\Delta$CSF solutions;
-and the reference excitation energies and corresponding methods.
+reference excitation energies and corresponding methods;
+\fk{and complementary statistical metrics}.
 For a subset of 27 excited states for closed-shell molecules and 15 excited states for radicals,

### 2023-10-22 Manuscript/sup.tex
saving work

@@
 \begin{table}[ht!]
-\caption{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
-with the cc-pVDZ basis set and frozen-core approximation.}
+\caption{\fk{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
+with the cc-pVDZ basis set and frozen-core approximation.}}
 \label{tab:ndet_open}
+{\color{orange}
 \begin{ruledtabular}
@@
 \end{ruledtabular}
+}
 \end{table}
@@
 \begin{table}[ht!]
-\caption{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{H4}, \ce{N2}, \ce{F2}, \ce{HF}, \ce{H8}, and ethylene,
-with the cc-pVDZ basis set and frozen-core approximation.}
+\caption{\fk{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{H4}, \ce{N2}, \ce{F2}, \ce{HF}, \ce{H8}, and ethylene,
+with the cc-pVDZ basis set and frozen-core approximation.}}
 \label{tab:ndet_closed}
+{\color{orange}
 \begin{ruledtabular}

### 2023-10-23 Manuscript/Response_Letter/Response_Letter.tex
saving work

@@
 
+\clearpage
+
 {4) The authors need to mention (or draw) the molecules of their excited state study.
@@
 ``Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
-and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.''
+and 2 doubly-excited states (glyoxal and nitroxyl). It does not include large molecules nor charge transfer states.''
 The details on each specific excited state can be found in The Supporting Information.
-New additions to text can be found on page x:
-``aa''
-The new subsection (IV.E) also discusses how hCI+EN2 models perform for different types of excited states.
+The revised manuscript now mentions the performance of the hCI models for different types of excited state, on page x:
+``$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}),
+whereas Rydberg states are better described than valence states (MAEs of \SI{0.14}{\eV} and \SI{0.24}{\eV}).
+The same trends are found for the low-order models ($\Delta$hCI1 and $\Delta$hCI1.5), and also for $\Delta$CISD, which slightly outperforms $\Delta$hCI2 for each type of excitation.
+The specific MAEs can be found in the Supporting Information.''
+This aspect is also discussed for the hCI+EN2 models in the new subsection (IV.E).
 }
-
-\clearpage
 

### 2023-10-23 Manuscript/manuscript.tex
saving work

@@
 \fk{Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
-and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.}
+and 2 doubly-excited states (glyoxal and nitroxyl). It does not include large molecules nor charge transfer states.}
 The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.
@@
 \begin{table}[ht!]
-\caption{\fk{Systems Considered in the Present Excited State Calculations.}}
+\caption{\fk{Systems Considered in the Excited State Calculations.}}
 \label{tab:systems}
@@
 \hline
-CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  &\fk{CIS+EN2}               & \fk{48} &       &      &      &      \\
+CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  & & & & & & \\
 CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84  & & & & & & \\
@@
 \hline
-$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  &\fk{$\Delta$CSF+EN2}       & \fk{50}  &        &      &      &      \\
+$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  & & & & & & \\
 $\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
@@
 The slightly worse performance of $\Delta$hCI2 probably stems from a residual unbalance associated with the state-specific classes of determinants.
+\fk{$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}), 
+whereas Rydberg states are better described than valence states (MAEs of \SI{0.14}{\eV} and \SI{0.24}{\eV}).
+The same trends are found for the low-order models ($\Delta$hCI1 and $\Delta$hCI1.5), and also for $\Delta$CISD, which slightly outperforms $\Delta$hCI2 for each type of excitation.
+The specific MAEs can be found in the {\SupInf}.}
 
@@
 as also discussed above for $\Delta$CISD and $\Delta$hCI2.
-
-\fk{[Performance of the hCI models for the different types of transitions.]}
 
@@
 \hline
-CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50  & \fk{CIS+EN2}            & \fk{19} &       &      &      &      \\
+CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50  & & & & & & \\
 CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17  & & & & & & \\
@@
 \hline
-$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59  & \fk{$\Delta$CSF+EN2}    & \fk{19} &       &      &      &      \\
+$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59  & & & & & & \\
 $\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21  & \fk{$\Delta$CISD+EN2}   & \fk{19} & \fk{-0}.\fk{02} & \fk{0}.\fk{04} & \fk{0}.\fk{06} & \fk{0}.\fk{06} \\
@@
 Ultimately, we did not find a combination of sCI models and orbitals that produced reasonable excitation energies for closed-shell systems.
-
-\fk{[Performance of the sCI models for the different types of transitions.]}
 
@@
 the EN2 corrected models show a MAE of \SI{0.04}{\eV} and \SI{0.05}{\eV}, respectively.
-Furthermore, the accuracy of $\Delta$hCI2+EN2 is comparable for singlets and triplets (MAE of \SIrange{0.06}{0.07}{\eV}),
-and somewhat superior for Rydberg states when compared to valence states (MAE of \SI{0.04}{\eV} and \SI{0.08}{\eV}).
-The same trend is found for $\Delta$CISD+EN2, which slightly outperforms $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above.
+Furthermore, the accuracy of $\Delta$hCI2+EN2 is comparable for singlets and triplets (MAEs of \SI{0.07}{\eV} and \SI{0.06}{\eV}),
+and superior for Rydberg states when compared to valence states (MAE of \SI{0.04}{\eV} and \SI{0.08}{\eV}).
+$\Delta$CISD+EN2 displays the same trends, slightly outperforming $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above, just as discussed above for the unperturbed case.
 }
@@
 For the open-shell transitions, the outcome is more encouraging.
-\fk{Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than eCI or hCI options.}
+\fk{Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than the eCI or hCI options.}
 The relevance of these sCI models lies in their possible connection with CC methods.

### 2023-10-24 Manuscript/Response_Letter/Response_Letter.tex
cyclobutadiene

@@
 \alert{
-aaa
+This is a pertinent remark.
+From the examples suggested by the reviewer, we decided to investigate the automerization problem of cyclobutadiene.
+Our results can be found in a new paragraph on page x, starting as:
+``We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene...''
+On the other hand, the case of benzynes is left as a perspective for the future, otherwise the manuscript would be excessively lengthy in our opinion.
+We notice that extensive CI+EN2 additional calculations have been carried out and a corresponding new subsection (IV.E) has been added to the manuscript, based on a suggestion from Reviewer \#1.
+We further highlight the point raised by Reviewer \#2 in the Conclusion, reading
+``hCI also produced reasonable energies for the automerization barrier of cyclobutadiene.
+It remains to be seen how these models perform for other strongly correlated systems, larger molecules, and charge transfer excited states.''
 }

### 2023-10-24 Manuscript/manuscript.tex
cyclobutadiene

@@
 
+\fk{
+We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene,
+a well-known and challenging problem that involves strong correlation. \cite{Monino_2022}
+High-level calculations are needed to achieve quantitative values for the height of the barrier that 
+connects the two equivalent rectangular $D_{2h}$ equilibrium geometries through the square $D_{4h}$ transition state geometry. \cite{Monino_2022}
+Here we employ the geometries presented in Ref.~\onlinecite{Monino_2022}, the 6-31+G(d) basis set and the frozen-core approximation.
+We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry, from which the excited determinants were generated in the CI calculations.
+For this basis set, the reference value of \SI{7.51}{kcal/mol} was obtained at CC with single, double, triple, and quadruple excitations. \cite{Monino_2022}
+With respect to this reference value, the errors for the automerization barrier are
+\SI{+3.55}{kcal/mol} with our two-determinant calculations, \SI{-2.28}{kcal/mol} with hCI2, \SI{-1.07}{kcal/mol} with hCI2+EN2, \SI{+0.61}{kcal/mol} with CISD, and \SI{-0.49}{kcal/mol} with CISD+EN2.
+These values are comparable to those obtained with other methods, extensively discussed in Ref.~\onlinecite{Monino_2022}.
+Here we just mention the errors of CC with singles and doubles, \SI{+0.80}{kcal/mol}, and of spin-flip equation-of-motion CC (EOM-CC) with singles and doubles, \SI{-1.65}{kcal/mol},
+which share the same $\order*{N^6}$ computational scaling as CISD and hCI2.
+This specific example serves to illustrate that the accuracy of hCI models can be similar to that of other methods.
+As such, hCI can be taken into consideration in future systematic studies that tackle challenging chemical problems where strong correlation comes into play.
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In this regard, rotations within occupied and virtual orbitals represent yet another factor worth exploring in the future developments of hCI models.
+\fk{hCI also produced reasonable energies for the automerization barrier of cyclobutadiene.
+It remains to be seen how these models perform for other strongly correlated systems, larger molecules, and charge transfer excited states.}
 

### 2023-10-24 Manuscript/manuscript.tex
CBD references

@@
 We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene,
-a well-known and challenging problem that involves strong correlation. \cite{Monino_2022}
-High-level calculations are needed to achieve quantitative values for the height of the barrier that 
-connects the two equivalent rectangular $D_{2h}$ equilibrium geometries through the square $D_{4h}$ transition state geometry. \cite{Monino_2022}
+which connects the two equivalent rectangular $D_{2h}$ equilibrium geometries through the square $D_{4h}$ transition state geometry. \cite{Bally_1980,Tallarico_1996}
+This is a well-known and challenging problem, requiring high-level calculations to achieve quantitative values for the height of the barrier (see Ref.~\onlinecite{Monino_2022} and references therein).
 Here we employ the geometries presented in Ref.~\onlinecite{Monino_2022}, the 6-31+G(d) basis set and the frozen-core approximation.
-We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry, from which the excited determinants were generated in the CI calculations.
+We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry,
+from which the excited determinants were generated in the CI calculations.
 For this basis set, the reference value of \SI{7.51}{kcal/mol} was obtained at CC with single, double, triple, and quadruple excitations. \cite{Monino_2022}

### 2023-10-25 Manuscript/Response_Letter/Response_Letter.tex
CBD references
corrections T2

@@
 \alert{
-We thank the reviwer for the supportive comments.
+We thank the reviewer for the supportive comments.
 }
@@
 \alert{
-The number of determinants in FCI and in each approximate CI model is now presented in Tables S1 and S2 of the Supporting Information.
+The number of determinants in FCI and in each approximate CI model is now presented in Tables S1, S2, and S3 of the Supporting Information.
 }
@@
 We have experimented including the values of $n$ to the figures, as suggested by the reviewer.
-However, we consider that they become somewhat clutered and less readable, and therefore we prefer to leave them unchanged.
-With the intention of helping the reader, we have been supplemented the following sentence to the captions of Figures 3 (4):
+However, we consider that they become somewhat cluttered and less readable, and therefore we prefer to leave them unchanged.
+With the intention of helping the reader, we have supplemented the following sentence to the captions of Figures 3 (4):
 ``Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (sCI1, sCI3, sCI5) (blue).''
@@
 \alert{
-We thank the reviwer for the positive appreciation of our work.
+We thank the reviewer for the positive appreciation of our work.
 }
@@
 \alert{
-The set of systems we considered is now presented in Table II, whreas on page x we describe the character of the excited states, reading
+The set of systems we considered is now presented in Table II, whereas on page x we describe the character of the excited states, reading
 ``Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
@@
 The details on each specific excited state can be found in The Supporting Information.
-The revised manuscript now mentions the performance of the hCI models for different types of excited state, on page x:
+The revised manuscript now mentions the performance of the hCI models for different types of excited states, on page x:
 ``$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}),
@@
 \alert{
-The statistical measures are now defined on page x of the Supporting Information, which are refered to on page x of the main text, reading
-``For completeness, the defition of these statistical measures can be found in the Supporting Information.''
+The statistical measures are now defined on page x of the Supporting Information, which are referred to on page x of the main text, reading
+``For completeness, the definition of these statistical measures can be found in the Supporting Information.''
 }
@@
 ``We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene...''
-On the other hand, the case of benzynes is left as a perspective for the future, otherwise the manuscript would be excessively lengthy in our opinion.
+On the other hand, the case of benzynes is left as a perspective for the future, otherwise, the manuscript would be excessively lengthy in our opinion.
+Moreover, our code would need further optimizations in order to tackle efficiently these larger systems.
 We notice that extensive CI+EN2 additional calculations have been carried out and a corresponding new subsection (IV.E) has been added to the manuscript, based on a suggestion from Reviewer \#1.

### 2023-10-25 Manuscript/manuscript.tex
corrections T2

@@
 \usepackage[normalem]{ulem}
-\newcommand{\fk}[1]{\textcolor{orange}{#1}}
+\newcommand{\fk}[1]{\textcolor{red}{#1}}
 \newcommand{\titou}[1]{\textcolor{red}{#1}}
@@
 \fk{unless supplemented by the EN2 correction, which substantially improves their accuracy.}
-In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
+In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled-cluster methods.
 \bigskip
@@
 quantum Monte Carlo, \cite{Scemama_2018a,Scemama_2018b,Dash_2018,Dash_2019,Dash_2021,Cuzzocrea_2022,Shepard_2022,Otis_2020,Otis_2023}
-and coupled cluster (CC)
+and coupled-cluster (CC)
 \cite{Piecuch_2000,Mayhall_2010,Lee_2019,Kossoski_2021,Marie_2021,Rishi_2023,Tuckman_2023}
@@
 Our fifth goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state-based and state-specific sCI models for excited states.
-\fk{Finally, in Sec.~\ref{sec:res_E} we assess how the excitation energies are impacted by the EN2 perturbative correction to hCI and sCI models, our sixth and last goal.}
+\fk{Finally, in Sec.~\ref{sec:res_E}, we assess how the excitation energies are impacted by the EN2 perturbative correction to hCI and sCI models, our sixth and last goal.}
 Section \ref{sec:conclusion} closes the present contribution with the main conclusions and perspectives.
@@
 \label{tab:systems}
-{\color{orange}
 \begin{ruledtabular}
-\begin{tabular}{lll|lll}
-\mc{3}{c|}{Closed-shells}           & \mc{3}{c}{Open-shells} \\
-\hline
+\begin{tabular}{llllll}
+\mc{3}{c}{Closed-shells}           & \mc{3}{c}{Open-shells} \\
+\cline{1-3} \cline{4-6}
 glyoxal      & \ce{HCF}  & \ce{H2O} & \ce{BeH} & \ce{NH2} & \ce{HCO} \\
@@
 \end{ruledtabular}
-}
 \end{table}
@@
 The computed excitation energies were benchmarked against the reference theoretical values provided in the QUEST database. \cite{Veril_2021}
-\fk{For each excited state considered here, the reference value and the method used to obtain it (of high-order coupled cluster or extrapolated FCI quality) can be found in the {\SupInf}.}
+\fk{For each excited state considered here, the reference value and the method used to obtain it (of high-order CC or extrapolated FCI quality) can be found in the {\SupInf}.}
 
@@
 This is a well-known and challenging problem, requiring high-level calculations to achieve quantitative values for the height of the barrier (see Ref.~\onlinecite{Monino_2022} and references therein).
-Here we employ the geometries presented in Ref.~\onlinecite{Monino_2022}, the 6-31+G(d) basis set and the frozen-core approximation.
+Here we employ the geometries presented in Ref.~\onlinecite{Monino_2022}, the 6-31+G(d) basis set, and the frozen-core approximation.
 We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry,
@@
 the development of polynomial scaling CC methods based on the concept of seniority.
-For closed-shell systems, DOCI (here referred to as sCI0) energies can be very well reproduced with state-specific pair coupled cluster doubles (pCCD), a method that has mean-field cost,
+For closed-shell systems, DOCI (here referred to as sCI0) energies can be very well reproduced with state-specific pair coupled-cluster doubles (pCCD), a method that has mean-field cost,
 for both ground \cite{Bytautas_2011,Allen_1962,Smith_1965,Veillard_1967} and excited \cite{Kossoski_2021,Marie_2021} states.
@@
 In turn, $\Delta$hCI1.5+EN2 is significantly more accurate than $\Delta$hCI2.5 for closed-shells,
-with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, whreas for open-shells both models produce a MAE of \SI{0.10}{\eV}.
-The same is found at the next order.
+with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, whereas for open-shells both models produce a MAE of \SI{0.10}{\eV}.
+The same trend is found at the next order.
 $\Delta$hCI2+EN2 is much more accurate than $\Delta$hCI3 for closed-shells, with MAEs of \SI{0.06}{\eV} and \SI{0.22}{\eV},
@@
 \fk{
-In another contrast to eCI and hCI routes, the EN2 correction has a more limited effect to the sCI route.
-There is little to no advance for the models based on ground-state HF orbitals.
-The outcome is more favourable with state-specific orbitals,
+In another contrast to eCI and hCI routes, the EN2 correction has a more limited effect on the sCI route.
+There is little to no advantage for the models based on ground-state HF orbitals.
+The outcome is more favorable with state-specific orbitals,
 where, for the closed-shells, $\Delta$sCI2/sCI2+EN2 and $\Delta$sCI2/sCI0+EN2 show comparable MAEs (\SI{0.26}{\eV} and \SI{0.23}{\eV}, respectively).
-For the open-shells, the MAE decrease from \SI{0.28}{\eV} ($\Delta$sCI1) to \SI{0.12}{\eV} ($\Delta$sCI1+EN2),
+For the open-shells, the MAE decreases from \SI{0.28}{\eV} ($\Delta$sCI1) to \SI{0.12}{\eV} ($\Delta$sCI1+EN2),
 the latter error being comparable to that of the much less expensive $\Delta$hCI1.5+EN2 model (\SI{0.10}{\eV}).

### 2023-10-25 Manuscript/sup.tex
corrections T2

@@
 
-\newcommand{\fk}[1]{\textcolor{orange}{#1}}
+\newcommand{\fk}[1]{\textcolor{red}{#1}}
 
@@
 the corresponding mean signed error (MSE), mean absolute error (MAE), root-mean-square error (RMSE), and standard deviation of the errors (SDE) are defined as:
-}
-
-\begin{equation}
-	\fk{\text{MSE} = \frac{1}{N} \sum_{i=1}^{N} (x_i - y_i),}
-\end{equation}
-
-\begin{equation}
-	\fk{\text{MAE} = \frac{1}{N} \sum_{i=1}^{N} |x_i - y_i|,}
-\end{equation}
-
-\begin{equation}
-	\fk{\text{RMSE} = \sqrt{ \frac{1}{N} \sum_{i=1}^{N} (x_i - y_i)^2 },}
-\end{equation}
-
-\begin{equation}
-	\fk{\text{SDE} = \sqrt{ \frac{1}{N} \sum_{i=1}^{N} (x_i - \mu)^2 },}
-\end{equation}
-
-\fk{where}
+\begin{subequations}
+\begin{align}
+	\text{MSE} & = \frac{1}{N} \sum_{i=1}^{N} (x_i - y_i),
+	\\
+	\text{MAE} & = \frac{1}{N} \sum_{i=1}^{N} \abs{x_i - y_i},
+	\\
+	\text{RMSE} & = \sqrt{ \frac{1}{N} \sum_{i=1}^{N} (x_i - y_i)^2 },
+	\\
+	\text{SDE} & = \sqrt{ \frac{1}{N} \sum_{i=1}^{N} (x_i - \mu)^2 },
+\end{align}
+\end{subequations}
+where
 \begin{equation}
 	\fk{\mu = \frac{1}{N} \sum_{i=1}^{N} x_i.}
-\end{equation}
+\end{equation}}
 
@@
 Since the fitting is restricted to the equilibrium region, employing more elaborate functions would be inconsequential, such that the adopted Morse potential is well justified.
-The fitting was performed with the Marquardt-Levenberg algorithm implemented in Gnuplot.
+The fitting was performed with the Marquardt-Levenberg algorithm implemented in \textsc{gnuplot}.
 }
@@
 \label{tab:ndet_open}
-{\color{orange}
 \begin{ruledtabular}
@@
 \begin{table}[ht!]
-\caption{\fk{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{H4}, \ce{N2}, \ce{F2}, \ce{HF}, \ce{H8}, and ethylene,
+\caption{\fk{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{H4}, \ce{N2}, and \ce{F2},
 with the cc-pVDZ basis set and frozen-core approximation.}}
 \label{tab:ndet_closed}
-{\color{orange}
 \begin{ruledtabular}
@@
 FCI    &  36 100 & 4 327 008 400 & 432 700 840 000 \\
-\hline
-\hline
-& & & \\
-\hline
-\hline
+\end{tabular}
+\end{ruledtabular}
+\end{table}
+
+\begin{table}[ht!]
+\caption{\fk{Number of Determinants in Hierarchy-, Excitation, and Seniority-Based CI Routes, as well as in Full CI, for \ce{HF}, \ce{H8}, and ethylene,
+with the cc-pVDZ basis set and frozen-core approximation.}}
+\label{tab:ndet_closed}
+\begin{ruledtabular}
+\begin{tabular}{lrrr}
 method &    \ce{HF} &       \ce{H8} &           ethylene \\

### 2023-10-25 Manuscript/manuscript.tex
Delta + update table

@@
 We optimize the orbitals for two closed-shell determinants in the $D_{2h}$ geometry, and for two spin-flipped open-shell determinants in the $D_{4h}$ geometry,
-from which the excited determinants were generated in the CI calculations.
+from which the excited determinants were generated in the subsequent CI calculations.
+Even though the calculations target the ground state, they are labelled $\Delta$CI because different references are employed for each geometry.
 For this basis set, the reference value of \SI{7.51}{kcal/mol} was obtained at CC with single, double, triple, and quadruple excitations. \cite{Monino_2022}
 With respect to this reference value, the errors for the automerization barrier are
-\SI{+3.55}{kcal/mol} with our two-determinant calculations, \SI{-2.28}{kcal/mol} with hCI2, \SI{-1.07}{kcal/mol} with hCI2+EN2, \SI{+0.61}{kcal/mol} with CISD, and \SI{-0.49}{kcal/mol} with CISD+EN2.
+\SI{+3.55}{kcal/mol} with our two-determinant calculations, \SI{-2.28}{kcal/mol} with $\Delta$hCI2, \SI{-1.07}{kcal/mol} with $\Delta$hCI2+EN2, 
+\SI{+0.61}{kcal/mol} with $\Delta$CISD, and \SI{-0.49}{kcal/mol} with $\Delta$CISD+EN2.
 These values are comparable to those obtained with other methods, extensively discussed in Ref.~\onlinecite{Monino_2022}.
 Here we just mention the errors of CC with singles and doubles, \SI{+0.80}{kcal/mol}, and of spin-flip equation-of-motion CC (EOM-CC) with singles and doubles, \SI{-1.65}{kcal/mol},
-which share the same $\order*{N^6}$ computational scaling as CISD and hCI2.
+which share the same $\order*{N^6}$ computational scaling as $\Delta$CISD and $\Delta$hCI2.
 This specific example serves to illustrate that the accuracy of hCI models can be similar to that of other methods.

### 2023-10-25 Manuscript/manuscript.tex
updated Figs 3 and 4

@@
 as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red), and sCI (blue) routes,
-with (light tones) and without (dark tones) the standard EN2 perturbative correction.
+with (light tone crosses) and without (dark tone circles) the standard EN2 perturbative correction.
 \fk{Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI1, sCI3, and sCI5 (blue).}
@@
 Non-parallelity error (NPE), distance error, equilibrium distance, and vibrational frequency (or force constant), for \ce{HF}, \ce{F2}, ethylene, \ce{N2}, \ce{H4}, and \ce{H8}, as functions of the number of determinants ($\Ndet$), according to the hCI (green), eCI (red) and sCI (blue) routes,
-with (light tones) and without (dark tones) the standard EN2 perturbative correction.
+with (light tone crosses) and without (dark tone circles) the standard EN2 perturbative correction.
 \fk{Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (blue).}

### 2023-10-25 Manuscript/Response_Letter/Response_Letter.tex
ready to submit

@@
 We have performed additional calculations for the full set of states with the different CI models surveyed here (except for the more computationally demanding higher-order models).
-These new results were included in Tables II, III, and in the Supporting Information, while a new subsection (IV.E), starting on page x, is devoted to their discussion.
-Additions to the Conclusion (page x) and the abstract have also been made.
+These new results were included in Tables II, III, and in the Supporting Information, while a new subsection (IV.E), starting on page 13, is devoted to their discussion.
+Additions to the Conclusion, on page 14, and to the abstract have also been made.
 }
@@
 \alert{
-This sentence has been rephrased and now reads, on page x,
+This sentence has been rephrased and now reads, on page 2,
 ``The number of determinants $\Ndet$ within a given class scales polynomially with the number of basis functions $N$, with the exponent depending on the specific class.
@@
 \alert{
-The reviewer's questions are addressed on page x of the revised Supporting Information, reading
+The reviewer's questions are addressed on page 3 of the revised Supporting Information, reading
 ``By fitting the computed potential energy curves with a Morse potential $V(r) = D [1 - e^{-a(r-r_e)}]^2 + C $,
@@
 \ce{CN} is the only system for which some of the potential energy curves are not smooth around the equilibrium geometry.
-This is mentioned on page x of the revised manuscript, reading
+This is mentioned on page 6 of the revised manuscript, reading
 ``For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
@@
 \alert{
-The set of systems we considered is now presented in Table II, whereas on page x we describe the character of the excited states, reading
+The set of systems we considered is now presented in Table II, whereas on page 5 we describe the character of the excited states, reading
 ``Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
@@
 The details on each specific excited state can be found in The Supporting Information.
-The revised manuscript now mentions the performance of the hCI models for different types of excited states, on page x:
+The revised manuscript now mentions the performance of the hCI models for different types of excited states, on page 11:
 ``$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}),
@@
 \alert{
-The statistical measures are now defined on page x of the Supporting Information, which are referred to on page x of the main text, reading
+The statistical measures are now defined on page 2 of the Supporting Information, which are referred to on page 9 of the main text, reading
 ``For completeness, the definition of these statistical measures can be found in the Supporting Information.''
@@
 \alert{
-This information is provided in the Supporting Information, which is stated on page x of the revised manuscript, reading
+This information is provided in the Supporting Information, which is stated on page 5 of the revised manuscript, reading
 ``For each excited state considered here, the reference value and the corresponding method to obtain it (of high-order coupled cluster or extrapolated FCI quality) can be found in the Supporting Information.''
@@
 From the examples suggested by the reviewer, we decided to investigate the automerization problem of cyclobutadiene.
-Our results can be found in a new paragraph on page x, starting as:
+Our results can be found in a new paragraph on page 9, starting as:
 ``We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene...''

### 2023-10-25 Manuscript/manuscript.tex
ready to submit

@@
 We find that hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
-Employing the second-order Epstein-Nesbet \fk{(EN2)} perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
+Employing the second-order Epstein-Nesbet (EN2) perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
 We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
@@
 State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results,
-\fk{unless supplemented by the EN2 correction, which substantially improves their accuracy.}
+unless supplemented by the EN2 correction, which substantially improves their accuracy.
 In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled-cluster methods.
@@
 Our fifth goal, detailed in Sec.~\ref{sec:res_D}, is therefore to define and gauge ground-state-based and state-specific sCI models for excited states.
-\fk{Finally, in Sec.~\ref{sec:res_E}, we assess how the excitation energies are impacted by the EN2 perturbative correction to hCI and sCI models, our sixth and last goal.}
+Finally, in Sec.~\ref{sec:res_E}, we assess how the excitation energies are impacted by the EN2 perturbative correction to hCI and sCI models, our sixth and last goal.
 Section \ref{sec:conclusion} closes the present contribution with the main conclusions and perspectives.
@@
 we refer to the class of an excited determinant as the combination of the number of electrons and seniority in each orbital subspace.
-%hCI was defined such that all classes of determinants whose number $\Ndet$ share the same scaling with the number of basis functions $N$ are included at the same order of truncation $h$,
-\fk{The number of determinants $\Ndet$ within a given class scales polynomially with the number of basis functions $N$, with the exponent depending on the specific class.
+The number of determinants $\Ndet$ within a given class scales polynomially with the number of basis functions $N$, with the exponent depending on the specific class.
 For instance, for the class of doubly-excited determinants with no unpaired electrons ($e=2$ and $s=0$), $\Ndet = \order*{N^2}$.
-hCI was defined such that, at a given hierarchy $h$, all classes of determinants presenting a scaling of $\Ndet = \order*{N^{2h}}$ or less are accounted for.}
+hCI was defined such that, at a given hierarchy $h$, all classes of determinants presenting a scaling of $\Ndet = \order*{N^{2h}}$ or less are accounted for.
 This means moving diagonally in the seniority-excitation map (denoted by the color tones in Fig.~\ref{fig:hCI}).
@@
 The various CI models introduced here were further assessed based on calculated vertical excitation energies for 69 electronic states,
-from 17 closed-shell systems and 13 radicals \fk{(shown in Table~\ref{tab:systems})}, with geometries extracted from the QUEST database. \cite{Veril_2021}
-\fk{Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
-and 2 doubly-excited states (glyoxal and nitroxyl). It does not include large molecules nor charge transfer states.}
+from 17 closed-shell systems and 13 radicals (shown in Table~\ref{tab:systems}), with geometries extracted from the QUEST database. \cite{Veril_2021}
+Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
+and 2 doubly-excited states (glyoxal and nitroxyl). It does not include large molecules nor charge transfer states.
 The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.
@@
 \begin{table}[ht!]
-\caption{\fk{Systems Considered in the Excited State Calculations.}}
+\caption{Systems Considered in the Excited State Calculations.}
 \label{tab:systems}
@@
 The computed excitation energies were benchmarked against the reference theoretical values provided in the QUEST database. \cite{Veril_2021}
-\fk{For each excited state considered here, the reference value and the method used to obtain it (of high-order CC or extrapolated FCI quality) can be found in the {\SupInf}.}
+For each excited state considered here, the reference value and the method used to obtain it (of high-order CC or extrapolated FCI quality) can be found in the {\SupInf}.
 
@@
 with (light tone crosses) and without (dark tone circles) the standard EN2 perturbative correction.
-\fk{Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI1, sCI3, and sCI5 (blue).}
+Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI1, sCI3, and sCI5 (blue).
 The dashed lines represent the FCI results.}
@@
 with (light tone crosses) and without (dark tone circles) the standard EN2 perturbative correction.
-\fk{Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (blue).}
+Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (blue).
 The dashed lines represent the FCI results.}
@@
 Finally, hCI systematically outperforms eCI in the calculation of vibrational frequencies, except for \ce{H7}, where no big difference is found.
-\fk{For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
-and somewhat less reliable equilibrium properties in these specific cases.}
+For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
+and somewhat less reliable equilibrium properties in these specific cases.
 Given the overall slight superiority of hCI for the equilibrium properties, it also manages to account for more dynamic correlation than eCI,
@@
 
-\fk{
 We further explored hCI to describe the automerization barrier in the ground state of cyclobutadiene,
@@
 As such, hCI can be taken into consideration in future systematic studies that tackle challenging chemical problems where strong correlation comes into play.
-}
 
@@
 with respect to the reference theoretical values for the excitation energies.
-\fk{For completeness, the defition of these statistical measures can be found in the {\SupInf}.}
+For completeness, the defition of these statistical measures can be found in the {\SupInf}.
 For the lower-order models, the calculations were performed for 50 (closed-shell) and 19 (open-shell) excited states.
@@
 and on a similar comparison between ground-state and state-specific approaches for hCI and sCI.
-\fk{The bare (unperturbed) hCI and sCI models are discussed in this section and in Sec.~\ref{sec:res_D}, respectively,
-whereas their EN2 perturbed analogues are left for Sec.~\ref{sec:res_E}.}
+The bare (unperturbed) hCI and sCI models are discussed in this section and in Sec.~\ref{sec:res_D}, respectively,
+whereas their EN2 perturbed analogues are left for Sec.~\ref{sec:res_E}.
 
@@
 \begin{tabular}{ldddddlddddd}
-method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
+method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & method & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
 \hline
@@
 \hline
-hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  &\fk{hCI1+EN2}              & \fk{50}  & \fk{+0}.\fk{65}  & \fk{0}.\fk{65} & \fk{0}.\fk{84} & \fk{0}.\fk{53} \\
-hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  &\fk{hCI1.5+EN2}            & \fk{50}  & \fk{+0}.\fk{62}  & \fk{0}.\fk{63} & \fk{0}.\fk{76} & \fk{0}.\fk{44} \\
+hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  & hCI1+EN2              & 50  & +0.65  & 0.65 & 0.84 & 0.53 \\
+hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  & hCI1.5+EN2            & 50  & +0.62  & 0.63 & 0.76 & 0.44 \\
 hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76  & & & & & & \\
@@
 $\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  & & & & & & \\
-$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
+$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  & $\Delta$CISD+EN2      & 50  & +0.02  & 0.06 & 0.09 & 0.09 \\
 $\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11  & & & & & & \\
@@
 \hline
-$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  &\fk{$\Delta$hCI1+EN2}      & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{24} & \fk{0}.\fk{30} & \fk{0}.\fk{27} \\
-$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
-$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{01}  & \fk{0}.\fk{07} & \fk{0}.\fk{10} & \fk{0}.\fk{10} \\
+$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  & $\Delta$hCI1+EN2      & 50  & +0.12  & 0.24 & 0.30 & 0.27 \\
+$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  & $\Delta$hCI1.5+EN2    & 50  & -0.01  & 0.13 & 0.19 & 0.19 \\
+$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  & $\Delta$hCI2+EN2      & 50  & +0.01  & 0.07 & 0.10 & 0.10 \\
 $\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
@@
 \hline
-sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  &\fk{sCI2/sCI2+EN2}         & \fk{50}  & \fk{+0}.\fk{60}  & \fk{0}.\fk{60} & \fk{0}.\fk{79} & \fk{0}.\fk{51} \\
-sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  &\fk{sCI2/sCI0+EN2}         & \fk{50}  & \fk{+0}.\fk{51}  & \fk{0}.\fk{52} & \fk{0}.\fk{69} & \fk{0}.\fk{47} \\
-$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  &\fk{$\Delta$sCI2/sCI2+EN2} & \fk{50}  & \fk{+0}.\fk{21}  & \fk{0}.\fk{26} & \fk{0}.\fk{35} & \fk{0}.\fk{28} \\
-$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  &\fk{$\Delta$sCI2/sCI0+EN2} & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{23} & \fk{0}.\fk{27} & \fk{0}.\fk{24} \\
+sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  & sCI2/sCI2+EN2         & 50  & +0.60  & 0.60 & 0.79 & 0.51 \\
+sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  & sCI2/sCI0+EN2         & 50  & +0.51  & 0.52 & 0.69 & 0.47 \\
+$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  & $\Delta$sCI2/sCI2+EN2 & 50  & +0.21  & 0.26 & 0.35 & 0.28 \\
+$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  & $\Delta$sCI2/sCI0+EN2 & 50  & +0.12  & 0.23 & 0.27 & 0.24 \\
 %
@@
 The slightly worse performance of $\Delta$hCI2 probably stems from a residual unbalance associated with the state-specific classes of determinants.
-\fk{$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}),
+$\Delta$hCI2 is more accurate for singlets than for triplets (MAEs of \SI{0.16}{\eV} and \SI{0.25}{\eV}),
 whereas Rydberg states are better described than valence states (MAEs of \SI{0.14}{\eV} and \SI{0.24}{\eV}).
 The same trends are found for the low-order models ($\Delta$hCI1 and $\Delta$hCI1.5), and also for $\Delta$CISD, which slightly outperforms $\Delta$hCI2 for each type of excitation.
-The specific MAEs can be found in the {\SupInf}.}
+The specific MAEs can be found in the {\SupInf}.
 
@@
 \hline
-hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64  & \fk{hCI1+EN2}           & \fk{19} & \fk{+0}.\fk{27} & \fk{0}.\fk{28} & \fk{0}.\fk{36} & \fk{0}.\fk{24} \\
-hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68  & \fk{hCI1.5+EN2}         & \fk{19} & \fk{+0}.\fk{22} & \fk{0}.\fk{25} & \fk{0}.\fk{33} & \fk{0}.\fk{24} \\
+hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64  & hCI1+EN2           & 19 & +0.27 & 0.28 & 0.36 & 0.24 \\
+hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68  & hCI1.5+EN2         & 19 & +0.22 & 0.25 & 0.33 & 0.24 \\
 hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90  & & & & & & \\
@@
 $\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59  & & & & & & \\
-$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21  & \fk{$\Delta$CISD+EN2}   & \fk{19} & \fk{-0}.\fk{02} & \fk{0}.\fk{04} & \fk{0}.\fk{06} & \fk{0}.\fk{06} \\
+$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21  & $\Delta$CISD+EN2   & 19 & -0.02 & 0.04 & 0.06 & 0.06 \\
 $\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08  & & & & & & \\
@@
 \hline
-$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44  & \fk{$\Delta$hCI1+EN2}   & \fk{19} & \fk{+0}.\fk{02} & \fk{0}.\fk{17} & \fk{0}.\fk{29} & \fk{0}.\fk{29} \\
-$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69  & \fk{$\Delta$hCI1.5+EN2} & \fk{19} & \fk{+0}.\fk{00} & \fk{0}.\fk{10} & \fk{0}.\fk{15} & \fk{0}.\fk{15} \\
-$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16  & \fk{$\Delta$hCI2+EN2}   & \fk{19} & \fk{-0}.\fk{00} & \fk{0}.\fk{05} & \fk{0}.\fk{08} & \fk{0}.\fk{08} \\
+$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44  & $\Delta$hCI1+EN2   & 19 & +0.02 & 0.17 & 0.29 & 0.29 \\
+$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69  & $\Delta$hCI1.5+EN2 & 19 & +0.00 & 0.10 & 0.15 & 0.15 \\
+$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16  & $\Delta$hCI2+EN2   & 19 & -0.00 & 0.05 & 0.08 & 0.08 \\
 $\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10  & & & & & & \\
@@
 \hline
-sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53  & \fk{sCI1+EN2}           & \fk{19} & \fk{+0}.\fk{23} & \fk{0}.\fk{25} & \fk{0}.\fk{32} & \fk{0}.\fk{21} \\
-$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37  & \fk{$\Delta$sCI1+EN2}   & \fk{19} & \fk{+0}.\fk{06} & \fk{0}.\fk{12} & \fk{0}.\fk{20} & \fk{0}.\fk{19} \\
+sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53  & sCI1+EN2           & 19 & +0.23 & 0.25 & 0.32 & 0.21 \\
+$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37  & $\Delta$sCI1+EN2   & 19 & +0.06 & 0.12 & 0.20 & 0.19 \\
 \end{tabular}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\fk{hCI and sCI plus EN2 correction for excited states}}
+\subsection{hCI and sCI plus EN2 correction for excited states}
 \label{sec:res_E}
@@
 
-\fk{
 The renormalized EN2 correction significantly reduces the errors across most CI models, for both closed- and open-shell systems.
@@
 We have not considered the EN2 correction for the higher-order models.
-}
-
-\fk{
+
 For a given scaling (see Table~\ref{tab:scaling}), whether the bare CI model or the lower-order CI model plus EN2 correction produces the smaller errors depends on the truncation order,
@@
 We recall that the calculation of the EN2 correction has a small associated prefactor, which makes the above $\Delta$hCI+EN2 models attractive alternatives to their higher-order unperturbed $\Delta$hCI counterparts.
-}
-
-\fk{
+
 Among the different CI models, it is worth highlighting the huge improvement observed from $\Delta$hCI1.5 to $\Delta$hCI1.5+EN2.
@@
 In addition, the individual errors for singlets, triplets, Rydberg, and valence states are comparable, lying in the range \SIrange{0.11}{0.15}{\eV}.
-}
-
-\fk{
+
 We found a similar impact of the EN2 correction for $\Delta$CISD and $\Delta$hCI2.
@@
 $\Delta$CISD+EN2 displays the same trends, slightly outperforming $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above, just as discussed above for the unperturbed case.
-}
-
-\fk{
+
 In another contrast to eCI and hCI routes, the EN2 correction has a more limited effect on the sCI route.
@@
 Despite the improvement, the EN2 correction is not enough to render the sCI models competitive.
-}
 
@@
 In this regard, rotations within occupied and virtual orbitals represent yet another factor worth exploring in the future developments of hCI models.
-\fk{hCI also produced reasonable energies for the automerization barrier of cyclobutadiene.
-It remains to be seen how these models perform for other strongly correlated systems, larger molecules, and charge transfer excited states.}
-
-\fk{The accuracy of the CI models is significantly enhanced thanks to the EN2 correction,
+hCI also produced reasonable energies for the automerization barrier of cyclobutadiene.
+It remains to be seen how these models perform for other strongly correlated systems, larger molecules, and charge transfer excited states.
+
+The accuracy of the CI models is significantly enhanced thanks to the EN2 correction,
 the case of $\Delta$hCI1.5 for closed-shells being the most striking, with a MAE plummeting from \SI{2.80}{\eV} to \SI{0.13}{\eV}.
 Low-orders of state-specific CI supplemented with a perturbative correction can thus provide a cost-effective option for the computation of excitation energies.
-}
 
@@
 For the open-shell transitions, the outcome is more encouraging.
-\fk{Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than the eCI or hCI options.}
+Despite the improvement brought about by the EN2 correction, the sCI models remain less accurate and more expensive than the eCI or hCI options.
 The relevance of these sCI models lies in their possible connection with CC methods.
@@
 
-\fk{Definition of the statistical measures;}
+Definition of the statistical measures;
 equilibrium geometries of ethylene and vinyl; details about the fitting of potential energy curves;
-\fk{number of determinants of each CI model;}
+number of determinants of each CI model;
 potential energy curves for \ce{OH}, \ce{CN}, vinyl, and \ce{H7},
@@
 reference excitation energies and corresponding methods;
-\fk{and complementary statistical metrics}.
+and complementary statistical metrics.
 For a subset of 27 excited states for closed-shell molecules and 15 excited states for radicals,
